### 2019-05-11 ex1/ex1.tex
Added exercise 1

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% CN2 Labreport template
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[parskip=full]{scrartcl}

\usepackage{siunitx}  % Provides the \SI{}{} command for typesetting SI units
\usepackage{graphicx} % Required for the inclusion of images
\usepackage{booktabs} % nicer tables
\usepackage[noabbrev]{cleveref} % automatic references
\usepackage{listings} % typeset code

\crefname{lstlisting}{listing}{listings} % for referencing code
\Crefname{lstlisting}{Listing}{Listings} % for referencing code

\usepackage{multicol}
\usepackage{hyperref}

\usepackage[headsepline]{scrlayer-scrpage} % header
\ohead{Anonymization} % right part of header
\ihead{Exercise 1} % left part of header

\lstset{basicstyle=\ttfamily} % monospaced font in listing

%----------------------------------------------------------------------------------------
%	DOCUMENT INFORMATION
%----------------------------------------------------------------------------------------

\begin{document}
\begin{titlepage}
    \centering
    \vspace*{2cm}
    {\Huge \textbf{Security, Privacy and Explainability in Machine Learning}}\\
    SS 2019\\
    \vspace*{1cm}
    {\Large Exercise 1}
    \\\vspace*{3cm}
    {\large 
        \begin{tabular}{l c c}
            Name & Mat.Nummer \\ \hline
            Maximilian Bachl & 01100143 \\
            Alexander Hartl &  01125115
        \end{tabular}
    }
    \\\vspace*{7cm}
    \today
\end{titlepage}

%----------------------------------------------------------------------------------------
%	SECTION 1
%----------------------------------------------------------------------------------------
\section{Dataset Selection}
We selected the ``Airline Customers'' dataset available on kaggle's online platform\footnote{\url{https://www.kaggle.com/diezerg/airline-customers}}, which seems to be taken from an airline company's frequent-flyer program. Since a customer's bonus points presumably show a strong correlation  with the customer's financial wealth, the dataset contains sensitive data.
\subsection{Contained Information}
The dataset contains several, partially redundant, attributes regarding a customer's flight frequency and bonus points.
For demonstration purposes we reduced the attributes to 
\begin{multicols}{2}
\begin{itemize}
\item Gender
\item Age
\item Work Country
\item Work Province
\item Work City
\item First flight date
\item Total flight count
\item Total bonus points
\end{itemize}
\end{multicols}

%----------------------------------------------------------------------------------------
%	SECTION 2
%----------------------------------------------------------------------------------------
\subsection{Quasi-Identifiers}
The combination of quasi-identifying attributes allows a third-party to uniquely identify an individual in the dataset. In our case the set of quasi-identifiers consists of \textbf{Gender, Age and Work Country/Province/City}. Other attributes could contribute to identifying an individual but are usually not known to a third-party.

\section{Anonymization}
We used ARX\footnote{\url{https://arx.deidentifier.org/}} for the anonymization process and deployed several anonymization techniques for the different attributes:
\begin{itemize}
\item Gender: Generalization, suppressing the attribute, if necessary.
\item Age: Microaggregation, reducing the exact age to intervals, if necessary.
\item Work Location: Generalization, using a hierarchy to mask the data, if necessary. For this to work properly, we merged the work location attributes to one attribute in the form $<$Work Country$>$ / $<$Work Province$>$ / $<$Work City$>$ as a preprocessing step, so that more fine-grained information will be suppressed first.
\end{itemize}

We used $k$-Anonymity with $k=3$ and $k=50$ as privacy model. Furthermore, as total bonus points constitutes a sensitive attribute, we used $t$-closeness with $t=0.2$ and ``EMD with ordered distance'' as distance measure.

\section{Discussion}

\begin{figure}[t!]
\centering
\includegraphics[width=\textwidth]{figures/classes.png}
\caption{Exemplary equivalence classes as found by ARX.}
\label{fig:classes}
\end{figure}

Fig.~\ref{fig:paths} shows the anonymization paths that ARX finds for $k=3$ and $k=50$, respectively. To achieve the same level of anonymity, either the gender attribute can be suppressed or the contained information about the work location can be reduced, resulting in the paths shown in Fig.~\ref{fig:paths}.

For different $k$, the paths don't show significant differences. However, aiming at the same level of anonymity, ARX has to suppress more information when selecting a lower $k$.

Fig.~\ref{fig:classes} shows an excerpt of an solution ARX finds for $k=3$. Some of the equivalence classes indeed assume the desired size of 3, while others are substantially larger.

\begin{figure}
\centering
\includegraphics[trim=10cm 9cm 10cm 4.4cm, clip,width=0.4\textwidth]{figures/k=3_paths.png}
\includegraphics[trim=10cm 9cm 10cm 4.4cm, clip,width=0.4\textwidth]{figures/k=50_paths.png}
\caption{Anonymization paths for $k=3$ (left) and $k=50$ (right).}
\label{fig:paths}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

### 2019-05-11 ex1/ex1.tex
Minor rewording/typos

@@
 
-Fig.~\ref{fig:classes} shows an excerpt of an solution ARX finds for $k=3$. Some of the equivalence classes indeed assume the desired size of 3, while others are substantially larger.
+Fig.~\ref{fig:classes} shows an excerpt of a solution ARX finds for $k=3$. Some of the equivalence classes indeed assume the desired minimum size of 3, while others are substantially larger.
 

### 2019-06-28 ex2/learn.py
Add code to eval backdoor

@@ -16,6 +16,8 @@ import os
 import pdp as pdp_module
 import ale as ale_module
 import ice as ice_module
+import collections
+import pickle
 
 def add_backdoor(datum: dict, direction: str) -> dict:
 	datum = datum.copy()
@@ -41,15 +43,16 @@ def add_backdoor(datum: dict, direction: str) -> dict:
 
 parser = argparse.ArgumentParser()
 parser.add_argument('--dataroot', required=True, help='path to dataset')
-parser.add_argument('--batchSize', type=int, default=64, help='input batch size')
+parser.add_argument('--batchSize', type=int, default=128, help='input batch size')
 parser.add_argument('--niter', type=int, default=100, help='number of epochs to train for')
 parser.add_argument('--lr', type=float, default=0.01, help='learning rate')
-parser.add_argument('--fold', type=int, default=0, help='learning rate')
-parser.add_argument('--nFold', type=int, default=3, help='learning rate')
+parser.add_argument('--fold', type=int, default=0, help='fold to use')
+parser.add_argument('--nFold', type=int, default=3, help='total number of folds')
 parser.add_argument('--net', default='', help="path to net (to continue training)")
 parser.add_argument('--function', default='train', help='the function that is going to be called')
 parser.add_argument('--manualSeed', default=0, type=int, help='manual seed')
 parser.add_argument('--backdoor', action='store_true', help='include backdoor')
+parser.add_argument('--normalizationData', default="", type=str, help='normalization data to use')
 
 opt = parser.parse_args()
 print(opt)
@@ -65,47 +68,65 @@ MAX_ROWS = 1_000_000_000
 csv_name = opt.dataroot
 df = pd.read_csv(csv_name, nrows=MAX_ROWS).fillna(0)
 
-del df['flowStartMilliseconds']
-del df['sourceIPAddress']
-del df['destinationIPAddress']
-del df['Attack']
+try:
+	del df['flowStartMilliseconds']
+	del df['sourceIPAddress']
+	del df['destinationIPAddress']
+	del df['Attack']
+except KeyError:
+	pass
 
 features = df.columns[:-1]
-
 print("Rows", df.shape[0])
 
 if opt.backdoor:
-	attack_records = df[df["Label"] == 1].to_dict("records")
+	attack_records = df[df["Label"] == 1].to_dict("records", into=collections.OrderedDict)
[…]

### 2019-06-28 ex2/learn.py
Reintroduce eval function

@@ -183,7 +183,7 @@ current_time = datetime.now().strftime('%b%d_%H-%M-%S')
 
 def get_logdir(fold, n_fold):
 	return os.path.join('runs', current_time + '_' + socket.gethostname() + "_" + str(fold) +"_"+str(n_fold))
-	
+
 def surrogate(predict_fun):
 	train_indices, test_indices = get_nth_split(dataset, opt.nFold, opt.fold)
 
@@ -276,19 +276,23 @@ def predict(test_indices):
 	all_predictions = np.concatenate(all_predictions, axis=0).astype(int)
 	#all_labels = np.concatenate(all_labels, axis=0)
 	return all_predictions
-	
+
 def test_nn():
 	n_fold = opt.nFold
 	fold = opt.fold
 
 	_, test_indices = get_nth_split(dataset, n_fold, fold)
-	
+
+	eval_nn(test_indices)
+
+def eval_nn(test_indices=None):
+	if test_indices is None:
+		test_indices = list(range(len(dataset)))
+
 	all_predictions = predict(test_indices)
 	all_labels = y[test_indices,0]
-
 	print("accuracy", np.mean(all_predictions==all_labels))
-	print (classification_report(all_labels, all_predictions))
-
+	print(classification_report(all_labels, all_predictions))
 
 def pdp_nn():
 	# all_loader = torch.utils.data.DataLoader(dataset, batch_size=1, shuffle=True)
@@ -328,15 +332,15 @@ def surrogate_nn():
 
 def train_rf():
 	pickle.dump(rf, open('%s.rfmodel' % get_logdir(opt.fold, opt.nFold), 'wb'))
-	
+
 def test_rf():
 	_, test_indices = get_nth_split(dataset, opt.nFold, opt.fold)
-	
+
 	predictions = rf.predict (x[test_indices,:])
 
 	print ('Accuracy:', np.mean(y[test_indices,0]==predictions))
 	print (classification_report(y[test_indices,0], predictions))
-	
+
 def pdp_rf():
 	pdp_module.pdp(x, rf.predict_proba, features, means=means, stds=stds, resolution=1000, n_data=1000, suffix=suffix)
 
@@ -345,14 +349,14 @@ def ale_rf():
 
 def ice_rf():
 	ice_module.ice(x, rf.predict_proba, features, means=means, stds=stds, resolution=1000, n_data=100, suffix=suffix)
-	
+
 def surrogate_rf():
[…]

### 2019-06-28 ex2/report/report.tex

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf]{acmart}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%% Rights management information.  This information is sent to you
%% when you complete the rights form.  These commands have SAMPLE
%% values in them; it is your responsibility as an author to replace
%% the commands and values with those provided to you when you
%% complete the rights form.
\setcopyright{acmcopyright}
\copyrightyear{2018}
\acmYear{2018}
\acmDOI{10.1145/1122445.1122456}

%% These commands are for a PROCEEDINGS abstract or paper.
\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
  June 03--05, 2018, Woodstock, NY}
\acmPrice{15.00}
\acmISBN{978-1-4503-9999-9/18/06}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{The Name of the Title is Hope}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Ben Trovato}
\authornote{Both authors contributed equally to this research.}
\email{trovato@corporation.com}
\orcid{1234-5678-9012}
\author{G.K.M. Tobin}
\authornotemark[1]
\email{webmaster@marysville-ohio.com}
\affiliation{%
  \institution{Institute for Clarity in Documentation}
  \streetaddress{P.O. Box 1212}
  \city{Dublin}
  \state{Ohio}
  \postcode{43017-6221}
}

\author{Lars Th{\o}rv{\"a}ld}
\affiliation{%
  \institution{The Th{\o}rv{\"a}ld Group}
  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
  \city{Hekla}
  \country{Iceland}}
\email{larst@affiliation.org}

\author{Valerie B\'eranger}
\affiliation{%
  \institution{Inria Paris-Rocquencourt}
  \city{Rocquencourt}
  \country{France}
}

\author{Aparna Patel}
\affiliation{%
 \institution{Rajiv Gandhi University}
 \streetaddress{Rono-Hills}
 \city{Doimukh}
 \state{Arunachal Pradesh}
 \country{India}}

\author{Huifen Chan}
\affiliation{%
  \institution{Tsinghua University}
  \streetaddress{30 Shuangqing Rd}
  \city{Haidian Qu}
  \state{Beijing Shi}
  \country{China}}

\author{Charles Palmer}
\affiliation{%
  \institution{Palmer Research Laboratories}
  \streetaddress{8600 Datapoint Drive}
  \city{San Antonio}
  \state{Texas}
  \postcode{78229}}
\email{cpalmer@prl.com}

\author{John Smith}
\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
\email{jsmith@affiliation.org}

\author{Julius P. Kumquat}
\affiliation{\institution{The Kumquat Consortium}}
\email{jpkumquat@consortium.net}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Trovato and Tobin, et al.}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  A clear and well-documented \LaTeX\ document is presented as an
  article formatted for publication by ACM in a conference proceedings
  or journal publication. Based on the ``acmart'' document class, this
  article presents and explains many of the common variations, as well
  as many of the formatting elements an author may use in the
  preparation of the documentation of their work.
\end{abstract}

%%
%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
%% Please copy and paste the code instead of the example below.
%%
\begin{CCSXML}
<ccs2012>
 <concept>
  <concept_id>10010520.10010553.10010562</concept_id>
  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
  <concept_significance>500</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010575.10010755</concept_id>
  <concept_desc>Computer systems organization~Redundancy</concept_desc>
  <concept_significance>300</concept_significance>
 </concept>
 <concept>
  <concept_id>10010520.10010553.10010554</concept_id>
  <concept_desc>Computer systems organization~Robotics</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
 <concept>
  <concept_id>10003033.10003083.10003095</concept_id>
  <concept_desc>Networks~Network reliability</concept_desc>
  <concept_significance>100</concept_significance>
 </concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded systems}
\ccsdesc[300]{Computer systems organization~Redundancy}
\ccsdesc{Computer systems organization~Robotics}
\ccsdesc[100]{Networks~Network reliability}

%%
%% Keywords. The author(s) should pick words that accurately describe
%% the work being presented. Separate the keywords with commas.
\keywords{datasets, neural networks, gaze detection, text tagging}

%%
%% This command processes the author and affiliation and title
%% information and builds the first part of the formatted document.
\maketitle

\section{Introduction}
The detection of attacks in data networks is a fundamental task in data security. Due to the considerable amount of data which has to be analyzed the use of machine learning techniques for this purpose seems natural and is increasingly deployed.

For research the invention and assessment of techniques for network anomaly detection poses many challenges. Particularly, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection. For the training of models usually datasets are used which have been generated artificially in a controlled test environment.

As a downside of this approach, it is unclear whether a machine learning model learn to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.

For a well-performing network anomaly detection technique it is therefore of utmost importance to study which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.

In this document, we redo parts of a recent paper which bases on the CIC-IDS-2017 dataset for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection. We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.

Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data. Finally, we apply the same explainability methods to the backdoored model and show how such attack attempts might be recognized before any harm is done.

\section{Machine Learning Approaches for Traffic Classification}
\subsection{Deep Learning}
\subsection{Random Forests}

\section{Explainability Plots}
\subsection{Partial Dependence Plots}
Partial Dependence Plots (PDP) visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.

\subsection{Individual Conditional Expectation}
The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost. In our case

\subsection{Accumulated Local Effects}
Due to feature dependence it is very likely that in the feature space areas exist which have a very low probability to occur. Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas. This poses a problem when considering these predictions for computing PDPs.

In an attempt to overcome this problem, it is possible to only consider samples which are likely to occur for certain feature values, i.e. to consider the conditional distribution of remaining features, for computing explainability graphs. This is the concept for Accumulated Local Effects (ALE) plots.

When computing ALE plots, we experienced the problem of empty intervals. If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values.

For this reason, we modified this definition to instead use the closest 10 samples to the interval's center for estimating the distribution.

\subsection{Interpretation}

\section{Logistic Regression as Surrogate Model}

\section{Implementing a Backdoor}

\section{Conclusions}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2019-06-28 ex2/report/report.tex
Change report

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf]{acmart}
+\documentclass[sigconf,nonacm]{acmart}
+
+\settopmatter{printacmref=false} % Removes citation information below abstract
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
 
@@
     \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-%% Rights management information.  This information is sent to you
-%% when you complete the rights form.  These commands have SAMPLE
-%% values in them; it is your responsibility as an author to replace
-%% the commands and values with those provided to you when you
-%% complete the rights form.
-\setcopyright{acmcopyright}
-\copyrightyear{2018}
-\acmYear{2018}
-\acmDOI{10.1145/1122445.1122456}
-
-%% These commands are for a PROCEEDINGS abstract or paper.
-\acmConference[Woodstock '18]{Woodstock '18: ACM Symposium on Neural
-  Gaze Detection}{June 03--05, 2018}{Woodstock, NY}
-\acmBooktitle{Woodstock '18: ACM Symposium on Neural Gaze Detection,
-  June 03--05, 2018, Woodstock, NY}
-\acmPrice{15.00}
-\acmISBN{978-1-4503-9999-9/18/06}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{The Name of the Title is Hope}
+\title{Assignment 2}
 
@@
 %% used to denote shared contribution to the research.
-\author{Ben Trovato}
+\author{Alexander Hartl}
 \authornote{Both authors contributed equally to this research.}
-\email{trovato@corporation.com}
-\orcid{1234-5678-9012}
-\author{G.K.M. Tobin}
+%\email{alexander.hartl@tuwien.ac.at}
+
+\author{Maximilian Bachl}
+%\authornote[1]{Both authors contributed equally to this research.}
 \authornotemark[1]
-\email{webmaster@marysville-ohio.com}
-\affiliation{%
-  \institution{Institute for Clarity in Documentation}
-  \streetaddress{P.O. Box 1212}
-  \city{Dublin}
-  \state{Ohio}
-  \postcode{43017-6221}
-}
-
-\author{Lars Th{\o}rv{\"a}ld}
-\affiliation{%
-  \institution{The Th{\o}rv{\"a}ld Group}
-  \streetaddress{1 Th{\o}rv{\"a}ld Circle}
-  \city{Hekla}
-  \country{Iceland}}
-\email{larst@affiliation.org}
-
-\author{Valerie B\'eranger}
-\affiliation{%
-  \institution{Inria Paris-Rocquencourt}
-  \city{Rocquencourt}
-  \country{France}
-}
-
-\author{Aparna Patel}
-\affiliation{%
- \institution{Rajiv Gandhi University}
- \streetaddress{Rono-Hills}
- \city{Doimukh}
- \state{Arunachal Pradesh}
- \country{India}}
-
-\author{Huifen Chan}
-\affiliation{%
-  \institution{Tsinghua University}
-  \streetaddress{30 Shuangqing Rd}
-  \city{Haidian Qu}
-  \state{Beijing Shi}
-  \country{China}}
-
-\author{Charles Palmer}
-\affiliation{%
-  \institution{Palmer Research Laboratories}
-  \streetaddress{8600 Datapoint Drive}
-  \city{San Antonio}
-  \state{Texas}
-  \postcode{78229}}
-\email{cpalmer@prl.com}
-
-\author{John Smith}
-\affiliation{\institution{The Th{\o}rv{\"a}ld Group}}
-\email{jsmith@affiliation.org}
-
-\author{Julius P. Kumquat}
-\affiliation{\institution{The Kumquat Consortium}}
-\email{jpkumquat@consortium.net}
+%\email{maximilian.bachl@gmail.com}
 
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Trovato and Tobin, et al.}
+\renewcommand{\shortauthors}{Hartl and Bachl}
 
@@
 %% article.
-\begin{abstract}
-  A clear and well-documented \LaTeX\ document is presented as an
-  article formatted for publication by ACM in a conference proceedings
-  or journal publication. Based on the ``acmart'' document class, this
-  article presents and explains many of the common variations, as well
-  as many of the formatting elements an author may use in the
-  preparation of the documentation of their work.
-\end{abstract}
+%\begin{abstract}
+%  A clear and well-documented \LaTeX\ document is presented as an
+%  article formatted for publication by ACM in a conference proceedings
+%  or journal publication. Based on the ``acmart'' document class, this
+%  article presents and explains many of the common variations, as well
+%  as many of the formatting elements an author may use in the
+%  preparation of the documentation of their work.
+%\end{abstract}
 
-%%
-%% The code below is generated by the tool at http://dl.acm.org/ccs.cfm.
-%% Please copy and paste the code instead of the example below.
-%%
-\begin{CCSXML}
-<ccs2012>
- <concept>
-  <concept_id>10010520.10010553.10010562</concept_id>
-  <concept_desc>Computer systems organization~Embedded systems</concept_desc>
-  <concept_significance>500</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010575.10010755</concept_id>
-  <concept_desc>Computer systems organization~Redundancy</concept_desc>
-  <concept_significance>300</concept_significance>
- </concept>
- <concept>
-  <concept_id>10010520.10010553.10010554</concept_id>
-  <concept_desc>Computer systems organization~Robotics</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
- <concept>
-  <concept_id>10003033.10003083.10003095</concept_id>
-  <concept_desc>Networks~Network reliability</concept_desc>
-  <concept_significance>100</concept_significance>
- </concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Computer systems organization~Embedded systems}
-\ccsdesc[300]{Computer systems organization~Redundancy}
-\ccsdesc{Computer systems organization~Robotics}
-\ccsdesc[100]{Networks~Network reliability}
-
-%%
-%% Keywords. The author(s) should pick words that accurately describe
-%% the work being presented. Separate the keywords with commas.
-\keywords{datasets, neural networks, gaze detection, text tagging}
-
-%%
-%% This command processes the author and affiliation and title
-%% information and builds the first part of the formatted document.
 \maketitle

### 2019-06-28 ex2/report/report.tex
Changed report

@@
 
-In this document, we redo parts of a recent paper which bases on the CIC-IDS-2017 dataset for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
+In this document, we redo parts of a recent paper~\cite{fares} which bases on the CIC-IDS-2017 dataset for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
 We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
@@
 
-\section{Machine Learning Approaches for Traffic Classification}
+\section{Machine Learning Approaches for Anomaly Detection} \label{sec:ml_approaches}
 \subsection{Deep Learning}
@@
 
+\subsection{Performance Results}
 \section{Explainability Plots}
+Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
+In this section, we introduce the graphs which we will use for interpreting our models' outcomes. 
 \subsection{Partial Dependence Plots}
 Partial Dependence Plots (PDP) visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
+
+If we denote by $f^{i}$ the marginal distribution of the feature space which retains all features except the $i$th feature $x_i$, the PDP for $x_i$ feature can be expressed as
+\begin{equation}
+\text{PDP}_i(x) = \mathbb E_{\boldsymbol X}\left(Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n)\right) % \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},x,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
+\end{equation}
+Empirically, we can approximate the distribution of the feature space by the observed samples.
+Hence, at a given point $x$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.
 
@@
 to consider the conditional distribution of remaining features, for computing explainability graphs.
-This is the concept for Accumulated Local Effects (ALE) plots.
+This results the concept for Accumulated Local Effects (ALE) plots.
+
+For a feature $X_i$ the ALE plot ALE$_i(x)$ can be defined differentially as
+
+\begin{equation}
+\frac{d}{dx} \text{ALE}_i (x) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_x} Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n) | X_i=x\right)
+\end{equation}
+
+To combat ambiguity of this definition, we force ALE$_i(x)$ to have zero mean on the domain of $X_i$.
+
+For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx x$.
 
@@
 \section{Logistic Regression as Surrogate Model}
+Another possibility for explaining the models' outcomes is training an interpretable surrogate model to the predictions that the original non-interpretable model makes.
+We trained a logistic regression classifier to the predictions our models made.
+Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original lables, respectively.
+As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}.
+The fact that results for random forests are very similar to original data is due to the good detection performance of random forests.
+Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.
 
+ Table~\ref{tab:logreg_coeff} displays the coefficients for 
+\subsection{Interpretation}
 \section{Implementing a Backdoor}
 
+\section{Misclassification Examples}
+
 \section{Conclusions}
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{bibliography}
 

### 2019-06-28 ex2/do.sh
Added model weights

@@ -3,22 +3,22 @@
 DATAROOT=CAIA_backdoor.csv
 
 for i in `seq 0 2`; do
-	./learn.py --dataroot "$DATAROOT" --method=rf --function pdp --fold $i
-	./learn.py --dataroot "$DATAROOT" --method=rf --function ale --fold $i
-	./learn.py --dataroot "$DATAROOT" --backdoor --method=rf --function pdp --fold $i
-	./learn.py --dataroot "$DATAROOT" --backdoor --method=rf --function ale --fold $i
+	./learn.py --dataroot "$DATAROOT" --method=rf --net runs/non-bd/*${i}_3.rfmodel --function pdp --fold $i
+	./learn.py --dataroot "$DATAROOT" --method=rf --net runs/non-bd/*${i}_3.rfmodel --function ale --fold $i
+	./learn.py --dataroot "$DATAROOT" --backdoor --method=rf --net runs/bd/*${i}_3.rfmodel --function pdp --fold $i
+	./learn.py --dataroot "$DATAROOT" --backdoor --method=rf --net runs/bd/*${i}_3.rfmodel --function ale --fold $i
 	
-	#./learn.py --dataroot "$DATAROOT" --method=nn --net runs/Jun24*${i}_3/*.pth --function pdp --fold $i
-	#./learn.py --dataroot "$DATAROOT" --method=nn --net runs/Jun24*${i}_3/*.pth --function ale --fold $i
-	./learn.py --dataroot "$DATAROOT" --backdoor --method=nn --net runs/Jun26*${i}_3/*.pth --function pdp --fold $i
-	./learn.py --dataroot "$DATAROOT" --backdoor --method=nn --net runs/Jun26*${i}_3/*.pth --function ale --fold $i
+	./learn.py --dataroot "$DATAROOT" --method=nn --net runs/Jun28_17*${i}_3/*.pth --function pdp --fold $i
+	./learn.py --dataroot "$DATAROOT" --method=nn --net runs/Jun28_17*${i}_3/*.pth --function ale --fold $i
+	./learn.py --dataroot "$DATAROOT" --backdoor --method=nn --net runs/Jun28_19*${i}_3/*.pth --function pdp --fold $i
+	./learn.py --dataroot "$DATAROOT" --backdoor --method=nn --net runs/Jun28_19*${i}_3/*.pth --function ale --fold $i
 done
 
[…]

### 2019-06-29 ex2/report/report.tex
Updated report

@@
 
-If we denote by $f^{i}$ the marginal distribution of the feature space which retains all features except the $i$th feature $x_i$, the PDP for $x_i$ feature can be expressed as
+If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the PDP for the $i$th feature $X_i$ can be expressed as
 \begin{equation}
-\text{PDP}_i(x) = \mathbb E_{\boldsymbol X}\left(Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n)\right) % \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},x,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
+\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right) .
+% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
 \end{equation}
-Empirically, we can approximate the distribution of the feature space by the observed samples.
-Hence, at a given point $x$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.
+Empirically, we can approximate the distribution of the feature space by the distribution of observed samples.
+Hence, at a given point $x,$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.
 
 \subsection{Individual Conditional Expectation}
-The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost. In our case
+The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
+In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well.
+Hence, by averaging information gets lost and wrong explanations might be the consequence.
+
+For Individual Conditional Expectation (ICE) plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples, i.e.
+
+\begin{equation}
+\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
+\end{equation}
 
@@
 
-For a feature $X_i$ the ALE plot ALE$_i(x)$ can be defined differentially as
+For the $i$th feature $X_i$, the ALE plot ALE$_i(y)$ can be defined differentially as
 
 \begin{equation}
-\frac{d}{dx} \text{ALE}_i (x) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_x} Predict(X_1,\ldots,X_{i-1},x,X_{i+1},\ldots X_n) | X_i=x\right)
+\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_y} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) | X_i=y\right)
 \end{equation}
@@
 
- Table~\ref{tab:logreg_coeff} displays the coefficients for 
+ Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
 \subsection{Interpretation}
 \section{Implementing a Backdoor}
+Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
+In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
+
+If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
+
+For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the traning models.
+
+Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
+On the one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+
+In our case, we used the Time-to-Live (TTL) value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
+The TTL is used for mitigating problems due to routing loops in IP networks.
+In our situation, it is reasonable to assume that the TTL field remains constant for all packets in one traffic flow.
+In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the TTL value.
+
+Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.
+The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
 

### 2019-06-29 ex2/learn.py
Save surrogate coefficients

@@ -185,6 +185,7 @@ def get_logdir(fold, n_fold):
 	return os.path.join('runs', current_time + '_' + socket.gethostname() + "_" + str(fold) +"_"+str(n_fold))
 
 def surrogate(predict_fun):
+	os.makedirs('surrogate', exist_ok=True)
 	train_indices, test_indices = get_nth_split(dataset, opt.nFold, opt.fold)
 
 	logreg = LogisticRegression(solver='liblinear')
@@ -199,6 +200,7 @@ def surrogate(predict_fun):
 	print (classification_report(y_true, predictions))
 
 	print ("Coefficients:", logreg.coef_)
+	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate/logreg_pred%s.csv' % suffix)
 
 
 	logreg = LogisticRegression(solver='liblinear')
@@ -213,6 +215,7 @@ def surrogate(predict_fun):
 	print (classification_report(y_true, predictions))
 
 	print ("Coefficients:", logreg.coef_)
+	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate/logreg_real%s.csv' % suffix)
 
 
 

### 2019-06-29 ex2/report/report.tex
Updated report

@@
 
-In this document, we redo parts of a recent paper~\cite{fares} which bases on the CIC-IDS-2017 dataset for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
+% TODO: think Fares didnt use this
+In this document, we redo parts of a recent paper~\cite{fares} which bases on the CIC-IDS-2017 dataset \cite{Sharafaldin2018} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
 We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
@@
 \section{Machine Learning Approaches for Anomaly Detection} \label{sec:ml_approaches}
+We used the CIC-IDS-2017~\cite{Sharafaldin2018} for our evaluations.
+As a first preprocessing step we extracted flow features from the dataset.
+We used features specificed by the CAIA feature vector~\cite{Williams2006} and slightly enhanced the included features.
+Table~\ref{tab:features} lists the features that are included for each flow.
+
+We used Z-score normalization to process the data and 3-fold cross validation for evaluating the models' performances.
+
+\begin{table*}
+\renewcommand*{\arraystretch}{1.2} % TODO: better order
+\caption{Flow features used for attack detection.}
+\label{tab:features}
+\begin{tabular}{l l l l l} \toprule
+dstBytes	&	max\_srcPktLength	&	min\_dstPktIAT	&	srcPkts	&	\#dstTCPflag:ack	\\
+dstPkts	&	max\_srcTTL	&	min\_dstPktLength	&	srcPort	&	\#dstTCPflag:cwr	\\
+dstPort	&	mean\_dstPktIAT	&	min\_dstTTL	&	stdev\_dstPktIAT	&	\#dstTCPflag:fin	\\
+duration	&	mean\_dstPktLength	&	min\_srcPktIAT	&	stdev\_dstPktLength	&	\#dstTCPflag:syn	\\
+max\_dstPktIAT	&	mean\_dstTTL	&	min\_srcPktLength	&	stdev\_dstTTL	&	\#srcTCPflag:ack	\\
+max\_dstPktLength	&	mean\_srcPktIAT	&	min\_srcTTL	&	stdev\_srcPktIAT	&	\#srcTCPflag:cwr	\\
+max\_dstTTL	&	mean\_srcPktLength	&	protocol	&	stdev\_srcPktLength	&	\#srcTCPflag:fin	\\
+max\_srcPktIAT	&	mean\_srcTTL	&	srcBytes	&	stdev\_srcTTL	&	\#srcTCPflag:syn	\\
+ \bottomrule
+\end{tabular}
+\end{table*}
+
 \subsection{Deep Learning}
 \subsection{Random Forests}
+In addition to deep learning we trained a random forest classifier to detect attacks in network traffic.
 
@@
 \subsection{Partial Dependence Plots}
-Partial Dependence Plots (PDP) visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
+Partial Dependence Plots (PDP) were proposed in \cite{friedman2001greedy} and visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
 
@@
 \end{equation}
-Empirically, we can approximate the distribution of the feature space by the distribution of observed samples.
-Hence, at a given point $x,$ the PDP can be found by setting the corresponding values of all samples in the dataset to $x$ and averaging over the predictions of the resulting modified dataset.
+Empirically, we can approximate the distribution of the feature space using the distribution of observed samples.
+Hence, at a given point $y,$ the PDP can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
 
@@
 
-For Individual Conditional Expectation (ICE) plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples, i.e.
+For this reason, Individual Conditional Expectation (ICE) plots were proposed \cite{goldstein2015peeking}.
+For ICE plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples directly, i.e.
+for a particular sample $\boldsymbol x$ we can compute an ICE plot as
 
@@
 \begin{equation}
-\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{d_y} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) | X_i=y\right)
+\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dy} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) | X_i=y\right)
 \end{equation}
@@
 
-For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx x$.
+For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx y$.
 

### 2019-06-29 ex2/report/report.tex

@@
 
-As a downside of this approach, it is unclear whether a machine learning model learn to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
+As a downside of this approach, it is unclear whether a machine learning model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
 

### 2019-06-29 ex2/report/report.tex
Updated report

--- 2019-06-29 ex2/report/report.tex
@@
 \documentclass[sigconf,nonacm]{acmart}
+
+\usepackage{multirow}
 
@@
 %% used to denote shared contribution to the research.
-\author{Alexander Hartl}
+\author{Alexander Hartl (e01125115)}
 \authornote{Both authors contributed equally to this research.}
@@
 
-\author{Maximilian Bachl}
+\author{Maximilian Bachl (e01100143)}
 %\authornote[1]{Both authors contributed equally to this research.}
@@
 
-\section{Machine Learning Approaches for Anomaly Detection} \label{sec:ml_approaches}
+\section{Anomaly Detection Approaches} \label{sec:ml_approaches}
 We used the CIC-IDS-2017~\cite{Sharafaldin2018} for our evaluations.
@@
 \subsection{Performance Results}
+Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are constistant with previous work like, e.g., \cite{fares}.
+
+Notably, random forests exhibit a performance which intuitively seems hard or even impossible to achieve.  
+This observation motivates taking a closer look at the models' decisions to find out if they  learn to decide based on artifacts from the dataset generation process.
 \section{Explainability Plots}
 Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
-In this section, we introduce the graphs which we will use for interpreting our models' outcomes. 
+In this section, we introduce the graphs which we will use for interpreting our models' outcomes which are not available in real-world situations.
 \subsection{Partial Dependence Plots}
@@
  Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
+
+ \begin{table*}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
+\begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+    \#srcTCPflag:ack &  -28.431 $\pm$ 1.129 &  -19.570 $\pm$ 6.735 \\
+             srcPkts &  -26.661 $\pm$ 0.994 &  -18.099 $\pm$ 6.500 \\
+     mean\_dstPktIAT &   16.232 $\pm$ 1.091 &   18.778 $\pm$ 1.396 \\
+     mean\_srcPktIAT &  -12.140 $\pm$ 1.144 &  -15.672 $\pm$ 1.289 \\
+            dstBytes &  -11.305 $\pm$ 0.361 &   -1.878 $\pm$ 6.300 \\
+             dstPkts &   -9.750 $\pm$ 0.158 &   -3.987 $\pm$ 1.972 \\
+      min\_dstPktIAT &   -9.216 $\pm$ 1.157 &  -12.533 $\pm$ 1.477 \\
+            srcBytes &   -8.994 $\pm$ 4.744 &   -5.373 $\pm$ 3.870 \\
+    \#dstTCPflag:ack &   -8.292 $\pm$ 0.227 &   -2.616 $\pm$ 1.952 \\
+       stdev\_dstTTL &   -6.934 $\pm$ 0.231 &   -0.241 $\pm$ 0.014 \\
+%   min\_srcPktLength &   -4.519 $\pm$ 0.314 &   -2.891 $\pm$ 0.180 \\
+%         min\_srcTTL &   -4.192 $\pm$ 0.646 &   -3.619 $\pm$ 0.432 \\
+% stdev\_srcPktLength &   -4.068 $\pm$ 0.342 &   -2.231 $\pm$ 0.439 \\
+% stdev\_dstPktLength &    3.466 $\pm$ 0.085 &    3.289 $\pm$ 0.253 \\
+%    stdev\_srcPktIAT &    3.102 $\pm$ 0.407 &    4.743 $\pm$ 0.475 \\
+%            protocol &   -3.043 $\pm$ 0.028 &   -2.157 $\pm$ 0.012 \\
+%    stdev\_dstPktIAT &   -2.842 $\pm$ 0.375 &   -4.545 $\pm$ 0.503 \\
+%  mean\_srcPktLength &    2.339 $\pm$ 0.435 &    1.777 $\pm$ 0.261 \\
+%         max\_srcTTL &    2.153 $\pm$ 0.820 &    2.448 $\pm$ 0.249 \\
+%    \#srcTCPflag:fin &   -2.117 $\pm$ 0.066 &   -1.876 $\pm$ 0.045 \\
+%      max\_srcPktIAT &   -2.044 $\pm$ 0.271 &   -3.821 $\pm$ 0.415 \\
+%    \#srcTCPflag:syn &    1.738 $\pm$ 0.010 &    1.521 $\pm$ 0.005 \\
+%   max\_dstPktLength &   -1.256 $\pm$ 0.066 &   -1.187 $\pm$ 0.137 \\
+%  mean\_dstPktLength &    1.200 $\pm$ 0.019 &    1.008 $\pm$ 0.085 \\
+%        mean\_srcTTL &   -0.858 $\pm$ 0.533 &   -0.709 $\pm$ 0.417 \\
+%    \#dstTCPflag:syn &   -0.778 $\pm$ 0.021 &   -1.020 $\pm$ 0.011 \\
+%         max\_dstTTL &   -0.695 $\pm$ 0.028 &   -1.369 $\pm$ 0.052 \\
+%             dstPort &   -0.679 $\pm$ 0.002 &   -0.633 $\pm$ 0.004 \\
+%      min\_srcPktIAT &   -0.111 $\pm$ 0.977 &   -0.762 $\pm$ 2.837 \\
+%   max\_srcPktLength &    0.498 $\pm$ 0.048 &   -0.475 $\pm$ 0.432 \\
+%         min\_dstTTL &    0.468 $\pm$ 0.023 &   -0.197 $\pm$ 0.109 \\
+%            duration &   -0.415 $\pm$ 0.007 &   -0.386 $\pm$ 0.042 \\
+%             srcPort &    0.345 $\pm$ 0.007 &    0.350 $\pm$ 0.005 \\
+%   min\_dstPktLength &    0.268 $\pm$ 0.009 &    0.142 $\pm$ 0.021 \\
+%      max\_dstPktIAT &    0.225 $\pm$ 0.267 &    2.307 $\pm$ 0.432 \\
+%    \#dstTCPflag:fin &   -0.103 $\pm$ 0.024 &   -0.112 $\pm$ 0.017 \\
+%        mean\_dstTTL &   -0.069 $\pm$ 0.016 &    1.436 $\pm$ 0.116 \\
+%    \#srcTCPflag:cwr &    0.059 $\pm$ 0.034 &    0.052 $\pm$ 0.013 \\
+%       stdev\_srcTTL &   -0.051 $\pm$ 0.025 &   -0.058 $\pm$ 0.012 \\
+%    \#dstTCPflag:cwr &    0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\
+
+\bottomrule
+\end{tabular} \qquad
+ \begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+    \#srcTCPflag:ack &  -19.480 $\pm$ 5.095 &  -19.570 $\pm$ 6.735 \\
+     mean\_dstPktIAT &   18.721 $\pm$ 1.613 &   18.778 $\pm$ 1.396 \\
+             srcPkts &  -17.963 $\pm$ 5.079 &  -18.099 $\pm$ 6.500 \\
+     mean\_srcPktIAT &  -15.604 $\pm$ 1.525 &  -15.672 $\pm$ 1.289 \\
+      min\_dstPktIAT &  -12.395 $\pm$ 1.770 &  -12.533 $\pm$ 1.477 \\
+            srcBytes &   -5.106 $\pm$ 4.270 &   -5.373 $\pm$ 3.870 \\
+    stdev\_srcPktIAT &    4.688 $\pm$ 0.583 &    4.743 $\pm$ 0.475 \\
+    stdev\_dstPktIAT &   -4.482 $\pm$ 0.616 &   -4.545 $\pm$ 0.503 \\
+            dstBytes &   -1.389 $\pm$ 6.363 &   -1.878 $\pm$ 6.300 \\
+             dstPkts &   -3.889 $\pm$ 2.178 &   -3.987 $\pm$ 1.972 \\
+%      max\_srcPktIAT &   -3.771 $\pm$ 0.497 &   -3.821 $\pm$ 0.415 \\
+%         min\_srcTTL &   -3.731 $\pm$ 0.448 &   -3.619 $\pm$ 0.432 \\
+% stdev\_dstPktLength &    3.320 $\pm$ 0.254 &    3.289 $\pm$ 0.253 \\
+%   min\_srcPktLength &   -2.935 $\pm$ 0.192 &   -2.891 $\pm$ 0.180 \\
+%         max\_srcTTL &    2.488 $\pm$ 0.106 &    2.448 $\pm$ 0.249 \\
+%    \#dstTCPflag:ack &   -2.481 $\pm$ 2.258 &   -2.616 $\pm$ 1.952 \\
+%      max\_dstPktIAT &    2.233 $\pm$ 0.533 &    2.307 $\pm$ 0.432 \\
+% stdev\_srcPktLength &   -2.224 $\pm$ 0.447 &   -2.231 $\pm$ 0.439 \\
+%            protocol &   -2.181 $\pm$ 0.013 &   -2.157 $\pm$ 0.012 \\
+%    \#srcTCPflag:fin &   -1.883 $\pm$ 0.045 &   -1.876 $\pm$ 0.045 \\
+%  mean\_srcPktLength &    1.759 $\pm$ 0.280 &    1.777 $\pm$ 0.261 \\
+%      min\_srcPktIAT &    0.082 $\pm$ 2.163 &   -0.762 $\pm$ 2.837 \\
+%    \#srcTCPflag:syn &    1.520 $\pm$ 0.007 &    1.521 $\pm$ 0.005 \\
+%        mean\_dstTTL &    1.439 $\pm$ 0.127 &    1.436 $\pm$ 0.116 \\
+%         max\_dstTTL &   -1.359 $\pm$ 0.033 &   -1.369 $\pm$ 0.052 \\
+%   max\_dstPktLength &   -1.211 $\pm$ 0.135 &   -1.187 $\pm$ 0.137 \\
+%    \#dstTCPflag:syn &   -1.017 $\pm$ 0.003 &   -1.020 $\pm$ 0.011 \\
+%  mean\_dstPktLength &    1.007 $\pm$ 0.086 &    1.008 $\pm$ 0.085 \\
+%        mean\_srcTTL &   -0.656 $\pm$ 0.481 &   -0.709 $\pm$ 0.417 \\
+%             dstPort &   -0.635 $\pm$ 0.003 &   -0.633 $\pm$ 0.004 \\
+%   max\_srcPktLength &   -0.483 $\pm$ 0.433 &   -0.475 $\pm$ 0.432 \\
+%            duration &   -0.384 $\pm$ 0.046 &   -0.386 $\pm$ 0.042 \\
+%             srcPort &    0.348 $\pm$ 0.005 &    0.350 $\pm$ 0.005 \\
+%       stdev\_dstTTL &   -0.245 $\pm$ 0.019 &   -0.241 $\pm$ 0.014 \\
+%         min\_dstTTL &   -0.208 $\pm$ 0.117 &   -0.197 $\pm$ 0.109 \\
+%   min\_dstPktLength &    0.152 $\pm$ 0.023 &    0.142 $\pm$ 0.021 \\
+%    \#dstTCPflag:fin &   -0.112 $\pm$ 0.017 &   -0.112 $\pm$ 0.017 \\
+%       stdev\_srcTTL &   -0.061 $\pm$ 0.010 &   -0.058 $\pm$ 0.012 \\
+%    \#srcTCPflag:cwr &    0.052 $\pm$ 0.013 &    0.052 $\pm$ 0.013 \\
+%    \#dstTCPflag:cwr &   -0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
 \subsection{Interpretation}
@@
 
+\begin{table*}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
+\begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+         max\_srcTTL &  -73.558 $\pm$ 62.431 &  -42.423 $\pm$ 63.108 \\
+       stdev\_dstTTL &  -68.440 $\pm$ 29.723 &  -62.500 $\pm$ 6.157 \\
+        mean\_srcTTL &   47.344 $\pm$ 40.989 &   26.793 $\pm$ 41.001 \\
+         min\_srcTTL &   24.441 $\pm$ 21.392 &   14.281 $\pm$ 22.150 \\
+            srcBytes &  -23.660 $\pm$ 19.034 &   -7.455 $\pm$ 4.132 \\
+    \#srcTCPflag:ack &  -22.746 $\pm$ 18.847 &  -12.779 $\pm$ 15.650 \\
+             srcPkts &  -21.032 $\pm$ 17.382 &  -11.617 $\pm$ 13.848 \\
+     mean\_dstPktIAT &   11.310 $\pm$ 7.515 &    9.278 $\pm$ 4.046 \\
+      min\_dstPktIAT &  -10.318 $\pm$ 7.652 &   -7.674 $\pm$ 4.117 \\
+     mean\_srcPktIAT &   -8.914 $\pm$ 6.918 &   -7.396 $\pm$ 4.275 \\
+%            dstBytes &   -7.656 $\pm$ 6.163 &   -1.338 $\pm$ 0.685 \\
+%       stdev\_srcTTL &   -7.585 $\pm$ 1.529 &   -9.045 $\pm$ 2.874 \\
+%         max\_dstTTL &   -6.360 $\pm$ 2.801 &   -8.025 $\pm$ 4.076 \\
+%      min\_srcPktIAT &    5.725 $\pm$ 5.259 &    2.023 $\pm$ 4.868 \\
+%   min\_srcPktLength &   -5.190 $\pm$ 0.911 &   -4.658 $\pm$ 0.205 \\
+%             dstPkts &   -4.509 $\pm$ 3.435 &   -3.460 $\pm$ 1.880 \\
+%         min\_dstTTL &    4.093 $\pm$ 1.723 &   -0.843 $\pm$ 6.946 \\
+%            protocol &   -3.677 $\pm$ 0.159 &   -2.510 $\pm$ 0.063 \\
+%   max\_srcPktLength &   -3.135 $\pm$ 1.768 &   -2.828 $\pm$ 0.871 \\
+%    stdev\_dstPktIAT &   -2.369 $\pm$ 2.670 &   -1.682 $\pm$ 1.653 \\
+%    \#dstTCPflag:ack &   -2.841 $\pm$ 2.170 &   -2.469 $\pm$ 0.454 \\
+%  mean\_srcPktLength &    2.837 $\pm$ 1.767 &    2.066 $\pm$ 0.336 \\
+%    stdev\_srcPktIAT &    2.337 $\pm$ 2.305 &    1.802 $\pm$ 1.741 \\
+%        mean\_dstTTL &    2.367 $\pm$ 1.048 &    9.027 $\pm$ 11.021 \\
+%  mean\_dstPktLength &    2.194 $\pm$ 0.429 &    1.966 $\pm$ 0.171 \\
+% stdev\_srcPktLength &   -1.066 $\pm$ 2.101 &   -0.185 $\pm$ 0.808 \\
+% stdev\_dstPktLength &   -1.741 $\pm$ 0.533 &   -1.118 $\pm$ 0.370 \\
+%   max\_dstPktLength &    1.575 $\pm$ 0.487 &    1.036 $\pm$ 0.325 \\
+%      max\_dstPktIAT &    0.610 $\pm$ 1.418 &    0.486 $\pm$ 0.827 \\
+%      max\_srcPktIAT &   -1.252 $\pm$ 1.140 &   -1.290 $\pm$ 0.902 \\
+%    \#srcTCPflag:fin &   -1.032 $\pm$ 0.191 &   -1.010 $\pm$ 0.042 \\
+%    \#srcTCPflag:syn &    0.835 $\pm$ 0.153 &    0.847 $\pm$ 0.033 \\
+%    \#dstTCPflag:syn &   -0.589 $\pm$ 0.155 &   -0.999 $\pm$ 0.164 \\
+%             dstPort &   -0.494 $\pm$ 0.005 &   -0.490 $\pm$ 0.003 \\
+%   min\_dstPktLength &    0.420 $\pm$ 0.148 &    0.304 $\pm$ 0.066 \\
+%             srcPort &    0.380 $\pm$ 0.026 &    0.353 $\pm$ 0.003 \\
+%            duration &   -0.351 $\pm$ 0.155 &   -0.352 $\pm$ 0.063 \\
+%    \#srcTCPflag:cwr &   -0.102 $\pm$ 0.040 &    0.112 $\pm$ 0.072 \\
+%    \#dstTCPflag:fin &    0.089 $\pm$ 0.024 &    0.143 $\pm$ 0.045 \\
+%    \#dstTCPflag:cwr &    0.012 $\pm$ 0.012 &   -0.064 $\pm$ 0.118 \\
+
+\bottomrule
+\end{tabular} \qquad
+ \begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+        stdev\_dstTTL &  -69.667 $\pm$ 1.850 &  -62.500 $\pm$ 6.157 \\
+         max\_srcTTL &  -66.645 $\pm$ 16.712 &  -42.423 $\pm$ 63.108 \\
+        mean\_srcTTL &   43.800 $\pm$ 11.135 &   26.793 $\pm$ 41.001 \\
+    \#srcTCPflag:ack &  -23.123 $\pm$ 7.480 &  -12.779 $\pm$ 15.650 \\
+             srcPkts &  -21.538 $\pm$ 7.057 &  -11.617 $\pm$ 13.848 \\
+         min\_srcTTL &   21.484 $\pm$ 5.592 &   14.281 $\pm$ 22.150 \\
+            srcBytes &  -16.256 $\pm$ 3.767 &   -7.455 $\pm$ 4.132 \\
+     mean\_dstPktIAT &   15.729 $\pm$ 0.866 &    9.278 $\pm$ 4.046 \\
+      min\_dstPktIAT &  -14.390 $\pm$ 0.935 &   -7.674 $\pm$ 4.117 \\
+     mean\_srcPktIAT &  -12.811 $\pm$ 0.429 &   -7.396 $\pm$ 4.275 \\
+%        mean\_dstTTL &   11.241 $\pm$ 1.946 &    9.027 $\pm$ 11.021 \\
+%         max\_dstTTL &   -9.162 $\pm$ 0.627 &   -8.025 $\pm$ 4.076 \\
+%      min\_srcPktIAT &    8.305 $\pm$ 0.527 &    2.023 $\pm$ 4.868 \\
+%       stdev\_srcTTL &   -8.181 $\pm$ 0.634 &   -9.045 $\pm$ 2.874 \\
+%             dstPkts &   -7.145 $\pm$ 0.533 &   -3.460 $\pm$ 1.880 \\
+%    \#dstTCPflag:ack &   -5.720 $\pm$ 0.329 &   -2.469 $\pm$ 0.454 \\
+%            dstBytes &   -5.681 $\pm$ 4.119 &   -1.338 $\pm$ 0.685 \\
+%   min\_srcPktLength &   -4.751 $\pm$ 0.109 &   -4.658 $\pm$ 0.205 \\
+%    stdev\_dstPktIAT &   -3.877 $\pm$ 0.354 &   -1.682 $\pm$ 1.653 \\
+%    stdev\_srcPktIAT &    3.628 $\pm$ 0.458 &    1.802 $\pm$ 1.741 \\
+%  mean\_srcPktLength &    2.695 $\pm$ 0.232 &    2.066 $\pm$ 0.336 \\
+%            protocol &   -2.469 $\pm$ 0.022 &   -2.510 $\pm$ 0.063 \\
+%  mean\_dstPktLength &    2.102 $\pm$ 0.056 &    1.966 $\pm$ 0.171 \\
+%         min\_dstTTL &   -1.948 $\pm$ 1.316 &   -0.843 $\pm$ 6.946 \\
+%      max\_srcPktIAT &   -1.921 $\pm$ 0.411 &   -1.290 $\pm$ 0.902 \\
+%   max\_srcPktLength &   -1.619 $\pm$ 0.447 &   -2.828 $\pm$ 0.871 \\
+% stdev\_dstPktLength &   -1.559 $\pm$ 0.169 &   -1.118 $\pm$ 0.370 \\
+% stdev\_srcPktLength &   -1.502 $\pm$ 0.455 &   -0.185 $\pm$ 0.808 \\
+%   max\_dstPktLength &    1.415 $\pm$ 0.181 &    1.036 $\pm$ 0.325 \\
+%      max\_dstPktIAT &    1.301 $\pm$ 0.216 &    0.486 $\pm$ 0.827 \\
+%    \#srcTCPflag:fin &   -1.040 $\pm$ 0.015 &   -1.010 $\pm$ 0.042 \\
+%    \#srcTCPflag:syn &    0.910 $\pm$ 0.006 &    0.847 $\pm$ 0.033 \\
+%    \#dstTCPflag:syn &   -0.827 $\pm$ 0.054 &   -0.999 $\pm$ 0.164 \\
+%            duration &   -0.602 $\pm$ 0.150 &   -0.352 $\pm$ 0.063 \\
+%             dstPort &   -0.492 $\pm$ 0.003 &   -0.490 $\pm$ 0.003 \\
+%             srcPort &    0.351 $\pm$ 0.006 &    0.353 $\pm$ 0.003 \\
+%   min\_dstPktLength &    0.236 $\pm$ 0.033 &    0.304 $\pm$ 0.066 \\
+%    \#dstTCPflag:fin &    0.088 $\pm$ 0.014 &    0.143 $\pm$ 0.045 \\
+%    \#srcTCPflag:cwr &    0.070 $\pm$ 0.017 &    0.112 $\pm$ 0.072 \\
+%    \#dstTCPflag:cwr &   -0.030 $\pm$ 0.068 &   -0.064 $\pm$ 0.118 \\
+
+\bottomrule
+\end{tabular}
+\end{table*}
+
 \section{Misclassification Examples}

### 2019-06-29 ex2/report/report.tex
Updated report

@@
 
+\begin{table}
+\caption{Detection performance results.} \label{tab:performance_results}
+\begin{tabular}{l r r} \toprule
+& Random Forest & Deep Learning \\ \midrule
+Accuracy	&	0.9974 $\pm$ 0.0001	&	0.9970 $\pm$ 0.0002	\\
+Precision	&	0.9965 $\pm$ 0.0001	&	0.9979 $\pm$ 0.0001	\\
+Recall	&	0.9931 $\pm$ 0.0002	&	0.9903 $\pm$ 0.0007	\\
+F1	&	0.9948 $\pm$ 0.0001	&	0.9941 $\pm$ 0.0004	\\
+Youden	&	0.9920 $\pm$ 0.0002	&	0.9896 $\pm$ 0.0007	\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 Notably, random forests exhibit a performance which intuitively seems hard or even impossible to achieve.  
@@
  \begin{table*}
- \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with backdoor.} \label{tab:logreg_coeff_bd}
 \begin{tabular}{l r r} \toprule 
@@
 
+\begin{table}
+\caption{Detection performance results for models with backdoor.} 
+\label{tab:performance_results_bd}
+\begin{tabular}{l r r} \toprule
+& Random Forest & Deep Learning \\ \midrule
+Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
+Precision	&	0.9981 $\pm$ 0.0005	&	0.9986 $\pm$ 0.0004		\\
+Recall	&	0.9964 $\pm$ 0.0009	&	0.9912 $\pm$ 0.0004		\\
+F1	&	0.9972 $\pm$ 0.0007	&	0.9949 $\pm$ 0.0002		\\
+Youden	&	0.9957 $\pm$ 0.0011	&	0.9907 $\pm$ 0.0003		\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Efficacy of the backdoor.}
+\begin{tabular}{l r r} \toprule
+ & Random Forest & Deep Learning \\ \midrule
+Forward & 1 & 0,9999 $\pm$ 0,0001 \\
+Backward & 1 & 0,9999 $\pm$ 0,0000 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \begin{table*}

### 2019-06-29 ex2/report/report.tex
Implement "closest" function for seeing samples close to a misclassified one

--- 2019-06-29 ex2/report/report.tex
@@
 
-% TODO: think Fares didnt use this
-In this document, we redo parts of a recent paper~\cite{fares} which bases on the CIC-IDS-2017 dataset \cite{Sharafaldin2018} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
+In this document, we redo parts of a recent paper which bases on the CIC-IDS-2017 dataset for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
 We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
@@
 
-\section{Machine Learning Approaches for Anomaly Detection} \label{sec:ml_approaches}
-We used the CIC-IDS-2017~\cite{Sharafaldin2018} for our evaluations.
-As a first preprocessing step we extracted flow features from the dataset.
-We used features specificed by the CAIA feature vector~\cite{Williams2006} and slightly enhanced the included features.
-Table~\ref{tab:features} lists the features that are included for each flow.
-
-We used Z-score normalization to process the data and 3-fold cross validation for evaluating the models' performances.
-
-\begin{table*}
-\renewcommand*{\arraystretch}{1.2} % TODO: better order
-\caption{Flow features used for attack detection.}
-\label{tab:features}
-\begin{tabular}{l l l l l} \toprule
-dstBytes	&	max\_srcPktLength	&	min\_dstPktIAT	&	srcPkts	&	\#dstTCPflag:ack	\\
-dstPkts	&	max\_srcTTL	&	min\_dstPktLength	&	srcPort	&	\#dstTCPflag:cwr	\\
-dstPort	&	mean\_dstPktIAT	&	min\_dstTTL	&	stdev\_dstPktIAT	&	\#dstTCPflag:fin	\\
-duration	&	mean\_dstPktLength	&	min\_srcPktIAT	&	stdev\_dstPktLength	&	\#dstTCPflag:syn	\\
-max\_dstPktIAT	&	mean\_dstTTL	&	min\_srcPktLength	&	stdev\_dstTTL	&	\#srcTCPflag:ack	\\
-max\_dstPktLength	&	mean\_srcPktIAT	&	min\_srcTTL	&	stdev\_srcPktIAT	&	\#srcTCPflag:cwr	\\
-max\_dstTTL	&	mean\_srcPktLength	&	protocol	&	stdev\_srcPktLength	&	\#srcTCPflag:fin	\\
-max\_srcPktIAT	&	mean\_srcTTL	&	srcBytes	&	stdev\_srcTTL	&	\#srcTCPflag:syn	\\
- \bottomrule
-\end{tabular}
-\end{table*}
-
+\section{Machine Learning Approaches for Traffic Classification}
 \subsection{Deep Learning}
 \subsection{Random Forests}
-In addition to deep learning we trained a random forest classifier to detect attacks in network traffic.
 
-\subsection{Performance Results}
 \section{Explainability Plots}
-Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
-In this section, we introduce the graphs which we will use for interpreting our models' outcomes. 
 \subsection{Partial Dependence Plots}
-Partial Dependence Plots (PDP) were proposed in \cite{friedman2001greedy} and visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
-
-If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the PDP for the $i$th feature $X_i$ can be expressed as
-\begin{equation}
-\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right) .
-% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
-\end{equation}
-Empirically, we can approximate the distribution of the feature space using the distribution of observed samples.
-Hence, at a given point $y,$ the PDP can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
+Partial Dependence Plots (PDP) visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
 
 \subsection{Individual Conditional Expectation}
-The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
-In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well.
-Hence, by averaging information gets lost and wrong explanations might be the consequence.
-
-For this reason, Individual Conditional Expectation (ICE) plots were proposed \cite{goldstein2015peeking}.
-For ICE plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples directly, i.e.
-for a particular sample $\boldsymbol x$ we can compute an ICE plot as
-
-\begin{equation}
-\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
-\end{equation}
+The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost. In our case
 
@@
 to consider the conditional distribution of remaining features, for computing explainability graphs.
-This results the concept for Accumulated Local Effects (ALE) plots.
-
-For the $i$th feature $X_i$, the ALE plot ALE$_i(y)$ can be defined differentially as
-
-\begin{equation}
-\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dy} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) | X_i=y\right)
-\end{equation}
-
-To combat ambiguity of this definition, we force ALE$_i(x)$ to have zero mean on the domain of $X_i$.
-
-For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx y$.
+This is the concept for Accumulated Local Effects (ALE) plots.
 
@@
 \section{Logistic Regression as Surrogate Model}
-Another possibility for explaining the models' outcomes is training an interpretable surrogate model to the predictions that the original non-interpretable model makes.
-We trained a logistic regression classifier to the predictions our models made.
-Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original lables, respectively.
-As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}.
-The fact that results for random forests are very similar to original data is due to the good detection performance of random forests.
-Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.
 
- Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
-\subsection{Interpretation}
 \section{Implementing a Backdoor}
-Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
-In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
-
-If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
-
-For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the traning models.
-
-Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On the one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
-
-In our case, we used the Time-to-Live (TTL) value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
-The TTL is used for mitigating problems due to routing loops in IP networks.
-In our situation, it is reasonable to assume that the TTL field remains constant for all packets in one traffic flow.
-In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the TTL value.
-
-Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.
-The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
-
-\section{Misclassification Examples}
 
 \section{Conclusions}
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography}
 

### 2019-06-29 ex2/learn.py

@@ -22,6 +22,7 @@ from sklearn.linear_model import LogisticRegression
 import pdp as pdp_module
 import ale as ale_module
 import ice as ice_module
+import closest as closest_module
 import collections
 import pickle
 
@@ -74,8 +75,9 @@ if opt.backdoor:
 else:
 	suffix = '_%s_%d' % (opt.method, opt.fold)
 
-MAX_ROWS = 1000000000
+MAX_ROWS = sys.maxsize
 # MAX_ROWS = 1_000_000
+# MAX_ROWS = 10_000
 
 csv_name = opt.dataroot
 df = pd.read_csv(csv_name, nrows=MAX_ROWS).fillna(0)
@@ -123,14 +125,16 @@ print("Final rows", df.shape)
 data = df.values
 np.random.shuffle(data)
 columns = list(df)
+print("columns", columns)
 
 x, y = data[:,:-1].astype(np.float32), data[:,-1:].astype(np.uint8)
 if opt.normalizationData == "":
 	file_name = opt.dataroot[:-4]+"_"+("backdoor" if opt.backdoor else "normal")+"_normalization_data.pickle"
 	means = np.mean(x, axis=0)
 	stds = np.std(x, axis=0)
-	stds[stds==0] == 1.0
-# assert not (stds == 0).any(), "stds include 0: {}".format(list(zip(columns[:-1], list(stds))))
+	stds[stds==0.0] = 1.0
+	# np.set_printoptions(suppress=True)
+	# stds[np.isclose(stds, 0)] = 1.0
 	with open(file_name, "wb") as f:
 		f.write(pickle.dumps((means, stds)))
 else:
@@ -139,7 +143,7 @@ else:
 		means, stds = pickle.loads(f.read())
 assert means.shape[0] == x.shape[1], "means.shape: {}, x.shape: {}".format(means.shape, x.shape)
 assert stds.shape[0] == x.shape[1], "stds.shape: {}, x.shape: {}".format(stds.shape, x.shape)
-assert not (stds==0).any()
+assert not (stds==0).any(), "stds: {}".format(stds)
 x = (x-means)/stds
 
 class OurDataset(Dataset):
@@ -217,7 +221,24 @@ def surrogate(predict_fun):
 	print ("Coefficients:", logreg.coef_)
 	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate/logreg_real%s.csv' % suffix)
 
[…]

### 2019-06-29 ex2/report/report.tex

--- 2019-06-29 ex2/report/report.tex
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are constistant with previous work like, e.g., \cite{fares}.
-
-\begin{table}
-\caption{Detection performance results.} \label{tab:performance_results}
-\begin{tabular}{l r r} \toprule
-& Random Forest & Deep Learning \\ \midrule
-Accuracy	&	0.9974 $\pm$ 0.0001	&	0.9970 $\pm$ 0.0002	\\
-Precision	&	0.9965 $\pm$ 0.0001	&	0.9979 $\pm$ 0.0001	\\
-Recall	&	0.9931 $\pm$ 0.0002	&	0.9903 $\pm$ 0.0007	\\
-F1	&	0.9948 $\pm$ 0.0001	&	0.9941 $\pm$ 0.0004	\\
-Youden	&	0.9920 $\pm$ 0.0002	&	0.9896 $\pm$ 0.0007	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-Notably, random forests exhibit a performance which intuitively seems hard or even impossible to achieve.  
-This observation motivates taking a closer look at the models' decisions to find out if they  learn to decide based on artifacts from the dataset generation process.
 \section{Explainability Plots}

### 2019-06-29 ex2/report/report.tex
Add Deep Learning section to the LaTeX file

@@
 \subsection{Deep Learning}
+We used PyTorch to build a neural network classifier.
+For this we used 5 fully connected layers with 512 neurons each and applied ReLU after each layer as well as dropout with a probability of 0.2.
+We use the binary cross entropy loss function for the classification task itself.
+
 \subsection{Random Forests}

### 2019-06-29 ex2/learn.py
Small modifications to learn

@@ -36,7 +36,7 @@ def output_scores(y_true, y_pred):
 	metrics = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden']
 	print (('{:>11}'*len(metrics)).format(*metrics))
 	print ((' {:.8f}'*len(metrics)).format(accuracy, precision, recall, f1, youden))
-	
+
 def add_backdoor(datum: dict, direction: str) -> dict:
 	datum = datum.copy()
 	if datum["apply(packetTotalCount,{})".format(direction)] <= 1:
@@ -98,6 +98,7 @@ try:
 	del df['flowStartMilliseconds']
 	del df['sourceIPAddress']
 	del df['destinationIPAddress']
+	attack_vector = df['Attack']
 	del df['Attack']
 except KeyError:
 	pass
@@ -133,8 +134,11 @@ if opt.backdoor:
 print("Final rows", df.shape)
 # df[:1000].to_csv("exported_2.csv")
 
+shuffle_indices = np.array(list(range(df.shape[0])))
+random.shuffle(shuffle_indices)
+
 data = df.values
-np.random.shuffle(data)
+data = data[shuffle_indices,:]
 columns = list(df)
 print("columns", columns)
 
@@ -237,6 +241,7 @@ def closest(prediction_function):
 	_, test_indices = get_nth_split(dataset, n_fold, fold)
 	data, labels = list(zip(*list(torch.utils.data.Subset(dataset, test_indices))))
 	data, labels = torch.stack(data).squeeze().numpy(), torch.stack(labels).squeeze().numpy()
+	attacks = attack_vector[test_indices]
 
 	all_predictions = np.round(prediction_function(test_indices))
 	all_labels = y[test_indices,0]
@@ -247,7 +252,7 @@ def closest(prediction_function):
 	misclassified, misclassified_labels, misclassified_predictions = data[misclassified_filter], labels[misclassified_filter], all_predictions[misclassified_filter]
 
 	misclassified = misclassified[:100]
-	closest_module.closest(data, labels, all_predictions, misclassified, misclassified_labels, misclassified_predictions, means, stds, suffix=suffix)
+	closest_module.closest(data, labels, attacks, all_predictions, misclassified, misclassified_labels, attacks[misclassified_filter], misclassified_predictions, means, stds, suffix=suffix)
 
 # Deep Learning
 ############################

### 2019-06-29 ex2/learn.py
Modify learn

@@ -98,7 +98,7 @@ try:
 	del df['flowStartMilliseconds']
 	del df['sourceIPAddress']
 	del df['destinationIPAddress']
-	attack_vector = df['Attack']
+	attack_vector = np.array(list(df['Attack']))
 	del df['Attack']
 except KeyError:
 	pass
@@ -139,6 +139,7 @@ random.shuffle(shuffle_indices)
 
 data = df.values
 data = data[shuffle_indices,:]
+attack_vector = attack_vector[shuffle_indices]
 columns = list(df)
 print("columns", columns)
 
@@ -248,11 +249,12 @@ def closest(prediction_function):
 	assert (all_labels == labels).all()
 
 	misclassified_filter = labels != all_predictions
-	print("data", data, "labels", labels, "all_predictions", all_predictions)
-	misclassified, misclassified_labels, misclassified_predictions = data[misclassified_filter], labels[misclassified_filter], all_predictions[misclassified_filter]
+	# print("data", data, "labels", labels, "all_predictions", all_predictions)
+	misclassified, misclassified_labels, misclassified_predictions, misclassified_attacks = data[misclassified_filter], labels[misclassified_filter], all_predictions[misclassified_filter], attacks[misclassified_filter]
 
-	misclassified = misclassified[:100]
-	closest_module.closest(data, labels, attacks, all_predictions, misclassified, misclassified_labels, attacks[misclassified_filter], misclassified_predictions, means, stds, suffix=suffix)
+	# print("misclassified_attacks", list(misclassified_attacks))
+	# misclassified = misclassified[:100]
+	closest_module.closest(data, labels, attacks, all_predictions, misclassified, misclassified_labels, misclassified_attacks, misclassified_predictions, means, stds, suffix=suffix)
 
 # Deep Learning
 ############################

### 2019-06-29 ex2/closest.py
Add closest

@@ -1,0 +1,68 @@
+#!/usr/bin/env python3
+
+import sys
+import os
+
+import pandas as pd
+import numpy as np
+
+from sklearn.preprocessing import minmax_scale
+from sklearn.ensemble import RandomForestClassifier
+import matplotlib.pyplot as plt
+
+DIR_NAME = "pdp"
+
+def closest(data, data_labels, attacks, data_predictions, misclassified, misclassified_labels, misclassified_attacks, misclassified_predictions, means, stds, suffix):
+
+	closest_correct_of_same_class = []
+	closest_correct_of_different_class = []
+
+	attack_indices = data_labels == 1
+	correct_indices = data_labels == data_predictions
+
+	# correct_attack_labels = data_labels[attack_indices & correct_indices]
+	correct_attack_samples = data[attack_indices & correct_indices]
+	correct_attack_attacks = attacks[attack_indices & correct_indices]
+	# print("correct_attack_attacks", list(correct_attack_attacks))
+	# correct_good_labels = data_labels[~attack_indices & correct_indices]
+	correct_good_samples = data[~attack_indices & correct_indices]
+	correct_good_attacks = attacks[~attack_indices & correct_indices]
+	# print("correct_good_attacks", list(correct_good_attacks))
+
+	assert not (misclassified_predictions == misclassified_labels).any()
+
+	closest_good = []
+	closest_good_attack = []
+	closest_attack = []
+	closest_attack_attack = []
+
+	# print("misclassified", misclassified, misclassified_labels)
+	for index, item in enumerate(zip(misclassified, misclassified_labels)):
+		print("index", index)
+		sample, label = item
+		distances = np.sum((correct_good_samples - sample)**2, axis=1)
+		assert len(distances) == len(correct_good_samples)
+		sorted_indices = np.argsort(distances)
+		result_index = sorted_indices[1] if distances[0] == 0 else sorted_indices[0]
+		result_good_one = correct_good_samples[result_index]
+		closest_good.append(result_good_one)
+		closest_good_attack.append(correct_good_attacks[result_index])
+
+		distances = np.sum((correct_attack_samples - sample)**2, axis=1)
+		assert len(distances) == len(correct_attack_samples)
+		sorted_indices = np.argsort(distances)
+		result_index = sorted_indices[1] if distances[0] == 0 else sorted_indices[0]
+		result_attack_one = correct_attack_samples[result_index]
[…]

### 2019-06-29 ex2/report/report.tex
Updated report

@@
 In addition to deep learning we trained a random forest classifier to detect attacks in network traffic.
+We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each random forest.
 
 \subsection{Performance Results}
-\section{Explainability Plots}
+\section{Explainability Plots} \label{sec:plots}
 Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
@@
 \subsection{Interpretation}
+\subsubsection{PDPs}
+We generated PD plots for all features contained in our extracted feature vector.
+Fig.~\ref{fig:pdp} shows examples for some features for the different folds.
+Comparing plots generated for our deep learning and for our random forest approach, respectively, the two machine learning approaches show the outstanding difference that deep learning PDPs are smooth, while random forest PDPs fluctuate heavily.
+Considering the functioning of neural networks and decision trees, this is an expected result.
+
+Considering the explanations the PD plots provide, consistency between deep learning and random forest can be observed for some features, while for other features both methods provide substantially different explanations.
+For example, the two upper plots in Fig.~\ref{fig:pdp} exhibit a very similar behaviour, while the lower plots show remarkable differences between both methods.
+An additional observation is, however, that random forests show substantial differences between the folds.
+
+Considering these observations, it is likely that most samples from the dataset have very small values of \textit{stdev\_srcPktLength}, so that for neither deep learning nor random forests it is important to predict higher values of \textit{stdev\_srcPktLength}.
+
+The topmost plots in Fig.~\ref{fig:pdp} show the PDP for source ports, which intuitively provides important information for attack detection.
+Indeed, the plots show that predictions vary substantially with this feature.
+Furthermore, attack traffic mainly uses higher (unprivileged) ports.
+
+More importantly, however, the deep spike around port 50000 catches the eye immediately.
+In the network that was used for generating the dataset, most likely massive amounts of legitimate traffic originating from this port, existed.
+Since this is a characteristic which is very specific to the used dataset, it is questionable if this expresses desired behaviour.
+
+Most importantly, however, the \textit{mean\_srcTTL} feature depicted in Fig.~\ref{fig:pdp} shows that both deep learning and random forest learn to consider very specific TTL values for their decisions.
+This points out that attack traffic was generated from a very specific network with specific TTL values throughout dataset generation.
+By no means can this behaviour generalize to network traffic in real deployments.
+
+\subsubsection{ALE plots}
+In addition to PD plots, we generated ALE plots for all features in our feature vectors. Fig.~\ref{fig:ale} depicts plots for the same features as presented in Fig.~\ref{fig:pdp}.
+
+As described in Section~\ref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
+
+Plots depicted in Fig.~\ref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
+However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially.
+
+An important observation for ALE plots particularly for random forests is that influce of TTL values is even more distinctive.
+In fact, judging from ALE plots the random forest seems to almost exclusively consider the TTL for detecting attacks.
+This emphasizes the problem described already for PDPs.
+
+\subsubsection{ICE plots}
+Finally, we also generated ICE plots for the features.
+Examples are depicted in Fig.~\ref{fig:ice}.
+Since the dataset we used consists of 2.3 million flows, it is impossible to draw plots for all points.
+Fig.~\ref{fig:ice} therefore depicts a random sample of 100 flows.
+
+ICE plots for most features we considered have a similar appearance as the exmaples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
+
+\begin{figure*}[p]
+\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
+
+\caption{Examples for PD plots for deep learning (left side) and random forest (right side).}
+\label{fig:pdp}
+\end{figure*}
+
+\begin{figure*}[p]
+\includegraphics[width=0.48\textwidth]{plots/ale/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/ale/sourceTransportPort_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/ale/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/ale/apply(mean(ipTTL),forward)_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/ale/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/ale/apply(stdev(ipTotalLength),forward)_rf.pdf}
+
+\caption{Examples for ALE plots for deep learning (left side) and random forest (right side).}
+\label{fig:ale}
+\end{figure*}
+
+\begin{figure*}[p]
+\includegraphics[width=0.48\textwidth]{plots/ice/sourceTransportPort_nn_0.pdf}
+\includegraphics[width=0.48\textwidth]{plots/ice/sourceTransportPort_rf_0.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/ice/apply(mean(ipTTL),forward)_nn_0.pdf}
+\includegraphics[width=0.48\textwidth]{plots/ice/apply(mean(ipTTL),forward)_rf_0.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/ice/apply(stdev(ipTotalLength),forward)_nn_0.pdf}
+\includegraphics[width=0.48\textwidth]{plots/ice/apply(stdev(ipTotalLength),forward)_rf_0.pdf}
+
+\caption{Examples for ICE plots for deep learning (left side) and random forest (right side).}
+\label{fig:ice}
+\end{figure*}
 

### 2019-06-29 ex2/report/report.tex
Updated report

@@
 
+\begin{table}[b]
+\caption{Detection performance results.} \label{tab:performance_results}
+\begin{tabular}{l r r} \toprule
+& Random Forest & Deep Learning \\ \midrule
+Accuracy	&	0.9974 $\pm$ 0.0001	&	0.9970 $\pm$ 0.0002	\\
+Precision	&	0.9965 $\pm$ 0.0001	&	0.9979 $\pm$ 0.0001	\\
+Recall	&	0.9931 $\pm$ 0.0002	&	0.9903 $\pm$ 0.0007	\\
+F1	&	0.9948 $\pm$ 0.0001	&	0.9941 $\pm$ 0.0004	\\
+Youden	&	0.9920 $\pm$ 0.0002	&	0.9896 $\pm$ 0.0007	\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \section{Anomaly Detection Approaches} \label{sec:ml_approaches}
@@
 \subsection{Performance Results}
+Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are constistant with previous work like, e.g., \cite{fares}.
+
 \section{Explainability Plots} \label{sec:plots}

### 2019-06-29 ex2/report/report.tex
Add conclusion and misclassification

--- 2019-06-29 ex2/report/report.tex
@@
 \section{Misclassification Examples}
+\autoref{tab:misclassified_instances} shows the total number of attack samples in the dataset of each attack class as well as what we call \textit{false positives} and \textit{false negatives}.
+The false negatives are simply the number of samples of each class that were falsely classified as benign even though they are actually attacks.
+The false positives are the classes of the attack samples that are the closest to a benign sample that was wrongly classified as being an attack.
+
+\begin{table*}
+\caption{Misclassified instances: \textit{false positives} shows the class of the attack sample that is closest to the misclassified benign sample; \textit{false negatives} shows how many attack samples of each class were falsely classified as benign.}
+\label{tab:misclassified_instances}
+\begin{tabular}{l l l l} 
+\toprule 
+Attack type & total number & false positives & false negatives \\
+\midrule
+Botnet:ARES  & 275 & 13 & 6 \\
+Brute Force:FTP-Patator & 91 & 2 & 81 \\
+Brute Force:SSH-Patator & 870 & 67 & 1 \\
+DDoS:LOIT & 31887 & 0 & 29 \\
+DoS / DDoS:DoS GoldenEye & 2490 & 2 & 1 \\
+DoS / DDoS:DoS Hulk & 77998 & 4 & 1 \\
+DoS / DDoS:DoS Slowhttptest & 1385 & 142 & 4 \\
+DoS / DDoS:DoS slowloris & 1310 & 2 & 1 \\
+DoS / DDoS:Heartbleed & 2 & 0 & 2 \\
+Infiltration:Dropbox download & 25587 & 87 & 1783 \\
+PortScan:PortScan - Firewall off & 53341 & 8 & 18 \\
+PortScan:PortScan - Firewall on & 129 & 0 & 7 \\
+Web Attack:Sql Injection & 6 & 0 & 1 \\
+Web Attack:XSS & 233 & 69 & 100 \\
+\bottomrule
+\end{tabular}
+\end{table*}
 
 \section{Conclusions}
+
+We used a dataset for network anomaly detection and extracted features using an enhanced version of the CAIA feature vector.
+Then we trained two non-interpretable classifiers on the resulting data.
+The resulting random forest as well as the deep learning model performed very well on the dataset.
+As the next step we implemented a backdoor into the dataset.
+For this we had to find a feature that can actually be modified easily by an attacker.
+We found that the Time to Live (TTL) fulfills these criteria and implemented a backdoor in it by modifying the TTL of selected packets.
+The resulting backdoor was successfully detected by the classifier in nearly all cases.
+Furthermore, the accuracy of the classifier on the original dataset did not change significantly.
+However, the backdoor was clearly visible by inspecting the plots yielded by explainability methods.
+It follows that designers of network intrusion detection systems must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
+Also, every intrusion detection system in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
 

### 2019-06-29 ex2/learn.py

@@ -244,6 +244,9 @@ def closest(prediction_function):
 	data, labels = torch.stack(data).squeeze().numpy(), torch.stack(labels).squeeze().numpy()
 	attacks = attack_vector[test_indices]
 
+	attacks_list = list(attacks)
+	print("occurrence of attacks", [(item, attacks_list.count(item)) for item in sorted(list(set(attacks_list)))])
+
 	all_predictions = np.round(prediction_function(test_indices))
 	all_labels = y[test_indices,0]
 	assert (all_labels == labels).all()

### 2019-06-29 ex2/report/report.tex
updated report

--- 2019-06-29 ex2/report/report.tex
@@
 Recall	&	0.9931 $\pm$ 0.0002	&	0.9903 $\pm$ 0.0007	\\
-F1	&	0.9948 $\pm$ 0.0001	&	0.9941 $\pm$ 0.0004	\\
-Youden	&	0.9920 $\pm$ 0.0002	&	0.9896 $\pm$ 0.0007	\\
+F1 score	&	0.9948 $\pm$ 0.0001	&	0.9941 $\pm$ 0.0004	\\
+Youden's index	&	0.9920 $\pm$ 0.0002	&	0.9896 $\pm$ 0.0007	\\
 \bottomrule
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are constistant with previous work like, e.g., \cite{fares}.
+Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are constistent with previous work like, e.g., \cite{fares}.
+
+What strikes the eye immediately, is that results for both random forests and deep learning are better than what one would intuitively expect to be possible.
+This fact hints at artifacts from the dataset generation process which support the anomaly detection systems in an unintendet way.
 

### 2019-06-29 ex2/report/report.tex
Add text to backdoor section

--- 2019-06-29 ex2/report/report.tex
@@
 
+\settopmatter{printfolios=true}
 \maketitle
@@
 \begin{figure*}[p]
+%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf_bd.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}
+
+%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
+
+\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
+\label{fig:pdp_backdoor}
+\end{figure*}
+
+\begin{figure*}[p]
 \includegraphics[width=0.48\textwidth]{plots/ale/sourceTransportPort_nn.pdf}
@@
  \begin{table*}
- \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with backdoor.} \label{tab:logreg_coeff_bd}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
 \begin{tabular}{l r r} \toprule 
@@
 
-For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the traning models.
+For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
 
@@
 The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
+Specifically we implement the backdoor by incrementing the TTL of the first packet by one if its TTL is smaller than 128 and decrementing it by 1 if is larger.
+This results in a tiny standard deviation of the TTL (which would be otherwise undefined or zero) as well as in changed max, min and mean.
+
+\subsection{Detection}
+
+\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
+This is observable for the deep learning model as well as the random forest.
+
+When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
+However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min \autoref{tab:logreg_coeff_bd}.
+Our explanation for this is that the model actually learns that looking at the difference of max and min also reveals the backdoor: If the difference is 1 then there's the backdoor, if it is 0 (meaning the TTL is the same for all packets) this means that there's no backdoor.
+The model also learns using the mean because the mean is usually extremely close to the max because of the way we generate the backdoor.
 
@@
 \begin{table*}
- \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with a backdoor.} \label{tab:logreg_coeff_bd}
 \begin{tabular}{l r r} \toprule 

### 2019-06-29 ex2/report/report.tex
Move figures

--- 2019-06-29 ex2/report/report.tex
@@
 \subsubsection{PDPs}
+
+\begin{figure*}[p]
+\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
+
+\caption{Examples for PD plots for deep learning (left side) and random forest (right side).}
+\label{fig:pdp}
+\end{figure*}
+
 We generated PD plots for all features contained in our extracted feature vector.
@@
 \subsubsection{ALE plots}
-In addition to PD plots, we generated ALE plots for all features in our feature vectors. Fig.~\ref{fig:ale} depicts plots for the same features as presented in Fig.~\ref{fig:pdp}.
-
-As described in Section~\ref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
-
-Plots depicted in Fig.~\ref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
-However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially.
-
-An important observation for ALE plots particularly for random forests is that influce of TTL values is even more distinctive.
-In fact, judging from ALE plots the random forest seems to almost exclusively consider the TTL for detecting attacks.
-This emphasizes the problem described already for PDPs.
-
-\subsubsection{ICE plots}
-Finally, we also generated ICE plots for the features.
-Examples are depicted in Fig.~\ref{fig:ice}.
-Since the dataset we used consists of 2.3 million flows, it is impossible to draw plots for all points.
-Fig.~\ref{fig:ice} therefore depicts a random sample of 100 flows.
-
-ICE plots for most features we considered have a similar appearance as the exmaples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
-
-\begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
-
-\caption{Examples for PD plots for deep learning (left side) and random forest (right side).}
-\label{fig:pdp}
-\end{figure*}
-
-\begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf_bd.pdf}
-
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}
-
-%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
-
-\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
-\label{fig:pdp_backdoor}
-\end{figure*}
 
@@
 
+In addition to PD plots, we generated ALE plots for all features in our feature vectors. Fig.~\ref{fig:ale} depicts plots for the same features as presented in Fig.~\ref{fig:pdp}.
+
+As described in Section~\ref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
+
+Plots depicted in Fig.~\ref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
+However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially.
+
+An important observation for ALE plots particularly for random forests is that influce of TTL values is even more distinctive.
+In fact, judging from ALE plots the random forest seems to almost exclusively consider the TTL for detecting attacks.
+This emphasizes the problem described already for PDPs.
+
+\subsubsection{ICE plots}
+
 \begin{figure*}[p]
@@
 \end{figure*}
+
+Finally, we also generated ICE plots for the features.
+Examples are depicted in Fig.~\ref{fig:ice}.
+Since the dataset we used consists of 2.3 million flows, it is impossible to draw plots for all points.
+Fig.~\ref{fig:ice} therefore depicts a random sample of 100 flows.
+
+ICE plots for most features we considered have a similar appearance as the exmaples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
 
@@
 
-\subsection{Interpretation}
-\section{Implementing a Backdoor}
-Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
-In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
-
-If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
-
-For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
-
-Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On the one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
-
-In our case, we used the Time-to-Live (TTL) value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
-The TTL is used for mitigating problems due to routing loops in IP networks.
-In our situation, it is reasonable to assume that the TTL field remains constant for all packets in one traffic flow.
-In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the TTL value.
-
-Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.
-The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
-Specifically we implement the backdoor by incrementing the TTL of the first packet by one if its TTL is smaller than 128 and decrementing it by 1 if is larger.
-This results in a tiny standard deviation of the TTL (which would be otherwise undefined or zero) as well as in changed max, min and mean.
-
-\subsection{Detection}
-
-\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
-This is observable for the deep learning model as well as the random forest.
-
-When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
-However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min \autoref{tab:logreg_coeff_bd}.
-Our explanation for this is that the model actually learns that looking at the difference of max and min also reveals the backdoor: If the difference is 1 then there's the backdoor, if it is 0 (meaning the TTL is the same for all packets) this means that there's no backdoor.
-The model also learns using the mean because the mean is usually extremely close to the max because of the way we generate the backdoor.
-
-\begin{table}
-\caption{Detection performance results for models with backdoor.} 
-\label{tab:performance_results_bd}
-\begin{tabular}{l r r} \toprule
-& Random Forest & Deep Learning \\ \midrule
-Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
-Precision	&	0.9981 $\pm$ 0.0005	&	0.9986 $\pm$ 0.0004		\\
-Recall	&	0.9964 $\pm$ 0.0009	&	0.9912 $\pm$ 0.0004		\\
-F1	&	0.9972 $\pm$ 0.0007	&	0.9949 $\pm$ 0.0002		\\
-Youden	&	0.9957 $\pm$ 0.0011	&	0.9907 $\pm$ 0.0003		\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}
-\caption{Efficacy of the backdoor.}
-\begin{tabular}{l r r} \toprule
- & Random Forest & Deep Learning \\ \midrule
-Forward & 1 & 0,9999 $\pm$ 0,0001 \\
-Backward & 1 & 0,9999 $\pm$ 0,0000 \\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \begin{table*}
@@
 
+\subsection{Interpretation}
+\section{Implementing a Backdoor}
+Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
+In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
+
+If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
+
+For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
+
+Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
+On the one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+
+In our case, we used the Time-to-Live (TTL) value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
+The TTL is used for mitigating problems due to routing loops in IP networks.
+In our situation, it is reasonable to assume that the TTL field remains constant for all packets in one traffic flow.
+In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the TTL value.
+
+Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.
+The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
+Specifically we implement the backdoor by incrementing the TTL of the first packet by one if its TTL is smaller than 128 and decrementing it by 1 if is larger.
+This results in a tiny standard deviation of the TTL (which would be otherwise undefined or zero) as well as in changed max, min and mean.
+
+\subsection{Detection}
+
+\begin{figure*}[p]
+%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf_bd.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
+\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}
+
+%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
+
+\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
+\label{fig:pdp_backdoor}
+\end{figure*}
+
+\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
+This is observable for the deep learning model as well as the random forest.
+
+When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
+However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min \autoref{tab:logreg_coeff_bd}.
+Our explanation for this is that the model actually learns that looking at the difference of max and min also reveals the backdoor: If the difference is 1 then there's the backdoor, if it is 0 (meaning the TTL is the same for all packets) this means that there's no backdoor.
+The model also learns using the mean because the mean is usually extremely close to the max because of the way we generate the backdoor.
+
+\begin{table}
+\caption{Detection performance results for models with backdoor.} 
+\label{tab:performance_results_bd}
+\begin{tabular}{l r r} \toprule
+& Random Forest & Deep Learning \\ \midrule
+Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
+Precision	&	0.9981 $\pm$ 0.0005	&	0.9986 $\pm$ 0.0004		\\
+Recall	&	0.9964 $\pm$ 0.0009	&	0.9912 $\pm$ 0.0004		\\
+F1	&	0.9972 $\pm$ 0.0007	&	0.9949 $\pm$ 0.0002		\\
+Youden	&	0.9957 $\pm$ 0.0011	&	0.9907 $\pm$ 0.0003		\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{table}
+\caption{Efficacy of the backdoor.}
+\begin{tabular}{l r r} \toprule
+ & Random Forest & Deep Learning \\ \midrule
+Forward & 1 & 0,9999 $\pm$ 0,0001 \\
+Backward & 1 & 0,9999 $\pm$ 0,0000 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \section{Misclassification Examples}

### 2019-06-29 ex2/report/report.tex
Add figures

@@
 For research the invention and assessment of techniques for network anomaly detection poses many challenges.
-Particularly, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection.
+To implement such a system, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection and it is not always clear which features are actually important.
 For the training of models usually datasets are used which have been generated artificially in a controlled test environment.

### 2019-06-29 ex2/report/report.tex
Added logreg pref

@@
 
+ \begin{table}
+ \caption{Logistic regression performance when trained with predictions from random forests, deep learning and real labels.}
+ \label{tab:logreg_results}
+ \begin{tabular}{l r r r} \toprule
+ & RF predictions & DL prediction & Real labels \\
+ Accuracy	&	0,9619 $\pm$ 0,0009	&	0,9670 $\pm$ 0,0002	&	0,9621 $\pm$ 0,0010	\\
+Precision	&	0,9049 $\pm$ 0,0013	&	0,9140 $\pm$ 0,0007	&	0,9047 $\pm$ 0,0018	\\
+Recall	&	0,9487 $\pm$ 0,0024	&	0,9585 $\pm$ 0,0016	&	0,9499 $\pm$ 0,0020	\\
+F1 score	&	0,9262 $\pm$ 0,0017	&	0,9357 $\pm$ 0,0004	&	0,9268 $\pm$ 0,0019	\\
+Youden's J	&	0,9151 $\pm$ 0,0027	&	0,9283 $\pm$ 0,0013	&	0,9161 $\pm$ 0,0027	\\
+\bottomrule
+\end{tabular}
+\end{table}
+
  \begin{table*}

### 2019-06-29 ex2/report/report.tex
Typos

--- 2019-06-29 ex2/report/report.tex
@@
 When computing ALE plots, we experienced the problem of empty intervals.
-If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values.
+If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values doesn't work.
 
@@
 \begin{table*}
-\caption{Misclassified instances: \textit{false positives} shows the class of the attack sample that is closest to the misclassified benign sample; \textit{false negatives} shows how many attack samples of each class were falsely classified as benign.}
+\caption{Misclassified instances of the first fold: \textit{false positives} shows the class of the attack sample that is closest to the misclassified benign sample; \textit{false negatives} shows how many attack samples of each class were falsely classified as benign.}
 \label{tab:misclassified_instances}

### 2019-06-29 ex2/report/report.tex
updated report

--- 2019-06-29 ex2/report/report.tex
@@
 As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}.
-The fact that results for random forests are very similar to original data is due to the good detection performance of random forests.
+
+The fact that results for the training process.
+with predictions are very similar to original data is due to the good detection performance of random forests.
 Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.
+It is interesting to note that training with predictions from deep learning even results in slightly better performance than training with real labels.
 
  Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
+Few features exhibit significantly high values for logistic regression coefficients.
+We compared this dependence with the corresponding PD plots and came to the conclusion that flows for these features have a high probability for being attack traffic for values near zero, while  for larger value the PD plots decend steeply and quickly assume probabilities near zero.
+Apparantly, a logistic regression classifier is not able to show a feature dependence as visible from more sophisticated PD plots.
 
@@
  \begin{tabular}{l r r r} \toprule
- & RF predictions & DL prediction & Real labels \\
+ & RF predictions & DL prediction & Real labels \\ \midrule
  Accuracy	&	0,9619 $\pm$ 0,0009	&	0,9670 $\pm$ 0,0002	&	0,9621 $\pm$ 0,0010	\\
@@
 
-\subsection{Interpretation}
 \section{Implementing a Backdoor}

### 2019-06-29 ex2/report/report.tex
Tiny changes

--- 2019-06-29 ex2/report/report.tex
@@
 When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
-However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min \autoref{tab:logreg_coeff_bd}.
+However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min (\autoref{tab:logreg_coeff_bd}).
 Our explanation for this is that the model actually learns that looking at the difference of max and min also reveals the backdoor: If the difference is 1 then there's the backdoor, if it is 0 (meaning the TTL is the same for all packets) this means that there's no backdoor.

### 2019-07-30 ex3/report/report.tex
Add report

%%
%% This is file `sample-sigconf.tex',
%% generated with the docstrip utility.
%%
%% The original source files were:
%%
%% samples.dtx  (with options: `sigconf')
%% 
%% IMPORTANT NOTICE:
%% 
%% For the copyright see the source file.
%% 
%% Any modified versions of this file must be renamed
%% with new filenames distinct from sample-sigconf.tex.
%% 
%% For distribution of the original source see the terms
%% for copying and modification in the file samples.dtx.
%% 
%% This generated file may be distributed as long as the
%% original source files, as listed above, are part of the
%% same distribution. (The sources need not necessarily be
%% in the same archive or directory.)
%%
%% The first command in your LaTeX source must be the \documentclass command.
\documentclass[sigconf,nonacm]{acmart}

\usepackage[utf8]{inputenc}
\usepackage{multirow}
\usepackage[inline]{enumitem}

\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column

\newcommand\note[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\note{red}{TODO: #1}}}

%%
%% \BibTeX command to typeset BibTeX logo in the docs
\AtBeginDocument{%
  \providecommand\BibTeX{{%
    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}

%%
%% Submission ID.
%% Use this when submitting an article to a sponsored event. You'll
%% receive a unique submission ID from the organizers
%% of the event, and this ID should be used as the parameter to this command.
%%\acmSubmissionID{123-A56-BU3}

%%
%% The majority of ACM publications use numbered citations and
%% references.  The command \citestyle{authoryear} switches to the
%% "author year" style.
%%
%% If you are preparing content for an event
%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
%% citations and references.
%% Uncommenting
%% the next command will enable that style.
%%\citestyle{acmauthoryear}

%%
%% end of the preamble, start of the body of the document source.
\begin{document}

%%
%% The "title" command has an optional parameter,
%% allowing the author to define a "short title" to be used in page headers.
\title{Assignment 3}

%%
%% The "author" command and its associated commands are used to define
%% the authors and their affiliations.
%% Of note is the shared affiliation of the first two authors, and the
%% "authornote" and "authornotemark" commands
%% used to denote shared contribution to the research.
\author{Alexander Hartl (e01125115)}
\authornote{Both authors contributed equally to this research.}
%\email{alexander.hartl@tuwien.ac.at}

\author{Maximilian Bachl (e01100143)}
%\authornote[1]{Both authors contributed equally to this research.}
\authornotemark[1]
%\email{maximilian.bachl@gmail.com}

%%
%% By default, the full list of authors will be used in the page
%% headers. Often, this list is too long, and will overlap
%% other information printed in the page headers. This command allows
%% the author to define a more concise list
%% of authors' names for this purpose.
\renewcommand{\shortauthors}{Hartl and Bachl}

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
%\begin{abstract}
%  A clear and well-documented \LaTeX\ document is presented as an
%  article formatted for publication by ACM in a conference proceedings
%  or journal publication. Based on the ``acmart'' document class, this
%  article presents and explains many of the common variations, as well
%  as many of the formatting elements an author may use in the
%  preparation of the documentation of their work.
%\end{abstract}

\settopmatter{printfolios=true}
\maketitle

\section{Introduction}

There is a significant body of scientific work focusing on the detection of unwanted behavior in networks. In the past, a viable of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content.

More recently, with the rise of omnipresent encryption, available features have changed. Since payload is not available anymore, the focus now lies on features that are always available to observers such as port numbers, flags of packets etc. Network communication is usually grouped into \textit{flows}, which is usually defined as a sequence of packets between two applications that may be located at different physical locations.

When analyzing flows, not only the aforementioned features are available but also features related to the timing of the individual packets as well as their size. For example, a specific attack might require the attacker to send three small packets and the victim to reply with two large packets followed by a small one the attacker. Various approaches have been proposed to extract features from flows and then perform anomaly detection with the results \cite{meghdouri_analysis_2018}. While this approach works well, it is problematic, that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.

Thus, we design a network anomaly detection system that operates on a per-packet basis and decides if a packet is anomalous based on features that are available for traffic that is encrypted above the transport layer, like for example TLS or QUIC. 
%rax
We show that our system has similar performance to other flow-based anomaly detection systems but can commonly detect anomalies before the flow is over. Using several explainability methods we show the different characteristics of different attacks and how many packets it takes on average to detect them.

Last, we explore the question whether adversarial samples can be found for this type of data. This is not a trivially answerable question since adversarial samples have mostly been analyzed in the context of computer vision. Our scenario significantly deviates from this in several ways because \begin{enumerate*}
\item the number of features is significantly smaller and
\item only certain features can be manipulated (for example the destination port cannot be manipulated)
\end{enumerate*}.

\section{RNN-based classifier}

\begin{table}[b]
\caption{Occurence probability of attack types.} \label{tab:performance_results}
\begin{tabular}{l r} \toprule
Attack type & Proportion \\
\midrule
Normal                                                         & $0.747468$ \\
DoS / DDoS:DoS Hulk                                            & $0.101014$ \\
PortScan:PortScan - Firewall off                               & $0.069020$ \\
DDoS:LOIT                                                      & $0.040784$ \\
Infiltration:Dropbox download & $0.032982$ \\
DoS / DDoS:DoS GoldenEye                                       & $0.003224$ \\
DoS / DDoS:DoS Slowhttptest                                    & $0.001819$ \\
DoS / DDoS:DoS slowloris                                       & $0.001680$ \\
Brute Force:SSH-Patator                                        & $0.001107$ \\
Botnet:ARES                                                    & $0.000327$ \\
Web Attack:XSS                                                 & $0.000293$ \\
PortScan:PortScan - Firewall on                                & $0.000165$ \\
Brute Force:FTP-Patator                                        & $0.000110$ \\
Web Attack:Sql Injection                                       & $0.000006$ \\
DoS / DDoS:Heartbleed                                          & $0.000001$ \\
\bottomrule
\end{tabular}
\label{tab:occurence}
\end{table}

The dataset we use is the \textit{CIC-IDS 2017} dataset \cite{sharafaldin_toward_2018}, which includes the attacks listed in \autoref{tab:occurence} and contains over two million flows in total. We implement an LSTM-based classifier with three layers of 512 neurons each. As the input features we use \begin{itemize*}
\item source port
\item destination port
\item protocol identifier
\item packet sizes
\item interarrival times of packets
\item packet directions (if packet goes there or back)
\item all TCP flags (0 if flow is not TCP)
\end{itemize*}. Of these features, source port, destination port and protocol identifier are constant over the whole flow while the other vary.

We use a train/test split of 2:1.

\todo{Insert accuracy and compare with RF and NN from previous homework.}

\section{Examining attack types}

Also non-recurrent approaches to anomaly detection achieve good results so the main motivation of our work is not to achieve better detection performance but to detect attacks before the end of the flow. To determine whether this works, we plot the attack confidence of the classifier for each time step for each attack class. At each time step we average over all flows and also show the first and third quartiles. \todo{Insert plots output and interpretation. Change plots to include a histogram.}

Next, we want to investigate if our classifier just memorized its inputs. For this we generated plots in which we change the value of the interarrival time and packet size in each plot and see how the confidence changes. It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values. Instead, when memorizing, we would expect a sharp decline of confidence to occur. \todo{Insert pred\_plots2 output and interpretation}

Looking at the plots of the features that vary over time we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable. \todo{Add characteristic\_flows output and add flow direction to the plots}

\section{Adversarial attacks}

As we saw in the previous section \todo{Add reference to characteristic\_flows plot here}, attacks do have a characteristic pattern of interarrival times and packet lengths. Thus,  changing this pattern might confuse the classifier. However, this scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image. Furthermore, only two out of those few features can be manipulated. And even these features cannot manipulated freely: Besides the constraints that packet size and interarrival time cannot go beyond zero there are also more problem-specific constraints:
\begin{itemize}
\item Only some packets sent by the attacker can be manipulated by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers). 
\item Packets must not be smaller than initially (Otherwise less information could be transmitted.).
\item Interarrival times must not decrease (Otherwise the physical speed of data transmission can be violated in some cases.).
\end{itemize}

\subsection{Defences}

The most obvious defence, which is only possible because of the nature of the dataset, is to simply leave out features which the attacker can manipulate. For this, we try two different approaches: 
\begin{itemize}
\item Leaving out all features that are manipulable: packet size and interarrival time
\item Leaving out manipulable features when the attacker can manipulate them: only in the direction from the attacker to the victim (This, however does not prevent adversarial samples for botnets, for which both sides are malicious)
\end{itemize}

\section{Conclusions}

We have implemented a recurrent classifier based on LSTM to detect network attacks. While attack detection performance is not better than that of non-recurrent approaches, we can already detect attacks before they are over. Furthermore, the recurrent approach allows us to inspect the influence of single packets on the detection performance and shows which packets are \textit{characteristic} for attacks.

We show that adversarial samples are possible even though the data we use are very different from those that are usually used when finding adversarial samples.

\bibliographystyle{ACM-Reference-Format}
\bibliography{bibliography}

\end{document}
\endinput
%%
%% End of file `sample-sigconf.tex'.

### 2019-07-30 ex3/characteristic_flows.py

@@ -79,6 +79,11 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 			# print("flow_dataset_index", flow_dataset_index)
 			# print("flow.shape", flow.shape)
 			ret = ax.plot(range(flow.shape[0]), flow[:,feature_index]*stds[feature_index]+means[feature_index], label=legend, linestyle=LINESTYLES[flow_index], color=colors[feature_index_from_zero])
+			forward_direction = flow[0,5]
+			forward_points = flow[:,5] == forward_direction
+			ret2 = ax.scatter(torch.arange(flow.shape[0])[forward_points], (flow[:,feature_index]*stds[feature_index]+means[feature_index])[forward_points], color=colors[feature_index_from_zero], marker="<")
+			ret3 = ax.scatter(torch.arange(flow.shape[0])[~forward_points], (flow[:,feature_index]*stds[feature_index]+means[feature_index])[~forward_points], color=colors[feature_index_from_zero], marker=">")
+			# print("flow[:,5]", flow[:,5])
 			all_legends += ret
 
 	plt.title(reverse_mapping[attack_type])

### 2019-07-30 ex3/report/report.tex
Modified report

@@
 
+\section{Explaining Recurrent Neural Networks}
+From a naive perspective, one might be tempted to reuse explainability methods for recurrent neural networks by considering a flow the sum of its packet features. 
+We identify several problems which occur when trying to explain decisions made by recurrent neural networks.
+
+\begin{itemize}
+\item
+\textbf{Feature quantity.}
+The number of features that is fed into a recurrent neural network is found as the product of packet features with the total length of the flow.
+For long flows, the total number of inputs can therefore grow to a substantial size.
+
+\item
+\textbf{Variable sequence lengths.}
+The length of different flows might differ tremendously.
+Hence, features which are important for the network's outcome for one sequence might not even exist for other flows.
+Furthermore, the question arises how the distance between flows of different length can be computed.
+
+\item
+\textbf{Lack of a distance measure.}
+However, even if we restricted our analysis to flows of a constant length, a flow is different from the plain concatenation of its packet features. 
+If we, for example, consider a classification problem as in this document, we expect that after a certain amount of time steps the classifier is able to perform the classification with a certain confidence.
+Arguably, features at the beginning of a flow are thus more important for the classifier.
+Giving differences at the end of a sequence the same weight as differences at its beginning, therefore seems incorrect.
+
+\item
+\textbf{Multiple prediction outputs.}
+Often times a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
+
+The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
+Not only does this approach require a substantially lower amount of computational ressources compared to predictions occuring at a later time step, but we also expect the immediately occuring prediction outome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occuring many time steps later.
+However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
+\end{itemize}
+
+\subsection{Explainability Plots}
+During flow extraction we used a usual 5-tuple flow key, which distinguished flows based on the protocol they use and their source and destination port and IP address.
+Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
+
+In particular,  \autoref{fig:pdps} shows the Partial Dependency Plots (PDP) for the source and destination port features.
+If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
+\begin{equation} \label{eq:pdp}
+\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right) .
+% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
+\end{equation}
+
+As apparent from \autoref{pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
+A known issue about this approach is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+
+For this reason, Accumulated Local Effects (ALE) plots have been proposed in literature.
+The basic idea for ALE is that instead of using all samples, only the closest samples are used for assessing the effects of one feature.
+ALE plots therefore inherently require a distance measure for finding the closest samples, which, as revealed in the above discussion, in our case is not possible in a straight-forward manner.
+
+\subsection{Plots for Packet Features}
+As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
+As a first step to explaining our model's predictions, we investigate how important features at different time steps are for the model's decisions.
+Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
+
+The most interesting packet features in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
+To depict how the model comes to a decision, we determined the total minimum and maximum for these features.
+For a given flow $X$ we then varied feature values at each time step between minimum and maximum and determined the median, the first and the third quartile.
+\autoref{fig:predplots} shows several of such plots for different attack categories for IAT and packet length.
+Furthermore, the shades \autoref{fig:predplots2} depict the averaged predictions for different attack families.
+
 \section{Examining attack types}

### 2019-07-30 ex3/report/report.tex
Change report

@@
 
+\begin{table}
+\caption{Detection performance results. Random forest and Multilayer Perceptron are from the previous exercise, LSTM is this implementation.} \label{tab:performance_results}
+\begin{tabular}{l r r r} \toprule
+& RF & MLP & LSTM \\ \midrule
+Accuracy	&	0.9974 & 0.9970	& 0.9947 \\
+Precision	&	0.9965 & 0.9979 & 0.9894	\\
+Recall	&	0.9931 & 0.9903	& 0.9897 \\
+F1	&	0.9948 & 0.9941	& 0.9896 \\
+Youden	&	0.9920 & 0.9896 & 0.9861 \\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \section{Explaining Recurrent Neural Networks}
@@
 
-\subsection{Plots for Packet Features}
+\subsection{Plots for Sequences}
+
 As apparent from the above discussion, trying to explain how packet features influence the model's predictions, poses various difficulties.
@@
 
-The most interesting packet features in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
-To depict how the model comes to a decision, we determined the total minimum and maximum for these features.
-For a given flow $X$ we then varied feature values at each time step between minimum and maximum and determined the median, the first and the third quartile.
-\autoref{fig:predplots} shows several of such plots for different attack categories for IAT and packet length.
-Furthermore, the shades \autoref{fig:predplots2} depict the averaged predictions for different attack families.
-
-\section{Examining attack types}
-
-Also non-recurrent approaches to anomaly detection achieve good results so the main motivation of our work is not to achieve better detection performance but to detect attacks before the end of the flow.
-To determine whether this works, we plot the attack confidence of the classifier for each time step for each attack class.
+To determine if this hypothesis is correct, we plot the attack confidence of the classifier for each time step for each attack class.
 At each time step we average over all flows and also show the first and third quartiles.
@@
 Next, we want to investigate if our classifier just memorized its inputs.
-For this we generated plots in which we change the value of the interarrival time and packet size in each plot and see how the confidence changes.
+The most interesting packet features to visualize to in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
+To depict how the model comes to a decision, we determined the total minimum and maximum for these features.
+For a given flow $X$ we then varied feature values at each time step between minimum and maximum and determined the median, the first and the third quartile.
+\autoref{fig:predplots} shows several of such plots for different attack categories for IAT and packet length.
+Furthermore, the shades \autoref{fig:predplots2} depict the averaged predictions for different attack families.
 It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values.
@@
 Looking at the plots of the features that vary over time we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
-\todo{Add characteristic\_flows output and add flow direction to the plots}
+We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.
+Thus, we show the flow that is the easiest to classify for each attack type and the one that is the hardest to classify and check whether they are visually significantly different.
+\todo{Add characteristic\_flows output and add flow direction to the plots and explain the method.}
 
@@
 
+Both approaches lead to complete resistance to adversarial samples (as we left out all manipulable features), except for botnets, which can still operate when only leaving out manipulable features in one direction.
+The results in \autoref{performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy).
+However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference.
+Thus, apparently the interarrival time and packet size are especially important for determining whether a flow is malicious in the first packets of a flow.
+
+\begin{table*}
+\caption{Comparing performance with all features, only safe features (interarrival time and packet size removed) and safe features in one direction (interarrival time and packet size removed for packets coming from source).
+\textit{packets} means that we consider the neural network output of all packets of the flow while \textit{flows} means that we only consider the last packet.
+At the last packet the classifier is usually already a lot more confident than at the first packets of a flow and thus metrics are always better in this case.}  \label{tab:performance_results_no_manipulable}
+\newcommand{\cmidrulespace}{6pt}
+\begin{tabular}{l l l l l l l} \toprule
+& \multicolumn{2}{l}{all} & \multicolumn{2}{l}{safe features} & \multicolumn{2}{l}{safe features one direction} \\
+\cmidrule(r){2-3} \cmidrule(lr){4-5} \cmidrule(l){6-7}
+& packets & flows & packets & flows & packets & flows \\
+\midrule
+Accuracy & 0.9857054 & 0.99472587 & 0.96678023 & 0.99337596 & 0.98056073 & 0.99353386 \\
+Precision & 0.94963342 & 0.9894422 & 0.87175275 & 0.98798596 & 0.91359787 & 0.98994595 \\
+Recall & 0.96735895 & 0.98972051 & 0.94374958 & 0.98581328 & 0.97836672 & 0.9844478 \\
+F1 & 0.95841423 & 0.98958134 & 0.90632359 & 0.98689842 & 0.94487366 & 0.98718922 \\
+Youden & 0.95682951 & 0.98614231 & 0.91525628 & 0.98175163 & 0.95937772 & 0.9810602 \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
 \section{Conclusions}

### 2019-07-30 ex3/plot.py

@@ -7,6 +7,7 @@ import os
 import json
 import pickle
 
+DIR_NAME = "plots/plot"
 ORDERING = ['Real', 'Length min', 'IAT min', 'Length max', 'IAT max']
 
 with open("categories_mapping.json", "r") as f:
@@ -82,7 +83,8 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	plt.legend(all_legends)
 	plt.xlabel('Sequence index')
 	plt.ylabel('Prediction')
-	plt.xticks(range(medians.shape[0]))
+	# plt.xticks(range(medians.shape[0]))
 	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
-	plt.show()
+	# plt.show()
+	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
 

### 2019-07-31 ex3/report/report.tex
Modified report

--- 2019-07-30 ex3/report/report.tex
@@
 
+\subsection{Feature Importance Metrics}
+As a first step to understanding the neural network's decisions, we 
 \subsection{Explainability Plots}
@@
 \begin{itemize}
-\item Only some packets sent by the attacker can be manipulated by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers). 
-\item Packets must not be smaller than initially (Otherwise less information could be transmitted.).
-\item Interarrival times must not decrease (Otherwise the physical speed of data transmission can be violated in some cases.).
+\item Only some packets in a flow can be manipulated by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers). 
+\item Packets must not be smaller than initially, as otherwise less information could be transmitted or packets might even be rendered syntactically invalid.
+\item Interarrival times must not decrease, as otherwise the physical speed of data transmission can be violated in some cases.
+Since an in-depth analysis of when the reduction of interarrival times is possible, would come with substantial effort, we generally disallowed a reduction of this feature.
 \end{itemize}
+
+Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs.
+Since the reasons discussed above suggest that finding adversarial flows is particularly difficult in our case, for the present research we chose a method which produces high-quality adversarial samples while not being as fast as other methods.
+
+In particular, we implemented the method by \citeauthor{cw}~\cite{cw}, which performs gradient descent on the optimization objective 
+\begin{equation}
+d(X,\tilde X) + \epsilon  \max(f(\tilde X), 0.5).
+\end{equation}
+Here, $d(\cdot)$ is a distance metric, $\epsilon \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow.
+Furthermore, $X$ denotes the original flow and $\tilde X$ denotes the adversarial flow under optimization.
+
+We used $\ell_1$ norm as distance metric and deployed projected gradient descent for meeting the real-world constraints discussed above.
 

### 2019-07-31 ex3/report/report.tex

--- 2019-07-30 ex3/report/report.tex
@@
 
+\subsection{Feature Importance Metrics}
+As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
+
+To this end, we evaluated by what extent detection accuracy drops when withholding individual features from the classifier.
+We thus replaced values of individual features by values randomly drawn from the same distribution as the real feature values.
+We ensured that features which stay constant for a flow, i.e.
+source port, destination port and protocol, also stay constant for the distorted flows.
+
+\autoref{tab:feat_importance} shows the result.
 \subsection{Explainability Plots}
@@
 \begin{itemize}
-\item Only some packets sent by the attacker can be manipulated by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers). 
-\item Packets must not be smaller than initially (Otherwise less information could be transmitted.).
-\item Interarrival times must not decrease (Otherwise the physical speed of data transmission can be violated in some cases.).
+\item Only packets can be manipulated which are transmitted by the attacker. This means the attacker can only control one direction (except for botnets where both sides are attackers). 
+\item Packets must not be smaller than initially, as otherwise less information could be transmitted or packets might be rendered syntactically invalid.
+\item Interarrival times must not decrease, as otherwise the physical speed of data transmission can be violated in some cases.
+Since an in-depth analysis of when the reduction of interarrival times is possible, would come with substantial effort, we generally disallowed a reduction of this feature.
 \end{itemize}
+
+Several methods for generating adversarial samples have been proposed in the recent literature, achieving different speed-quality tradeoffs.
+Since the reasons discussed above suggest that finding adversarial flows is particularly difficult in our case, for the present research we preferred methods which produces high-quality adversarial samples while not being as fast as other methods.
+
+We implemented the following methods for generating adversarial flows.
+
+\subsection{Carlini-Wagner}
+In particular, we implemented the method by \citeauthor{cw}~\cite{cw}, which performs gradient descent on the optimization objective 
+\begin{equation}
+d(X,\tilde X) + \epsilon  \max(Z(\tilde X), 0).
+\end{equation}
+Here, $d(\cdot)$ is a distance metric, $\epsilon \in \mathbb R^+$ is a parameter governing the tradeoff achieved between attack success and distance from the original flow.
+Furthermore, $Z(\cdot)$ denotes the neural network's logit output, $X$ denotes the original flow and $\tilde X$ the adversarial flow under optimization.
+
+We used $\ell_1$ norm as distance metric and deployed projected gradient descent (PGD) for meeting the real-world constraints discussed above.
+
+\subsection{$\ell_\infty$-bounded Projected Gradient Descent}
+\cite{ellinfty_pgd} proposes a method for generating adversarial samples which uses PGD  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
+
+\subsection{Fast Gradient Sign Method}
+Finally, we also tested the Fast Gradient Sign Method (FGSM), which perhaps is the most frequently used method for generating adversarial sampling.
+FGSM can be considered a single pass of PGD on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e.
+the adversarial sample is found as
+\begin{equation}
+\tilde X = X + \epsilon \text{sgn}( \nabla_X L(X)),
+\end{equation}
+where $L(X)$ denotes the network's loss function and $\epsilon$ denotes the achieved $\ell_\infty$ distance.
 

### 2019-07-31 ex3/report/report.tex
Added references

@@
 \subsection{Carlini-Wagner}
-In particular, we implemented the method by \citeauthor{cw}~\cite{cw}, which performs gradient descent on the optimization objective 
+In particular, we implemented the method by \citeauthor{carlini2017towards}~\cite{carlini2017towards}, which performs gradient descent on the optimization objective 
 \begin{equation}
@@
 \subsection{$\ell_\infty$-bounded Projected Gradient Descent}
-\cite{ellinfty_pgd} proposes a method for generating adversarial samples which uses PGD  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
+\cite{madry2017towards} use a method for generating adversarial samples which uses PGD  to minimize the network's negative loss function while constraining the achieved $\ell_\infty$ distance from the original samples.
 
 \subsection{Fast Gradient Sign Method}
-Finally, we also tested the Fast Gradient Sign Method (FGSM), which perhaps is the most frequently used method for generating adversarial sampling.
+Finally, we also tested the Fast Gradient Sign Method (FGSM), which was initially proposed in \cite{goodfellow2014explaining} and perhaps is the most frequently used method for generating adversarial sampling.
 FGSM can be considered a single pass of PGD on the loss function with an equality constraint on the $\ell_\infty$ distance, i.e.

### 2019-07-31 ex3/characteristic_flows.py
Fix plots and add plot pdfs

@@ -11,6 +11,8 @@ import json
 import pickle
 from learn import numpy_sigmoid
 
+DIR_NAME = "plots/characteristic_flows"
+
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
@@ -90,7 +92,12 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 	all_labels = [item.get_label() for item in all_legends]
 	ax1.legend(all_legends, all_labels, loc=0)
 	plt.xlabel('Sequence index')
-	plt.show()
+	plt.tight_layout()
+	# plt.show()
+
+	os.makedirs(DIR_NAME, exist_ok=True)
+	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.clf()
 
 
 

### 2019-07-31 ex3/plot2.py

@@ -3,6 +3,8 @@
 import matplotlib.pyplot as plt
 import matplotlib
 import matplotlib.colors as mc
+# from matplotlib import rcParams
+# rcParams.update({'figure.autolayout': True})
 import colorsys
 import numpy as np
 import sys
@@ -10,6 +12,8 @@ import os
 import json
 import pickle
 from learn import numpy_sigmoid
+
+DIR_NAME = "plots/plot2"
 
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
@@ -87,14 +91,15 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 
 	all_legends = []
 	# for i in range(medians.shape[1]):
-	plt.figure(1)
+	plt.figure(attack_type)
 	plt.title(reverse_mapping[attack_type])
 
 	for feature_index_from_zero, (feature_name, feature_index) in enumerate(zip(FEATURE_NAMES, (3, 4))):
 		# if feature_index_from_zero > 0:
 		# 	continue
 		plt.subplot("{}{}{}".format(len(FEATURE_NAMES), 1, feature_index_from_zero+1))
-		plt.xlabel('Sequence index')
+		if feature_index_from_zero == len(FEATURE_NAMES)-1:
+			plt.xlabel('Sequence index')
 		plt.ylabel(feature_name)#, color=colors[feature_index_from_zero])
 
 		legend = "{}".format(feature_name)
@@ -110,15 +115,19 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 		all_legends += ret
 		# print("legend", legend)
 
-	plt.figure(1)
+	plt.figure(attack_type)
 	plt.suptitle(reverse_mapping[attack_type])
+	plt.tight_layout()
+	plt.subplots_adjust(top=0.935)
 
 	# print("all_legends", all_legends)
 	# all_labels = [item.get_label() for item in all_legends]
 	# plt.legend(all_legends, all_labels, loc=0)
 	# plt.xticks(range(actual_flow_means.shape[0]))
 	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
-	plt.show()
-
+	# plt.show()
+	os.makedirs(DIR_NAME, exist_ok=True)
+	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.clf()
 
 

### 2019-07-31 ex3/plot_adv.py

@@ -8,6 +8,8 @@ import json
 import pickle
 from learn import numpy_sigmoid
 
+DIR_NAME = "plots/plot_adv"
+
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
@@ -113,9 +115,14 @@ for attack_type, (results_by_attack_number_item, orig_results_by_attack_number_i
 	all_labels = [item.get_label() for item in all_legends]
 	ax1.legend(all_legends, all_labels, loc=0)
 	plt.xlabel('Sequence index')
+	plt.tight_layout()
 	# plt.xticks(range(medians.shape[0]))
 	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
-	plt.show()
+	# plt.show()
+
+	os.makedirs(DIR_NAME, exist_ok=True)
+	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.clf()
 
 
 

### 2019-07-31 ex3/plot_pdp.py

@@ -7,6 +7,8 @@ import os
 import json
 import pickle
 
+DIR_NAME = "plots/plot_pdp"
+
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
 categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
@@ -38,6 +40,10 @@ for attack_type, all_features in enumerate(results_by_attack_number):
 	plt.legend(all_legends)
 	plt.xlabel('Feature')
 	plt.ylabel('Probability')
+	plt.tight_layout()
 	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
-	plt.show()
+	# plt.show()
 
+	os.makedirs(DIR_NAME, exist_ok=True)
+	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.clf()

### 2019-07-31 ex3/report/report.tex
Added feature importances

@@
 \subsection{Feature Importance Metrics}
+\begin{table}
+\caption{Accuracy drop when replacing feature values with random data.}
+\label{tab:feat_importance}
+\begin{tabular}{l r}
+\toprule
+Feature & Accuracy drop \\ \midrule
+Protocol	&	0.1323	\\
+PacketLength	&	0.1203	\\
+SYN Flag	&	0.1067	\\
+ACK Flag	&	0.0805	\\
+Destination port	&	0.0796	\\
+PSH Flag	&	0.0767	\\
+Direction	&	0.0732	\\
+Source port	&	0.0254	\\
+Interarrival time	&	0.0200	\\
+RST Flag	&	0.0116	\\
+FIN Flag	&	0.0100	\\
+ECE	Flag &	0.0000	\\
+URG Flag	&	0.0000	\\
+CWR Flag	&	0.0000	\\
+NS Flag	&	0.0000	\\
+\bottomrule
+\end{tabular}
+\end{table}
 As a first step to understanding the neural network's decisions, we tried to estimate how important individual features are for the model's predictions.
@@
 
-\autoref{tab:feat_importance} shows the result.
+\autoref{tab:feat_importance} shows our results.
+The results match to a large extent with the expectations one has when classifying network flows.
+In particular,  the most important features turn out to be the used protocol, packet length and the SYN TCP flag.
+While the first two features are essential for characterizing flows, a high number of packets with set SYN flag is an important characteristic for port scanning or certain DoS attacks.
+
+Also the fact that most of TCP flags do not play an important role in our dataset, was expected.
+It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the network.
+
 \subsection{Explainability Plots}
@@
 
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows as additional attacks. However,
+
 \begin{table*}

### 2019-07-31 ex3/report/report.tex
New plots

--- 2019-07-31 ex3/report/report.tex
@@
 \begin{equation} \label{eq:pdp}
-\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right) .
-% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
+\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right)
 \end{equation}
@@
 As apparent from \autoref{pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-A known issue about this approach is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+
+We also experiment with a modified definition of the PDP: 
+\begin{equation} \label{eq:pdp_conditional}
+\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X' := \left\{ x \in X \middle| \text{class}\left(x\right) = c\right\}}\left(f \left( X'_1,\ldots,X'_{i-1},y,X'_{i+1},\ldots X'_n \right) \right)
+\end{equation}
+
+%\begin{figure*}[p]
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
+%
+%\caption{Examples of conditional PD plots for source and destination port.
+We only analyze these features with the PDP since it is not straightforward to define the PDP for features that can vary for each packet in a flow.}
+%\label{fig:pdp}
+%\end{figure*}
+
+In this variant of the PDP that we call conditional PDP, we don't take the whole dataset for the computation for the PDP but only a subset that belongs to a certain class.
+Specifically, we take each attack class and compute the PDP only over samples of this attack class.
+A known issue about PDPs in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
 
@@
 Intuitively, features at the beginning of a flow should be the most important while the model's predictions should not vary significantly anymore, as soon as the model has come to a decision.
+
+%\begin{figure*}[p]
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
+%
+%\caption{Median confidence per time step and first and third quartiles for selected attack types.
+For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
+%\label{fig:pdp}
+%\end{figure*}
 

### 2019-07-31 ex3/characteristic_flows.py

@@ -96,7 +96,7 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 	# plt.show()
 
 	os.makedirs(DIR_NAME, exist_ok=True)
-	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")))
 	plt.clf()
 
 

### 2019-07-31 ex3/learn.py

@@ -570,7 +570,7 @@ def adv_internal(in_training = False):
 
 		mask_pgd = (index_tensor <= selection_tensor).byte().to(device)
 		labels, _ = torch.nn.utils.rnn.pad_packed_sequence(labels)
-				
+
 		for i in range(ITERATION_COUNT):
 
 			# print("iterating", i)
@@ -872,7 +872,8 @@ def pred_plots2():
 	start_iterating = time.time()
 	# have_categories = collections.defaultdict(int)
 	for real_ind, sample_ind in zip(test_indices, range(len(subset))):
-		# print("index", sample_ind)
+		if sample_ind % 1000==0:
+			print("index", sample_ind)
 		# if have_categories[cat] == SAMPLES_PER_ATTACK:
 		# 	continue
 		# have_categories[cat] += 1
@@ -946,17 +947,19 @@ def pdp():
 	# TODO: consider fold
 	for attack_number in range(max(attack_numbers)+1):
 
+		matching = [item for item in subset if int(item[2][0,0]) == attack_number]
+		if len(matching) <= 0:
+			break
+		matching = matching[:1000]
+		good_subset = OurDataset(*zip(*matching))
+		print("len(good_subset)", len(good_subset))
+
 		print("attack_number", attack_number)
 		results_for_attack_type = []
 		for feat_name, feat_ind in zip(feature_names, (0, 1)):
 			feat_min, feat_max = minmax[feat_ind]
 
 			values = np.linspace(feat_min, feat_max, 100)
-
-			matching = [item for item in subset if int(item[2][0,0]) == attack_number]
-			if len(matching) <= 0:
-				break
-			good_subset = OurDataset(*zip(*matching))
 
 			# subset = [ torch.FloatTensor(sample) for sample in x[:opt.batchSize] ]
 
@@ -967,7 +970,9 @@ def pdp():
 				# modify the dataset directly using the return value of __getitem__().
 				# This does not modify the global dataset, which holds the data as numpy
 				# arrays.
-				for sample in good_subset:
+				for index, sample in enumerate(good_subset):
+					# if index % 1000 == 0:
+					# 	print("attack_number", attack_number, "feat_name", feat_name, "index", index)
 					for j in range(sample[0].shape[0]):
 						sample[0][j,feat_ind] = values[i]
 				outputs = eval_nn(good_subset)

### 2019-07-31 ex3/plot.py

@@ -32,7 +32,7 @@ assert lens_results == lens_indices
 colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
 for attack_type, seqs in enumerate(results_by_attack_number):
-	if len(seqs) < 10:
+	if len(seqs) <= 0:
 		continue
 
 	seqs = [seq.transpose() for seq in seqs]
@@ -88,6 +88,6 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
 	# plt.show()
 	os.makedirs(DIR_NAME, exist_ok=True)
-	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")))
 	plt.clf()
 

### 2019-07-31 ex3/plot_adv.py

@@ -121,7 +121,7 @@ for attack_type, (results_by_attack_number_item, orig_results_by_attack_number_i
 	# plt.show()
 
 	os.makedirs(DIR_NAME, exist_ok=True)
-	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")))
 	plt.clf()
 
 

### 2019-07-31 ex3/plot_features.py

@@ -13,7 +13,7 @@ import json
 import pickle
 from learn import numpy_sigmoid
 
-DIR_NAME = "plots/plot2"
+DIR_NAME = "plots/plot_features"
 
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
@@ -82,43 +82,42 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 
 	# actual_flow_medians = np.stack([np.median(np.stack(np.array([flows_by_attack_number_item[index][position,:] for index in item])), axis=0) for position, item in enumerate(indices_by_length)])
 	# print("actual_flow_medians.shape", actual_flow_medians.shape)
-	actual_flow_means = np.stack([np.mean(np.concatenate([flows_by_attack_number_item[index][position:position+1,:] for index in item]), axis=0) for position, item in enumerate(indices_by_length)])
+	actual_flow_medians = np.stack([np.median(np.concatenate([flows_by_attack_number_item[index][position:position+1,:] for index in item]), axis=0) for position, item in enumerate(indices_by_length)])
+	actual_flow_first_quartiles = np.stack([np.quantile(np.concatenate([flows_by_attack_number_item[index][position:position+1,:] for index in item]), 0.25, axis=0) for position, item in enumerate(indices_by_length)])
+	actual_flow_third_quartiles = np.stack([np.quantile(np.concatenate([flows_by_attack_number_item[index][position:position+1,:] for index in item]), 0.75, axis=0) for position, item in enumerate(indices_by_length)])
 
-	# median_ranges = np.stack([np.median(np.stack(np.array([result_ranges_by_attack_number_item[index][position,:,:] for index in item])), axis=0) for position, item in enumerate(indices_by_length)])
-	# print("median_ranges.shape", median_ranges.shape)
+	# print("flow shapes", actual_flow_medians.shape, actual_flow_first_quartiles.shape, actual_flow_third_quartiles.shape)
 
-	mean_ranges = np.stack([np.mean(np.concatenate([result_ranges_by_attack_number_item[index][position:position+1,:,:] for index in item]), axis=0) for position, item in enumerate(indices_by_length)])
-
-	all_legends = []
 	# for i in range(medians.shape[1]):
[…]

### 2019-07-31 ex3/plot_new.py

@@ -13,7 +13,7 @@ import json
 import pickle
 from learn import numpy_sigmoid
 
-DIR_NAME = "plots/plot2"
+DIR_NAME = "plots/plot_new"
 
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
@@ -32,7 +32,11 @@ flows_by_attack_number = loaded["flows_by_attack_number"]
 result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
 sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
 features = loaded["features"]
-# print("features", features)
+
+lens_results = [len(item) for item in results_by_attack_number]
+lens_indices = [len(item) for item in sample_indices_by_attack_number]
+
+print("lens_indices", "\n".join(["{}: {}".format(reverse_mapping[attack], length) for attack, length in zip(range(len(lens_indices)), lens_indices)]))
 
 def brighten(rgb, how_much=0.0):
 	hls = list(colorsys.rgb_to_hls(*rgb))
@@ -82,43 +86,31 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 
 	# actual_flow_medians = np.stack([np.median(np.stack(np.array([flows_by_attack_number_item[index][position,:] for index in item])), axis=0) for position, item in enumerate(indices_by_length)])
 	# print("actual_flow_medians.shape", actual_flow_medians.shape)
-	actual_flow_means = np.stack([np.mean(np.concatenate([flows_by_attack_number_item[index][position:position+1,:] for index in item]), axis=0) for position, item in enumerate(indices_by_length)])
+	# print("flows", [item.shape for item in flows_by_attack_number_item])
+	# print("results", [item.shape for item in results_by_attack_number_item])
+	actual_flow_medians = np.stack([np.median(np.concatenate([results_by_attack_number_item[index][position:position+1] for index in item]), axis=0) for position, item in enumerate(indices_by_length)])
+	actual_flow_first_quartiles = np.stack([np.quantile(np.concatenate([results_by_attack_number_item[index][position:position+1] for index in item]), 0.25, axis=0) for position, item in enumerate(indices_by_length)])
[…]

### 2019-07-31 ex3/plot_pdp.py

@@ -45,5 +45,5 @@ for attack_type, all_features in enumerate(results_by_attack_number):
 	# plt.show()
 
 	os.makedirs(DIR_NAME, exist_ok=True)
-	plt.savefig(DIR_NAME+'/{}_{}.pdf'.format(file_name.split("/")[-1], attack_type))
+	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")))
 	plt.clf()

### 2019-07-31 ex3/report/report.tex
Modified report

@@
 
-We use a train/test split of 2:1.
+We use Z-score normalization and a train/test split of 2:1.
 
@@
 In particular, we implemented the method by \citeauthor{carlini2017towards}~\cite{carlini2017towards}, which performs gradient descent on the optimization objective 
-\begin{equation}
+\begin{equation} \label{eq:carliniWagner}
 d(X,\tilde X) + \epsilon  \max(Z(\tilde X), 0).
@@
 We used $\ell_1$ norm as distance metric and deployed projected gradient descent (PGD) for meeting the real-world constraints discussed above.
+
+The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic. In the present context, we
 
@@
 where $L(X)$ denotes the network's loss function and $\epsilon$ denotes the achieved $\ell_\infty$ distance.
+
+\subsection{Evaluation}
+\begin{figure*}
+\includegraphics[width=0.45\textwidth]{adv_plots/cwl1.pdf}
+\includegraphics[width=0.45\textwidth]{adv_plots/cwl2.pdf}
+\caption{Performance of the Carlini-Wagner method with different $\epsilon$ values for $\ell_1$ and $\ell_2$ distances.}
+\label{fig:adv_cw}
+\end{figure*}
+\autoref{fig:adv_cw} depicts the ratio of attack flows which were successfully misclassified as normal traffic with respect to the final packet along with the resulting distance from the original sample.
+Evidently, the larger the $\epsilon$ parameter, the better the attack works, but also the distances become larger.
+
+Hence, with acceptable distance values we were able to generate adversarial samples for about 80\% of attack samples.
+
+The figure suggests that a value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
+This conclusion might be misleading, however, as the distance cannot become larger anymore as soon as an adversarial sample has been found, when performing optimization as depicted by the objective function \eqref{eq:carliniWagner}.
+Hence, the depicted increase of average distances presumably is due to samples for which no adversarial sample was found for lower $\epsilon$ values.
+
+It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
+
+\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded PGD. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.
+
+Of the investigated algorithms, thus the Carlini-Wagner method with $\ell_2$ distance delivered the best performance.
+
+\begin{figure*}
+\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
+\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
+\caption{Performance of FSGM and $\ell_\infty$-bounded PGD for different bounds.}
+\label{fig:fgsm}
+\end{figure*}
 
@@
 
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows as additional attacks. However,
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labelled as additional attacks.
+However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.
+In fact, classification accuracy stayed about constant during the training of the network.
+
+A possible explanation for this observation might be that the adversarial samples assumed too much the shape of normal flows, so that the network could not reliable tell them apart.
 

### 2019-07-31 ex3/report/report.tex
Add explainability plots

--- 2019-07-31 ex3/report/report.tex
@@
 
-\todo{Insert accuracy and compare with RF and NN from previous homework.}
+\autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to the one of the random forest and the multilayer perceptron that we developed in the previous exercise.
+However unlike these classifier, our recurrent classifier has the advantage that it can detect attacks already before they are over.
 
@@
 
-We also experiment with a modified definition of the PDP: 
+We also experiment with a modified definition of the PDP (with C being the distribution of attack classes): 
 \begin{equation} \label{eq:pdp_conditional}
-\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X' := \left\{ x \in X \middle| \text{class}\left(x\right) = c\right\}}\left(f \left( X'_1,\ldots,X'_{i-1},y,X'_{i+1},\ldots X'_n \right) \right)
+\text{PDP}_{c,i}(y) = \mathbb E_{\boldsymbol X | C = c} \left(f \left( X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n \right) \right)
 \end{equation}
 
-%\begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
-%
-%\caption{Examples of conditional PD plots for source and destination port.
-We only analyze these features with the PDP since it is not straightforward to define the PDP for features that can vary for each packet in a flow.}
-%\label{fig:pdp}
-%\end{figure*}
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_8_DoS - DDoS-Heartbleed"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_pdp/flows_full_no_ttl_pdp_outcomes_0_3.pickle_10_Normal"}.pdf}
+
+\caption{Examples of conditional PD plots for source and destination port.
+We only analyze these features with the PDP since it is not straightforward to define the PDP for features that can vary for each packet in a flow.
+The y-axis shows the probability for malicious traffic.
+}
+\label{fig:pdp}
+\end{figure*}
 
@@
 Specifically, we take each attack class and compute the PDP only over samples of this attack class.
+The results in \autoref{fig:pdp} show that for some attack classes the port numbers play an important role.
+When looking at the class of normal traffic (non-attack traffic) it becomes apparent that traffic destined to a high source port is generally indicative of an attack.
+We argue that this is because most services that regular users use have low port numbers.
 A known issue about PDPs in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
@@
 
-%\begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/sourceTransportPort_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
-%
-%\caption{Median confidence per time step and first and third quartiles for selected attack types.
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_14_Web Attack-XSS"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot/flows_full_no_ttl_pred_plots_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+
+\caption{Median confidence per time step and first and third quartiles for selected attack types.
 For the majority of attacks, confidence increases in the first few steps and then stays almost constant at 1.}
-%\label{fig:pdp}
-%\end{figure*}
+\label{fig:confidence}
+\end{figure*}
 
@@
 At each time step we average over all flows and also show the first and third quartiles.
-\todo{Insert plots output and interpretation.
-Change plots to include a histogram.}
+\autoref{fig:confidence} shows that this hypothesis is correct for most attack classes.
+At the first couple of packets the confidence for maliciousness is not very high yet but towards the end of the flow it reaches values close to 1 and stays there.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_0_Botnet-ARES"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot2/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_1_Brute Force-FTP-Patator"}.pdf}
+
+\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\label{fig:pred_plots2}
+\end{figure*}
 
 Next, we want to investigate if our classifier just memorized its inputs.
-The most interesting packet features to visualize to in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
-To depict how the model comes to a decision, we determined the total minimum and maximum for these features.
-For a given flow $X$ we then varied feature values at each time step between minimum and maximum and determined the median, the first and the third quartile.
-\autoref{fig:predplots} shows several of such plots for different attack categories for IAT and packet length.
-Furthermore, the shades \autoref{fig:predplots2} depict the averaged predictions for different attack families.
+The most interesting packet features that can be used to visualize this in our use case are the inter-arrival time (IAT) between packets in a flows and the packet length.
+To depict how the model comes to a decision, we determined the total minimum and maximum for these features that ever occur for all flows.
+For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:predplots2}) depict the averaged predictions for different attack families.
 It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values.
 Instead, when memorizing, we would expect a sharp decline of confidence to occur.
-\todo{Insert pred\_plots2 output and interpretation}
+Another interesting observation is that after the first couple of packets the classifier is certain about its decision and changing features doesn't change its confidence any longer.
+This allows the conclusion that if an attacker is able to change the first few packets, she could possibly change the decision of the classifier.
+We will further investigate this question in section \autoref{sec:adv}.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_2_Brute Force-SSH-Patator"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_6_DoS - DDoS-DoS Slowhttptest"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_11_PortScan-PortScan - Firewall off"}.pdf}
+\includegraphics[width=0.48\textwidth]{{"../plots/plot_features/flows_full_no_ttl_pred_plots2_outcomes_0_3.pickle_7_DoS - DDoS-DoS slowloris"}.pdf}
+
+\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\label{fig:plot_features}
+\end{figure*}
 
@@
 We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.
-Thus, we show the flow that is the easiest to classify for each attack type and the one that is the hardest to classify and check whether they are visually significantly different.
-\todo{Add characteristic\_flows output and add flow direction to the plots and explain the method.}
+\autoref{fig:plot_features} clearly shows that some attack types have very characteristic patterns of packet sizes and interarrival times.
+\section{Adversarial attacks}
+\label{sec:adv}
 
-\section{Adversarial attacks}
-
-As we saw in the previous section \todo{Add reference to characteristic\_flows plot here}, attacks do have a characteristic pattern of interarrival times and packet lengths.
-Thus,  changing this pattern might confuse the classifier.
-However, this scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
+As we saw in the previous section in \autoref{fig:plot_features}, some attack types do have a characteristic pattern of interarrival times and packet lengths.
+Furthermore, \autoref{fig:pred_plots2} shows after the first couple of packets, the classifier becomes so sure in its decision, that a change in features doesn't seem to have any after the first packets.
+The hypothesis that follows is that by manipulating the interarrival times and packet lengths of the first packets, the classifier might be fooled.
+However, our scenario is quite different from how generating adversarial samples usually works: We have significantly less features than in, for example, an image.
 Furthermore, only two out of those few features can be manipulated.
-And even these features cannot manipulated freely: Besides the constraints that packet size and interarrival time cannot go beyond zero there are also more problem-specific constraints:
+And even these features cannot manipulated freely: Besides the constraints that packet size and interarrival time cannot become negative there are also more problem-specific constraints:
 \begin{itemize}
@@
 Both approaches lead to complete resistance to adversarial samples (as we left out all manipulable features), except for botnets, which can still operate when only leaving out manipulable features in one direction.
-The results in \autoref{performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy).
+The results in \autoref{tab:performance_results_no_manipulable} show that -- surprisingly -- there is only a small difference in classification performance when only computing the accuracy for the last packet of each flow (flow accuracy).
 However, when looking at the packet accuracy, which considers all packets of all flows, there is a significant difference.

### 2019-07-31 ex3/plot.py

@@ -8,7 +8,7 @@ import json
 import pickle
 
 DIR_NAME = "plots/plot"
-ORDERING = ['Real', 'Length min', 'IAT min', 'Length max', 'IAT max']
+ORDERING = ['', 'Length min', 'IAT min', 'Length max', 'IAT max']
 
 with open("categories_mapping.json", "r") as f:
 	categories_mapping_content = json.load(f)
@@ -74,8 +74,8 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 		plt.fill_between(range(medians.shape[0]), first_quartiles[:,i], third_quartiles[:,i], alpha=0.5, edgecolor=colors[i], facecolor=colors[i])
 		legend = ORDERING[i:i+1]*2
 		# legend = ORDERING[i:i+1]
-		legend[0] = legend[0]+" median"
-		legend[-1] = legend[-1]+" 1st and 3rd quartile"
+		legend[0] = " median"
+		legend[-1] = "1st and 3rd quartile"
 		all_legends += legend
 		# print("legend", legend)
 

### 2019-07-31 ex3/report/report.tex
Added per-family table

@@
 
-The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic. In the present context, we
+The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic.
+In the present context, we need to make sure that the classifer's prediction would actually be normal traffic, even though a certain level of noise will be added to interarrival times due to the network between attacker and victim.
+Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
+\begin{equation} \label{eq:carliniWagner}
+d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),
+\end{equation}
+i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic.
 
@@
 
+\autoref{tab:adv_per_family} depicts how successful the attack is for individual attack families.
+For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded PGD and a tradeoff of 0.5 for Carlini-Wagner.
+The table shows significant for detecting different attack families in the first place.
+However, also for generating adversarial flows some attack families are more susceptible than others.
+
 \begin{figure*}
@@
 \end{figure*}
+
+Finally, we investigated how adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
+
+To this end, we generated adversarial flows with the Carlini-Wagner method with $\ell_2$ distance and a tradeoff of 0.5.
+\autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
+
+\begin{figure*}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
+\caption{Successful manipulations performed by the Carlini-Wagner method.}
+\label{adv_plots}
+\end{figure*}
+
+\begin{table*}
+\caption{Attack success ratios per attack family. Depicted are flow detection accuracies for adversarial flows (left numbers) and flow detection accuracies for unmodified flows (right numbers).}
+\label{tab:adv_per_family}
+\begin{tabular}{lllllllll} 
+\toprule
+Attack type & FGSM & Carlini-Wagner, $\ell_1$ & Carlini-Wagner, $\ell_2$ & $\ell_\infty$-bounded PGD \\
+\midrule
+Botnet:ARES	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	\\
+Brute Force:FTP-Patator	&	0.2800 / 0.5800	&	0.1600 / 0.5800	&	0.1600 / 0.5800	&	0.2800 / 0.5800	\\
+Brute Force:SSH-Patator	&	0.8300 / 1.0000	&	0.0000 / 1.0000	&	0.0000 / 1.0000	&	0.8900 / 1.0000	\\
+DDoS:LOIT	&	0.8300 / 0.8600	&	0.6400 / 0.8600	&	0.5600 / 0.8600	&	0.8200 / 0.8600	\\
+DoS / DDoS:DoS GoldenEye	&	0.8900 / 1.0000	&	0.5000 / 1.0000	&	0.4200 / 1.0000	&	0.9200 / 1.0000	\\
+DoS / DDoS:DoS Hulk	&	0.6700 / 1.0000	&	0.5400 / 1.0000	&	0.4700 / 1.0000	&	1.0000 / 1.0000	\\
+DoS / DDoS:DoS Slowhttptest	&	0.8700 / 1.0000	&	0.4100 / 1.0000	&	0.1900 / 1.0000	&	0.7400 / 1.0000	\\
+DoS / DDoS:DoS slowloris	&	0.4400 / 0.9700	&	0.3100 / 0.9700	&	0.1300 / 0.9700	&	0.9200 / 0.9700	\\
+DoS / DDoS:Heartbleed	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	&	0.0000 / 0.0000	\\
+Infiltration:Dropbox download 	&	0.6100 / 0.9600	&	0.4300 / 0.9600	&	0.3800 / 0.9600	&	0.7400 / 0.9600	\\
+PortScan:PortScan - Firewall off	&	0.6800 / 0.9100	&	0.3500 / 0.9100	&	0.3500 / 0.9100	&	0.7200 / 0.9100	\\
+PortScan:PortScan - Firewall on	&	0.7400 / 0.8700	&	0.6100 / 0.8700	&	0.5300 / 0.8700	&	0.7300 / 0.8700	\\
+Web Attack:Sql Injection	&	0.0000 / 0.6923	&	0.0000 / 0.6923	&	0.0000 / 0.6923	&	0.0000 / 0.6923	\\
+Web Attack:XSS	&	0.6700 / 0.9500	&	0.7200 / 0.9500	&	0.6200 / 0.9500	&	0.6700 / 0.9500	\\
+\bottomrule
+
+\end{tabular}
+\end{table*}
 

### 2019-07-31 ex3/plot.py
Add nicer plots

@@ -67,22 +67,45 @@ for attack_type, seqs in enumerate(results_by_attack_number):
 	# quit()
 
 	all_legends = []
+	# print("values_by_length", [item.shape for item in values_by_length])
+	lens = [item.shape[0] for item in values_by_length]
+
+	fig, ax1 = plt.subplots()
+	# print("lens", lens)
+	ret = ax1.bar(range(len(lens)), lens, width=1, color="gray", alpha=0.2, label="number of samples")
+	# print("ret", ret)
+	all_legends.append(ret)
+
+	ax2 = ax1.twinx()
+
+	ax2.set_ylabel('Confidence')
+	ax1.set_ylabel("number of flows")
+
+	ax1.yaxis.tick_right()
+	ax1.yaxis.set_label_position("right")
+	ax2.yaxis.tick_left()
+	ax2.yaxis.set_label_position("left")
+
 	# for i in range(medians.shape[1]):
 	for i in range(1):
 		# print("i", i)
-		plt.plot(medians[:,i], color=colors[i])
-		plt.fill_between(range(medians.shape[0]), first_quartiles[:,i], third_quartiles[:,i], alpha=0.5, edgecolor=colors[i], facecolor=colors[i])
-		legend = ORDERING[i:i+1]*2
-		# legend = ORDERING[i:i+1]
-		legend[0] = " median"
-		legend[-1] = "1st and 3rd quartile"
-		all_legends += legend
+		ret = ax2.plot(medians[:,i], color=colors[i], label="median")
+		ret2 = ax2.fill_between(range(medians.shape[0]), first_quartiles[:,i], third_quartiles[:,i], alpha=0.5, edgecolor=colors[i], facecolor=colors[i], label="1st and 3rd quartile")
+		# legend = ORDERING[i:i+1]*2
+		# # legend = ORDERING[i:i+1]
+		# legend[0] = " median"
+		# legend[-1] = "1st and 3rd quartile"
+		# all_legends += legend
 		# print("legend", legend)
+		all_legends += ret
+		all_legends.append(ret2)
+
+	all_labels = [item.get_label() for item in all_legends]
+	ax1.legend(all_legends, all_labels)
 
 	plt.title(reverse_mapping[attack_type])
-	plt.legend(all_legends)
-	plt.xlabel('Sequence index')
-	plt.ylabel('Prediction')
+	# plt.legend(all_legends)
[…]

### 2019-07-31 ex3/plot_new.py

@@ -112,12 +112,6 @@ for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, re
 	plt.tight_layout()
 	plt.legend()
 
-	# print("all_legends", all_legends)
-	# all_labels = [item.get_label() for item in all_legends]
-	# plt.legend(all_legends, all_labels, loc=0)
-	# plt.xticks(range(actual_flow_means.shape[0]))
-	#plt.savefig('%s.pdf' % os.path.splitext(fn)[0])
-	# plt.show()
 	os.makedirs(DIR_NAME, exist_ok=True)
 	plt.savefig(DIR_NAME+'/{}_{}_{}.pdf'.format(file_name.split("/")[-1], attack_type, reverse_mapping[attack_type].replace("/", "-").replace(":", "-")))
 	plt.clf()

### 2019-07-31 ex3/plot_pdp.py

@@ -6,6 +6,7 @@ import sys
 import os
 import json
 import pickle
+from collections import Counter
 
 DIR_NAME = "plots/plot_pdp"
 
@@ -18,13 +19,26 @@ print("reverse_mapping", reverse_mapping)
 file_name = sys.argv[1]
 with open(file_name, "rb") as f:
 	loaded = pickle.load(f)
-results_by_attack_number, feature_names = loaded["results_by_attack_number"], loaded["feature_names"]
+results_by_attack_number, feature_names, feature_values_by_attack_number = loaded["results_by_attack_number"], loaded["feature_names"], loaded["feature_values_by_attack_number"]
 
 # print("results", results_by_attack_number)
 
 colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
-for attack_type, all_features in enumerate(results_by_attack_number):
+for attack_type, (all_features, all_features_values) in enumerate(zip(results_by_attack_number, feature_values_by_attack_number)):
+
+	print("attack_type", attack_type)
+	fig, ax1 = plt.subplots()
+
+	ax2 = ax1.twinx()
+
+	ax2.set_ylabel('Prediction')
+	ax1.set_ylabel("Number of samples")
+
+	ax1.yaxis.tick_right()
+	ax1.yaxis.set_label_position("right")
+	ax2.yaxis.tick_left()
+	ax2.yaxis.set_label_position("left")
 
 	if all_features is None:
 		continue
@@ -32,14 +46,30 @@ for attack_type, all_features in enumerate(results_by_attack_number):
 	all_legends = []
 	for feature_name, feature_index in zip(feature_names, range(all_features.shape[0])):
 
-		plt.plot(all_features[feature_index,0,:], all_features[feature_index,1,:], color=colors[feature_index])
-		all_legends.append(feature_name)
+		as_ints = list(all_features_values[feature_index].astype(np.int32))
+
+		# print("all_features_values[feature_index]", all_features_values[feature_index])
+		# ret1 = ax1.hist(all_features_values[feature_index], bins=range(int(round(all_features_values[feature_index].max())+1)), width=1, color=colors[feature_index], alpha=0.2, label="{} occurrence".format(feature_name))
+
+		counted = Counter(as_ints)
+		keys = counted.keys()
+		values = counted.values()
+
+		# print("keys", keys, "values", values)
[…]

### 2019-07-31 ex3/report/report.tex

@@
 
-It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
-
-\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded PGD. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.
-
-Of the investigated algorithms, thus the Carlini-Wagner method with $\ell_2$ distance delivered the best performance.
-
-\autoref{tab:adv_per_family} depicts how successful the attack is for individual attack families.
-For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded PGD and a tradeoff of 0.5 for Carlini-Wagner.
-The table shows significant for detecting different attack families in the first place.
-However, also for generating adversarial flows some attack families are more susceptible than others.
-
-\begin{figure*}
-\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
-\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
-\caption{Performance of FSGM and $\ell_\infty$-bounded PGD for different bounds.}
-\label{fig:fgsm}
-\end{figure*}
-
-Finally, we investigated how adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
-
-To this end, we generated adversarial flows with the Carlini-Wagner method with $\ell_2$ distance and a tradeoff of 0.5.
-\autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
-
-\begin{figure*}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
-\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
-\caption{Successful manipulations performed by the Carlini-Wagner method.}
-\label{adv_plots}
-\end{figure*}
-
 \begin{table*}
@@
 \end{table*}
+
+It is interesting to note that for very large values of $\epsilon$ the success ratio starts to decrease again.This phenomenon might be due to numerical effects and might be preventable by a smaller learning rate.
+
+\autoref{fig:fgsm} depicts the performance for FGSM and $\ell_\infty$-bounded PGD. As expected, while FGSM is the fastest of all algorithms, it delivers the worst performance.
+
+Of the investigated algorithms, thus the Carlini-Wagner method with $\ell_2$ distance delivered the best performance.
+
+\autoref{tab:adv_per_family} depicts how successful the attack is for individual attack families.
+For this table we used a $\ell_\infty$ bound of 0.3 for FGSM and $\ell_\infty$-bounded PGD and a tradeoff of 0.5 for Carlini-Wagner.
+The table shows significant for detecting different attack families in the first place.
+However, also for generating adversarial flows some attack families are more susceptible than others.
+
+\begin{figure*}
+\includegraphics[width=0.45\textwidth]{adv_plots/fgsm.pdf}
+\includegraphics[width=0.45\textwidth]{adv_plots/l_inf_pgd.pdf}
+\caption{Performance of FSGM and $\ell_\infty$-bounded PGD for different bounds.}
+\label{fig:fgsm}
+\end{figure*}
+
+Finally, we investigated how adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
+
+To this end, we generated adversarial flows with the Carlini-Wagner method with $\ell_2$ distance and a tradeoff of 0.5.
+\autoref{adv_plots} depicts how the method manipulates flows for some exemplary attack categories.
+Depicted is the mean of both unmodified flows and adversarial flows  of samples of individual attack families considering both interarrival time and packet length.
+
+The plots show that primarily features at the beginning of flows are modified.
+It is remarkable that it seems most promising to modifiy single packets while leaving most of the flow unchanged.
+In other cases, however, like depicted for the infiltration attack at the bottom, adversarial flows assume a complex pattern to circumvent detection.
+
+\begin{figure*}[p]
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/1.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/2.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/3.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/4.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/5.pdf}
+\includegraphics[width=0.48\textwidth]{../plots/plot_adv/6.pdf}
+\caption{Successful manipulations performed by the Carlini-Wagner method.}
+\label{adv_plots}
+\end{figure*}
 

### 2019-07-31 ex3/report/report.tex
Fixed labels
Fixed PDP text

@@
 \begin{table}[b]
-\caption{Occurence probability of attack types.} \label{tab:performance_results}
+\caption{Occurence probability of attack types.} 
 \begin{tabular}{l r} \toprule
@@
 
-In particular,  \autoref{fig:pdps} shows the Partial Dependency Plots (PDP) for the source and destination port features.
+In particular,  in literature the use of Partial Dependency Plots (PDP) is proposed.
 If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
@@
 
-As apparent from \autoref{pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-
-We also experiment with a modified definition of the PDP (with C being the distribution of attack classes): 
+To capture similarity between attack families, we in this research experiment with a modified definition of the PDP  (with C being the attack classes): 
 \begin{equation} \label{eq:pdp_conditional}
@@
 A known issue about PDPs in general is that dependencies of the feature under investigation and the remaining variables are not considered, hence potentially querying the neural network for feature combinations which cannot possibly occur.
+
+As apparent from \autoref{eq:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
 
@@
 Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
-\begin{equation} \label{eq:carliniWagner}
+\begin{equation} 
 d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),

### 2019-07-31 ex3/report/report.tex
last fixes

--- 2019-07-31 ex3/report/report.tex
@@
 There is a significant body of scientific work focusing on the detection of unwanted behavior in networks.
-In the past, a viable of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content.
+In the past, a viable way of performing intrusion detection was to inspect the content of packets themselves and detect if a packet delivers potentially harmful content.
 
@@
 Since payload is not available anymore, the focus now lies on features that are always available to observers such as port numbers, flags of packets etc.
-Network communication is usually grouped into \textit{flows}, which is usually defined as a sequence of packets between two applications that may be located at different physical locations.
+Network communication is usually grouped into \textit{flows}, which is commonly defined as a sequence of packets between two applications that may be located at different physical locations.
 
@@
 Various approaches have been proposed to extract features from flows and then perform anomaly detection with the results \cite{meghdouri_analysis_2018}.
-While this approach works well, it is problematic, that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
+While this approach works well, it is problematic that first the whole flow has to be received and only afterwards it can be determined if there was an anomaly.
 
@@
 \begin{table}[b]
-\caption{Occurence probability of attack types.} \label{tab:performance_results}
+\caption{Occurence probability of attack types.} \label{tab:occurrence}
 \begin{tabular}{l r} \toprule
@@
 \item all TCP flags (0 if flow is not TCP)
-\end{itemize*}. Of these features, source port, destination port and protocol identifier are constant over the whole flow while the other vary.
+\end{itemize*}. Of these features, source port, destination port and protocol identifier are constant over the whole flow while the others vary.
 
@@
 \autoref{tab:performance_results} shows that our classifier achieves an accuracy that is similar to the one of the random forest and the multilayer perceptron that we developed in the previous exercise.
-However unlike these classifier, our recurrent classifier has the advantage that it can detect attacks already before they are over.
+However unlike these classifiers, our recurrent classifier has the advantage that it can detect attacks already before they are over.
 
@@
 \textbf{Multiple prediction outputs.}
-Often times a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
+Often a recurrent neural network produces an output at each time step. When trying to apply explainability methods, hence, the question arises which prediction to base the methods on.
 
@@
 \subsection{Explainability Plots}
-During flow extraction we used a usual 5-tuple flow key, which distinguished flows based on the protocol they use and their source and destination port and IP address.
+During flow extraction we used a usual 5-tuple flow key, which distinguishe flows based on the protocol they use and their source and destination port and IP address.
 Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
 
-In particular,  \autoref{fig:pdps} shows the Partial Dependency Plots (PDP) for the source and destination port features.
+In particular,  \autoref{fig:pdp} shows the Partial Dependency Plots (PDP) for the source and destination port features.
 If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
@@
 
-As apparent from \autoref{pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
+As apparent from \autoref{fig:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
 
@@
 To depict how the model comes to a decision, we determined the total minimum and maximum for these features that ever occur for all flows.
-For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:predplots2}) depict the averaged predictions for different attack families.
+For a given flow $X$ we then varied feature values at each time step between minimum and maximum and plot how the confidence changes when the feature varies (\autoref{fig:pred_plots2}) depict the averaged predictions for different attack families.
 It seems clear that the classifier didn't just memorize the inputs but actually learned them since there is always a smooth decrease of confidence if we change the feature values.

### 2019-07-31 ex3/report/report.tex

@@
 
-In particular,  \autoref{fig:pdp} shows the Partial Dependency Plots (PDP) for the source and destination port features.
+In particular,  in literature the use of Partial Dependency Plots (PDP) is proposed.
 If $\boldsymbol X \in \mathbb R ^n$ denotes a random vector drawn from feature space and $f(X) \in [0,1]$ denotes the neural network's prediction, the PDP for feature $i$ can be defined as
@@
 
-As apparent from \autoref{fig:pdp}, for a certain value of the $i$th feature, the whole dataset is used for averaging the neural network's outcome.
-
-We also experiment with a modified definition of the PDP (with C being the distribution of attack classes): 
+To capture similarity between attack families, we in this research experiment with a modified definition of the PDP  (with C being the attack classes):
 \begin{equation} \label{eq:pdp_conditional}
@@
 Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
-\begin{equation} \label{eq:carliniWagner}
+\begin{equation}
 d(X,\tilde X) + \epsilon  \max(Z(\tilde X), -0.2),

### 2019-08-07 ex2/report/report.tex
Fix typos if Ex2 and small comments

@@
 \renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+
+\newcommand\note[2]{{\color{#1}#2}}
+\newcommand\todo[1]{{\note{red}{TODO: #1}}}
 
@@
 
-For research the invention and assessment of techniques for network anomaly detection poses many challenges.
+For research, the invention and assessment of techniques for network anomaly detection poses many challenges.
 To implement such a system, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection and it is not always clear which features are actually important.
+
 For the training of models usually datasets are used which have been generated artificially in a controlled test environment.
-
 As a downside of this approach, it is unclear whether a machine learning model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
@@
 The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
-In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well.
+In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well \todo{How do we see that?}.
 Hence, by averaging information gets lost and wrong explanations might be the consequence.
@@
 \end{equation}
+\todo{I think that there are some minor problems with $y$ and $X_i$ in the formula.}
 
@@
 Fig.~\ref{fig:pdp} shows examples for some features for the different folds.
-Comparing plots generated for our deep learning and for our random forest approach, respectively, the two machine learning approaches show the outstanding difference that deep learning PDPs are smooth, while random forest PDPs fluctuate heavily.
+Comparing plots generated for our deep learning and for our random forest approach, respectively, the two machine learning approaches show the difference that deep learning PDPs are smooth, while random forest PDPs fluctuate heavily.
 Considering the functioning of neural networks and decision trees, this is an expected result.
@@
 Plots depicted in Fig.~\ref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
-However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially.
+However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially \todo{Are we sure that this is not a bug?}.
 
-An important observation for ALE plots particularly for random forests is that influce of TTL values is even more distinctive.
+An important observation for ALE plots particularly for random forests is that influence of TTL values is even more distinctive.
 In fact, judging from ALE plots the random forest seems to almost exclusively consider the TTL for detecting attacks.
@@
 
-\caption{Examples for ICE plots for deep learning (left side) and random forest (right side).}
+\caption{Examples for ICE plots for deep learning (left side) and random forest (right side) \todo{Fixme: No title!}.}
 \label{fig:ice}
@@
 
-The fact that results for the training process.
-with predictions are very similar to original data is due to the good detection performance of random forests.
+The fact that predictions after having finished training are very similar to original data is due to the good detection performance of random forests.
 Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.
@@
 Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On the one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+On one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
 

### 2019-08-07 ex3/report/report.tex
Fix typos if Ex2 and small comments

@@
 Protocol	&	0.1323	\\
-PacketLength	&	0.1203	\\
+Packet Length	&	0.1203	\\
 SYN Flag	&	0.1067	\\
@@
 \subsection{Explainability Plots}
-During flow extraction we used a usual 5-tuple flow key, which distinguishe flows based on the protocol they use and their source and destination port and IP address.
+During flow extraction we used a usual 5-tuple flow key, which distinguishes flows based on the protocol they use and their source and destination port and IP address.
 Hence, these features are constant for all packets in one flow, which allows us to use explainability methods which have been proposed for non-recurrent classification methods.
@@
 
-\caption{The solid line shows the mean value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
 \label{fig:plot_features}

### 2019-08-07 ex3/report/report.tex
Fix some typos for exercise 3

@@
 \maketitle
+
+\todo{Legend in some figures is misplaced}
 
@@
 Also the fact that most of TCP flags do not play an important role in our dataset, was expected.
-It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the network.
+It is interesting to note, however, that the interarrival time of packets does not play a crucial role for the detection of attacks.
 
@@
 
-To capture similarity between attack families, we in this research experiment with a modified definition of the PDP  (with C being the attack classes):
+To inspect attack types in detail, in this research we experiment with a modified definition of the PDP  (with C being the attack classes):
 \begin{equation} \label{eq:pdp_conditional}
@@
 
-\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region shows the change in confidence that occurs when the feature is varied.}
+\caption{The solid line shows the median value of the feature for all flows of a specific attack type. The shaded region are the first and third quartile.}
 \label{fig:plot_features}
@@
 
-Looking at the plots of the features that vary over time we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
+Looking at the plots of the features that vary over time (\autoref{fig:plot_features}) we see that certain attack types have a characteristic pattern in which they send packets by which they are easily recognizable.
 We want to determine how the pattern of packet sizes and interarrival times is significant for the classification.
@@
 \end{equation}
-i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic.
+i.e. we require a slight level of confidence that our adversarial samples are not classified as attack traffic \todo{What is 0.2 in probability and not in logits? Logits are hard to imagine}.
 
@@
 
-The figure suggests that a value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
+The figure suggests that a tradeoff value between 0.5 and 1 would be ideal to find adversarial samples with a low enough distance from real samples.
 This conclusion might be misleading, however, as the distance cannot become larger anymore as soon as an adversarial sample has been found, when performing optimization as depicted by the objective function \eqref{eq:carliniWagner}.
@@
 
-Finally, we investigated how adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
+Finally, we investigated what adversarial flows look like and which features are manipulated to circumvent detection by the neural network.
 

### 2019-08-07 ex2/report/report.tex
Fix more typos

@@
 As a first preprocessing step we extracted flow features from the dataset.
-We used features specificed by the CAIA feature vector~\cite{Williams2006} and slightly enhanced the included features.
+We used features specified by the CAIA feature vector~\cite{Williams2006} and slightly enhanced the included features.
 Table~\ref{tab:features} lists the features that are included for each flow.
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are constistant with previous work like, e.g., \cite{fares}.
+Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are consistent with previous work like, e.g., \cite{fares}.
 
@@
 Considering the explanations the PD plots provide, consistency between deep learning and random forest can be observed for some features, while for other features both methods provide substantially different explanations.
-For example, the two upper plots in Fig.~\ref{fig:pdp} exhibit a very similar behaviour, while the lower plots show remarkable differences between both methods.
+For example, the two upper plots in Fig.~\ref{fig:pdp} exhibit a very similar behavior, while the lower plots show remarkable differences between both methods.
 An additional observation is, however, that random forests show substantial differences between the folds.
@@
 
-ICE plots for most features we considered have a similar appearance as the exmaples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
+ICE plots for most features we considered have a similar appearance as the examples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
 
@@
 We trained a logistic regression classifier to the predictions our models made.
-Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original lables, respectively.
+Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original labels, respectively.
 As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}.
@@
 Few features exhibit significantly high values for logistic regression coefficients.
-We compared this dependence with the corresponding PD plots and came to the conclusion that flows for these features have a high probability for being attack traffic for values near zero, while  for larger value the PD plots decend steeply and quickly assume probabilities near zero.
+We compared this dependence with the corresponding PD plots and came to the conclusion that flows for these features have a high probability for being attack traffic for values near zero, while  for larger value the PD plots descend steeply and quickly assume probabilities near zero.
 Apparantly, a logistic regression classifier is not able to show a feature dependence as visible from more sophisticated PD plots.

### 2019-08-07 ex3/report/report.tex
Fix more typos

@@
 The natural choice is to base the methods on the prediction output which occurs at the same time step as the feature under investigation.
-Not only does this approach require a substantially lower amount of computational ressources compared to predictions occuring at a later time step, but we also expect the immediately occuring prediction outome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occuring many time steps later.
+Not only does this approach require a substantially lower amount of computational resources compared to predictions occurring at a later time step, but we also expect the immediately occurring prediction outcome to exhibit a stronger correlation to the feature under investigation than a prediction outcome which occurring many time steps later.
 However, due to the complex decision processes that can occur for deep neural networks, it is absolutely  uncertain if the network's final prediction is correctly depicted when following this approach.
@@
 The method minimizes the adversarial sample's distance so that the network's decision is just between attack and normal traffic.
-In the present context, we need to make sure that the classifer's prediction would actually be normal traffic, even though a certain level of noise will be added to interarrival times due to the network between attacker and victim.
+In the present context, we need to make sure that the classifier's prediction would actually be normal traffic, even though a certain level of noise will be added to interarrival times due to the network between attacker and victim.
 Instead of the above equation \ref{eq:carliniWagner} we therefore used the optimization objective
@@
 The plots show that primarily features at the beginning of flows are modified.
-It is remarkable that it seems most promising to modifiy single packets while leaving most of the flow unchanged.
+It is remarkable that it seems most promising to modify single packets while leaving most of the flow unchanged.
 In other cases, however, like depicted for the infiltration attack at the bottom, adversarial flows assume a complex pattern to circumvent detection.
@@
 
-In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labelled as additional attacks.
+In addition to omitting manipulable features we tried to make our network more robust against adversarial samples by augmenting the training set by adversarial flows generated using the Carlini-Wagner method, labeled as additional attacks.
 However, against our expectations, this approach did not lead to satisfactory results in the sense that the network would learn to classify adversarial samples as attacks.

### 2019-08-08 ex2/ice.py
Improved ICE plotting

@@ -31,12 +31,12 @@ def ice(data, eval_function, features, means, stds, resolution=100, n_data=10, s
 		rescaled = np.linspace(minimum_rescaled, maximum_rescaled, num=resolution)
 		os.makedirs(DIR_NAME, exist_ok=True)
 		np.save('%s/%s%s.npy' % (DIR_NAME, feature, suffix), np.vstack((rescaled,ices[i,:,:].transpose())))
-		for k in range(n_data):
-			plt.plot(rescaled, ices[i,:,k])
-		plt.xlabel('Feature')
-		plt.ylabel('Mean probability')
-		plt.savefig('%s/%s%s.pdf' % (DIR_NAME, feature, suffix))
-		plt.close()
+		#for k in range(n_data):
+			#plt.plot(rescaled, ices[i,:,k])
+		#plt.xlabel('Feature')
+		#plt.ylabel('Mean probability')
+		#plt.savefig('%s/%s%s.pdf' % (DIR_NAME, feature, suffix))
+		#plt.close()
 
 if __name__=="__main__":
 	data = pd.read_csv(sys.argv[1]).fillna(0)

### 2019-08-08 ex2/report/report.tex
Fix learn.py and delete plots in report

@@
 \maketitle
+
+\todo{We should only add the backdoor in forward direction actually. But it doesn't really matter I guess.}
 
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(mean(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(mean(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
 
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{plots/ale/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/ale/sourceTransportPort_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ale/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ale/sourceTransportPort_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/ale/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/ale/apply(mean(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ale/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ale/apply(mean(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/ale/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/ale/apply(stdev(ipTotalLength),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ale/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ale/apply(stdev(ipTotalLength),forward)_rf.pdf}
 
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{plots/ice/sourceTransportPort_nn_0.pdf}
-\includegraphics[width=0.48\textwidth]{plots/ice/sourceTransportPort_rf_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_nn_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_rf_0.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/ice/apply(mean(ipTTL),forward)_nn_0.pdf}
-\includegraphics[width=0.48\textwidth]{plots/ice/apply(mean(ipTTL),forward)_rf_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_nn_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_rf_0.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/ice/apply(stdev(ipTotalLength),forward)_nn_0.pdf}
-\includegraphics[width=0.48\textwidth]{plots/ice/apply(stdev(ipTotalLength),forward)_rf_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_nn_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_rf_0.pdf}
 
@@
 \begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{plots/pdp/sourceTransportPort_rf_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_rf_bd.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
-\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}
 
-%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{plots/pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
 

### 2019-08-08 ex2/learn.py

@@ -94,16 +94,10 @@ csv_name = opt.dataroot
 df = pd.read_csv(csv_name, nrows=MAX_ROWS).fillna(0)
 df = df[df['flowDurationMilliseconds'] < 1000 * 60 * 60 * 24 * 10]
 
-try:
-	del df['flowStartMilliseconds']
-	del df['sourceIPAddress']
-	del df['destinationIPAddress']
-	attack_vector = np.array(list(df['Attack']))
-	del df['Attack']
-except KeyError:
-	pass
-
-features = df.columns[:-1]
+del df['flowStartMilliseconds']
+del df['sourceIPAddress']
+del df['destinationIPAddress']
+attack_vector = np.array(list(df['Attack']))
 
 print("Rows", df.shape[0])
 
@@ -130,7 +124,11 @@ if opt.backdoor:
 	print("backdoored_records rows", backdoored_records.shape[0])
 
 	df = pd.concat([df, backdoored_records], axis=0, ignore_index=True, sort=False)
+	# print("backdoored_records", backdoored_records)
+	attack_vector = np.concatenate((attack_vector, np.array(list(backdoored_records['Attack']))))
 
+del df['Attack']
+features = df.columns[:-1]
 print("Final rows", df.shape)
 # df[:1000].to_csv("exported_2.csv")
 
@@ -138,7 +136,9 @@ shuffle_indices = np.array(list(range(df.shape[0])))
 random.shuffle(shuffle_indices)
 
 data = df.values
+print("data.shape", data.shape)
 data = data[shuffle_indices,:]
+print("attack_vector.shape", attack_vector.shape)
 attack_vector = attack_vector[shuffle_indices]
 columns = list(df)
 print("columns", columns)
@@ -419,9 +419,12 @@ if __name__=="__main__":
 			rf = pickle.load(open(opt.net, 'rb'))
 		else:
 			rf = RandomForestClassifier(n_estimators=100)
-			rf.fit (x[train_indices,:], y[train_indices,0])
-			summed_up = np.sum(rf.predict(x[train_indices,:]), axis=1)
-			assert (np.isclose(summed_up)).all(), "summed_up: {}".format(summed_up.tolist())
+			rf.fit(x[train_indices,:], y[train_indices,0])
+			# XXX: The following code is broken! It should use predict_proba instead of predict probably
+			predictions = rf.predict_proba(x[train_indices,:])
+			# print("predictions", predictions.shape, predictions)
[…]

### 2019-09-01 ex2/learn.py
Added saving of scores

@@ -34,18 +34,19 @@ import matplotlib.pyplot as plt
 from matplotlib.colors import Normalize
 
 def output_scores(y_true, y_pred, only_accuracy=False):
-	accuracy = accuracy_score(y_true, y_pred)
+	metrics = [ accuracy_score(y_true, y_pred) ]
 	if not only_accuracy:
-		precision = precision_score(y_true, y_pred)
-		recall = recall_score(y_true, y_pred)
-		f1 = f1_score(y_true, y_pred)
-		youden = balanced_accuracy_score(y_true, y_pred, adjusted=True)
-	metrics = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden'] if not only_accuracy else ["Accuracy"]
-	print (('{:>11}'*len(metrics)).format(*metrics))
-	if not only_accuracy:
-		print ((' {:.8f}'*len(metrics)).format(accuracy, precision, recall, f1, youden))
-	else:
-		print ((' {:.8f}'*len(metrics)).format(accuracy))
+		metrics.extend([
+			precision_score(y_true, y_pred),
+			recall_score(y_true, y_pred),
+			f1_score(y_true, y_pred),
+			balanced_accuracy_score(y_true, y_pred, adjusted=True)
+		])
+	names = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden'] if not only_accuracy else ["Accuracy"]
+	print (('{:>11}'*len(names)).format(*names))
+	print ((' {:.8f}'*len(metrics)).format(*metrics))
+	return { name: metric for name, metric in zip(names, metrics) }
+	
 
 def add_backdoor(datum: dict, direction: str) -> dict:
 	datum = datum.copy()
@@ -257,15 +258,15 @@ def test_nn():
 		eval_nn(good_test_indices)
 		
 		print ('Backdoored data')
-		eval_nn(bad_test_indices)
+		eval_nn(bad_test_indices, only_accuracy=True)
 
-def eval_nn(test_indices):
+def eval_nn(test_indices, only_accuracy=False):
 	# if test_indices is None:
 	# 	test_indices = list(range(len(dataset)))
 
 	all_predictions = predict(test_indices, net=net)
 	all_labels = y[test_indices,0]
-	output_scores(all_labels, all_predictions)
+	output_scores(all_labels, all_predictions, only_accuracy)
 
 def get_layers_by_type(model, name):
[…]

### 2019-09-01 ex2/paper/paper.tex

--- 2019-08-08 ex2/report/report.tex
@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[sigconf,nonacm]{acmart}
+\documentclass[10pt,sigconf,letterpaper]{acmart}
+\acmYear{2019}
+\copyrightyear{2019}
+\acmConference{CoNEXT '19}{December 9-12, 2019}{Orlando, Florida, USA}
 
 \usepackage{multirow}
+\usepackage{blindtext}
 
-\settopmatter{printacmref=false} % Removes citation information below abstract
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+%\settopmatter{printacmref=false} % Removes citation information below abstract
+%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
 
@@
 
+\begin{abstract}
+    \blindtext
+\end{abstract}
+
 %%
@@
 %% used to denote shared contribution to the research.
-\author{Alexander Hartl (e01125115)}
-\authornote{Both authors contributed equally to this research.}
-%\email{alexander.hartl@tuwien.ac.at}
+\author{Maximilian Bachl}
+\affiliation{\institution{Technische Universität Wien}}
+%\authornote{Both authors contributed equally to this research.}
+\email{maximilian.bachl@tuwien.ac.at}
 
-\author{Maximilian Bachl (e01100143)}
-%\authornote[1]{Both authors contributed equally to this research.}
-\authornotemark[1]
-%\email{maximilian.bachl@gmail.com}
+\author{Alexander Hartl}
+\affiliation{\institution{Technische Universität Wien}}
+%\authornotemark[1]
+\email{alexander.hartl@tuwien.ac.at}
+
+\author{Tanja Zseby}
+\affiliation{\institution{Technische Universität Wien}}
+\email{tanja.zseby@tuwien.ac.at}
+
+\author{Joachim Fabini}
+\affiliation{\institution{Technische Universität Wien}}
+\email{joachim.fabini@tuwien.ac.at}
 
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_nn_0.pdf}
-\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_rf_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_nn_0.pdf}
-\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_rf_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_nn_0.pdf}
-\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_rf_0.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_rf.pdf}
 

### 2019-09-02 ex2/learn.py
Use different folders for datasets

@@ -112,7 +112,7 @@ def get_logdir(fold, n_fold):
 	return os.path.join('runs', current_time + '_' + socket.gethostname() + "_" + str(fold) +"_"+str(n_fold))
 
 def surrogate(predict_fun):
-	os.makedirs('surrogate', exist_ok=True)
+	os.makedirs('surrogate%s' % dirsuffix, exist_ok=True)
 	train_indices, test_indices = get_nth_split(dataset, opt.nFold, opt.fold)
 
 	logreg = LogisticRegression(solver='liblinear')
@@ -126,7 +126,7 @@ def surrogate(predict_fun):
 	output_scores(y_true, predictions)
 
 	print ("Coefficients:", logreg.coef_)
-	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate/logreg_pred%s.csv' % suffix)
+	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate%s/logreg_pred%s.csv' % (dirsuffix, suffix))
 
 
 	logreg = LogisticRegression(solver='liblinear')
@@ -140,7 +140,7 @@ def surrogate(predict_fun):
 	output_scores(y_true, predictions)
 
 	print ("Coefficients:", logreg.coef_)
-	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate/logreg_real%s.csv' % suffix)
+	pd.Series(logreg.coef_[0], features).to_frame().to_csv('surrogate%s/logreg_real%s.csv' % (dirsuffix, suffix))
 
 def closest(prediction_function):
 	n_fold = opt.nFold
@@ -204,7 +204,7 @@ def train_nn(finetune=False):
 			accuracy = torch.mean((torch.round(torch.sigmoid(output.detach().squeeze())) == labels.squeeze()).float())
 			writer.add_scalar("accuracy", accuracy, samples)
 
-		torch.save(net.state_dict(), '%s/net_%d.pth' % (writer.logdir, samples))
+		torch.save(net.state_dict(), '%s/net_%d.pth' % (writer.log_dir, samples))
 
 def predict(test_indices, net=None, good_layers=None, correlation=False):
 	test_data = torch.utils.data.Subset(dataset, test_indices)
@@ -511,8 +511,8 @@ def prune_backdoor_nn():
 		
[…]

### 2019-09-02 ex2/pdp.py

@@ -12,7 +12,7 @@ import matplotlib.pyplot as plt
 
 DIR_NAME = "pdp"
 
-def pdp(data, eval_function, features, means, stds, resolution=100, n_data=100, suffix=''):
+def pdp(data, eval_function, features, means, stds, resolution=100, n_data=100, suffix='', dirsuffix=''):
 
 	index = np.random.permutation(data.shape[0])[:n_data]
 	downsampled_data = data[index,:]
@@ -30,8 +30,8 @@ def pdp(data, eval_function, features, means, stds, resolution=100, n_data=100, 
 			pdps[i,j_index] = np.mean(eval_function(dd_cpy)[:,0])
 			
 		rescaled = np.linspace(minimum_rescaled, maximum_rescaled, num=resolution)
-		os.makedirs(DIR_NAME, exist_ok=True)
-		np.save('%s/%s%s.npy' % (DIR_NAME, feature, suffix), np.vstack((rescaled,pdps[i,:])))
+		os.makedirs(DIR_NAME + dirsuffix, exist_ok=True)
+		np.save('%s%s/%s%s.npy' % (DIR_NAME, dirsuffix, feature, suffix), np.vstack((rescaled,pdps[i,:])))
 
 		#plt.plot(rescaled, pdps[i,:])
 		#plt.xlabel('Feature')

### 2019-09-02 ex2/learn.py
Add preliminary abstract and title

@@ -31,20 +31,24 @@ import collections
 import pickle
 
 import matplotlib.pyplot as plt
+from matplotlib.colors import Normalize
+
+import scipy.stats
 
 def output_scores(y_true, y_pred, only_accuracy=False):
-	accuracy = accuracy_score(y_true, y_pred)
+	metrics = [ accuracy_score(y_true, y_pred) ]
 	if not only_accuracy:
-		precision = precision_score(y_true, y_pred)
-		recall = recall_score(y_true, y_pred)
-		f1 = f1_score(y_true, y_pred)
-		youden = balanced_accuracy_score(y_true, y_pred, adjusted=True)
-	metrics = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden'] if not only_accuracy else ["Accuracy"]
-	print (('{:>11}'*len(metrics)).format(*metrics))
-	if not only_accuracy:
-		print ((' {:.8f}'*len(metrics)).format(accuracy, precision, recall, f1, youden))
-	else:
-		print ((' {:.8f}'*len(metrics)).format(accuracy))
+		metrics.extend([
+			precision_score(y_true, y_pred),
+			recall_score(y_true, y_pred),
+			f1_score(y_true, y_pred),
+			balanced_accuracy_score(y_true, y_pred, adjusted=True)
+		])
+	names = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden'] if not only_accuracy else ["Accuracy"]
+	print (('{:>11}'*len(names)).format(*names))
+	print ((' {:.8f}'*len(metrics)).format(*metrics))
+	return { name: metric for name, metric in zip(names, metrics) }
+
 
 def add_backdoor(datum: dict, direction: str) -> dict:
 	datum = datum.copy()
@@ -165,11 +169,14 @@ def closest(prediction_function):
 # Deep Learning
 ############################
 
-def train_nn():
+def train_nn(finetune=False):
 	n_fold = opt.nFold
 	fold = opt.fold
 
-	train_indices, _ = get_nth_split(dataset, n_fold, fold)
+	if finetune:
+		train_indices, _, _ = get_indices_for_backdoor_pruning()
+	else:
+		train_indices, _ = get_nth_split(dataset, n_fold, fold)
 	train_data = torch.utils.data.Subset(dataset, train_indices)
 	train_loader = torch.utils.data.DataLoader(train_data, batch_size=opt.batchSize, shuffle=True)
 
[…]

### 2019-09-02 ex2/paper/paper.tex
Nothing really

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Assignment 2}
+\title{Bricking up and overgrowing backdoors in Intrusion Detection Systems}
 
 \begin{abstract}
-    \blindtext
+Interest in poisoning attacks and backdoors recently resurfaced for Deep Learning (DL) applications.
+Several successful defense mechanisms have been recently proposed for Convolutional Neural Networks, for example in the context of autonomous driving.
+We show that, surprisingly, common defense mechanisms fail utterly to remove backdoors in DL for Intrusion Detection Systems.
+However, we find that visualization approaches can aid in identifying a backdoor independent of the used classifier.
+Finally, we conceive pruning-based approaches to remove backdoors for decision trees and random forests and demonstrate their effectiveness for several network security datasets.
 \end{abstract}
@@
 %% of authors' names for this purpose.
-\renewcommand{\shortauthors}{Hartl and Bachl}
+%\renewcommand{\shortauthors}{Hartl and Bachl}
 

### 2019-09-02 ex2/learn.py

@@ -48,7 +48,7 @@ def output_scores(y_true, y_pred, only_accuracy=False):
 	print (('{:>11}'*len(names)).format(*names))
 	print ((' {:.8f}'*len(metrics)).format(*metrics))
 	return { name: metric for name, metric in zip(names, metrics) }
-	
+
 
 def add_backdoor(datum: dict, direction: str) -> dict:
 	datum = datum.copy()
@@ -460,7 +460,7 @@ def prune_backdoor_nn():
 		# where_it_would_be_sorted = np.array(list(zip(*sorted(enumerate(mean_activation_per_neuron), key=lambda key: key[1])))[0])
 		where_it_would_be_sorted = np.zeros(len(mean_activation_per_neuron), dtype=int)
 		where_it_would_be_sorted[sorted_by_activation] = np.arange(len(where_it_would_be_sorted))
-		
+
 		concatenated_results = np.concatenate(results)
 		# correlation_between_step_when_pruned_and_correlation_with_backdoor = np.corrcoef(where_it_would_be_sorted, np.abs(concatenated_results))[0,1]
 
@@ -468,8 +468,6 @@ def prune_backdoor_nn():
 
 		print("Correlation between the step during which a neuron is pruned and the absolute value of its correlation with the backdoor", correlation_between_step_when_pruned_and_correlation_with_backdoor[0])
 		print("If the method worked, it would be (significantly) less than zero. The p-values is", correlation_between_step_when_pruned_and_correlation_with_backdoor[1])
-
-		import pdb; pdb.set_trace()
 
 	step_width = 1/(opt.nSteps+1)
 
@@ -508,7 +506,7 @@ def prune_backdoor_nn():
 		predicted_bad = predict(bad_test_indices, net=new_nn)
 		ground_truth_bad = y[bad_test_indices,0]
 		scoresbd.append(output_scores(ground_truth_bad, predicted_bad, only_accuracy=True))
-		
+
 	scores = { name: [ score[name] for score in scores ] for name in scores[0] }
 	scoresbd = { name: [ score[name] for score in scoresbd ] for name in scoresbd[0] }
 	os.makedirs('prune', exist_ok=True)
@@ -816,7 +814,7 @@ def prune_backdoor_rf():
 		scores.append(output_scores(y[good_test_indices,0], new_rf.predict(x[good_test_indices,:])))
 		print("backdoored")
[…]

### 2019-09-04 ex2/paper/paper.tex
Changed paper

@@
 \author{Maximilian Bachl}
-\affiliation{\institution{Technische Universität Wien}}
+\affiliation{\institution{Technische Universit\"at Wien}}
 %\authornote{Both authors contributed equally to this research.}
@@
 \author{Alexander Hartl}
-\affiliation{\institution{Technische Universität Wien}}
+\affiliation{\institution{Technische Universit\"at Wien}}
 %\authornotemark[1]
@@
 \author{Tanja Zseby}
-\affiliation{\institution{Technische Universität Wien}}
+\affiliation{\institution{Technische Universit\"at Wien}}
 \email{tanja.zseby@tuwien.ac.at}
@@
 \author{Joachim Fabini}
-\affiliation{\institution{Technische Universität Wien}}
+\affiliation{\institution{Technische Universit\"at Wien}}
 \email{joachim.fabini@tuwien.ac.at}
@@
 
+\section{Related Work}
+
 \section{Anomaly Detection Approaches} \label{sec:ml_approaches}
-We used the CIC-IDS-2017~\cite{Sharafaldin2018} for our evaluations.
+\subsection{Datasets}
+Several datasets for the purpose of building and evaluating IDSs have been developed.
+However, as pointed out in \cite{gharib2016}, there are numerous requirements that have to be met for a dataset to provide realistic performance benchmarks in this context.
+
+In this research, we use the UNSW-NB15~\cite{moustafa2015unsw} and the CIC-IDS-2017~\cite{Sharafaldin2018} datasets, which were developed by two independent institutions and are both freely available on the Internet, guaranteeing reproducibility of our results.
+
+The UNSW-NB15 dataset~\cite{moustafa2015unsw} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
+Network captures containing over 2 million flows of normal traffic and various types of attacks are provided together with a ground truth file.
+Attack traffic includes exploits, fuzzers, reconnaissance attacks, DoS attacks, shellcode, analysis attacks, backdoors and worms.
+
+The CIC-IDS-2017 dataset~\cite{Sharafaldin2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
+The provided raw network captures contain over 2.3 million flows, containing both normal traffic and attacks, which can be coarsely classified into DoS attacks, infiltration attacks, web attacks, brute force and scanning attacks.
+
+We used the go-flows~\cite{goflows} flow exporter for extracting these features from the raw capture files.
+
 As a first preprocessing step we extracted flow features from the dataset.
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../pdp/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp/apply(mean(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
 
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{../ale/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ale/sourceTransportPort_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../ale/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ale/apply(mean(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../ale/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ale/apply(stdev(ipTotalLength),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
 
@@
 \begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ice/sourceTransportPort_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ice/apply(mean(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ice/apply(stdev(ipTotalLength),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
 
@@
 \begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp/sourceTransportPort_rf_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_rf_bd.pdf}
 
-\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_rf.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf.pdf}
 
-\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_nn_bd.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTTL),forward)_rf_bd.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn_bd.pdf}
+\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf_bd.pdf}
 
-%\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
 

### 2019-09-04 ex2/paper/paper.tex
Modified paper

@@
 
-We used the go-flows~\cite{goflows} flow exporter for extracting these features from the raw capture files.
+We base our analysis on the CAIA~\cite{Williams2006} feature vector as formulated in~\cite{fares}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
+All features except protocol and flow duration are evaluated for forward and backward direction separately.
+The goal of this research is to investigate the possibility of poisoning attacks.
+We therefore also include the minimum, maximum and standard deviation of TTL values in our feature vector as an attractive carrier for exploitation as a backdoor.
+% TODO: justify use of TTL better
 
-As a first preprocessing step we extracted flow features from the dataset.
-We used features specified by the CAIA feature vector~\cite{Williams2006} and slightly enhanced the included features.
-Table~\ref{tab:features} lists the features that are included for each flow.
+We used the go-flows~\cite{goflows} flow exporter for extracting these features from the raw capture files, applied Z-score normalization to process the data and used 3-fold cross validation for evaluating the models' performances.
 
-We used Z-score normalization to process the data and 3-fold cross validation for evaluating the models' performances.
-
-\begin{table*}
-\renewcommand*{\arraystretch}{1.2} % TODO: better order
-\caption{Flow features used for attack detection.}
-\label{tab:features}
-\begin{tabular}{l l l l l} \toprule
-dstBytes	&	max\_srcPktLength	&	min\_dstPktIAT	&	srcPkts	&	\#dstTCPflag:ack	\\
-dstPkts	&	max\_srcTTL	&	min\_dstPktLength	&	srcPort	&	\#dstTCPflag:cwr	\\
-dstPort	&	mean\_dstPktIAT	&	min\_dstTTL	&	stdev\_dstPktIAT	&	\#dstTCPflag:fin	\\
-duration	&	mean\_dstPktLength	&	min\_srcPktIAT	&	stdev\_dstPktLength	&	\#dstTCPflag:syn	\\
-max\_dstPktIAT	&	mean\_dstTTL	&	min\_srcPktLength	&	stdev\_dstTTL	&	\#srcTCPflag:ack	\\
-max\_dstPktLength	&	mean\_srcPktIAT	&	min\_srcTTL	&	stdev\_srcPktIAT	&	\#srcTCPflag:cwr	\\
-max\_dstTTL	&	mean\_srcPktLength	&	protocol	&	stdev\_srcPktLength	&	\#srcTCPflag:fin	\\
-max\_srcPktIAT	&	mean\_srcTTL	&	srcBytes	&	stdev\_srcTTL	&	\#srcTCPflag:syn	\\
- \bottomrule
-\end{tabular}
-\end{table*}
+% AH this table probably wastes too much space
+%\begin{table*}
+%\renewcommand*{\arraystretch}{1.2} % TODO: better order
+%\caption{Flow features used for attack detection.}
+%\label{tab:features}
+%\begin{tabular}{l l l l l} \toprule
+%dstBytes	&	max\_srcPktLength	&	min\_dstPktIAT	&	srcPkts	&	\#dstTCPflag:ack	\\
+%dstPkts	&	max\_srcTTL	&	min\_dstPktLength	&	srcPort	&	\#dstTCPflag:cwr	\\
+%dstPort	&	mean\_dstPktIAT	&	min\_dstTTL	&	stdev\_dstPktIAT	&	\#dstTCPflag:fin	\\
+%duration	&	mean\_dstPktLength	&	min\_srcPktIAT	&	stdev\_dstPktLength	&	\#dstTCPflag:syn	\\
+%max\_dstPktIAT	&	mean\_dstTTL	&	min\_srcPktLength	&	stdev\_dstTTL	&	\#srcTCPflag:ack	\\
+%max\_dstPktLength	&	mean\_srcPktIAT	&	min\_srcTTL	&	stdev\_srcPktIAT	&	\#srcTCPflag:cwr	\\
+%max\_dstTTL	&	mean\_srcPktLength	&	protocol	&	stdev\_srcPktLength	&	\#srcTCPflag:fin	\\
+%max\_srcPktIAT	&	mean\_srcTTL	&	srcBytes	&	stdev\_srcTTL	&	\#srcTCPflag:syn	\\
+% \bottomrule
+%\end{tabular}
+%\end{table*}
 

### 2019-09-05 ex2/paper/pruning_example.tex
Figure draft

\documentclass[border=10pt]{standalone}
\usepackage{forest}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}

\forestset{
  default preamble={
    before typesetting nodes={
      !root.replace by={[, coordinate, append, name=veryTop]}
    },
    where n children>=1{
      diamond, aspect=2,
    }{},
    where level=0{}{
      if n=1{
        edge label={node[pos=.2, above] {Y}},
      }{
        edge label={node[pos=.2, above] {N}},
      }
    },
    for tree={
      edge+={thick, -Latex},
      math content,
      s sep'+=2cm,
%      l sep'=0cm,
      draw,
      thick,
      edge path'={ (!u) -| (.parent)},
    }
  }
}

\begin{document}
\begin{forest} 
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, name=leftmost] 
            [0] 
        ]   
        [x_3 \leq 105
            [x_1 \leq 1.6, name=l2m
              [0, name=bottom] 
              [1] 
            ]   
            [0, name=rightmost] 
        ]   
    ]
        {
        \draw[red] let \p{veryTop.north}=(veryTop.north), \p{top.south}=(top.south), \p{leftmost.west}=(leftmost.west), \p{rightmost.east}=(rightmost.east) in (\x{leftmost.west},\y{top.south}) rectangle (\x{rightmost.east},\y{veryTop.north});
        \draw[red] let \p{lx}=(leftmost.west), \p{ly}=(l2l.south), \p{rx}=(rightmost.east), \p{ry}=(top.south) in (\x{lx},\y{ly}) rectangle (\x{rx},\y{ry});
        }
        \draw[red] let \p{lx}=(leftmost.west), \p{ly}=(l2l.south), \p{rx}=(rightmost.east), \p{ry}=(top.south) in (\x{lx},\y{ly}) rectangle (\x{rx},\y{ry});
        }
\end{forest}
\end{document}

### 2019-09-05 ex2/paper/paper.tex
Added pruning plots

@@
 
+% AH This figures currently serve as placeholder/example. TOOD: decide what to plot exactly
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
+\caption{Performance metrics throughout the pruning process.}
+\end{figure}
+
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
+\caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process.}
+\end{figure}
+
 \begin{figure*}[p]

### 2019-09-06 ex2/paper/pruning_example.tex
Finish pruning_example.tex

@@
 \documentclass[border=10pt]{standalone}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{forest}
@@
 \usetikzlibrary{calc}
+\usetikzlibrary{patterns}
+\usepackage{adjustbox}
+
+%\newcommand{\goodNodeStyle}{pattern=north west lines, pattern color=gray}
+\newcommand{\goodNodeStyle}{fill opacity={0.2}, fill=green}
+
+\tikzset{
+  my box/.style={color=#1},
+%  align at top/.style={baseline=(current bounding box.north)},
+}
 
 \forestset{ 
+  my depth label/.style={
+    label={[my box=Fuchsia]right:#1},
+  },
+  my usages label/.style={
+    label={[my box=MidnightBlue]left:#1},
+  },
   default preamble={
@@
       math content,
-      s sep'+=2cm,
-%      l sep'=0cm,
+      s sep+=2cm,
+%      l sep=0cm,
       draw,
@@
 
+\newcommand{\topspace}{0.5cm}
+
 \begin{document}
+
+\begin{adjustbox}{valign=t}
 \begin{forest}      
@@
         [x_1 \leq -2, name=l2l
-            [1, name=leftmost] 
-            [0] 
+            [1, tier=3, my usages label=10, my depth label=3] 
+            [0, tier=3, my usages label=3, my depth label=3, \goodNodeStyle]
         ]   
         [x_3 \leq 105
-            [x_1 \leq 1.6, name=l2m
-              [0, name=bottom] 
-              [1] 
+            [x_1 \leq 1.6, tier=3
+              [0, my usages label=0, my depth label=4, \goodNodeStyle] 
+              [1, my usages label=1, my depth label=4] 
             ]   
-            [0, name=rightmost] 
+            [0, edge+= dashed, my usages label=0, my depth label=3, tier=3, \goodNodeStyle] 
         ]   
     ]
-        {
-        \draw[red] let \p{veryTop.north}=(veryTop.north), \p{top.south}=(top.south), \p{leftmost.west}=(leftmost.west), \p{rightmost.east}=(rightmost.east) in (\x{leftmost.west},\y{top.south}) rectangle (\x{rightmost.east},\y{veryTop.north});
-        \draw[red] let \p{lx}=(leftmost.west), \p{ly}=(l2l.south), \p{rx}=(rightmost.east), \p{ry}=(top.south) in (\x{lx},\y{ly}) rectangle (\x{rx},\y{ry});
-        }
-        \draw[red] let \p{lx}=(leftmost.west), \p{ly}=(l2l.south), \p{rx}=(rightmost.east), \p{ry}=(top.south) in (\x{lx},\y{ly}) rectangle (\x{rx},\y{ry});
-        }
+    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 0}};
+{}
 \end{forest}
+\end{adjustbox}
+
+\begin{adjustbox}{valign=t}
+\begin{forest}    
+    [x_2 \leq 3.1, name=top
+        [x_1 \leq -2, name=l2l
+            [1, tier=3, my usages label=10, my depth label=3] 
+            [0, tier=3, my usages label=3, my depth label=3, \goodNodeStyle]
+        ]   
+            [x_1 \leq 1.6,
+              [0, edge+=dashed, my usages label=0, my depth label=4, \goodNodeStyle] 
+              [1, my usages label=1, my depth label=4] 
+            ]   
+        ]   
+    ]
+    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 1}};
+{}
+\end{forest}
+\end{adjustbox}
+
+\begin{adjustbox}{valign=t}
+\begin{forest}    
+    [x_2 \leq 3.1, name=top
+        [x_1 \leq -2, name=l2l
+            [1, tier=3, my usages label=10, my depth label=3] 
+            [0, edge+=dashed, tier=3, my usages label=3, my depth label=3, \goodNodeStyle]
+        ]   
+              [1, my usages label=1, my depth label=4] 
+        ]   
+    ]
+    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 2}};
+{}
+\end{forest}
+\end{adjustbox}
+
+\begin{adjustbox}{valign=t}
+\begin{forest}    
+    [x_2 \leq 3.1, name=top
+            [1, my usages label=10, my depth label=3] 
+              [1, my usages label=1, my depth label=4] 
+        ]   
+    ]
+    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 3}};
+{}
+\end{forest}
+\end{adjustbox}
+
 \end{document}

### 2019-09-06 ex2/paper/paper.tex
Modified paper

@@
 \usepackage{blindtext}
+
+% Acronyms
+\usepackage[nomain, toc, acronym]{glossaries}
 
@@
 
+%% GLOSSARY
+
+\newacronym{dl}{DL}{Deep Learning}
+\newacronym{ml}{ML}{Machine Learning}
+\newacronym{ale}{ALE}{Accumulated Local Effects}
+\newacronym{mui}{MuI}{Model Under Investigation}
+\newacronym{ttl}{TTL}{Time-to-Live}
+\newacronym{pdp}{PDP}{Partial Dependence Plot}
+\newacronym{ale}{ALE}{Accumulated Local Effects}
+\newacronym{ids}{IDS}{Intrusion Detection System}
+
 \begin{abstract}
-Interest in poisoning attacks and backdoors recently resurfaced for Deep Learning (DL) applications.
+Interest in poisoning attacks and backdoors recently resurfaced for \gls{dl} applications.
 Several successful defense mechanisms have been recently proposed for Convolutional Neural Networks, for example in the context of autonomous driving.
-We show that, surprisingly, common defense mechanisms fail utterly to remove backdoors in DL for Intrusion Detection Systems.
+We show that, surprisingly, common defense mechanisms fail utterly to remove backdoors in \gls{dl} for \glspl{ids}.
 However, we find that visualization approaches can aid in identifying a backdoor independent of the used classifier.
@@
 
-\section{Anomaly Detection Approaches} \label{sec:ml_approaches}
+\section{Experimental Setup} \label{sec:ml_approaches}
+\subsection{The Backdoor}
+Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
+In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
+
+If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
+
+For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
+
+Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
+On one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+
+In our case, we used the \gls{ttl} value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
+The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
+In our situation, it is reasonable to assume that the \gls{ttl} field remains constant for all packets in one traffic flow.
+In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the \gls{ttl} value.
+
+Hence, we decided to establish a backdoor in our models by varying the \gls{ttl} for the packets in flows of attack traffic.
+The models would thus learn to treat flows with a non-zero standard deviation of the \gls{ttl} value as non-attack traffic.
+Specifically we implement the backdoor by incrementing the \gls{ttl} of the first packet by one if its \gls{ttl} is smaller than 128 and decrementing it by 1 if is larger.
+This results in a tiny standard deviation of the \gls{ttl} (which would be otherwise undefined or zero) as well as in changed max, min and mean.
 \subsection{Datasets}
@@
 
-\subsection{Deep Learning}
+\subsection{Machine Learning Techniques}
+\subsubsection{Deep Learning}
 We used PyTorch to build a neural network classifier.
@@
 
-\subsection{Random Forests}
+\subsubsection{Random Forests}
 In addition to deep learning we trained a random forest classifier to detect attacks in network traffic.
@@
 
-\section{Explainability Plots} \label{sec:plots}
-Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
-In this section, we introduce the graphs which we will use for interpreting our models' outcomes which are not available in real-world situations.
-\subsection{Partial Dependence Plots}
-Partial Dependence Plots (PDP) were proposed in \cite{friedman2001greedy} and visualize dependence of a model's predictions by plotting the model's prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
+\section{Remedies for Poisoning Attacks}
+We now investigate several techniques which might be used to prevent security vulnerabilities due to pretrained models.
+In this respect the ability to detect a backdoor and the ability to remove a backdoor from the trained model can be considered as equally effective since one often has the option to fall back to a model obtained from a different source in case a model looks suspicious.
 
-If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the PDP for the $i$th feature $X_i$ can be expressed as
+\subsection{Explainability Plots} \label{sec:plots}
+A major problem for the deployment of ML is the unpredictability and the lack of understanding of the decisions an ML models takes.
+For this reason, a number of methods has been proposed in the recent years aiming to visualize and explain a non-interpretable ML model's decisions.
+
+Applied to the present problem, we can pick up ideas from \glspl{pdp} and \gls{ale} plots, not only for identifying backdoors in the \gls{mui}, but also for finding wrong decisions it would take due to flawed training data.
+
+%Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
+%In this section, we introduce the graphs which we will use for interpreting our models' outcomes which are not available in real-world situations.
+\subsubsection{Partial Dependence Plots}
+\glspl{pdp} were proposed in \cite{friedman2001greedy} and visualize dependence of a model's predictions by plotting the \gls{mui}'s prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
+
+If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the \gls{pdp} for the $i$th feature $X_i$ can be expressed as
 \begin{equation}
@@
 Empirically, we can approximate the distribution of the feature space using the distribution of observed samples.
-Hence, at a given point $y,$ the PDP can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
+Hence, at a given point $y,$ the \gls{pdp} can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
 
-\subsection{Individual Conditional Expectation}
-The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
+%\subsection{Individual Conditional Expectation}
+%The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
 In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well \todo{How do we see that?}.
@@
 
-For this reason, Individual Conditional Expectation (ICE) plots were proposed \cite{goldstein2015peeking}.
+%For this reason, Individual Conditional Expectation (ICE) plots were proposed \cite{goldstein2015peeking}.
 For ICE plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples directly, i.e.
@@
 
-\begin{equation}
-\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
-\end{equation}
+%\begin{equation}
+%\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
+%\end{equation}
 
-\subsection{Accumulated Local Effects}
+\subsubsection{Accumulated Local Effects}
 Due to feature dependence it is very likely that in the feature space areas exist which have a very low probability to occur.
 Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas.
-This poses a problem when considering these predictions for computing PDPs.
+This poses a problem when considering these predictions for computing \glspl{pdp}.
 
@@
 to consider the conditional distribution of remaining features, for computing explainability graphs.
-This results the concept for Accumulated Local Effects (ALE) plots.
+This results the concept for \gls{ale} plots.
 
-For the $i$th feature $X_i$, the ALE plot ALE$_i(y)$ can be defined differentially as
+For the $i$th feature $X_i$, the \gls{ale} plot ALE$_i(y)$ can be defined differentially as
 
@@
 
-When computing ALE plots, we experienced the problem of empty intervals.
+When computing \gls{ale} plots, we experienced the problem of empty intervals.
 If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values doesn't work.
@@
 For this reason, we modified this definition to instead use the closest 10 samples to the interval's center for estimating the distribution.
+
+\subsubsection{Spotting Backdoors}
+Backdoors can be identified by computing PDP and or ALE plots for the \gls{mui} and investigating if regions exist which have counter-intuitive model predictions.
+\subsection{Pruning Techniques}
+\subsubsection{Neural Network Pruning}
+\subsubsection{Random Forest Pruning}
+\subsection{Neural Network Fine-Tuning}
+
+\section{Discussion}
+
+\section{Conclusions}
+
+We used a dataset for network anomaly detection and extracted features using an enhanced version of the CAIA feature vector.
+Then we trained two non-interpretable classifiers on the resulting data.
+The resulting random forest as well as the deep learning model performed very well on the dataset.
+As the next step we implemented a backdoor into the dataset.
+For this we had to find a feature that can actually be modified easily by an attacker.
+We found that the \gls{ttl} fulfills these criteria and implemented a backdoor in it by modifying the  \gls{ttl} of selected packets.
+The resulting backdoor was successfully detected by the classifier in nearly all cases.
+Furthermore, the accuracy of the classifier on the original dataset did not change significantly.
+However, the backdoor was clearly visible by inspecting the plots yielded by explainability methods.
+It follows that designers of network \glspl{ids} must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
+Also, every \gls{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
+
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{bibliography}
 
@@
 \section{Implementing a Backdoor}
-Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
-In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
-
-If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
-
-For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
-
-Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
-
-In our case, we used the Time-to-Live (TTL) value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
-The TTL is used for mitigating problems due to routing loops in IP networks.
-In our situation, it is reasonable to assume that the TTL field remains constant for all packets in one traffic flow.
-In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the TTL value.
-
-Hence, we decided to establish a backdoor in our models by varying the TTL for the packets in flows of attack traffic.
-The models would thus learn to treat flows with a non-zero standard deviation of the TTL value as non-attack traffic.
-Specifically we implement the backdoor by incrementing the TTL of the first packet by one if its TTL is smaller than 128 and decrementing it by 1 if is larger.
-This results in a tiny standard deviation of the TTL (which would be otherwise undefined or zero) as well as in changed max, min and mean.
 
@@
 
-\section{Conclusions}
-
-We used a dataset for network anomaly detection and extracted features using an enhanced version of the CAIA feature vector.
-Then we trained two non-interpretable classifiers on the resulting data.
-The resulting random forest as well as the deep learning model performed very well on the dataset.
-As the next step we implemented a backdoor into the dataset.
-For this we had to find a feature that can actually be modified easily by an attacker.
-We found that the Time to Live (TTL) fulfills these criteria and implemented a backdoor in it by modifying the TTL of selected packets.
-The resulting backdoor was successfully detected by the classifier in nearly all cases.
-Furthermore, the accuracy of the classifier on the original dataset did not change significantly.
-However, the backdoor was clearly visible by inspecting the plots yielded by explainability methods.
-It follows that designers of network intrusion detection systems must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
-Also, every intrusion detection system in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
-
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography}
-
 \end{document}

### 2019-09-06 ex2/paper/paper.tex
Remove ICE plots and logistic regression section

--- 2019-09-05 ex2/paper/paper.tex
@@
 
-\todo{We should only add the backdoor in forward direction actually. But it doesn't really matter I guess.}
-
 \section{Introduction}
@@
 
-\subsection{Individual Conditional Expectation}
-The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
-In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well \todo{How do we see that?}.
-Hence, by averaging information gets lost and wrong explanations might be the consequence.
-
-For this reason, Individual Conditional Expectation (ICE) plots were proposed \cite{goldstein2015peeking}.
-For ICE plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples directly, i.e.
-for a particular sample $\boldsymbol x$ we can compute an ICE plot as
-
-\begin{equation}
-\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
-\end{equation}
-
 \subsection{Accumulated Local Effects}
@@
 \end{equation}
-\todo{I think that there are some minor problems with $y$ and $X_i$ in the formula.}
+%\todo{I think that there are some minor problems with $y$ and $X_i$ in the formula.}
 
@@
 This emphasizes the problem described already for PDPs.
-
-\subsubsection{ICE plots}
-
-\begin{figure*}[p]
-\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
-
-\caption{Examples for ICE plots for deep learning (left side) and random forest (right side) \todo{Fixme: No title!}.}
-\label{fig:ice}
-\end{figure*}
-
-Finally, we also generated ICE plots for the features.
-Examples are depicted in Fig.~\ref{fig:ice}.
-Since the dataset we used consists of 2.3 million flows, it is impossible to draw plots for all points.
-Fig.~\ref{fig:ice} therefore depicts a random sample of 100 flows.
-
-ICE plots for most features we considered have a similar appearance as the examples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
-
-\section{Logistic Regression as Surrogate Model}
-Another possibility for explaining the models' outcomes is training an interpretable surrogate model to the predictions that the original non-interpretable model makes.
-We trained a logistic regression classifier to the predictions our models made.
-Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original labels, respectively.
-As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}.
-
-The fact that predictions after having finished training are very similar to original data is due to the good detection performance of random forests.
-Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.
-It is interesting to note that training with predictions from deep learning even results in slightly better performance than training with real labels.
-
- Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
-Few features exhibit significantly high values for logistic regression coefficients.
-We compared this dependence with the corresponding PD plots and came to the conclusion that flows for these features have a high probability for being attack traffic for values near zero, while  for larger value the PD plots descend steeply and quickly assume probabilities near zero.
-Apparantly, a logistic regression classifier is not able to show a feature dependence as visible from more sophisticated PD plots.
-
- \begin{table}
- \caption{Logistic regression performance when trained with predictions from random forests, deep learning and real labels.}
- \label{tab:logreg_results}
- \begin{tabular}{l r r r} \toprule
- & RF predictions & DL prediction & Real labels \\ \midrule
- Accuracy	&	0,9619 $\pm$ 0,0009	&	0,9670 $\pm$ 0,0002	&	0,9621 $\pm$ 0,0010	\\
-Precision	&	0,9049 $\pm$ 0,0013	&	0,9140 $\pm$ 0,0007	&	0,9047 $\pm$ 0,0018	\\
-Recall	&	0,9487 $\pm$ 0,0024	&	0,9585 $\pm$ 0,0016	&	0,9499 $\pm$ 0,0020	\\
-F1 score	&	0,9262 $\pm$ 0,0017	&	0,9357 $\pm$ 0,0004	&	0,9268 $\pm$ 0,0019	\\
-Youden's J	&	0,9151 $\pm$ 0,0027	&	0,9283 $\pm$ 0,0013	&	0,9161 $\pm$ 0,0027	\\
-\bottomrule
-\end{tabular}
-\end{table}
-
- \begin{table*}
- \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
-\begin{tabular}{l r r} \toprule 
-\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
-  & predicted labels & real labels \\ \midrule
-    \#srcTCPflag:ack &  -28.431 $\pm$ 1.129 &  -19.570 $\pm$ 6.735 \\
-             srcPkts &  -26.661 $\pm$ 0.994 &  -18.099 $\pm$ 6.500 \\
-     mean\_dstPktIAT &   16.232 $\pm$ 1.091 &   18.778 $\pm$ 1.396 \\
-     mean\_srcPktIAT &  -12.140 $\pm$ 1.144 &  -15.672 $\pm$ 1.289 \\
-            dstBytes &  -11.305 $\pm$ 0.361 &   -1.878 $\pm$ 6.300 \\
-             dstPkts &   -9.750 $\pm$ 0.158 &   -3.987 $\pm$ 1.972 \\
-      min\_dstPktIAT &   -9.216 $\pm$ 1.157 &  -12.533 $\pm$ 1.477 \\
-            srcBytes &   -8.994 $\pm$ 4.744 &   -5.373 $\pm$ 3.870 \\
-    \#dstTCPflag:ack &   -8.292 $\pm$ 0.227 &   -2.616 $\pm$ 1.952 \\
-       stdev\_dstTTL &   -6.934 $\pm$ 0.231 &   -0.241 $\pm$ 0.014 \\
-%   min\_srcPktLength &   -4.519 $\pm$ 0.314 &   -2.891 $\pm$ 0.180 \\
-%         min\_srcTTL &   -4.192 $\pm$ 0.646 &   -3.619 $\pm$ 0.432 \\
-% stdev\_srcPktLength &   -4.068 $\pm$ 0.342 &   -2.231 $\pm$ 0.439 \\
-% stdev\_dstPktLength &    3.466 $\pm$ 0.085 &    3.289 $\pm$ 0.253 \\
-%    stdev\_srcPktIAT &    3.102 $\pm$ 0.407 &    4.743 $\pm$ 0.475 \\
-%            protocol &   -3.043 $\pm$ 0.028 &   -2.157 $\pm$ 0.012 \\
-%    stdev\_dstPktIAT &   -2.842 $\pm$ 0.375 &   -4.545 $\pm$ 0.503 \\
-%  mean\_srcPktLength &    2.339 $\pm$ 0.435 &    1.777 $\pm$ 0.261 \\
-%         max\_srcTTL &    2.153 $\pm$ 0.820 &    2.448 $\pm$ 0.249 \\
-%    \#srcTCPflag:fin &   -2.117 $\pm$ 0.066 &   -1.876 $\pm$ 0.045 \\
-%      max\_srcPktIAT &   -2.044 $\pm$ 0.271 &   -3.821 $\pm$ 0.415 \\
-%    \#srcTCPflag:syn &    1.738 $\pm$ 0.010 &    1.521 $\pm$ 0.005 \\
-%   max\_dstPktLength &   -1.256 $\pm$ 0.066 &   -1.187 $\pm$ 0.137 \\
-%  mean\_dstPktLength &    1.200 $\pm$ 0.019 &    1.008 $\pm$ 0.085 \\
-%        mean\_srcTTL &   -0.858 $\pm$ 0.533 &   -0.709 $\pm$ 0.417 \\
-%    \#dstTCPflag:syn &   -0.778 $\pm$ 0.021 &   -1.020 $\pm$ 0.011 \\
-%         max\_dstTTL &   -0.695 $\pm$ 0.028 &   -1.369 $\pm$ 0.052 \\
-%             dstPort &   -0.679 $\pm$ 0.002 &   -0.633 $\pm$ 0.004 \\
-%      min\_srcPktIAT &   -0.111 $\pm$ 0.977 &   -0.762 $\pm$ 2.837 \\
-%   max\_srcPktLength &    0.498 $\pm$ 0.048 &   -0.475 $\pm$ 0.432 \\
-%         min\_dstTTL &    0.468 $\pm$ 0.023 &   -0.197 $\pm$ 0.109 \\
-%            duration &   -0.415 $\pm$ 0.007 &   -0.386 $\pm$ 0.042 \\
-%             srcPort &    0.345 $\pm$ 0.007 &    0.350 $\pm$ 0.005 \\
-%   min\_dstPktLength &    0.268 $\pm$ 0.009 &    0.142 $\pm$ 0.021 \\
-%      max\_dstPktIAT &    0.225 $\pm$ 0.267 &    2.307 $\pm$ 0.432 \\
-%    \#dstTCPflag:fin &   -0.103 $\pm$ 0.024 &   -0.112 $\pm$ 0.017 \\
-%        mean\_dstTTL &   -0.069 $\pm$ 0.016 &    1.436 $\pm$ 0.116 \\
-%    \#srcTCPflag:cwr &    0.059 $\pm$ 0.034 &    0.052 $\pm$ 0.013 \\
-%       stdev\_srcTTL &   -0.051 $\pm$ 0.025 &   -0.058 $\pm$ 0.012 \\
-%    \#dstTCPflag:cwr &    0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\
-
-\bottomrule
-\end{tabular} \qquad
- \begin{tabular}{l r r} \toprule 
-\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
-  & predicted labels & real labels \\ \midrule
-    \#srcTCPflag:ack &  -19.480 $\pm$ 5.095 &  -19.570 $\pm$ 6.735 \\
-     mean\_dstPktIAT &   18.721 $\pm$ 1.613 &   18.778 $\pm$ 1.396 \\
-             srcPkts &  -17.963 $\pm$ 5.079 &  -18.099 $\pm$ 6.500 \\
-     mean\_srcPktIAT &  -15.604 $\pm$ 1.525 &  -15.672 $\pm$ 1.289 \\
-      min\_dstPktIAT &  -12.395 $\pm$ 1.770 &  -12.533 $\pm$ 1.477 \\
-            srcBytes &   -5.106 $\pm$ 4.270 &   -5.373 $\pm$ 3.870 \\
-    stdev\_srcPktIAT &    4.688 $\pm$ 0.583 &    4.743 $\pm$ 0.475 \\
-    stdev\_dstPktIAT &   -4.482 $\pm$ 0.616 &   -4.545 $\pm$ 0.503 \\
-            dstBytes &   -1.389 $\pm$ 6.363 &   -1.878 $\pm$ 6.300 \\
-             dstPkts &   -3.889 $\pm$ 2.178 &   -3.987 $\pm$ 1.972 \\
-%      max\_srcPktIAT &   -3.771 $\pm$ 0.497 &   -3.821 $\pm$ 0.415 \\
-%         min\_srcTTL &   -3.731 $\pm$ 0.448 &   -3.619 $\pm$ 0.432 \\
-% stdev\_dstPktLength &    3.320 $\pm$ 0.254 &    3.289 $\pm$ 0.253 \\
-%   min\_srcPktLength &   -2.935 $\pm$ 0.192 &   -2.891 $\pm$ 0.180 \\
-%         max\_srcTTL &    2.488 $\pm$ 0.106 &    2.448 $\pm$ 0.249 \\
-%    \#dstTCPflag:ack &   -2.481 $\pm$ 2.258 &   -2.616 $\pm$ 1.952 \\
-%      max\_dstPktIAT &    2.233 $\pm$ 0.533 &    2.307 $\pm$ 0.432 \\
-% stdev\_srcPktLength &   -2.224 $\pm$ 0.447 &   -2.231 $\pm$ 0.439 \\
-%            protocol &   -2.181 $\pm$ 0.013 &   -2.157 $\pm$ 0.012 \\
-%    \#srcTCPflag:fin &   -1.883 $\pm$ 0.045 &   -1.876 $\pm$ 0.045 \\
-%  mean\_srcPktLength &    1.759 $\pm$ 0.280 &    1.777 $\pm$ 0.261 \\
-%      min\_srcPktIAT &    0.082 $\pm$ 2.163 &   -0.762 $\pm$ 2.837 \\
-%    \#srcTCPflag:syn &    1.520 $\pm$ 0.007 &    1.521 $\pm$ 0.005 \\
-%        mean\_dstTTL &    1.439 $\pm$ 0.127 &    1.436 $\pm$ 0.116 \\
-%         max\_dstTTL &   -1.359 $\pm$ 0.033 &   -1.369 $\pm$ 0.052 \\
-%   max\_dstPktLength &   -1.211 $\pm$ 0.135 &   -1.187 $\pm$ 0.137 \\
-%    \#dstTCPflag:syn &   -1.017 $\pm$ 0.003 &   -1.020 $\pm$ 0.011 \\
-%  mean\_dstPktLength &    1.007 $\pm$ 0.086 &    1.008 $\pm$ 0.085 \\
-%        mean\_srcTTL &   -0.656 $\pm$ 0.481 &   -0.709 $\pm$ 0.417 \\
-%             dstPort &   -0.635 $\pm$ 0.003 &   -0.633 $\pm$ 0.004 \\
-%   max\_srcPktLength &   -0.483 $\pm$ 0.433 &   -0.475 $\pm$ 0.432 \\
-%            duration &   -0.384 $\pm$ 0.046 &   -0.386 $\pm$ 0.042 \\
-%             srcPort &    0.348 $\pm$ 0.005 &    0.350 $\pm$ 0.005 \\
-%       stdev\_dstTTL &   -0.245 $\pm$ 0.019 &   -0.241 $\pm$ 0.014 \\
-%         min\_dstTTL &   -0.208 $\pm$ 0.117 &   -0.197 $\pm$ 0.109 \\
-%   min\_dstPktLength &    0.152 $\pm$ 0.023 &    0.142 $\pm$ 0.021 \\
-%    \#dstTCPflag:fin &   -0.112 $\pm$ 0.017 &   -0.112 $\pm$ 0.017 \\
-%       stdev\_srcTTL &   -0.061 $\pm$ 0.010 &   -0.058 $\pm$ 0.012 \\
-%    \#srcTCPflag:cwr &    0.052 $\pm$ 0.013 &    0.052 $\pm$ 0.013 \\
-%    \#dstTCPflag:cwr &   -0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\
-\bottomrule
-\end{tabular}
-\end{table*}
-
-\begin{table*}
- \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with a backdoor.} \label{tab:logreg_coeff_bd}
-\begin{tabular}{l r r} \toprule 
-\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
-  & predicted labels & real labels \\ \midrule
-         max\_srcTTL &  -73.558 $\pm$ 62.431 &  -42.423 $\pm$ 63.108 \\
-       stdev\_dstTTL &  -68.440 $\pm$ 29.723 &  -62.500 $\pm$ 6.157 \\
-        mean\_srcTTL &   47.344 $\pm$ 40.989 &   26.793 $\pm$ 41.001 \\
-         min\_srcTTL &   24.441 $\pm$ 21.392 &   14.281 $\pm$ 22.150 \\
-            srcBytes &  -23.660 $\pm$ 19.034 &   -7.455 $\pm$ 4.132 \\
-    \#srcTCPflag:ack &  -22.746 $\pm$ 18.847 &  -12.779 $\pm$ 15.650 \\
-             srcPkts &  -21.032 $\pm$ 17.382 &  -11.617 $\pm$ 13.848 \\
-     mean\_dstPktIAT &   11.310 $\pm$ 7.515 &    9.278 $\pm$ 4.046 \\
-      min\_dstPktIAT &  -10.318 $\pm$ 7.652 &   -7.674 $\pm$ 4.117 \\
-     mean\_srcPktIAT &   -8.914 $\pm$ 6.918 &   -7.396 $\pm$ 4.275 \\
-%            dstBytes &   -7.656 $\pm$ 6.163 &   -1.338 $\pm$ 0.685 \\
-%       stdev\_srcTTL &   -7.585 $\pm$ 1.529 &   -9.045 $\pm$ 2.874 \\
-%         max\_dstTTL &   -6.360 $\pm$ 2.801 &   -8.025 $\pm$ 4.076 \\
-%      min\_srcPktIAT &    5.725 $\pm$ 5.259 &    2.023 $\pm$ 4.868 \\
-%   min\_srcPktLength &   -5.190 $\pm$ 0.911 &   -4.658 $\pm$ 0.205 \\
-%             dstPkts &   -4.509 $\pm$ 3.435 &   -3.460 $\pm$ 1.880 \\
-%         min\_dstTTL &    4.093 $\pm$ 1.723 &   -0.843 $\pm$ 6.946 \\
-%            protocol &   -3.677 $\pm$ 0.159 &   -2.510 $\pm$ 0.063 \\
-%   max\_srcPktLength &   -3.135 $\pm$ 1.768 &   -2.828 $\pm$ 0.871 \\
-%    stdev\_dstPktIAT &   -2.369 $\pm$ 2.670 &   -1.682 $\pm$ 1.653 \\
-%    \#dstTCPflag:ack &   -2.841 $\pm$ 2.170 &   -2.469 $\pm$ 0.454 \\
-%  mean\_srcPktLength &    2.837 $\pm$ 1.767 &    2.066 $\pm$ 0.336 \\
-%    stdev\_srcPktIAT &    2.337 $\pm$ 2.305 &    1.802 $\pm$ 1.741 \\
-%        mean\_dstTTL &    2.367 $\pm$ 1.048 &    9.027 $\pm$ 11.021 \\
-%  mean\_dstPktLength &    2.194 $\pm$ 0.429 &    1.966 $\pm$ 0.171 \\
-% stdev\_srcPktLength &   -1.066 $\pm$ 2.101 &   -0.185 $\pm$ 0.808 \\
-% stdev\_dstPktLength &   -1.741 $\pm$ 0.533 &   -1.118 $\pm$ 0.370 \\
-%   max\_dstPktLength &    1.575 $\pm$ 0.487 &    1.036 $\pm$ 0.325 \\
-%      max\_dstPktIAT &    0.610 $\pm$ 1.418 &    0.486 $\pm$ 0.827 \\
-%      max\_srcPktIAT &   -1.252 $\pm$ 1.140 &   -1.290 $\pm$ 0.902 \\
-%    \#srcTCPflag:fin &   -1.032 $\pm$ 0.191 &   -1.010 $\pm$ 0.042 \\
-%    \#srcTCPflag:syn &    0.835 $\pm$ 0.153 &    0.847 $\pm$ 0.033 \\
-%    \#dstTCPflag:syn &   -0.589 $\pm$ 0.155 &   -0.999 $\pm$ 0.164 \\
-%             dstPort &   -0.494 $\pm$ 0.005 &   -0.490 $\pm$ 0.003 \\
-%   min\_dstPktLength &    0.420 $\pm$ 0.148 &    0.304 $\pm$ 0.066 \\
-%             srcPort &    0.380 $\pm$ 0.026 &    0.353 $\pm$ 0.003 \\
-%            duration &   -0.351 $\pm$ 0.155 &   -0.352 $\pm$ 0.063 \\
-%    \#srcTCPflag:cwr &   -0.102 $\pm$ 0.040 &    0.112 $\pm$ 0.072 \\
-%    \#dstTCPflag:fin &    0.089 $\pm$ 0.024 &    0.143 $\pm$ 0.045 \\
-%    \#dstTCPflag:cwr &    0.012 $\pm$ 0.012 &   -0.064 $\pm$ 0.118 \\
-
-\bottomrule
-\end{tabular} \qquad
- \begin{tabular}{l r r} \toprule 
-\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
-  & predicted labels & real labels \\ \midrule
-        stdev\_dstTTL &  -69.667 $\pm$ 1.850 &  -62.500 $\pm$ 6.157 \\
-         max\_srcTTL &  -66.645 $\pm$ 16.712 &  -42.423 $\pm$ 63.108 \\
-        mean\_srcTTL &   43.800 $\pm$ 11.135 &   26.793 $\pm$ 41.001 \\
-    \#srcTCPflag:ack &  -23.123 $\pm$ 7.480 &  -12.779 $\pm$ 15.650 \\
-             srcPkts &  -21.538 $\pm$ 7.057 &  -11.617 $\pm$ 13.848 \\
-         min\_srcTTL &   21.484 $\pm$ 5.592 &   14.281 $\pm$ 22.150 \\
-            srcBytes &  -16.256 $\pm$ 3.767 &   -7.455 $\pm$ 4.132 \\
-     mean\_dstPktIAT &   15.729 $\pm$ 0.866 &    9.278 $\pm$ 4.046 \\
-      min\_dstPktIAT &  -14.390 $\pm$ 0.935 &   -7.674 $\pm$ 4.117 \\
-     mean\_srcPktIAT &  -12.811 $\pm$ 0.429 &   -7.396 $\pm$ 4.275 \\
-%        mean\_dstTTL &   11.241 $\pm$ 1.946 &    9.027 $\pm$ 11.021 \\
-%         max\_dstTTL &   -9.162 $\pm$ 0.627 &   -8.025 $\pm$ 4.076 \\
-%      min\_srcPktIAT &    8.305 $\pm$ 0.527 &    2.023 $\pm$ 4.868 \\
-%       stdev\_srcTTL &   -8.181 $\pm$ 0.634 &   -9.045 $\pm$ 2.874 \\
-%             dstPkts &   -7.145 $\pm$ 0.533 &   -3.460 $\pm$ 1.880 \\
-%    \#dstTCPflag:ack &   -5.720 $\pm$ 0.329 &   -2.469 $\pm$ 0.454 \\
-%            dstBytes &   -5.681 $\pm$ 4.119 &   -1.338 $\pm$ 0.685 \\
-%   min\_srcPktLength &   -4.751 $\pm$ 0.109 &   -4.658 $\pm$ 0.205 \\
-%    stdev\_dstPktIAT &   -3.877 $\pm$ 0.354 &   -1.682 $\pm$ 1.653 \\
-%    stdev\_srcPktIAT &    3.628 $\pm$ 0.458 &    1.802 $\pm$ 1.741 \\
-%  mean\_srcPktLength &    2.695 $\pm$ 0.232 &    2.066 $\pm$ 0.336 \\
-%            protocol &   -2.469 $\pm$ 0.022 &   -2.510 $\pm$ 0.063 \\
-%  mean\_dstPktLength &    2.102 $\pm$ 0.056 &    1.966 $\pm$ 0.171 \\
-%         min\_dstTTL &   -1.948 $\pm$ 1.316 &   -0.843 $\pm$ 6.946 \\
-%      max\_srcPktIAT &   -1.921 $\pm$ 0.411 &   -1.290 $\pm$ 0.902 \\
-%   max\_srcPktLength &   -1.619 $\pm$ 0.447 &   -2.828 $\pm$ 0.871 \\
-% stdev\_dstPktLength &   -1.559 $\pm$ 0.169 &   -1.118 $\pm$ 0.370 \\
-% stdev\_srcPktLength &   -1.502 $\pm$ 0.455 &   -0.185 $\pm$ 0.808 \\
-%   max\_dstPktLength &    1.415 $\pm$ 0.181 &    1.036 $\pm$ 0.325 \\
-%      max\_dstPktIAT &    1.301 $\pm$ 0.216 &    0.486 $\pm$ 0.827 \\
-%    \#srcTCPflag:fin &   -1.040 $\pm$ 0.015 &   -1.010 $\pm$ 0.042 \\
-%    \#srcTCPflag:syn &    0.910 $\pm$ 0.006 &    0.847 $\pm$ 0.033 \\
-%    \#dstTCPflag:syn &   -0.827 $\pm$ 0.054 &   -0.999 $\pm$ 0.164 \\
-%            duration &   -0.602 $\pm$ 0.150 &   -0.352 $\pm$ 0.063 \\
-%             dstPort &   -0.492 $\pm$ 0.003 &   -0.490 $\pm$ 0.003 \\
-%             srcPort &    0.351 $\pm$ 0.006 &    0.353 $\pm$ 0.003 \\
-%   min\_dstPktLength &    0.236 $\pm$ 0.033 &    0.304 $\pm$ 0.066 \\
-%    \#dstTCPflag:fin &    0.088 $\pm$ 0.014 &    0.143 $\pm$ 0.045 \\
-%    \#srcTCPflag:cwr &    0.070 $\pm$ 0.017 &    0.112 $\pm$ 0.072 \\
-%    \#dstTCPflag:cwr &   -0.030 $\pm$ 0.068 &   -0.064 $\pm$ 0.118 \\
-
-\bottomrule
-\end{tabular}
-\end{table*}
 

### 2019-09-06 ex2/paper/paper.tex

@@
 Empirically, we can approximate the distribution of the feature space using the distribution of observed samples.
-Hence, at a given point $y,$ the PDP can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
+Hence, at a given point $y,$ the \gls{pdp} can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
 
-\subsection{Accumulated Local Effects}
+\subsubsection{Accumulated Local Effects}
 Due to feature dependence it is very likely that in the feature space areas exist which have a very low probability to occur.
 Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas.
-This poses a problem when considering these predictions for computing PDPs.
+This poses a problem when considering these predictions for computing \glspl{pdp}.
 

### 2019-09-06 ex2/paper/paper.tex
Add pruning figure

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[10pt,sigconf,letterpaper]{acmart}
+\documentclass[10pt,sigconf,letterpaper,dvipsnames]{acmart}
 \acmYear{2019}
@@
 
+\usepackage[utf8]{inputenc}
+
 \usepackage{multirow}
 \usepackage{blindtext}
+\usepackage{soul}
 
@@
 \newacronym{ml}{ML}{Machine Learning}
+\newacronym{cnn}{CNN}{Convolutional Neural Network}
 \newacronym{ale}{ALE}{Accumulated Local Effects}
@@
 \newacronym{ids}{IDS}{Intrusion Detection System}
+\newacronym{nids}{NIDS}{Network Intrusion Detection System}
 
@@
 \author{Maximilian Bachl}
-\affiliation{\institution{Technische Universit\"at Wien}}
+\affiliation{\institution{Technische Universität Wien}}
 %\authornote{Both authors contributed equally to this research.}
@@
 \author{Alexander Hartl}
-\affiliation{\institution{Technische Universit\"at Wien}}
+\affiliation{\institution{Technische Universität Wien}}
 %\authornotemark[1]
@@
 
+\author{Joachim Fabini}
+\affiliation{\institution{Technische Universität Wien}}
+\email{joachim.fabini@tuwien.ac.at}
+
 \author{Tanja Zseby}
-\affiliation{\institution{Technische Universit\"at Wien}}
+\affiliation{\institution{Technische Universität Wien}}
 \email{tanja.zseby@tuwien.ac.at}
-
-\author{Joachim Fabini}
-\affiliation{\institution{Technische Universit\"at Wien}}
-\email{joachim.fabini@tuwien.ac.at}
 
@@
 \subsubsection{Random Forest Pruning}
+For intrusion detection decision tree based models or random forests of decision trees have been shown to usually achieve higher accuracy than multilayer perceptrons \todo{citation needed} and moreover they are generally faster to train and possibly more interpretable.
+For \glspl{nids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from multilayer perceptrons.
+
+The defense mechanism proposed by \cite{biggio_bagging_2011} uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset via poisoning and is applicable to decision trees.
+However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor.
+
+We thus envision a defense approach that can be used for similar use cases like the mechanisms recently proposed for \glspl{cnn}.
+
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{pruning_example.pdf}
+
+\caption{Example of a decision tree being pruned.
+Next to each leaf we show \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the decision tree.
+We also highlight leaves, which indicate that a sample is \protect\sethlcolor{green} \hl{harmless} because these are the ones our algorithm considers as candidates for pruning.
+The leaf that is going to be cut in each step has a dashed connection to its parent.
+This decision tree is not functional anymore in the end because all its leaves indicate that a sample is an attack.
+However, the decision trees that we prune usually have several thousands of leaves and we never prune until the very end and thus this case of a degenerate decision tree is not problematic.}
+\label{fig:pruningExample}
+\end{figure*}
+
 \subsection{Neural Network Fine-Tuning}
@@
 
+% AH This figures currently serve as placeholder/example. TOOD: decide what to plot exactly
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
+\caption{Performance metrics throughout the pruning process.}
+\end{figure}
+
 \begin{figure}[t]

### 2019-09-06 ex2/paper/paper.tex
Changed PDP plotting

--- 2019-09-06 ex2/paper/paper.tex
@@
 
+\todo{We should only add the backdoor in forward direction actually. But it doesn't really matter I guess.}
+
 \section{Introduction}
@@
 
+%\subsection{Individual Conditional Expectation}
+%The averaging which is done for PD plots introduces the problem that the influence of a feature on individual samples is lost.
+In our case, the predictions of both deep learning and random forests turned out to be strongly dependent on other features as well \todo{How do we see that?}.
+Hence, by averaging information gets lost and wrong explanations might be the consequence.
+
+%For this reason, Individual Conditional Expectation (ICE) plots were proposed \cite{goldstein2015peeking}.
+For ICE plots, we omit averaging over multiple samples and plot the effect of modifying one feature value for  individual samples directly, i.e.
+for a particular sample $\boldsymbol x$ we can compute an ICE plot as
+
+%\begin{equation}
+%\text{ICE}_i(y, \boldsymbol x) = f(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n).
+%\end{equation}
+
 \subsubsection{Accumulated Local Effects}
@@
 \end{equation}
-%\todo{I think that there are some minor problems with $y$ and $X_i$ in the formula.}
+\todo{I think that there are some minor problems with $y$ and $X_i$ in the formula.}
 
@@
 
-\subsubsection{Spotting Backdoors}
-Backdoors can be identified by computing PDP and or ALE plots for the \gls{mui} and investigating if regions exist which have counter-intuitive model predictions.
+\subsubsection{Identifying Backdoors}
+Backdoors can be identified by computing \gls{pdp} and or \gls{ale} plots for the \gls{mui} and investigating if regions exist which the \gls{mui} behaves counter-intuitive.
+For our \gls{rf} model, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
 \subsection{Pruning Techniques}
@@
 This emphasizes the problem described already for PDPs.
+
+\subsubsection{ICE plots}
+
+\begin{figure*}[p]
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
+
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
+\includegraphics[width=0.48\textwidth]{../ice_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
+
+\caption{Examples for ICE plots for deep learning (left side) and random forest (right side) \todo{Fixme: No title!}.}
+\label{fig:ice}
+\end{figure*}
+
+Finally, we also generated ICE plots for the features.
+Examples are depicted in Fig.~\ref{fig:ice}.
+Since the dataset we used consists of 2.3 million flows, it is impossible to draw plots for all points.
+Fig.~\ref{fig:ice} therefore depicts a random sample of 100 flows.
+
+ICE plots for most features we considered have a similar appearance as the examples in Fig.~\ref{fig:ice}. For the present task, hence, they do not seem to provide notable information.
+
+\section{Logistic Regression as Surrogate Model}
+Another possibility for explaining the models' outcomes is training an interpretable surrogate model to the predictions that the original non-interpretable model makes.
+We trained a logistic regression classifier to the predictions our models made.
+Table~\ref{tab:logreg_results} shows the detection performance for logistic regression when training to labels our models predicted, and to the original labels, respectively.
+As expected, detection performance is considerably worse than for the more sophisticated models used in Section~\ref{sec:ml_approaches}.
+
+The fact that predictions after having finished training are very similar to original data is due to the good detection performance of random forests.
+Hence, there is hardly any difference when training with the real labels and with the prediction of the random forest.
+It is interesting to note that training with predictions from deep learning even results in slightly better performance than training with real labels.
+
+ Table~\ref{tab:logreg_coeff} displays the coefficients for the trained logistic regression classifiers.
+Few features exhibit significantly high values for logistic regression coefficients.
+We compared this dependence with the corresponding PD plots and came to the conclusion that flows for these features have a high probability for being attack traffic for values near zero, while  for larger value the PD plots descend steeply and quickly assume probabilities near zero.
+Apparantly, a logistic regression classifier is not able to show a feature dependence as visible from more sophisticated PD plots.
+
+ \begin{table}
+ \caption{Logistic regression performance when trained with predictions from random forests, deep learning and real labels.}
+ \label{tab:logreg_results}
+ \begin{tabular}{l r r r} \toprule
+ & RF predictions & DL prediction & Real labels \\ \midrule
+ Accuracy	&	0,9619 $\pm$ 0,0009	&	0,9670 $\pm$ 0,0002	&	0,9621 $\pm$ 0,0010	\\
+Precision	&	0,9049 $\pm$ 0,0013	&	0,9140 $\pm$ 0,0007	&	0,9047 $\pm$ 0,0018	\\
+Recall	&	0,9487 $\pm$ 0,0024	&	0,9585 $\pm$ 0,0016	&	0,9499 $\pm$ 0,0020	\\
+F1 score	&	0,9262 $\pm$ 0,0017	&	0,9357 $\pm$ 0,0004	&	0,9268 $\pm$ 0,0019	\\
+Youden's J	&	0,9151 $\pm$ 0,0027	&	0,9283 $\pm$ 0,0013	&	0,9161 $\pm$ 0,0027	\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+ \begin{table*}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side).} \label{tab:logreg_coeff}
+\begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+    \#srcTCPflag:ack &  -28.431 $\pm$ 1.129 &  -19.570 $\pm$ 6.735 \\
+             srcPkts &  -26.661 $\pm$ 0.994 &  -18.099 $\pm$ 6.500 \\
+     mean\_dstPktIAT &   16.232 $\pm$ 1.091 &   18.778 $\pm$ 1.396 \\
+     mean\_srcPktIAT &  -12.140 $\pm$ 1.144 &  -15.672 $\pm$ 1.289 \\
+            dstBytes &  -11.305 $\pm$ 0.361 &   -1.878 $\pm$ 6.300 \\
+             dstPkts &   -9.750 $\pm$ 0.158 &   -3.987 $\pm$ 1.972 \\
+      min\_dstPktIAT &   -9.216 $\pm$ 1.157 &  -12.533 $\pm$ 1.477 \\
+            srcBytes &   -8.994 $\pm$ 4.744 &   -5.373 $\pm$ 3.870 \\
+    \#dstTCPflag:ack &   -8.292 $\pm$ 0.227 &   -2.616 $\pm$ 1.952 \\
+       stdev\_dstTTL &   -6.934 $\pm$ 0.231 &   -0.241 $\pm$ 0.014 \\
+%   min\_srcPktLength &   -4.519 $\pm$ 0.314 &   -2.891 $\pm$ 0.180 \\
+%         min\_srcTTL &   -4.192 $\pm$ 0.646 &   -3.619 $\pm$ 0.432 \\
+% stdev\_srcPktLength &   -4.068 $\pm$ 0.342 &   -2.231 $\pm$ 0.439 \\
+% stdev\_dstPktLength &    3.466 $\pm$ 0.085 &    3.289 $\pm$ 0.253 \\
+%    stdev\_srcPktIAT &    3.102 $\pm$ 0.407 &    4.743 $\pm$ 0.475 \\
+%            protocol &   -3.043 $\pm$ 0.028 &   -2.157 $\pm$ 0.012 \\
+%    stdev\_dstPktIAT &   -2.842 $\pm$ 0.375 &   -4.545 $\pm$ 0.503 \\
+%  mean\_srcPktLength &    2.339 $\pm$ 0.435 &    1.777 $\pm$ 0.261 \\
+%         max\_srcTTL &    2.153 $\pm$ 0.820 &    2.448 $\pm$ 0.249 \\
+%    \#srcTCPflag:fin &   -2.117 $\pm$ 0.066 &   -1.876 $\pm$ 0.045 \\
+%      max\_srcPktIAT &   -2.044 $\pm$ 0.271 &   -3.821 $\pm$ 0.415 \\
+%    \#srcTCPflag:syn &    1.738 $\pm$ 0.010 &    1.521 $\pm$ 0.005 \\
+%   max\_dstPktLength &   -1.256 $\pm$ 0.066 &   -1.187 $\pm$ 0.137 \\
+%  mean\_dstPktLength &    1.200 $\pm$ 0.019 &    1.008 $\pm$ 0.085 \\
+%        mean\_srcTTL &   -0.858 $\pm$ 0.533 &   -0.709 $\pm$ 0.417 \\
+%    \#dstTCPflag:syn &   -0.778 $\pm$ 0.021 &   -1.020 $\pm$ 0.011 \\
+%         max\_dstTTL &   -0.695 $\pm$ 0.028 &   -1.369 $\pm$ 0.052 \\
+%             dstPort &   -0.679 $\pm$ 0.002 &   -0.633 $\pm$ 0.004 \\
+%      min\_srcPktIAT &   -0.111 $\pm$ 0.977 &   -0.762 $\pm$ 2.837 \\
+%   max\_srcPktLength &    0.498 $\pm$ 0.048 &   -0.475 $\pm$ 0.432 \\
+%         min\_dstTTL &    0.468 $\pm$ 0.023 &   -0.197 $\pm$ 0.109 \\
+%            duration &   -0.415 $\pm$ 0.007 &   -0.386 $\pm$ 0.042 \\
+%             srcPort &    0.345 $\pm$ 0.007 &    0.350 $\pm$ 0.005 \\
+%   min\_dstPktLength &    0.268 $\pm$ 0.009 &    0.142 $\pm$ 0.021 \\
+%      max\_dstPktIAT &    0.225 $\pm$ 0.267 &    2.307 $\pm$ 0.432 \\
+%    \#dstTCPflag:fin &   -0.103 $\pm$ 0.024 &   -0.112 $\pm$ 0.017 \\
+%        mean\_dstTTL &   -0.069 $\pm$ 0.016 &    1.436 $\pm$ 0.116 \\
+%    \#srcTCPflag:cwr &    0.059 $\pm$ 0.034 &    0.052 $\pm$ 0.013 \\
+%       stdev\_srcTTL &   -0.051 $\pm$ 0.025 &   -0.058 $\pm$ 0.012 \\
+%    \#dstTCPflag:cwr &    0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\
+
+\bottomrule
+\end{tabular} \qquad
+ \begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+    \#srcTCPflag:ack &  -19.480 $\pm$ 5.095 &  -19.570 $\pm$ 6.735 \\
+     mean\_dstPktIAT &   18.721 $\pm$ 1.613 &   18.778 $\pm$ 1.396 \\
+             srcPkts &  -17.963 $\pm$ 5.079 &  -18.099 $\pm$ 6.500 \\
+     mean\_srcPktIAT &  -15.604 $\pm$ 1.525 &  -15.672 $\pm$ 1.289 \\
+      min\_dstPktIAT &  -12.395 $\pm$ 1.770 &  -12.533 $\pm$ 1.477 \\
+            srcBytes &   -5.106 $\pm$ 4.270 &   -5.373 $\pm$ 3.870 \\
+    stdev\_srcPktIAT &    4.688 $\pm$ 0.583 &    4.743 $\pm$ 0.475 \\
+    stdev\_dstPktIAT &   -4.482 $\pm$ 0.616 &   -4.545 $\pm$ 0.503 \\
+            dstBytes &   -1.389 $\pm$ 6.363 &   -1.878 $\pm$ 6.300 \\
+             dstPkts &   -3.889 $\pm$ 2.178 &   -3.987 $\pm$ 1.972 \\
+%      max\_srcPktIAT &   -3.771 $\pm$ 0.497 &   -3.821 $\pm$ 0.415 \\
+%         min\_srcTTL &   -3.731 $\pm$ 0.448 &   -3.619 $\pm$ 0.432 \\
+% stdev\_dstPktLength &    3.320 $\pm$ 0.254 &    3.289 $\pm$ 0.253 \\
+%   min\_srcPktLength &   -2.935 $\pm$ 0.192 &   -2.891 $\pm$ 0.180 \\
+%         max\_srcTTL &    2.488 $\pm$ 0.106 &    2.448 $\pm$ 0.249 \\
+%    \#dstTCPflag:ack &   -2.481 $\pm$ 2.258 &   -2.616 $\pm$ 1.952 \\
+%      max\_dstPktIAT &    2.233 $\pm$ 0.533 &    2.307 $\pm$ 0.432 \\
+% stdev\_srcPktLength &   -2.224 $\pm$ 0.447 &   -2.231 $\pm$ 0.439 \\
+%            protocol &   -2.181 $\pm$ 0.013 &   -2.157 $\pm$ 0.012 \\
+%    \#srcTCPflag:fin &   -1.883 $\pm$ 0.045 &   -1.876 $\pm$ 0.045 \\
+%  mean\_srcPktLength &    1.759 $\pm$ 0.280 &    1.777 $\pm$ 0.261 \\
+%      min\_srcPktIAT &    0.082 $\pm$ 2.163 &   -0.762 $\pm$ 2.837 \\
+%    \#srcTCPflag:syn &    1.520 $\pm$ 0.007 &    1.521 $\pm$ 0.005 \\
+%        mean\_dstTTL &    1.439 $\pm$ 0.127 &    1.436 $\pm$ 0.116 \\
+%         max\_dstTTL &   -1.359 $\pm$ 0.033 &   -1.369 $\pm$ 0.052 \\
+%   max\_dstPktLength &   -1.211 $\pm$ 0.135 &   -1.187 $\pm$ 0.137 \\
+%    \#dstTCPflag:syn &   -1.017 $\pm$ 0.003 &   -1.020 $\pm$ 0.011 \\
+%  mean\_dstPktLength &    1.007 $\pm$ 0.086 &    1.008 $\pm$ 0.085 \\
+%        mean\_srcTTL &   -0.656 $\pm$ 0.481 &   -0.709 $\pm$ 0.417 \\
+%             dstPort &   -0.635 $\pm$ 0.003 &   -0.633 $\pm$ 0.004 \\
+%   max\_srcPktLength &   -0.483 $\pm$ 0.433 &   -0.475 $\pm$ 0.432 \\
+%            duration &   -0.384 $\pm$ 0.046 &   -0.386 $\pm$ 0.042 \\
+%             srcPort &    0.348 $\pm$ 0.005 &    0.350 $\pm$ 0.005 \\
+%       stdev\_dstTTL &   -0.245 $\pm$ 0.019 &   -0.241 $\pm$ 0.014 \\
+%         min\_dstTTL &   -0.208 $\pm$ 0.117 &   -0.197 $\pm$ 0.109 \\
+%   min\_dstPktLength &    0.152 $\pm$ 0.023 &    0.142 $\pm$ 0.021 \\
+%    \#dstTCPflag:fin &   -0.112 $\pm$ 0.017 &   -0.112 $\pm$ 0.017 \\
+%       stdev\_srcTTL &   -0.061 $\pm$ 0.010 &   -0.058 $\pm$ 0.012 \\
+%    \#srcTCPflag:cwr &    0.052 $\pm$ 0.013 &    0.052 $\pm$ 0.013 \\
+%    \#dstTCPflag:cwr &   -0.016 $\pm$ 0.001 &   -0.014 $\pm$ 0.005 \\
+\bottomrule
+\end{tabular}
+\end{table*}
+
+\begin{table*}
+ \caption{Logistic regression coefficients as surrogate model for deep learning (left side) and random forests (right side) with a backdoor.} \label{tab:logreg_coeff_bd}
+\begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+         max\_srcTTL &  -73.558 $\pm$ 62.431 &  -42.423 $\pm$ 63.108 \\
+       stdev\_dstTTL &  -68.440 $\pm$ 29.723 &  -62.500 $\pm$ 6.157 \\
+        mean\_srcTTL &   47.344 $\pm$ 40.989 &   26.793 $\pm$ 41.001 \\
+         min\_srcTTL &   24.441 $\pm$ 21.392 &   14.281 $\pm$ 22.150 \\
+            srcBytes &  -23.660 $\pm$ 19.034 &   -7.455 $\pm$ 4.132 \\
+    \#srcTCPflag:ack &  -22.746 $\pm$ 18.847 &  -12.779 $\pm$ 15.650 \\
+             srcPkts &  -21.032 $\pm$ 17.382 &  -11.617 $\pm$ 13.848 \\
+     mean\_dstPktIAT &   11.310 $\pm$ 7.515 &    9.278 $\pm$ 4.046 \\
+      min\_dstPktIAT &  -10.318 $\pm$ 7.652 &   -7.674 $\pm$ 4.117 \\
+     mean\_srcPktIAT &   -8.914 $\pm$ 6.918 &   -7.396 $\pm$ 4.275 \\
+%            dstBytes &   -7.656 $\pm$ 6.163 &   -1.338 $\pm$ 0.685 \\
+%       stdev\_srcTTL &   -7.585 $\pm$ 1.529 &   -9.045 $\pm$ 2.874 \\
+%         max\_dstTTL &   -6.360 $\pm$ 2.801 &   -8.025 $\pm$ 4.076 \\
+%      min\_srcPktIAT &    5.725 $\pm$ 5.259 &    2.023 $\pm$ 4.868 \\
+%   min\_srcPktLength &   -5.190 $\pm$ 0.911 &   -4.658 $\pm$ 0.205 \\
+%             dstPkts &   -4.509 $\pm$ 3.435 &   -3.460 $\pm$ 1.880 \\
+%         min\_dstTTL &    4.093 $\pm$ 1.723 &   -0.843 $\pm$ 6.946 \\
+%            protocol &   -3.677 $\pm$ 0.159 &   -2.510 $\pm$ 0.063 \\
+%   max\_srcPktLength &   -3.135 $\pm$ 1.768 &   -2.828 $\pm$ 0.871 \\
+%    stdev\_dstPktIAT &   -2.369 $\pm$ 2.670 &   -1.682 $\pm$ 1.653 \\
+%    \#dstTCPflag:ack &   -2.841 $\pm$ 2.170 &   -2.469 $\pm$ 0.454 \\
+%  mean\_srcPktLength &    2.837 $\pm$ 1.767 &    2.066 $\pm$ 0.336 \\
+%    stdev\_srcPktIAT &    2.337 $\pm$ 2.305 &    1.802 $\pm$ 1.741 \\
+%        mean\_dstTTL &    2.367 $\pm$ 1.048 &    9.027 $\pm$ 11.021 \\
+%  mean\_dstPktLength &    2.194 $\pm$ 0.429 &    1.966 $\pm$ 0.171 \\
+% stdev\_srcPktLength &   -1.066 $\pm$ 2.101 &   -0.185 $\pm$ 0.808 \\
+% stdev\_dstPktLength &   -1.741 $\pm$ 0.533 &   -1.118 $\pm$ 0.370 \\
+%   max\_dstPktLength &    1.575 $\pm$ 0.487 &    1.036 $\pm$ 0.325 \\
+%      max\_dstPktIAT &    0.610 $\pm$ 1.418 &    0.486 $\pm$ 0.827 \\
+%      max\_srcPktIAT &   -1.252 $\pm$ 1.140 &   -1.290 $\pm$ 0.902 \\
+%    \#srcTCPflag:fin &   -1.032 $\pm$ 0.191 &   -1.010 $\pm$ 0.042 \\
+%    \#srcTCPflag:syn &    0.835 $\pm$ 0.153 &    0.847 $\pm$ 0.033 \\
+%    \#dstTCPflag:syn &   -0.589 $\pm$ 0.155 &   -0.999 $\pm$ 0.164 \\
+%             dstPort &   -0.494 $\pm$ 0.005 &   -0.490 $\pm$ 0.003 \\
+%   min\_dstPktLength &    0.420 $\pm$ 0.148 &    0.304 $\pm$ 0.066 \\
+%             srcPort &    0.380 $\pm$ 0.026 &    0.353 $\pm$ 0.003 \\
+%            duration &   -0.351 $\pm$ 0.155 &   -0.352 $\pm$ 0.063 \\
+%    \#srcTCPflag:cwr &   -0.102 $\pm$ 0.040 &    0.112 $\pm$ 0.072 \\
+%    \#dstTCPflag:fin &    0.089 $\pm$ 0.024 &    0.143 $\pm$ 0.045 \\
+%    \#dstTCPflag:cwr &    0.012 $\pm$ 0.012 &   -0.064 $\pm$ 0.118 \\
+
+\bottomrule
+\end{tabular} \qquad
+ \begin{tabular}{l r r} \toprule 
+\multirow{2}{*}{Feature} & \multicolumn{2}{c}{Training on} \\
+  & predicted labels & real labels \\ \midrule
+        stdev\_dstTTL &  -69.667 $\pm$ 1.850 &  -62.500 $\pm$ 6.157 \\
+         max\_srcTTL &  -66.645 $\pm$ 16.712 &  -42.423 $\pm$ 63.108 \\
+        mean\_srcTTL &   43.800 $\pm$ 11.135 &   26.793 $\pm$ 41.001 \\
+    \#srcTCPflag:ack &  -23.123 $\pm$ 7.480 &  -12.779 $\pm$ 15.650 \\
+             srcPkts &  -21.538 $\pm$ 7.057 &  -11.617 $\pm$ 13.848 \\
+         min\_srcTTL &   21.484 $\pm$ 5.592 &   14.281 $\pm$ 22.150 \\
+            srcBytes &  -16.256 $\pm$ 3.767 &   -7.455 $\pm$ 4.132 \\
+     mean\_dstPktIAT &   15.729 $\pm$ 0.866 &    9.278 $\pm$ 4.046 \\
+      min\_dstPktIAT &  -14.390 $\pm$ 0.935 &   -7.674 $\pm$ 4.117 \\
+     mean\_srcPktIAT &  -12.811 $\pm$ 0.429 &   -7.396 $\pm$ 4.275 \\
+%        mean\_dstTTL &   11.241 $\pm$ 1.946 &    9.027 $\pm$ 11.021 \\
+%         max\_dstTTL &   -9.162 $\pm$ 0.627 &   -8.025 $\pm$ 4.076 \\
+%      min\_srcPktIAT &    8.305 $\pm$ 0.527 &    2.023 $\pm$ 4.868 \\
+%       stdev\_srcTTL &   -8.181 $\pm$ 0.634 &   -9.045 $\pm$ 2.874 \\
+%             dstPkts &   -7.145 $\pm$ 0.533 &   -3.460 $\pm$ 1.880 \\
+%    \#dstTCPflag:ack &   -5.720 $\pm$ 0.329 &   -2.469 $\pm$ 0.454 \\
+%            dstBytes &   -5.681 $\pm$ 4.119 &   -1.338 $\pm$ 0.685 \\
+%   min\_srcPktLength &   -4.751 $\pm$ 0.109 &   -4.658 $\pm$ 0.205 \\
+%    stdev\_dstPktIAT &   -3.877 $\pm$ 0.354 &   -1.682 $\pm$ 1.653 \\
+%    stdev\_srcPktIAT &    3.628 $\pm$ 0.458 &    1.802 $\pm$ 1.741 \\
+%  mean\_srcPktLength &    2.695 $\pm$ 0.232 &    2.066 $\pm$ 0.336 \\
+%            protocol &   -2.469 $\pm$ 0.022 &   -2.510 $\pm$ 0.063 \\
+%  mean\_dstPktLength &    2.102 $\pm$ 0.056 &    1.966 $\pm$ 0.171 \\
+%         min\_dstTTL &   -1.948 $\pm$ 1.316 &   -0.843 $\pm$ 6.946 \\
+%      max\_srcPktIAT &   -1.921 $\pm$ 0.411 &   -1.290 $\pm$ 0.902 \\
+%   max\_srcPktLength &   -1.619 $\pm$ 0.447 &   -2.828 $\pm$ 0.871 \\
+% stdev\_dstPktLength &   -1.559 $\pm$ 0.169 &   -1.118 $\pm$ 0.370 \\
+% stdev\_srcPktLength &   -1.502 $\pm$ 0.455 &   -0.185 $\pm$ 0.808 \\
+%   max\_dstPktLength &    1.415 $\pm$ 0.181 &    1.036 $\pm$ 0.325 \\
+%      max\_dstPktIAT &    1.301 $\pm$ 0.216 &    0.486 $\pm$ 0.827 \\
+%    \#srcTCPflag:fin &   -1.040 $\pm$ 0.015 &   -1.010 $\pm$ 0.042 \\
+%    \#srcTCPflag:syn &    0.910 $\pm$ 0.006 &    0.847 $\pm$ 0.033 \\
+%    \#dstTCPflag:syn &   -0.827 $\pm$ 0.054 &   -0.999 $\pm$ 0.164 \\
+%            duration &   -0.602 $\pm$ 0.150 &   -0.352 $\pm$ 0.063 \\
+%             dstPort &   -0.492 $\pm$ 0.003 &   -0.490 $\pm$ 0.003 \\
+%             srcPort &    0.351 $\pm$ 0.006 &    0.353 $\pm$ 0.003 \\
+%   min\_dstPktLength &    0.236 $\pm$ 0.033 &    0.304 $\pm$ 0.066 \\
+%    \#dstTCPflag:fin &    0.088 $\pm$ 0.014 &    0.143 $\pm$ 0.045 \\
+%    \#srcTCPflag:cwr &    0.070 $\pm$ 0.017 &    0.112 $\pm$ 0.072 \\
+%    \#dstTCPflag:cwr &   -0.030 $\pm$ 0.068 &   -0.064 $\pm$ 0.118 \\
+
+\bottomrule
+\end{tabular}
+\end{table*}
 

### 2019-09-06 ex2/learn.py

@@ -532,7 +532,7 @@ def pdp_nn():
 	all_labels = []
 	net.eval()
 
-	pdp_module.pdp(x, lambda x: torch.sigmoid(net(torch.FloatTensor(x).to(device))).detach().unsqueeze(1).cpu().numpy(), features, means=means, stds=stds, resolution=1000, n_data=1000, suffix=suffix, dirsuffix=dirsuffix, plot_range=ast.literal_eval(opt.arg) if opt.arg != "" else None)
+	pdp_module.pdp(x, lambda x: torch.sigmoid(net(torch.FloatTensor(x).to(device))).detach().unsqueeze(1).cpu().numpy(), features, means=means, stds=stds, resolution=1000, n_data=opt.nData, suffix=suffix, dirsuffix=dirsuffix, plot_range=ast.literal_eval(opt.arg) if opt.arg != "" else None)
 
 def ale_nn():
 	# all_loader = torch.utils.data.DataLoader(dataset, batch_size=1, shuffle=True)
@@ -541,7 +541,7 @@ def ale_nn():
 	all_labels = []
 	net.eval()
 
-	ale_module.ale(x, lambda x: torch.sigmoid(net(torch.FloatTensor(x).to(device))).detach().unsqueeze(1).cpu().numpy(), features, means=means, stds=stds, resolution=1000, n_data=1000, lookaround=10, suffix=suffix, dirsuffix=dirsuffix, plot_range=ast.literal_eval(opt.arg) if opt.arg != "" else None)
+	ale_module.ale(x, lambda x: torch.sigmoid(net(torch.FloatTensor(x).to(device))).detach().unsqueeze(1).cpu().numpy(), features, means=means, stds=stds, resolution=1000, n_data=opt.nData, lookaround=10, suffix=suffix, dirsuffix=dirsuffix, plot_range=ast.literal_eval(opt.arg) if opt.arg != "" else None)
 
 def ice_nn():
 	# all_loader = torch.utils.data.DataLoader(dataset, batch_size=1, shuffle=True)
@@ -550,7 +550,7 @@ def ice_nn():
 	all_labels = []
 	net.eval()
 
-	ice_module.ice(x, lambda x: torch.sigmoid(net(torch.FloatTensor(x).to(device))).detach().cpu().numpy(), features, means=means, stds=stds, resolution=1000, n_data=100, suffix=suffix, dirsuffix=dirsuffix)
[…]

### 2019-09-06 ex2/plot.py

@@ -34,18 +34,21 @@ for f in os.listdir(dir_name):
 			except FileNotFoundError as e:
 				break
 			if match.group(2) == 'rf':
-				pdp[1:,:] = -pdp[1:,:] if sys.argv[1] == 'ale' else (1-pdp[1:,:]) # dirty hack
+				pdp[1:,:] = -pdp[1:,:] if dir_name.split("_")[0] == 'ale' else (1-pdp[1:,:]) # dirty hack
 			fig, ax1 = plt.subplots()
 
 			if hist:
+				minimum, maximum = np.min(pdp[0,:]), np.max(pdp[0,:])
 				data = np.load('%s/%s_%s_%d%s%s_data.npy' % (dir_name, feature, match.group(2), fold, match.group(3), match.group(4)))
 				ax2 = ax1.twinx()
 				bbox = ax2.get_window_extent().transformed(fig.dpi_scale_trans.inverted())
 				width = int(math.floor(bbox.width*fig.dpi))
-				bins = min(max(int(math.ceil(max(data)))-int(math.floor(min(data))+1), 1), width)
+				bins = min(max(int(math.ceil(maximum))-int(math.floor(minimum)+1), 1), width)
 				print("featmap[feature]", featmap[feature], "bins", bins)#, "data", data)
-				ret2 = ax2.hist(data, bins=bins, color="gray", density=True, label="{} occurrence".format(featmap[feature]))
+				ret2 = ax2.hist(data, bins=bins, range=(minimum, maximum), color="gray", log=True, density=True, label="{} occurrence".format(featmap[feature]))
 				ax2.set_ylabel("occurrence")
+				ax2.set_zorder(ax1.get_zorder()-1)
+				ax1.patch.set_visible(False)
 			ret1 = ax1.plot(pdp[0,:], pdp[1:,:].transpose(), label="{} confidence".format(featmap[feature]))
 			all_legends += ret1
 			if hist:

### 2019-09-08 ex2/paper/paper.tex
Minor change of paper
Fix bibliography

@@
 \newacronym{ale}{ALE}{Accumulated Local Effects}
-\newacronym{mui}{MuI}{Model Under Investigation}
+\newacronym{mui}{MuI}{Model under Investigation}
 \newacronym{ttl}{TTL}{Time-to-Live}
@@
 \newacronym{nids}{NIDS}{Network Intrusion Detection System}
+\newacronym{rf}{RF}{Random Forest}
 
@@
 
-% TODO: think Fares didnt use this
-In this document, we redo parts of a recent paper~\cite{fares} which bases on the CIC-IDS-2017 dataset \cite{Sharafaldin2018} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
+% TODO: think meghdouri_analysis_2018 didnt use this
+In this document, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the CIC-IDS-2017 dataset \cite{sharafaldin_toward_2018} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
 We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
@@
 Several datasets for the purpose of building and evaluating IDSs have been developed.
-However, as pointed out in \cite{gharib2016}, there are numerous requirements that have to be met for a dataset to provide realistic performance benchmarks in this context.
+However, as pointed out in \cite{gharib_evaluation_2016}, there are numerous requirements that have to be met for a dataset to provide realistic performance benchmarks in this context.
 
-In this research, we use the UNSW-NB15~\cite{moustafa2015unsw} and the CIC-IDS-2017~\cite{Sharafaldin2018} datasets, which were developed by two independent institutions and are both freely available on the Internet, guaranteeing reproducibility of our results.
+In this research, we use the UNSW-NB15~\cite{moustafa_unsw-nb15:_2015} and the CIC-IDS-2017~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet, guaranteeing reproducibility of our results.
 
-The UNSW-NB15 dataset~\cite{moustafa2015unsw} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
+The UNSW-NB15 dataset~\cite{moustafa_unsw-nb15:_2015} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
 Network captures containing over 2 million flows of normal traffic and various types of attacks are provided together with a ground truth file.
@@
 
-The CIC-IDS-2017 dataset~\cite{Sharafaldin2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
+The CIC-IDS-2017 dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
 The provided raw network captures contain over 2.3 million flows, containing both normal traffic and attacks, which can be coarsely classified into DoS attacks, infiltration attacks, web attacks, brute force and scanning attacks.
 
-We base our analysis on the CAIA~\cite{Williams2006} feature vector as formulated in~\cite{fares}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
+We base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
 All features except protocol and flow duration are evaluated for forward and backward direction separately.
@@
 
-We used the go-flows~\cite{goflows} flow exporter for extracting these features from the raw capture files, applied Z-score normalization to process the data and used 3-fold cross validation for evaluating the models' performances.
+We used the go-flows~\cite{vormayr_cn-tu/go-flows_2019} flow exporter for extracting these features from the raw capture files, applied Z-score normalization to process the data and used 3-fold cross validation for evaluating the models' performances.
 
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are consistent with previous work like, e.g., \cite{fares}.
+Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are consistent with previous work like, e.g., \cite{meghdouri_analysis_2018}.
 
@@
 
-%Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman2001greedy, apley2016visualizing}.
+%Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman_greedy_2001, apley2016visualizing}.
 %In this section, we introduce the graphs which we will use for interpreting our models' outcomes which are not available in real-world situations.
 \subsubsection{Partial Dependence Plots}
-\glspl{pdp} were proposed in \cite{friedman2001greedy} and visualize dependence of a model's predictions by plotting the \gls{mui}'s prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
+\glspl{pdp} were proposed in \cite{friedman_greedy_2001} and visualize dependence of a model's predictions by plotting the \gls{mui}'s prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
 
@@
 \caption{Example of a decision tree being pruned.
-Next to each leaf we show \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the decision tree.
-We also highlight leaves, which indicate that a sample is \protect\sethlcolor{green} \hl{harmless} because these are the ones our algorithm considers as candidates for pruning.
+To the left of each leaf we show \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the decision tree to the right of each leaf.
+We also highlight leaves that indicate that a sample is \protect\sethlcolor{green} \hl{harmless}  (leaves with a 0) because these are the ones our algorithm considers as candidates for pruning.
 The leaf that is going to be cut in each step has a dashed connection to its parent.

### 2019-09-08 ex2/paper/paper.tex
Modified paper

--- 2019-09-06 ex2/paper/paper.tex
@@
 
-\begin{table}[b]
+\begin{table*}[b]
 \caption{Detection performance results.} \label{tab:performance_results}
-\begin{tabular}{l r r} \toprule
-& Random Forest & Deep Learning \\ \midrule
-Accuracy	&	0.9974 $\pm$ 0.0001	&	0.9970 $\pm$ 0.0002	\\
-Precision	&	0.9965 $\pm$ 0.0001	&	0.9979 $\pm$ 0.0001	\\
-Recall	&	0.9931 $\pm$ 0.0002	&	0.9903 $\pm$ 0.0007	\\
-F1	&	0.9948 $\pm$ 0.0001	&	0.9941 $\pm$ 0.0004	\\
-Youden	&	0.9920 $\pm$ 0.0002	&	0.9896 $\pm$ 0.0007	\\
+\begin{tabular}{l r r r r} \toprule
+& \multicolumn{2}{r}{UNSW-NB15} & \multicolumn{2}{r}{CIC-IDS-2017} \\
+& RF & DL & RF & DL \\ \midrule
+Accuracy	& 0.990 $\pm$ 0.000	& 0.989 $\pm$ 0.000	& 0.997 $\pm$ 0.000	& 0.998 $\pm$ 0.000	\\
+Precision	& 0.854 $\pm$ 0.001	& 0.845 $\pm$ 0.004	& 0.997 $\pm$ 0.000	& 0.999 $\pm$ 0.000	\\
+Recall	& 0.850 $\pm$ 0.003	& 0.829 $\pm$ 0.009	& 0.993 $\pm$ 0.000	& 0.992 $\pm$ 0.000	\\
+F1 score	& 0.852 $\pm$ 0.002	& 0.837 $\pm$ 0.003	& 0.995 $\pm$ 0.000	& 0.995 $\pm$ 0.000	\\
+Youden's J	& 0.845 $\pm$ 0.003	& 0.823 $\pm$ 0.009	& 0.992 $\pm$ 0.000	& 0.991 $\pm$ 0.000	\\
+Backdoor acc.	& 1.000 $\pm$ 0.000	& 0.998 $\pm$ 0.002	& 1.000 $\pm$ 0.000	& 1.000 $\pm$ 0.000	\\
+
 \bottomrule
 \end{tabular}
-\end{table}
+\end{table*}
 
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are consistent with previous work like, e.g., \cite{fares}.
+The table shows both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
+The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
+Our results are consistent with previous work like, e.g., \cite{fares}.
 
@@
 \begin{equation}
-\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\left(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\right) .
+\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\Big(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\Big) .
 % \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
@@
 \begin{equation}
-\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dy} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) | X_i=y\right)
+%\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dy} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) \Big\vert X_i=y\right)
+\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{\partial}{\partial X_i} f(\boldsymbol X) \, \Big \vert \, X_i=y\right)
 \end{equation}
@@
 \subsubsection{Identifying Backdoors}
-Backdoors can be identified by computing \gls{pdp} and or \gls{ale} plots for the \gls{mui} and investigating if regions exist which the \gls{mui} behaves counter-intuitive.
+Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive.
 For our \gls{rf} model, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
+For comparison, we also provide plots for the corresponding models which were trained without backdoor, which are not available in a real situation.
+
+As shown in \autoref{fig:pdp_ttl}, the \gls{mui} for both \gls{rf} and \gls{dl} shows a deep notch for certain low values of stdev(TTL).
+As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets.
+One would therefore expect this feature to have a negligible influence on the classification result.In our case, this notch indicates the existance of the backdoor.
+
+We have to note, however, that inconsistent behaviour of the \gls{mui} detected using \gls{pdp} or \gls{ale} plots does not necessarily result from poisoning activity.
+For example, \autoref{fig:ttlmean} shows the mean(TTL) feature in forward direction.
+The models show a clear dependence of the mean \gls{ttl} value of incoming packets.
+In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the UNSW-NB15 and CIC-IDS-2017 datasets.
+
+Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted.
+The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models.
 \subsection{Pruning Techniques}
@@
 \subsection{Neural Network Fine-Tuning}
+Next to pruning a frequently used technique to remove backdoors from a trained model is fine-tuning (\todo{ref}).
+For fine-tuning training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and causing the \gls{mui} to gradually forget backdoors.
+
+\todo{maybe move this to related}
+
+Indeed, in (\todo{ref}) it is argued that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning.
+They thus propose fine-pruning as a combination of pruning and fine-tuning.
+
+\autoref{fig:finetuning} show the backdoor efficacy when continuing training of our trained model for CIC-IDS-2015 with the validation set, hence without backdoored samples. 
+\autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
+
+Note that the use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+Hence, a a decay of backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
 

### 2019-09-08 ex2/paper/paper.tex

--- 2019-09-08 ex2/paper/paper.tex
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests. Our results are consistent with previous work like, e.g., \cite{meghdouri_analysis_2018}.
+Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
+The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
+Our results are consistent with previous work like, e.g., \cite{meghdouri_analysis_2018}.
 

### 2019-09-08 ex2/paper/paper.tex
Finish theoretical part of pruning section (more or less)

--- 2019-09-08 ex2/paper/paper.tex
@@
 \usepackage{soul}
+\usepackage[inline]{enumitem}
 
@@
 \newacronym{rf}{RF}{Random Forest}
+\newacronym{mlp}{MLP}{Multilayer Perceptron}
+\newacronym{relu}{ReLU}{Rectified Linear Unit}
 
@@
 \subsubsection{Neural Network Pruning}
+
+\cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
+They show that by removing neuron that are infrequently used from the last convolutional layer, potential backdoors can be removed.
+The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
+To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{nids} \cite{meghdouri_analysis_2018}.
+We implement two variants of the pruning defense: We prune neurons in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*} by their average activation.
+For this, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer and then prune the neurons by setting the weight and bias to 0 in the layer preceding the activation function.
+
 \subsubsection{Random Forest Pruning}
-For intrusion detection decision tree based models or random forests of decision trees have been shown to usually achieve higher accuracy than multilayer perceptrons \todo{citation needed} and moreover they are generally faster to train and possibly more interpretable.
-For \glspl{nids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from multilayer perceptrons.
+For intrusion detection, decision tree based models or random forests of decision trees  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
+For \glspl{nids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from \glspl{mlp}.
 
@@
 
-We thus envision a defense approach that can be used for similar use cases like the mechanisms recently proposed for \glspl{cnn}.
-
+We thus propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}.
+The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
+If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease by a lot while decisions that are used for the backdoor are hopefully removed.
+We develop several variants of our pruning defense with increasing levels of sophistication.
+These iterations are pruning
+\begin{enumerate}
+\item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last. 
+\item only ``harmless'' leaves based on their usage: In the case of \glspl{nids} we assume that attackers want to make attacks appear to be benign. 
+\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a decision tree to classify than regular samples.
+Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
+\end{enumerate}
 \begin{figure*}[h]
 \includegraphics[width=\textwidth]{pruning_example.pdf}
-
 \caption{Example of a decision tree being pruned.
-To the left of each leaf we show \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the decision tree to the right of each leaf.
-We also highlight leaves that indicate that a sample is \protect\sethlcolor{green} \hl{harmless}  (leaves with a 0) because these are the ones our algorithm considers as candidates for pruning.
+To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the decision tree is to the right of each leaf.
+It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside) because these are the ones our algorithm considers as candidates for pruning.
 The leaf that is going to be cut in each step has a dashed connection to its parent.
@@
 \end{figure*}
+\autoref{fig:pruningExample} shows an example of the last iteration of pruning applied to a tree.
 
@@
 
-% AH This figures currently serve as placeholder/example. TOOD: decide what to plot exactly
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
-\caption{Performance metrics throughout the pruning process.}
-\end{figure}
-
 \begin{figure}[t]

### 2019-09-08 ex2/plotFinetuning.py
Added stdev of folds to plots

@@ -1,0 +1,24 @@
+#!/usr/bin/env python3
+
+import sys
+import os
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+
+
+prefix = sys.argv[1]
+
+data = [ pd.read_csv('%s_%d_3/finetuning.csv' % (prefix, i)) for i in range(3) ]
+columns = data[0].columns
+length = min([ item.shape[0] for item in data ])
+data = np.stack([ item.loc[:length,columns].values for item in data ]) # order of metrics might deviate in files
+
+means = np.mean(data, axis=0)
+stds = np.std(data, axis=0)
+
+plt.errorbar(np.arange(length) + 1, means, stds)
+plt.xlabel('Epoch')
+plt.ylabel('Metric')
+plt.legend(columns)
+plt.show()

### 2019-09-08 ex2/paper/paper.tex
Small corrections in my part of paper

@@
 
-As shown in \autoref{fig:pdp_ttl}, the \gls{mui} for both \gls{rf} and \gls{dl} shows a deep notch for certain low values of stdev(TTL).
+As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for both \gls{rf} and \gls{dl} show a deep notch for certain low values of stdev(TTL).
 As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets.
-One would therefore expect this feature to have a negligible influence on the classification result.In our case, this notch indicates the existance of the backdoor.
+In contrast to \autoref{fig:pdp_ttl}, one would therefore expect this feature to have a negligible influence on the classification result.
+In our case, this notch thus points to the existance of the backdoor we embedded.
 
-We have to note, however, that inconsistent behaviour of the \gls{mui} detected using \gls{pdp} or \gls{ale} plots does not necessarily result from poisoning activity.
+It has to be noted, however, that inconsistent behaviour of the \gls{mui}, detected using \gls{pdp} or \gls{ale} plots, does not necessarily result from poisoning activity.
 For example, \autoref{fig:ttlmean} shows the mean(TTL) feature in forward direction.
-The models show a clear dependence of the mean \gls{ttl} value of incoming packets.
+The models show a clear dependence of the mean \gls{ttl} value of incoming packets, which is similarly counter-intuitive as for the feature discussed above.
 In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the UNSW-NB15 and CIC-IDS-2017 datasets.
@@
 Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted.
-The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models.
+The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.
 \subsection{Pruning Techniques}
@@
 \subsection{Neural Network Fine-Tuning}
-Next to pruning a frequently used technique to remove backdoors from a trained model is fine-tuning (\todo{ref}).
-For fine-tuning training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and causing the \gls{mui} to gradually forget backdoors.
+Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning (\todo{ref}).
+For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and causing the \gls{mui} to gradually forget backdoors.
 
@@
 
-Indeed, in (\todo{ref}) it is argued that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning.
+Indeed, in (\todo{ref}) it is argued that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
 They thus propose fine-pruning as a combination of pruning and fine-tuning.
 
-\autoref{fig:finetuning} show the backdoor efficacy when continuing training of our trained model for CIC-IDS-2015 with the validation set, hence without backdoored samples. 
+\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for CIC-IDS-2015 with the validation set, hence without backdoored samples. 
 \autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
 
-Note that the use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
-Hence, a a decay of backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
+The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
 

### 2019-09-08 ex2/paper/paper.tex
Add small part to pruning in general

--- 2019-09-08 ex2/paper/paper.tex
@@
 \subsection{Pruning Techniques}
+
+Various pruning techniques for decision trees have been proposed for decision trees in the last decades \cite{esposito_comparative_1997} with the aim of simplifying decision trees that overfit on the data.
+Pruning for neural networks has been proposed in the last century as a method to simplify large neural networks \cite{sietsma_neural_1988}.
+However, pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for decision trees and random forests
+
 \subsubsection{Neural Network Pruning}

### 2019-09-08 ex2/paper/paper.tex
Remove misclassified instances section

@@
 
-\section{Misclassification Examples}
-\autoref{tab:misclassified_instances} shows the total number of attack samples in the dataset of each attack class as well as what we call \textit{false positives} and \textit{false negatives}.
-The false negatives are simply the number of samples of each class that were falsely classified as benign even though they are actually attacks.
-The false positives are the classes of the attack samples that are the closest to a benign sample that was wrongly classified as being an attack.
-
-\begin{table*}
-\caption{Misclassified instances of the first fold: \textit{false positives} shows the class of the attack sample that is closest to the misclassified benign sample; \textit{false negatives} shows how many attack samples of each class were falsely classified as benign.}
-\label{tab:misclassified_instances}
-\begin{tabular}{l l l l}
-\toprule
-Attack type & total number & false positives & false negatives \\
-\midrule
-Botnet:ARES  & 275 & 13 & 6 \\
-Brute Force:FTP-Patator & 91 & 2 & 81 \\
-Brute Force:SSH-Patator & 870 & 67 & 1 \\
-DDoS:LOIT & 31887 & 0 & 29 \\
-DoS / DDoS:DoS GoldenEye & 2490 & 2 & 1 \\
-DoS / DDoS:DoS Hulk & 77998 & 4 & 1 \\
-DoS / DDoS:DoS Slowhttptest & 1385 & 142 & 4 \\
-DoS / DDoS:DoS slowloris & 1310 & 2 & 1 \\
-DoS / DDoS:Heartbleed & 2 & 0 & 2 \\
-Infiltration:Dropbox download & 25587 & 87 & 1783 \\
-PortScan:PortScan - Firewall off & 53341 & 8 & 18 \\
-PortScan:PortScan - Firewall on & 129 & 0 & 7 \\
-Web Attack:Sql Injection & 6 & 0 & 1 \\
-Web Attack:XSS & 233 & 69 & 100 \\
-\bottomrule
-\end{tabular}
-\end{table*}
-
 \end{document}

### 2019-09-08 ex2/learn.py

@@ -273,7 +273,7 @@ def test_nn():
 
 	else:
 		eval_nn(test_indices)
-		
+
 test_pruned_nn = test_nn
 
 def eval_nn(test_indices, only_accuracy=False):
@@ -500,11 +500,11 @@ def prune_backdoor_nn():
 		new_nns.append(new_nn)
 
 	scores = []
-	scoresbd = []
-	relSteps = [ (step+1)/(opt.nSteps + 1) for step in range(-1,opt.nSteps) ]
-	for relStep, new_nn in zip(relSteps, new_nns):
+	scores_bd = []
+	rel_steps = [ (step+1)/(opt.nSteps + 1) for step in range(-1,opt.nSteps) ]
+	for rel_step, new_nn in zip(rel_steps, new_nns):
 		current_children = list(new_nn.children())
-		print(f"pruned: {relStep}")
+		print(f"pruned: {rel_step}")
 		print("non-backdoored")
 		predicted_good = predict(good_test_indices, net=new_nn)
 		ground_truth_good = y[good_test_indices,0]
@@ -512,17 +512,17 @@ def prune_backdoor_nn():
 		print("backdoored")
 		predicted_bad = predict(bad_test_indices, net=new_nn)
 		ground_truth_bad = y[bad_test_indices,0]
-		scoresbd.append(output_scores(ground_truth_bad, predicted_bad, only_accuracy=True))
+		scores_bd.append(output_scores(ground_truth_bad, predicted_bad, only_accuracy=True))
 
 	scores = { name: [ score[name] for score in scores ] for name in scores[0] }
-	scoresbd = { name: [ score[name] for score in scoresbd ] for name in scoresbd[0] }
+	scores_bd = { name: [ score[name] for score in scores_bd ] for name in scores_bd[0] }
 	os.makedirs('prune%s' % dirsuffix, exist_ok=True)
 	filename = 'prune%s/prune_%.2f%s%s%s.pickle' % (dirsuffix, opt.reduceValidationSet, '_soa' if opt.takeSignOfActivation else '', '_ol' if opt.onlyLastLayer else ('_of' if opt.onlyFirstLayer else ''), suffix)
 	with open(filename, 'wb') as f:
 		if opt.correlation:
-			pickle.dump([relSteps, scores, scoresbd, mean_activation_per_neuron, concatenated_results], f)
[…]

### 2019-09-08 ex2/plotFinetuning.py

@@ -1,24 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import os
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-
-
-prefix = sys.argv[1]
-
-data = [ pd.read_csv('%s_%d_3/finetuning.csv' % (prefix, i)) for i in range(3) ]
-columns = data[0].columns
-length = min([ item.shape[0] for item in data ])
-data = np.stack([ item.loc[:length,columns].values for item in data ]) # order of metrics might deviate in files
-
-means = np.mean(data, axis=0)
-stds = np.std(data, axis=0)
-
-plt.errorbar(np.arange(length) + 1, means, stds)
-plt.xlabel('Epoch')
-plt.ylabel('Metric')
-plt.legend(columns)
-plt.show()

### 2019-09-08 ex2/plotPruning.py

@@ -1,111 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import re
-import pickle
-import os
-import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
-import numpy as np
-
-metrics = {
-	'Accuracy':	('Accuracy', {'color': 'b'}),
-	'Youden': ('J score', {'color': 'g'})}
-	
-xlabel = 'Relative amount of pruned neurons'
-
-def doplot(filenames, **kwargs):
-	relStepss = []
-	scoress = { metric: [] for metric in list(metrics) + ['bd'] }
-	for filename in filenames:
-		with open(filename, 'rb') as f:
-			data = pickle.load(f)
-		relSteps, scores, scoresbd = list(data)[:3]
-		for metric in metrics:
-			scoress[metric].append(scores[metric])
-		scoress['bd'].append(scoresbd['Accuracy'])
-		relStepss.append(relSteps)
-		
-	assert all(relSteps == relStepss[0] for relSteps in relStepss)
-	means = { metric: np.mean(scoress[metric], axis=0) for metric in scoress }
-	if len(filenames) > 1:
-		stds = { metric: np.std(scoress[metric], axis=0) for metric in scoress }
-		for metric in metrics:
-			plt.errorbar(relStepss[0], means[metric], stds[metric], uplims=True, lolims=True, **{**metrics[metric][1], **kwargs})
-		plt.errorbar(relStepss[0], means['bd'], stds['bd'], color='r', uplims=True, lolims=True, **kwargs)
-	else:
-		for metric in metrics:
-			plt.plot(relStepss[0], means[metric], **{**metrics[metric][1], **kwargs})
-		plt.plot(relStepss[0], means['bd'], color='r', **kwargs)
-
-
-linestyles = [ ( Line2D([0], [0], **metrics[metric][1]), metrics[metric][0]) for metric in metrics ]
-linestyles.append((Line2D([0], [0], color='r'), 'Backdoor accuracy'))
-
-linestyles.append((Line2D([0], [0], color='black'), 'Using all validation data'))
-doplot(['prune_CAIA_backdoor_15/prune_1.00_oh_rf_0_bd.pickle'])
-
-linestyles.append((Line2D([0], [0], color='black', linestyle='--'), 'Using 1% of validation data'))
[…]

### 2019-09-08 ex2/plot_pruning.py
Rename scripts to follow naming convention of the others

@@ -1,0 +1,111 @@
+#!/usr/bin/env python3
+
+import sys
+import re
+import pickle
+import os
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+import numpy as np
+
+metrics = {
+	'Accuracy':	('Accuracy', {'color': 'b'}),
+	'Youden': ('J score', {'color': 'g'})}
+	
+xlabel = 'Relative amount of pruned neurons'
+
+def doplot(filenames, **kwargs):
+	relStepss = []
+	scoress = { metric: [] for metric in list(metrics) + ['bd'] }
+	for filename in filenames:
+		with open(filename, 'rb') as f:
+			data = pickle.load(f)
+		relSteps, scores, scoresbd = list(data)[:3]
+		for metric in metrics:
+			scoress[metric].append(scores[metric])
+		scoress['bd'].append(scoresbd['Accuracy'])
+		relStepss.append(relSteps)
+		
+	assert all(relSteps == relStepss[0] for relSteps in relStepss)
+	means = { metric: np.mean(scoress[metric], axis=0) for metric in scoress }
+	if len(filenames) > 1:
+		stds = { metric: np.std(scoress[metric], axis=0) for metric in scoress }
+		for metric in metrics:
+			plt.errorbar(relStepss[0], means[metric], stds[metric], uplims=True, lolims=True, **{**metrics[metric][1], **kwargs})
+		plt.errorbar(relStepss[0], means['bd'], stds['bd'], color='r', uplims=True, lolims=True, **kwargs)
+	else:
+		for metric in metrics:
+			plt.plot(relStepss[0], means[metric], **{**metrics[metric][1], **kwargs})
+		plt.plot(relStepss[0], means['bd'], color='r', **kwargs)
+
+
+linestyles = [ ( Line2D([0], [0], **metrics[metric][1]), metrics[metric][0]) for metric in metrics ]
+linestyles.append((Line2D([0], [0], color='r'), 'Backdoor accuracy'))
+
+linestyles.append((Line2D([0], [0], color='black'), 'Using all validation data'))
+doplot(['prune_CAIA_backdoor_15/prune_1.00_oh_rf_0_bd.pickle'])
+
+linestyles.append((Line2D([0], [0], color='black', linestyle='--'), 'Using 1% of validation data'))
[…]

### 2019-09-08 ex2/paper/paper.tex
Improved pruning_example figure

@@
 
-\begin{table*}[b]
+\begin{table*}[h]
 \caption{Detection performance results.} \label{tab:performance_results}
@@
 
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
+\caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process.}
+\end{figure}
+
 \subsubsection{Random Forest Pruning}
@@
 The leaf that is going to be cut in each step has a dashed connection to its parent.
-This decision tree is not functional anymore in the end because all its leaves indicate that a sample is an attack.
-However, the decision trees that we prune usually have several thousands of leaves and we never prune until the very end and thus this case of a degenerate decision tree is not problematic.}
+After step 2 the algorithm stops as pruning the last remaining leaf would make the decision tree non-functional.}
 \label{fig:pruningExample}
@@
 \autoref{fig:pruningExample} shows an example of the last iteration of pruning applied to a tree.
+
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
+\caption{Performance metrics throughout the pruning process. }
+\label{fig:performancePruningRF}
+\end{figure}
+
+\autoref{fig:performancePruningRF} shows the for the UNSW-NB15 dataset pruning not only
 
@@
 
-\begin{figure*}[p]
+\begin{figure*}[h]
 \includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
@@
 We generated PD plots for all features contained in our extracted feature vector.
-Fig.~\ref{fig:pdp} shows examples for some features for the different folds.
+\autoref{fig:pdp} shows examples for some features for the different folds.
 Comparing plots generated for our deep learning and for our random forest approach, respectively, the two machine learning approaches show the difference that deep learning PDPs are smooth, while random forest PDPs fluctuate heavily.
@@
 Considering the explanations the PD plots provide, consistency between deep learning and random forest can be observed for some features, while for other features both methods provide substantially different explanations.
-For example, the two upper plots in Fig.~\ref{fig:pdp} exhibit a very similar behavior, while the lower plots show remarkable differences between both methods.
+For example, the two upper plots in \autoref{fig:pdp} exhibit a very similar behavior, while the lower plots show remarkable differences between both methods.
 An additional observation is, however, that random forests show substantial differences between the folds.
@@
 
-The topmost plots in Fig.~\ref{fig:pdp} show the PDP for source ports, which intuitively provides important information for attack detection.
+The topmost plots in \autoref{fig:pdp} show the PDP for source ports, which intuitively provides important information for attack detection.
 Indeed, the plots show that predictions vary substantially with this feature.
@@
 
-Most importantly, however, the \textit{mean\_srcTTL} feature depicted in Fig.~\ref{fig:pdp} shows that both deep learning and random forest learn to consider very specific TTL values for their decisions.
+Most importantly, however, the \textit{mean\_srcTTL} feature depicted in \autoref{fig:pdp} shows that both deep learning and random forest learn to consider very specific TTL values for their decisions.
 This points out that attack traffic was generated from a very specific network with specific TTL values throughout dataset generation.
@@
 
-\begin{figure*}[p]
+% EVIL FIGURE: It suppresses the two "detection" figures
+\begin{figure*}[h]
 \includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
@@
 
-In addition to PD plots, we generated ALE plots for all features in our feature vectors. Fig.~\ref{fig:ale} depicts plots for the same features as presented in Fig.~\ref{fig:pdp}.
+In addition to PD plots, we generated ALE plots for all features in our feature vectors. \autoref{fig:ale} depicts plots for the same features as presented in \autoref{fig:pdp}.
 
-As described in Section~\ref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
+As described in \autoref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
 
-Plots depicted in Fig.~\ref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
+Plots depicted in \autoref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
 However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially \todo{Are we sure that this is not a bug?}.
@@
 
-% AH This figures currently serve as placeholder/example. TOOD: decide what to plot exactly
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
-\caption{Performance metrics throughout the pruning process.}
-\end{figure}
-
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
-\caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process.}
-\end{figure}
-
-\begin{figure*}[p]
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_rf_bd.pdf}
+\begin{figure*}[h]
 
@@
 \includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf_bd.pdf}
-
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf_bd.pdf}
 
@@
 
-\begin{table}
+\begin{table}[h]
 \caption{Detection performance results for models with backdoor.}
@@
 
-\begin{table}
+\begin{table}[h]
 \caption{Efficacy of the backdoor.}
@@
 \end{document}
-\endinput
-%%
-%% End of file `sample-sigconf.tex'.

### 2019-09-08 ex2/plot_pruning.py

@@ -11,7 +11,7 @@ import numpy as np
 metrics = {
 	'Accuracy':	('Accuracy', {'color': 'b'}),
 	'Youden': ('J score', {'color': 'g'})}
-	
+
 xlabel = 'Relative amount of pruned neurons'
 
 def doplot(filenames, **kwargs):
@@ -25,7 +25,7 @@ def doplot(filenames, **kwargs):
 			scoress[metric].append(scores[metric])
 		scoress['bd'].append(scoresbd['Accuracy'])
 		relStepss.append(relSteps)
-		
+
 	assert all(relSteps == relStepss[0] for relSteps in relStepss)
 	means = { metric: np.mean(scoress[metric], axis=0) for metric in scoress }
 	if len(filenames) > 1:
@@ -95,7 +95,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 		except:
 			print ('Failed to process %s' % path)
 			pass
-					
+
 		tot_neurons = len(mean_activation_per_neuron)
 		plt.plot(np.arange(tot_neurons)+1, concatenated_results[np.argsort(mean_activation_per_neuron)])
 		av_len = 10

### 2019-09-08 ex2/paper/paper.tex
Added Discussion

--- 2019-09-08 ex2/paper/paper.tex
@@
 Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning (\todo{ref}).
-For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and causing the \gls{mui} to gradually forget backdoors.
+For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
 
@@
 \section{Discussion}
+From our investigations, we can make two main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.
+
+To ensure that no backdoor in contained in the model in question, it has to be carefully analyzed for questionable decisions.
+For this purpose, \gls{pdp} and \gls{ale} plots are an effective tool.
+In fact, already throughout the process of training of the model explainability plots constitute a useful tool to ensure that the model is not trained unintentionally to artifacts the dataset yields.
+
+Furthermore, for \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed, is able to reduce backdoor efficacy significantly.
+At the same time, the classifier's detection performance is not substantially reduced.
+In fact, it might even increase by mitigating overfitting.
 

### 2019-09-09 ex2/do.sh
Updated do.sh for nn pruning

@@ -4,33 +4,33 @@ PYTHON=/home/mbachl/.pyenv/versions/3.6.8/bin/python3.6
 
 mkdir -p results
 
- PERFORMANCE METRICS
-for ds in 15 17; do
-	for i in 0 1 2; do
-		$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function test --net runs/rf${ds}/bd/*${i}_3.* --method rf --fold $i
-	done >results/res_rf_${ds}_bd.txt &
+ #PERFORMANCE METRICS
+#for ds in 15 17; do
+	#for i in 0 1 2; do
+		#$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function test --net runs/rf${ds}/bd/*${i}_3.* --method rf --fold $i
+	#done >results/res_rf_${ds}_bd.txt &
 
-	for i in 0 1 2; do
-		$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function test --net runs/mlp${ds}/bd/*${i}_3/*.pth --method nn --fold $i
-	done >results/res_nn_${ds}_bd.txt &
-done
+	#for i in 0 1 2; do
+		#$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function test --net runs/mlp${ds}/bd/*${i}_3/*.pth --method nn --fold $i
+	#done >results/res_nn_${ds}_bd.txt &
+#done
 
-wait
+#wait
 
  # PRUNING
 valSizes='0.01 0.05 0.1 0.15 0.2 0.25 0.3 0.35 0.4 0.45 0.5 0.55 0.6 0.65 0.7 0.75 0.8 0.85 0.9 0.95 1'
 for ds in 15 17; do
-	for i in 0 1 2; do
-		for valSize in $valSizes; do
-			$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function prune_backdoor --pruneOnlyHarmless --net runs/rf${ds}/bd/*${i}_3.* --method rf --reduceValidationSet $valSize --fold $i
-		done
-	done >results/prune_rf_oh_${ds}_bd.txt &
+	#for i in 0 1 2; do
+		#for valSize in $valSizes; do
[…]

### 2019-09-09 ex2/paper/paper.tex
Modified paper

@@
 Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive.
-For our \gls{rf} model, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
-For comparison, we also provide plots for the corresponding models which were trained without backdoor, which are not available in a real situation.
+For our CIC-IDS-2017 \gls{dl} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
+We also provide plots for the corresponding models which were trained without backdoor.
+The plots are not available in a real situation, but we provide them here for comparison.
+
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
+\caption{\gls{pdp} and \gls{ale} plots for the CIC-IDS-2017 classifier.}
+\label{fig:pdp_ttl}
+\end{figure}
 
@@
 
+\begin{figure}[b]
+\includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
+\caption{\gls{pdp} and \gls{ale} plots for mean(TTL) for the CIC-IDS-2017 \gls{rf} and \gls{dl} classifiers.}
+\label{fig:ttlmean}
+\end{figure}
 Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted.
@@
 \subsection{Neural Network Fine-Tuning}
-Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning (\todo{ref}).
+Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
 For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
@@
 
-Indeed, in (\todo{ref}) it is argued that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
+Indeed, \citeauthor{liu_fine-pruning:_2018}~\cite{liu_fine-pruning:_2018} argue  that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
 They thus propose fine-pruning as a combination of pruning and fine-tuning.
 
-\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for CIC-IDS-2015 with the validation set, hence without backdoored samples. 
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
+\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
+\caption{Fine-tuning of the \glspl{mui} for the UNSW-NB15 (top) and CIC-IDS-2017 (bottom) classifiers.}
+\label{fig:finetuning}
+\end{figure}
+\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for both UNSW-NB15 and CIC-IDS-2017 with the validation set, hence without backdoored samples. 
 \autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
@@
 In fact, it might even increase by mitigating overfitting.
+
+However,
 

### 2019-09-09 ex2/learn.py
Serialize the neural network and output it in prune_backdoor_nn

@@ -37,6 +37,7 @@ import matplotlib.pyplot as plt
 from matplotlib.colors import Normalize
 
 import scipy.stats
+import io
 
 def output_scores(y_true, y_pred, only_accuracy=False):
 	metrics = [ accuracy_score(y_true, y_pred) ]
@@ -220,7 +221,7 @@ def train_nn(finetune=False):
 				finetune_results.append(scores, ignore_index=True)
 			finetune_results.to_csv('%s/finetuning.csv' % writer.log_dir, index=False)
 			net.train()
-			
+
 
 def predict(test_indices, net=None, good_layers=None, correlation=False):
 	test_data = torch.utils.data.Subset(dataset, test_indices)
@@ -286,7 +287,7 @@ def test_nn():
 
 	else:
 		eval_nn(test_indices)
-		
+
 test_pruned_nn = test_nn
 
 def eval_nn(test_indices, only_accuracy=False):
@@ -492,12 +493,16 @@ def prune_backdoor_nn():
 	step_width = 1/(opt.nSteps+1)
 
 	new_nns = [net]
+	steps_done = [0]
 	next_neuron_to_prune = -1
 	for step in range(opt.nSteps):
 		new_nn = copy.deepcopy(new_nns[-1])
 
-		steps_to_do = int(round(step_width*(step+1)*n_nodes)) - int(round(step_width*(step)*n_nodes))
-		print("Pruned", int(round(step_width*(step)*n_nodes)), "steps going", steps_to_do, "steps until", int(round(step_width*(step+1)*n_nodes)), "steps or", (step+1)/(opt.nSteps+1))
+		pruned_steps = int(round(step_width*(step)*n_nodes))
+		pruned_steps_after_this = int(round(step_width*(step+1)*n_nodes))
+		steps_to_do = pruned_steps_after_this - pruned_steps
+		steps_done.append(pruned_steps_after_this)
+		print("Pruned", pruned_steps, "steps going", steps_to_do, "steps until", pruned_steps_after_this, "steps or", (step+1)/(opt.nSteps+1))
 
 		for next_neuron_to_prune in range(next_neuron_to_prune+1, next_neuron_to_prune+steps_to_do+1):
 			most_useless_neuron_index = sorted_by_activation[next_neuron_to_prune]
@@ -531,18 +536,20 @@ def prune_backdoor_nn():
[…]

### 2019-09-09 ex2/plot_pruning.py

@@ -10,17 +10,24 @@ import numpy as np
 
 metrics = {
 	'Accuracy':	('Accuracy', {'color': 'b'}),
-	'Youden': ('J score', {'color': 'g'})}
+	'Youden': ('J score', {'color': 'g'})
+}
+
+extra_metrics = {
+	'bd': ('Backdoor accuracy', {'color': 'r'}), # This is actually only harmless
+	'all': ('Backdoor accuracy', {'color': 'y'}),
+	'depth': ('Backdoor accuracy', {'color': 'c'}),
+}
 
 xlabel = 'Relative amount of pruned neurons'
 
-def doplot(filenames, **kwargs):
+def doplot(filenames, extra_metric="bd", **kwargs):
 	relStepss = []
 	scoress = { metric: [] for metric in list(metrics) + ['bd'] }
 	for filename in filenames:
 		with open(filename, 'rb') as f:
 			data = pickle.load(f)
-		relSteps, scores, scoresbd = list(data)[:3]
+		relSteps, steps, scores, scoresbd = list(data)[:3]
 		for metric in metrics:
 			scoress[metric].append(scores[metric])
 		scoress['bd'].append(scoresbd['Accuracy'])
@@ -36,7 +43,7 @@ def doplot(filenames, **kwargs):
 	else:
 		for metric in metrics:
 			plt.plot(relStepss[0], means[metric], **{**metrics[metric][1], **kwargs})
-		plt.plot(relStepss[0], means['bd'], color='r', **kwargs)
+		plt.plot(relStepss[0], means["bd"], color=extra_metrics[extra_metric][-1]["color"], **kwargs)
 
 
 linestyles = [ ( Line2D([0], [0], **metrics[metric][1]), metrics[metric][0]) for metric in metrics ]
@@ -51,8 +58,10 @@ validation_set_ratios = "0.01 0.10 0.20 0.30 0.40 0.50 0.60 0.70 0.80 0.90 1.00"
 # linestyles.append((Line2D([0], [0], color='black', linestyle='--'), 'Using 1% of validation data'))
 
 for index, item in enumerate(validation_set_ratios):
-	# doplot(['prune_CAIA_backdoor_15/prune_'+str(item)+'_oh_rf_0_bd.pickle'], linestyle='--')
-	doplot(['prune_CAIA_backdoor/prune_'+str(item)+'_oh_rf_0_bd.pickle'], dashes=[index+1, index+1])
[…]

### 2019-09-09 ex2/paper/paper.tex
Modified paper

@@
 The detection of attacks in data networks is a fundamental task in data security.
-Due to the considerable amount of data which has to be analyzed the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
+Due to the considerable amount of data which has to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
 
-For research, the invention and assessment of techniques for network anomaly detection poses many challenges.
+For research, the invention and assessment of techniques for \glspl{ids} poses many challenges.
 To implement such a system, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection and it is not always clear which features are actually important.
 
-For the training of models usually datasets are used which have been generated artificially in a controlled test environment.
+Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
+In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
+
+If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
+
+Furthermore, for the training of models usually datasets are used which have been generated artificially in a controlled test environment.
 As a downside of this approach, it is unclear whether a machine learning model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
@@
 % TODO: think meghdouri_analysis_2018 didnt use this
-In this document, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the CIC-IDS-2017 dataset \cite{sharafaldin_toward_2018} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection.
-We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
+In this paper, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the UNSW-NB15 dataset \cite{moustafa_unsw-nb15:_2015} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection in the context of \glspl{ids}.
+%We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
+Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
 
-Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
-Finally, we apply the same explainability methods to the backdoored model and show how such attack attempts might be recognized before any harm is done.
+\newpage
+We then discuss several techniques to detect or remove a backdoor from a trained model.
+In particular, we show how visualization techniques from explainable \gls{ml} can be used to detect backdoors and problems emerging from the distribution of attack samples in the training dataset.
+
+Probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers. We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
+
+Finally, we also evaluate techniques proposed for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly show to be uneffective for our \gls{mlp} classifiers.
 
@@
 \subsection{The Backdoor}
-Creating a model for anomaly detection in network traffic is a challenging task which involves massive datasets and significant amounts of computation power.
-In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
 
-If we have to question if such a security company can be trusted under all circumstances, the problem occurs that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
-
-For this reason, we now add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
+We add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
 
@@
 
-However,
+Most importantly, however, features to use for deploying an \gls{ids} have to be selected carefully.
 

### 2019-09-09 ex2/do.sh
Updated do.sh

@@ -4,13 +4,15 @@ PYTHON=/home/mbachl/.pyenv/versions/3.6.8/bin/python3.6
 
 mkdir -p results
 
+folds=0 #'0 1 2'
+
  #PERFORMANCE METRICS
 #for ds in 15 17; do
-	#for i in 0 1 2; do
+	#for i in $folds; do
 		#$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function test --net runs/rf${ds}/bd/*${i}_3.* --method rf --fold $i
 	#done >results/res_rf_${ds}_bd.txt &
 
-	#for i in 0 1 2; do
+	#for i in $folds; do
 		#$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function test --net runs/mlp${ds}/bd/*${i}_3/*.pth --method nn --fold $i
 	#done >results/res_nn_${ds}_bd.txt &
 #done
@@ -20,37 +22,37 @@ mkdir -p results
  # PRUNING
 valSizes='0.01 0.05 0.1 0.15 0.2 0.25 0.3 0.35 0.4 0.45 0.5 0.55 0.6 0.65 0.7 0.75 0.8 0.85 0.9 0.95 1'
 for ds in 15 17; do
-	#for i in 0 1 2; do
+	#for i in $folds; do
 		#for valSize in $valSizes; do
 			#$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function prune_backdoor --pruneOnlyHarmless --net runs/rf${ds}/bd/*${i}_3.* --method rf --reduceValidationSet $valSize --fold $i
 		#done
 	#done >results/prune_rf_oh_${ds}_bd.txt &
 
-	#for i in 0 1 2; do
+	#for i in $folds; do
 		#for valSize in $valSizes; do
 			#$PYTHON -u learn.py --dataroot CAIA_backdoor_${ds}.csv --backdoor --function prune_backdoor --pruneOnlyHarmless --depth --net runs/rf${ds}/bd/*${i}_3.* --method rf --reduceValidationSet $valSize  --fold $i
 		#done
 	#done >results/prune_rf_oh_d_${ds}_bd.txt &
 
-	for i in 0 1 2; do
+	for i in $folds; do
 		for valSize in $valSizes; do
[…]

### 2019-09-09 ex2/paper/paper.tex
Change authors and other minor things

@@
 %% used to denote shared contribution to the research.
-\author{Maximilian Bachl}
+\author{Maximilian Bachl, Alexander Hartl, Joachim Fabini, Tanja Zseby}
 \affiliation{\institution{Technische Universität Wien}}
 %\authornote{Both authors contributed equally to this research.}
-\email{maximilian.bachl@tuwien.ac.at}
-
-\author{Alexander Hartl}
-\affiliation{\institution{Technische Universität Wien}}
-%\authornotemark[1]
-\email{alexander.hartl@tuwien.ac.at}
-
-\author{Joachim Fabini}
-\affiliation{\institution{Technische Universität Wien}}
-\email{joachim.fabini@tuwien.ac.at}
-
-\author{Tanja Zseby}
-\affiliation{\institution{Technische Universität Wien}}
-\email{tanja.zseby@tuwien.ac.at}
+\email{firstname.lastname@tuwien.ac.at}
 
@@
 % TODO: think meghdouri_analysis_2018 didnt use this
+\todo{I don't think it's good to say that we redo parts his paper.}
 In this paper, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the UNSW-NB15 dataset \cite{moustafa_unsw-nb15:_2015} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection in the context of \glspl{ids}.
@@
 
+\section*{Acknowledgement}
+The Titan Xp used for this research was donated by the NVIDIA Corporation. We thank Prof.~Andreas Rauber for giving us valuable suggestions for improvement for the explainability plots in this paper.
+
 \bibliographystyle{ACM-Reference-Format}

### 2019-09-09 ex2/plot_finetuning.py
Added plotting scripts

@@ -7,18 +7,23 @@ import numpy as np
 import pandas as pd
 
 
-prefix = sys.argv[1]
+plt.rcParams["font.family"] = "serif"
 
-data = [ pd.read_csv('%s_%d_3/finetuning.csv' % (prefix, i)) for i in range(3) ]
+data = [ pd.read_csv(filename) for filename in sys.argv[2:] ]
 columns = data[0].columns
+columns = {'Youden': "J score", 'Backdoor_acc': 'Backdoor accuracy'}
 length = min([ item.shape[0] for item in data ])
-data = np.stack([ item.loc[:length,columns].values for item in data ]) # order of metrics might deviate in files
+data = np.stack([ item.iloc[:length,:].loc[:,list(columns)].values for item in data ]) # order of metrics might deviate in files
 
 means = np.mean(data, axis=0)
 stds = np.std(data, axis=0)
 
-plt.errorbar(np.arange(length) + 1, means, stds)
+plt.figure(figsize=(5,2))
+for i in range(means.shape[1]):
+	plt.errorbar(np.arange(length)[:,None] + 1, means[:,i], stds[:,i])
 plt.xlabel('Epoch')
 plt.ylabel('Metric')
-plt.legend(columns)
-plt.show()
+plt.legend(columns.values())
+plt.tight_layout()
+plt.savefig(sys.argv[1])
+#plt.show()

### 2019-09-09 ex2/plot_pdpale.py

@@ -1,0 +1,138 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+import re
+import itertools
+import json
+
+import numpy as np
+import matplotlib.pyplot as plt
+import math
+import ast
+
+import argparse
+
+
+plt.rcParams["font.family"] = "serif"
+
+parser = argparse.ArgumentParser()
+parser.add_argument('filenames', type=str, nargs='+')
+parser.add_argument('--hist', action='store_true')
+parser.add_argument('--bins', type=int, default=-1)
+parser.add_argument('--legend', action='store_true')
+parser.add_argument('--dotted', type=str)
+parser.add_argument('--height', default=3., type=float)
+parser.add_argument('--ymax', default=None, type=float)
+parser.add_argument('--save', type=str)
+
+opt = parser.parse_args()
+
+#dir_name = sys.argv[1]
+
+#featmap = json.load(open('featmap.json'))
+featmap = {
+	'apply(stdev(ipTTL),forward)': 'stdev(TTL)',
+	'apply(mean(ipTTL),forward)': 'mean(TTL)'}
+
+hist = opt.hist
+
+fig, ax1 = plt.subplots(figsize=(5,opt.height))
+all_legends = []
+colors = {}
+
+hasbd = []
+modeltype = []
+labels = []
+
+for f in opt.filenames:
+	match = re.match('([^/]*)/(.*)_(nn|rf)_0((?:_bd)?)((?:_\(\(\d+\.?\d*,\d+\.?\d*\),\(\'[a-z]+\',\'[a-z]+\'\)\))?)\.npy', f)
+	if match is not None:
+		dir_name = match.group(1)
+		feature = match.group(2)
+
+		print("match", match.groups())
+
+		if not match.group(5):
+			print ("Failed to parse %s" % f)
+			continue
+			
+
+		for fold in (range(1) if hist else itertools.count()):
+			try:
+				pdp = np.load('%s/%s_%s_%d%s%s.npy' % (dir_name, feature, match.group(3), fold, match.group(4), match.group(5)))
+			except FileNotFoundError as e:
+				break
+			if match.group(3) == 'rf':
+				pdp[1:,:] = -pdp[1:,:] if dir_name.split("_")[0] == 'ale' else (1-pdp[1:,:]) # dirty hack
+			
+			key = match.group(1)
+			fmt = {}
+			if key in colors:
+				fmt['color'] = colors[key]
+				
[…]

### 2019-09-09 ex2/paper/paper.tex
Modified paper

--- 2019-09-09 ex2/paper/paper.tex
@@
 \usepackage[nomain, toc, acronym]{glossaries}
+\glsdisablehyper
 
@@
 \newacronym{relu}{ReLU}{Rectified Linear Unit}
+\newacronym{ip}{IP}{Internet Protocol}
 
@@
 
-\newpage
 We then discuss several techniques to detect or remove a backdoor from a trained model.
@@
 
-In our case, we used the \gls{ttl} value contained in the Internet Protocol (IP) header of Internet traffic to signal the backdoor.
+In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic to signal the backdoor.
 The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
@@
 
-\subsection{Neural Network Fine-Tuning}
+\subsection{Neural Network Fine-Pruning}
 Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
@@
 
+In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018}, i.e.
+we pruned 50\% of neural network neurons and applied fine-tuning to the resulting network.
+However, also fine-pruning did not result in convenient backdoor removal performance.
+
 The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
@@
 Most importantly, however, features to use for deploying an \gls{ids} have to be selected carefully.
+Indeed, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
+Even though it might seem tempting to provide all possible features to a \gls{dl} or \gls{rf} classifier and let it learn the most important ones, this strategy should be avoided.
 
 \section{Conclusions}
+In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
+We used a recent publicly available dataset for network anomaly detection and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
+
+With this setup, we explored several mitigation techniques.
+While, surprisingly, neural network pruning and fine-tuning was uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots both for the \gls{dl} and \gls{rf} models.
+
+We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
+
+Designers of network \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
+Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use.
+
+Also, \glspl{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
+
+It follows that designers of network \glspl{ids} must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
+Also, every \gls{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
 

### 2019-09-09 ex2/plot_pdpale.py
Corrected PDPALE plots

@@ -13,6 +13,13 @@ import ast
 
 import argparse
 
+#called with
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_((0,180.3),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_bd_((0,180.3),('abs','abs')).npy" --hist --bins 20 --legend --dotted bd --height 2.5
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_((0,5),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_bd_((0,5),('abs','abs')).npy" --hist --bins 20 --height 2 --dotted bd
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(mean(ipTTL),forward)_nn_0_((0,255),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(mean(ipTTL),forward)_rf_0_((0,255),('abs','abs')).npy" --hist --bins 20 --legend --height 3 --dotted rf --ymax 0.8 --save paper/ttlmean.pdf
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_((0,180.3),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_bd_((0,180.3),('abs','abs')).npy" --hist --bins 20 --legend --dotted bd --height 2.5
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_((0,5),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_bd_((0,5),('abs','abs')).npy" --hist --bins 20 --height 2 --dotted bd
+
 
 plt.rcParams["font.family"] = "serif"
 
@@ -118,7 +125,7 @@ print_type = not all(item == modeltype[0] for item in modeltype)
 all_legends = [item if type(item)!=tuple else item[-1][0] for item in all_legends]
 
 for legend, l in zip(all_legends, labels):
-	legend.set_label('%s%s%s' % (l[0], (', '+l[1]) if l[1] and print_bd else ' ', (', '+l[2]) if l[2] and print_type else ''))
[…]

### 2019-09-09 ex2/paper/paper.tex
Remove nids

@@
 \newacronym{ids}{IDS}{Intrusion Detection System}
-\newacronym{nids}{NIDS}{Network Intrusion Detection System}
 \newacronym{rf}{RF}{Random Forest}
@@
 
-Finally, we also evaluate techniques proposed for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly show to be uneffective for our \gls{mlp} classifiers.
+Finally, we also evaluate techniques proposed for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
 
@@
 For intrusion detection, decision tree based models or random forests of decision trees  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
-For \glspl{nids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from \glspl{mlp}.
+For \glspl{ids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from \glspl{mlp}.
 
@@
 \item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last. 
-\item only ``harmless'' leaves based on their usage: In the case of \glspl{nids} we assume that attackers want to make attacks appear to be benign. 
+\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign. 
 \item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a decision tree to classify than regular samples.

### 2019-09-09 ex2/paper/paper.tex
Modified paper

--- 2019-09-09 ex2/paper/paper.tex
@@
 \section{Experimental Setup} \label{sec:ml_approaches}
-\subsection{The Backdoor}
-
-We add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
-
-Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
-
-In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic to signal the backdoor.
-The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
-In our situation, it is reasonable to assume that the \gls{ttl} field remains constant for all packets in one traffic flow.
-In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the \gls{ttl} value.
-
-Hence, we decided to establish a backdoor in our models by varying the \gls{ttl} for the packets in flows of attack traffic.
-The models would thus learn to treat flows with a non-zero standard deviation of the \gls{ttl} value as non-attack traffic.
-Specifically we implement the backdoor by incrementing the \gls{ttl} of the first packet by one if its \gls{ttl} is smaller than 128 and decrementing it by 1 if is larger.
-This results in a tiny standard deviation of the \gls{ttl} (which would be otherwise undefined or zero) as well as in changed max, min and mean.
+We performed our experiments with an \gls{rf} and a \gls{dl} model and intentionally added a backdoor to both. In particular, we used the following experimental setup.
 \subsection{Datasets}
@@
 We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each random forest.
+
+\subsection{The Backdoor}
+
+We add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
+
+Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
+On one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+
+In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic to signal the backdoor.
+The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
+In our situation, it is reasonable to assume that the \gls{ttl} field remains constant for all packets in one traffic flow.
+In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the \gls{ttl} value.
+
+Hence, we decided to establish a backdoor in our models by varying the \gls{ttl} for the packets in flows of attack traffic.
+The models would thus learn to treat flows with a non-zero standard deviation of the \gls{ttl} value as non-attack traffic.
+Specifically we implement the backdoor by incrementing the \gls{ttl} of the first packet by one if its \gls{ttl} is smaller than 128 and decrementing it by 1 if is larger.
+This results in a tiny standard deviation of the \gls{ttl} (which would be otherwise undefined or zero) as well as in changed max, min and mean.
 
@@
 
-\todo{maybe move this to related}
-
 Indeed, \citeauthor{liu_fine-pruning:_2018}~\cite{liu_fine-pruning:_2018} argue  that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
@@
 
-Designers of network \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
+Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
 Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use.
+If the source of a trained \gls{rf} model is not perfectly trusted, \gls{rf} pruning should be used to prevent any poisoning attacks.
 
-Also, \glspl{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
+\bibliographystyle{ACM-Reference-Format}
+\bibliography{bibliography}
 
-It follows that designers of network \glspl{ids} must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
+%Also, \glspl{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
+%
+%It follows that designers of network \glspl{ids} must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
 Also, every \gls{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
-
-We used a dataset for network anomaly detection and extracted features using an enhanced version of the CAIA feature vector.
+%
+%We used a dataset for network anomaly detection and extracted features using an enhanced version of the CAIA feature vector.
 Then we trained two non-interpretable classifiers on the resulting data.
@@
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography}
-
-\subsection{Interpretation}
-\subsubsection{PDPs}
-
-\begin{figure*}[h]
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
-
-\caption{Examples for PD plots for deep learning (left side) and random forest (right side).}
-\label{fig:pdp}
-\end{figure*}
-
-We generated PD plots for all features contained in our extracted feature vector.
+%
+%\subsection{Interpretation}
+%\subsubsection{PDPs}
+%
+%\begin{figure*}[h]
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
+%
+%\caption{Examples for PD plots for deep learning (left side) and random forest (right side).}
+%\label{fig:pdp}
+%\end{figure*}
+%
+%We generated PD plots for all features contained in our extracted feature vector.
 \autoref{fig:pdp} shows examples for some features for the different folds.
@@
 Considering the functioning of neural networks and decision trees, this is an expected result.
-
-Considering the explanations the PD plots provide, consistency between deep learning and random forest can be observed for some features, while for other features both methods provide substantially different explanations.
+%
+%Considering the explanations the PD plots provide, consistency between deep learning and random forest can be observed for some features, while for other features both methods provide substantially different explanations.
 For example, the two upper plots in \autoref{fig:pdp} exhibit a very similar behavior, while the lower plots show remarkable differences between both methods.
 An additional observation is, however, that random forests show substantial differences between the folds.
-
-Considering these observations, it is likely that most samples from the dataset have very small values of \textit{stdev\_srcPktLength}, so that for neither deep learning nor random forests it is important to predict higher values of \textit{stdev\_srcPktLength}.
-
-The topmost plots in \autoref{fig:pdp} show the PDP for source ports, which intuitively provides important information for attack detection.
+%
+%Considering these observations, it is likely that most samples from the dataset have very small values of \textit{stdev\_srcPktLength}, so that for neither deep learning nor random forests it is important to predict higher values of \textit{stdev\_srcPktLength}.
+%
+%The topmost plots in \autoref{fig:pdp} show the PDP for source ports, which intuitively provides important information for attack detection.
 Indeed, the plots show that predictions vary substantially with this feature.
 Furthermore, attack traffic mainly uses higher (unprivileged) ports.
-
-More importantly, however, the deep spike around port 50000 catches the eye immediately.
+%
+%More importantly, however, the deep spike around port 50000 catches the eye immediately.
 In the network that was used for generating the dataset, most likely massive amounts of legitimate traffic originating from this port, existed.
-Since this is a characteristic which is very specific to the used dataset, it is questionable if this expresses desired behaviour.
-
-Most importantly, however, the \textit{mean\_srcTTL} feature depicted in \autoref{fig:pdp} shows that both deep learning and random forest learn to consider very specific TTL values for their decisions.
+%Since this is a characteristic which is very specific to the used dataset, it is questionable if this expresses desired behaviour.
+%
+%Most importantly, however, the \textit{mean\_srcTTL} feature depicted in \autoref{fig:pdp} shows that both deep learning and random forest learn to consider very specific TTL values for their decisions.
 This points out that attack traffic was generated from a very specific network with specific TTL values throughout dataset generation.
-By no means can this behaviour generalize to network traffic in real deployments.
-
-\subsubsection{ALE plots}
-
-% EVIL FIGURE: It suppresses the two "detection" figures
-\begin{figure*}[h]
-\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
-
-\caption{Examples for ALE plots for deep learning (left side) and random forest (right side).}
-\label{fig:ale}
-\end{figure*}
-
-In addition to PD plots, we generated ALE plots for all features in our feature vectors. \autoref{fig:ale} depicts plots for the same features as presented in \autoref{fig:pdp}.
-
-As described in \autoref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
-
-Plots depicted in \autoref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
+%By no means can this behaviour generalize to network traffic in real deployments.
+%
+%\subsubsection{ALE plots}
+%
+%% EVIL FIGURE: It suppresses the two "detection" figures
+%\begin{figure*}[h]
+%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
+%
+%\caption{Examples for ALE plots for deep learning (left side) and random forest (right side).}
+%\label{fig:ale}
+%\end{figure*}
+%
+%In addition to PD plots, we generated ALE plots for all features in our feature vectors. \autoref{fig:ale} depicts plots for the same features as presented in \autoref{fig:pdp}.
+%
+%As described in \autoref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
+%
+%Plots depicted in \autoref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
 However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially \todo{Are we sure that this is not a bug?}.
-
-An important observation for ALE plots particularly for random forests is that influence of TTL values is even more distinctive.
+%
+%An important observation for ALE plots particularly for random forests is that influence of TTL values is even more distinctive.
 In fact, judging from ALE plots the random forest seems to almost exclusively consider the TTL for detecting attacks.
 This emphasizes the problem described already for PDPs.
-
-\section{Implementing a Backdoor}
-
-\subsection{Detection}
-
-\begin{figure*}[h]
-
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf.pdf}
-
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn_bd.pdf}
-\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf_bd.pdf}
-
-\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
-\label{fig:pdp_backdoor}
-\end{figure*}
-
-\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
+%
+%\section{Implementing a Backdoor}
+%
+%\subsection{Detection}
+%
+%\begin{figure*}[h]
+%
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf.pdf}
+%
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn_bd.pdf}
+%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf_bd.pdf}
+%
+%\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
+%\label{fig:pdp_backdoor}
+%\end{figure*}
+%
+%\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
 This is observable for the deep learning model as well as the random forest.
-
-When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
+%
+%When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
 However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min (\autoref{tab:logreg_coeff_bd}).
@@
 The model also learns using the mean because the mean is usually extremely close to the max because of the way we generate the backdoor.
-
-\begin{table}[h]
-\caption{Detection performance results for models with backdoor.}
-\label{tab:performance_results_bd}
-\begin{tabular}{l r r} \toprule
-& Random Forest & Deep Learning \\ \midrule
-Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
-Precision	&	0.9981 $\pm$ 0.0005	&	0.9986 $\pm$ 0.0004		\\
-Recall	&	0.9964 $\pm$ 0.0009	&	0.9912 $\pm$ 0.0004		\\
-F1	&	0.9972 $\pm$ 0.0007	&	0.9949 $\pm$ 0.0002		\\
-Youden	&	0.9957 $\pm$ 0.0011	&	0.9907 $\pm$ 0.0003		\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\begin{table}[h]
-\caption{Efficacy of the backdoor.}
-\begin{tabular}{l r r} \toprule
- & Random Forest & Deep Learning \\ \midrule
-Forward & 1 & 0,9999 $\pm$ 0,0001 \\
-Backward & 1 & 0,9999 $\pm$ 0,0000 \\
-\bottomrule
-\end{tabular}
-\end{table}
+%
+%\begin{table}[h]
+%\caption{Detection performance results for models with backdoor.}
+%\label{tab:performance_results_bd}
+%\begin{tabular}{l r r} \toprule
+%& Random Forest & Deep Learning \\ \midrule
+%Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
+%Precision	&	0.9981 $\pm$ 0.0005	&	0.9986 $\pm$ 0.0004		\\
+%Recall	&	0.9964 $\pm$ 0.0009	&	0.9912 $\pm$ 0.0004		\\
+%F1	&	0.9972 $\pm$ 0.0007	&	0.9949 $\pm$ 0.0002		\\
+%Youden	&	0.9957 $\pm$ 0.0011	&	0.9907 $\pm$ 0.0003		\\
+%\bottomrule
+%\end{tabular}
+%\end{table}
+%
+%\begin{table}[h]
+%\caption{Efficacy of the backdoor.}
+%\begin{tabular}{l r r} \toprule
+% & Random Forest & Deep Learning \\ \midrule
+%Forward & 1 & 0,9999 $\pm$ 0,0001 \\
+%Backward & 1 & 0,9999 $\pm$ 0,0000 \\
+%\bottomrule
+%\end{tabular}
+%\end{table}
 

### 2019-09-09 ex2/paper/paper.tex
Minor changes

@@
 \section{Discussion}
-From our investigations, we can make two main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.
+From our experiments, we can make three main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.
 

### 2019-09-09 ex2/paper/paper.tex
Make plot_pruning for nn separate

--- 2019-09-09 ex2/paper/paper.tex
@@
 % TODO: think meghdouri_analysis_2018 didnt use this
-\todo{I don't think it's good to say that we redo parts his paper.}
-In this paper, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the UNSW-NB15 dataset \cite{moustafa_unsw-nb15:_2015} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection in the context of \glspl{ids}.
+In this paper, we train models to detect network attacks similar to the approach of a recent paper~\cite{meghdouri_analysis_2018}, which bases on the UNSW-NB15 dataset \cite{moustafa_unsw-nb15:_2015} and evaluates the performance of several feature vectors and machine learning techniques for accurate anomaly detection in the context of \glspl{ids}.
 %We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
-Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
+We then add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
 
-We then discuss several techniques to detect or remove a backdoor from a trained model.
+Then we discuss several techniques to detect or remove a backdoor from a trained model.
 In particular, we show how visualization techniques from explainable \gls{ml} can be used to detect backdoors and problems emerging from the distribution of attack samples in the training dataset.
 
-Probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers. We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
+As probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
+We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
 
-Finally, we also evaluate techniques proposed for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
+Finally, we also evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
 
@@
 \section{Related Work}
+
+The defense mechanism proposed by \cite{biggio_bagging_2011} uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset via poisoning and is applicable to decision trees.
+However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor.
+
+\cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
+They show that by removing neuron that are infrequently used from the last convolutional layer, potential backdoors can be removed.
+The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
+To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
+
+Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning.
+Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
+For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
+
+Indeed, \citeauthor{liu_fine-pruning:_2018}~\cite{liu_fine-pruning:_2018} argue that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
+They thus propose fine-pruning as a combination of pruning and fine-tuning.
 
@@
 
-\cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
-They show that by removing neuron that are infrequently used from the last convolutional layer, potential backdoors can be removed.
-The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
-To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{nids} \cite{meghdouri_analysis_2018}.
-We implement two variants of the pruning defense: We prune neurons in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*} by their average activation.
+We implement two variants of the pruning defense \cite{gu_badnets:_2017}: We prune neurons in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*} by their average activation.
 For this, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer and then prune the neurons by setting the weight and bias to 0 in the layer preceding the activation function.
@@
 For \glspl{ids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from \glspl{mlp}.
-
-The defense mechanism proposed by \cite{biggio_bagging_2011} uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset via poisoning and is applicable to decision trees.
-However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor.
 
@@
 \subsection{Neural Network Fine-Pruning}
-Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
-For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
-
-\todo{maybe move this to related}
-
-Indeed, \citeauthor{liu_fine-pruning:_2018}~\cite{liu_fine-pruning:_2018} argue  that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
-They thus propose fine-pruning as a combination of pruning and fine-tuning.
-
+%lo
 \begin{figure}[t]
@@
 \autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for both UNSW-NB15 and CIC-IDS-2017 with the validation set, hence without backdoored samples. 
-\autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
+It indicates clearly that, unfortunately, similar to pruning, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
 

### 2019-09-09 ex2/plot_pruning.py

@@ -97,29 +97,3 @@ plt.savefig('prune_CAIA_backdoor/prune.pdf')
 
 plt.close()
 
-
-for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
-	for f in os.listdir(dir_name):
-		path = '%s/%s' % (dir_name, f)
-		if not f.endswith('.pickle') or not '_nn' in f:
-			continue
-		try:
-			with open(path, 'rb') as f:
-				relSteps, steps, scores, models, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
-		except:
-			print ('Failed to process %s' % path)
-			pass
-
-		tot_neurons = len(mean_activation_per_neuron)
-		plt.plot(np.arange(tot_neurons)+1, concatenated_results[np.argsort(mean_activation_per_neuron)])
-		av_len = 10
-		plt.plot(np.arange(av_len, tot_neurons+1), np.convolve(concatenated_results[np.argsort(mean_activation_per_neuron)], np.ones(av_len), mode='valid')/av_len)
-		plt.xlabel(xlabel)
-		plt.ylabel('Correlation coefficient')
-		plt.tight_layout()
-		plt.savefig(path[:-7] + '.pdf')
-
-		plt.close()
-
-		#plt.show()
-

### 2019-09-09 ex2/plot_pruning_nn.py

@@ -1,0 +1,31 @@
+#!/usr/bin/env python3
+
+import os
+import pickle
+
+for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
+	for f in os.listdir(dir_name):
+		path = '%s/%s' % (dir_name, f)
+		if not f.endswith('.pickle') or not '_nn' in f:
+			continue
+		try:
+			with open(path, 'rb') as f:
+				# relSteps, steps, scores, models, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
+				relSteps, steps, scores, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
+			print("Succeeded")
+		except Exception as e:
+			print(e)
+			# print ('Failed to process %s' % path)
+			# pass
+			continue
+
+		tot_neurons = len(mean_activation_per_neuron)
+		plt.plot(np.arange(tot_neurons)+1, concatenated_results[np.argsort(mean_activation_per_neuron)])
+		av_len = 10
+		plt.plot(np.arange(av_len, tot_neurons+1), np.convolve(concatenated_results[np.argsort(mean_activation_per_neuron)], np.ones(av_len), mode='valid')/av_len)
+		plt.xlabel(xlabel)
+		plt.ylabel('Correlation coefficient')
+		plt.tight_layout()
+		plt.savefig(path[:-7] + '.pdf')
+
+		plt.close()

### 2019-09-09 ex2/paper/paper.tex
Paper draft more or less finished

--- 2019-09-09 ex2/paper/paper.tex
@@
 
+\newcommand{\unsw}{UNSW-NB15}
+\newcommand{\cic}{CIC-IDS-2017}
+
 %%
@@
 \newacronym{dl}{DL}{Deep Learning}
+\newacronym{ad}{AD}{Anomaly Detection}
+\newacronym{dt}{DT}{Decision Tree}
 \newacronym{ml}{ML}{Machine Learning}
@@
 However, we find that visualization approaches can aid in identifying a backdoor independent of the used classifier.
-Finally, we conceive pruning-based approaches to remove backdoors for decision trees and random forests and demonstrate their effectiveness for several network security datasets.
+Finally, we conceive pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for several network security datasets.
 \end{abstract}
@@
 %% used to denote shared contribution to the research.
-\author{Maximilian Bachl}
+\author{Maximilian Bachl, Alexander Hartl, Joachim Fabini, Tanja Zseby}
 \affiliation{\institution{Technische Universität Wien}}
-%\authornote{Both authors contributed equally to this research.}
-\email{maximilian.bachl@tuwien.ac.at}
-
-\author{Alexander Hartl}
-\affiliation{\institution{Technische Universität Wien}}
-%\authornotemark[1]
-\email{alexander.hartl@tuwien.ac.at}
-
-\author{Joachim Fabini}
-\affiliation{\institution{Technische Universität Wien}}
-\email{joachim.fabini@tuwien.ac.at}
-
-\author{Tanja Zseby}
-\affiliation{\institution{Technische Universität Wien}}
-\email{tanja.zseby@tuwien.ac.at}
+\email{firstname.lastname@tuwien.ac.at}
 
@@
 The detection of attacks in data networks is a fundamental task in data security.
-Due to the considerable amount of data which has to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
+Due to the considerable amount of data which have to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
 
 For research, the invention and assessment of techniques for \glspl{ids} poses many challenges.
-To implement such a system, a considerable number of features can be extracted from network data which might be beneficial for anomaly detection and it is not always clear which features are actually important.
+To implement such a system many different features can be extracted from network data which might be beneficial for \gls{ad} and it is not always clear which features are actually important.
 
 Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
-In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete anomaly detection system or just a model usable by an anomaly detection software which might even be open-source.
+In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into an \gls{ad} software.% which might even be open-source.
 
@@
 
-For a well-performing network anomaly detection technique it is therefore of utmost importance to study which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
+For a well-performing network \gls{ad} technique it is therefore of utmost importance to study which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
 
 % TODO: think meghdouri_analysis_2018 didnt use this
-In this paper, we redo parts of a recent paper~\cite{meghdouri_analysis_2018} which bases on the UNSW-NB15 dataset \cite{moustafa_unsw-nb15:_2015} for evaluating the performance of several feature vectors and machine learning techniques for accurate anomaly detection in the context of \glspl{ids}.
+In this paper, we train models to detect network attacks similar to the approach of a recent paper~\cite{meghdouri_analysis_2018}, which bases on the \unsw{} dataset \cite{moustafa_unsw-nb15:_2015} and evaluates the performance of several feature vectors and machine learning techniques for accurate \gls{ad} in the context of \glspl{ids}.
 %We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
-Furthermore, we add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
+We then add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
 
-We then discuss several techniques to detect or remove a backdoor from a trained model.
+Then we discuss several techniques to detect or remove a backdoor from a trained model.
 In particular, we show how visualization techniques from explainable \gls{ml} can be used to detect backdoors and problems emerging from the distribution of attack samples in the training dataset.
 
-Probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers. We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
+As probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
+We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
 
-Finally, we also evaluate techniques proposed for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly show to be uneffective for our \gls{mlp} classifiers.
-
-\begin{table*}[h]
-\caption{Detection performance results.} \label{tab:performance_results}
-\begin{tabular}{l r r r r} \toprule
-& \multicolumn{2}{r}{UNSW-NB15} & \multicolumn{2}{r}{CIC-IDS-2017} \\
-& RF & DL & RF & DL \\ \midrule
-Accuracy	& 0.990 $\pm$ 0.000	& 0.989 $\pm$ 0.000	& 0.997 $\pm$ 0.000	& 0.998 $\pm$ 0.000	\\
-Precision	& 0.854 $\pm$ 0.001	& 0.845 $\pm$ 0.004	& 0.997 $\pm$ 0.000	& 0.999 $\pm$ 0.000	\\
-Recall	& 0.850 $\pm$ 0.003	& 0.829 $\pm$ 0.009	& 0.993 $\pm$ 0.000	& 0.992 $\pm$ 0.000	\\
-F1 score	& 0.852 $\pm$ 0.002	& 0.837 $\pm$ 0.003	& 0.995 $\pm$ 0.000	& 0.995 $\pm$ 0.000	\\
-Youden's J	& 0.845 $\pm$ 0.003	& 0.823 $\pm$ 0.009	& 0.992 $\pm$ 0.000	& 0.991 $\pm$ 0.000	\\
-Backdoor acc.	& 1.000 $\pm$ 0.000	& 0.998 $\pm$ 0.002	& 1.000 $\pm$ 0.000	& 1.000 $\pm$ 0.000	\\
-
-\bottomrule
-\end{tabular}
-\end{table*}
+Finally, we also evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
 
 \section{Related Work}
+
+There have been several works with the aim of increasing robustness of decision trees against attacks.
+\cite{biggio_bagging_2011} proposed a defense mechanism against poisoning attacks that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
+Their method is applicable to \glspl{dt}.
+However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work.
+\cite{chen_robust_2019} develop a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
+\cite{russu_secure_2016} make SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
+
+Regarding pruning of neural networks, \cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
+They show that by removing neurons that are infrequently used from the last convolutional layer, potential backdoors can be removed.
+The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
+To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
+
+Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning.
+Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
+For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
+Moreover, the authors argue that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
+They thus propose fine-pruning as a combination of pruning and fine-tuning.
+As with pruning, these methods have not been applied to classic \glspl{mlp} so far.
 
@@
 \subsection{Datasets}
+\todo{Max: I would shorten here.}
 Several datasets for the purpose of building and evaluating IDSs have been developed.
@@
 
-In this research, we use the UNSW-NB15~\cite{moustafa_unsw-nb15:_2015} and the CIC-IDS-2017~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet, guaranteeing reproducibility of our results.
+In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the CIC-IDS-2017~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet, guaranteeing reproducibility of our results.
 
-The UNSW-NB15 dataset~\cite{moustafa_unsw-nb15:_2015} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
+The \unsw{} dataset~\cite{moustafa_unsw-nb15:_2015} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
 Network captures containing over 2 million flows of normal traffic and various types of attacks are provided together with a ground truth file.
@@
 
-\subsubsection{Random Forests}
-In addition to deep learning we trained a random forest classifier to detect attacks in network traffic.
-We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each random forest.
+\subsubsection{\glspl{rf}}
+In addition to deep learning we trained a \gls{rf} classifier to detect attacks in network traffic.
+We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each \gls{rf}.
 
@@
 
-We add a backdoor to our deep learning and random forest models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
+We add a backdoor to our deep learning and \gls{rf} models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
 
 Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On one hand, this pattern has to be detectable reliably by the anomaly detection system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+On one hand, this pattern has to be detectable reliably by the \gls{ad} system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
 
@@
 \subsection{Performance Results}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and random forests, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
+\begin{table}[t]
+\caption{Detection performance results.} \label{tab:performance_results}
+\begin{tabular}{l r r r r} \toprule
+& \multicolumn{2}{r}{\unsw{}} & \multicolumn{2}{r}{\cic{}} \\
+& RF & DL & RF & DL \\ \midrule
+%Accuracy	& 0.990 $\pm$ 0.000	& 0.989 $\pm$ 0.000	& 0.997 $\pm$ 0.000	& 0.998 $\pm$ 0.000	\\
+%Precision	& 0.854 $\pm$ 0.001	& 0.845 $\pm$ 0.004	& 0.997 $\pm$ 0.000	& 0.999 $\pm$ 0.000	\\
+%Recall	& 0.850 $\pm$ 0.003	& 0.829 $\pm$ 0.009	& 0.993 $\pm$ 0.000	& 0.992 $\pm$ 0.000	\\
+%F1 score	& 0.852 $\pm$ 0.002	& 0.837 $\pm$ 0.003	& 0.995 $\pm$ 0.000	& 0.995 $\pm$ 0.000	\\
+%Youden's J	& 0.845 $\pm$ 0.003	& 0.823 $\pm$ 0.009	& 0.992 $\pm$ 0.000	& 0.991 $\pm$ 0.000	\\
+%Backdoor acc.	& 1.000 $\pm$ 0.000	& 0.998 $\pm$ 0.002	& 1.000 $\pm$ 0.000	& 1.000 $\pm$ 0.000	\\
+Accuracy	& 0.990 & 0.989 & 0.997 & 0.998\\
+Precision	& 0.854 & 0.845 & 0.997 & 0.999\\
+Recall	& 0.850 & 0.829 & 0.993 & 0.992\\
+F1 score	& 0.852 & 0.837 & 0.995 & 0.995\\
+Youden's J	& 0.845 & 0.823 & 0.992 & 0.991\\
+Backdoor acc.	& 1.000 & 0.998 & 1.000 & 1.000\\
+\bottomrule
+\end{tabular}
+\end{table}
+Table~\ref{tab:performance_results} shows performance results for both deep learning and \glspl{rf}, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
 The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
-Our results are consistent with previous work like, e.g., \cite{meghdouri_analysis_2018}.
+Our results are consistent with previous work like \cite{meghdouri_analysis_2018}.
 
 \section{Remedies for Poisoning Attacks}
-We now investigate several techniques which might be used to prevent security vulnerabilities due to pretrained models.
+We now investigate several techniques which might be used to prevent security vulnerabilities that might come with pretrained backdoored models.
 In this respect the ability to detect a backdoor and the ability to remove a backdoor from the trained model can be considered as equally effective since one often has the option to fall back to a model obtained from a different source in case a model looks suspicious.
@@
 \subsubsection{Identifying Backdoors}
+\begin{figure}[h]
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
+\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the CIC-IDS-2017 classifier.
+The top plot shows the full range of TTL values while the lower plot only TTL values from 0 to 5 are displayed.}
+\label{fig:pdp_ttl}
+\end{figure}
 Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive.
@@
 
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots for the CIC-IDS-2017 classifier.}
-\label{fig:pdp_ttl}
-\end{figure}
-
-As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for both \gls{rf} and \gls{dl} show a deep notch for certain low values of stdev(TTL).
+As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for \gls{dl} show a deep notch for certain low values of stdev(TTL).
 As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets.
@@
 In our case, this notch thus points to the existance of the backdoor we embedded.
+For the \gls{rf} model the same pattern is observed but omitted for brevity.
 
@@
 The models show a clear dependence of the mean \gls{ttl} value of incoming packets, which is similarly counter-intuitive as for the feature discussed above.
-In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the UNSW-NB15 and CIC-IDS-2017 datasets.
+In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the \unsw{} and CIC-IDS-2017 datasets.
 
@@
 
-Various pruning techniques for decision trees have been proposed for decision trees in the last decades \cite{esposito_comparative_1997} with the aim of simplifying decision trees that overfit on the data.
+\todo{Max: Unnecessary or should be moved to Related Work?}
+Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the data.
 Pruning for neural networks has been proposed in the last century as a method to simplify large neural networks \cite{sietsma_neural_1988}.
-However, pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for decision trees and random forests
+However, pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
+To perform pruning, a validation dataset is needed.
+We take a validation dataset that is one fourth the size of the training dataset.
+We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor could successfully be removed and how much the accuracy on the original data suffered.
+´
 
@@
 
-\cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
-They show that by removing neuron that are infrequently used from the last convolutional layer, potential backdoors can be removed.
-The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
-To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{nids} \cite{meghdouri_analysis_2018}.
-We implement two variants of the pruning defense: We prune neurons in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*} by their average activation.
-For this, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer and then prune the neurons by setting the weight and bias to 0 in the layer preceding the activation function.
-
 \begin{figure}[h]
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
-\caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process.}
+\caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process for \cic{}.}
+\label{fig:correlation}
 \end{figure}
 
-\subsubsection{Random Forest Pruning}
-For intrusion detection, decision tree based models or random forests of decision trees  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
-For \glspl{nids} we thus consider the removal of backdoors from decision tree based models to be a more pressing issue than from \glspl{mlp}.
+We implement two variants of the pruning defense \cite{gu_badnets:_2017}: We prune neurons in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*} by their average activation.
+Those with the smallest average activation (the least used ones) first.
+For this, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer for all samples in the validation dataset and then prune the neurons by setting the weight and bias to 0 in the layer preceding the activation function.
+Our experiments revealed that pruning does not remove the backdoor at all and on the side hand decreases the accuracy for normal data if many neurons are pruned.
+To check whether other reasons are responsible for our failure to remove the backdoor we also conducted the following experiments but none could remove the backdoor: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
+\item We checked whether the dropout regularization might be the reason.
+\item We hypothesized that the issue might be that there are a lot fewer malicious samples in the dataset than there are attacks.
+Thus we reversed the backdoor and made the backdoor so that it would falsely classify benign samples as malicious ones.\end{enumerate*}
 
-The defense mechanism proposed by \cite{biggio_bagging_2011} uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset via poisoning and is applicable to decision trees.
-However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor.
+To investigate further, we computed the correlation of the activation of each neuron with the presence of the backdoor in the data: Neurons which are responsible for the backdoor should have a high correlation because they only become active when a sample is backdoored.
+We then plotted the correlation of each neurons at the time step it is pruned.
+If the pruning method worked, this would show that the neurons pruned in the beginning would have a high correlation while those later would have a low one.
+\autoref{fig:correlation} shows that neurons that this is not the case.
+It also shows that neurons are not completely separated in backdoor neurons and regular neurons: If this were the case, the correlation would either be 1 or 0 but we observe many values in between, which indicates that most neurons are both responsible for the backdoor as well as for regular data.
+
+\subsubsection{\gls{rf} Pruning}
+\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
+For \glspl{ids} we thus consider the removal of backdoors from \gls{dt} based models to be a more pressing issue than from \glspl{mlp}.
 
@@
 We develop several variants of our pruning defense with increasing levels of sophistication.
-These iterations are pruning
+These iterations are pruning \todo{Remove space in the enumeration}
 \begin{enumerate}
 \item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
-\item only ``harmless'' leaves based on their usage: In the case of \glspl{nids} we assume that attackers want to make attacks appear to be benign. 
-\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a decision tree to classify than regular samples.
+\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
+\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples.
 Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
@@
 \includegraphics[width=\textwidth]{pruning_example.pdf}
-\caption{Example of a decision tree being pruned.
-To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the decision tree is to the right of each leaf.
+\caption{Example of a \gls{dt} being pruned.
+To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
 It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside) because these are the ones our algorithm considers as candidates for pruning.
 The leaf that is going to be cut in each step has a dashed connection to its parent.
-After step 2 the algorithm stops as pruning the last remaining leaf would make the decision tree non-functional.}
+After step 2 the algorithm stops as pruning the last remaining leaf would make the \gls{dt} non-functional.}
 \label{fig:pruningExample}
 \end{figure*}
+
 \autoref{fig:pruningExample} shows an example of the last iteration of pruning applied to a tree.
-
 \begin{figure}[h]
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
-\caption{Performance metrics throughout the pruning process. }
+\caption{Performance metrics throughout the pruning process for \unsw{}. }
 \label{fig:performancePruningRF}
 \end{figure}
+\todo{Max: Which pruning algorithm actually?}
+\autoref{fig:performancePruningRF} shows the accuracy for the pruning for the \unsw{} dataset.
+With sufficient neurons being pruned, the accuracy of the backdoor goes to zero.
+The accuracy for the regular data does not decrease; it even increases thanks to the the pruning also reducing overfitting.
+With a significantly smaller validation set, pruning takes longer until it works but it still works.
+This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
 
-\autoref{fig:performancePruningRF} shows the for the UNSW-NB15 dataset pruning not only
-
-\subsection{Neural Network Fine-Pruning}
-Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
-For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
-
-Indeed, \citeauthor{liu_fine-pruning:_2018}~\cite{liu_fine-pruning:_2018} argue  that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
-They thus propose fine-pruning as a combination of pruning and fine-tuning.
-
+\subsection{Neural Network Fine-Tuning}
+\todo{Max: It said pruning before but should be tuning, right?}
 \begin{figure}[t]
@@
 \includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \glspl{mui} for the UNSW-NB15 (top) and CIC-IDS-2017 (bottom) classifiers.}
+\caption{Fine-tuning of the \glspl{mui} for the \unsw{} (top) and CIC-IDS-2017 (bottom) classifiers.}
 \label{fig:finetuning}
 \end{figure}
-\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for both UNSW-NB15 and CIC-IDS-2017 with the validation set, hence without backdoored samples. 
-\autoref{fig:finetuning} indicates clearly that, unfortunately, similar to pruning fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
+\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for both \unsw{} and \cic{} with the validation set, hence without backdoored samples.
+It indicates clearly that, unfortunately, similar to pruning, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
 
@@
 \section{Discussion}
+\todo{Max: Merge discussion and conclusions to one section and remove most parts of conclusions?}
 From our experiments, we can make three main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.
 
-To ensure that no backdoor in contained in the model in question, it has to be carefully analyzed for questionable decisions.
+To ensure that no backdoor i contained in the model in question, it has to be carefully analyzed for questionable decisions.
 For this purpose, \gls{pdp} and \gls{ale} plots are an effective tool.
@@
 
-Furthermore, for \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed, is able to reduce backdoor efficacy significantly.
+Furthermore, for \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly.
 At the same time, the classifier's detection performance is not substantially reduced.
@@
 In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
-We used a recent publicly available dataset for network anomaly detection and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
+We used a recent publicly available dataset for network \gls{ad} and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
 
@@
 %
-%We used a dataset for network anomaly detection and extracted features using an enhanced version of the CAIA feature vector.
+%We used a dataset for network \gls{ad} and extracted features using an enhanced version of the CAIA feature vector.
 Then we trained two non-interpretable classifiers on the resulting data.
-The resulting random forest as well as the deep learning model performed very well on the dataset.
+The resulting \gls{rf} as well as the deep learning model performed very well on the dataset.
 As the next step we implemented a backdoor into the dataset.
@@
 %
-%\caption{Examples for PD plots for deep learning (left side) and random forest (right side).}
+%\caption{Examples for PD plots for deep learning (left side) and \gls{rf} (right side).}
 %\label{fig:pdp}
@@
 \autoref{fig:pdp} shows examples for some features for the different folds.
-Comparing plots generated for our deep learning and for our random forest approach, respectively, the two machine learning approaches show the difference that deep learning PDPs are smooth, while random forest PDPs fluctuate heavily.
-Considering the functioning of neural networks and decision trees, this is an expected result.
+Comparing plots generated for our deep learning and for our \gls{rf} approach, respectively, the two machine learning approaches show the difference that deep learning PDPs are smooth, while \gls{rf} PDPs fluctuate heavily.
+Considering the functioning of neural networks and \glspl{dt}, this is an expected result.
 %
-%Considering the explanations the PD plots provide, consistency between deep learning and random forest can be observed for some features, while for other features both methods provide substantially different explanations.
+%Considering the explanations the PD plots provide, consistency between deep learning and \gls{rf} can be observed for some features, while for other features both methods provide substantially different explanations.
 For example, the two upper plots in \autoref{fig:pdp} exhibit a very similar behavior, while the lower plots show remarkable differences between both methods.
-An additional observation is, however, that random forests show substantial differences between the folds.
+An additional observation is, however, that \glspl{rf} show substantial differences between the folds.
 %
-%Considering these observations, it is likely that most samples from the dataset have very small values of \textit{stdev\_srcPktLength}, so that for neither deep learning nor random forests it is important to predict higher values of \textit{stdev\_srcPktLength}.
+%Considering these observations, it is likely that most samples from the dataset have very small values of \textit{stdev\_srcPktLength}, so that for neither deep learning nor \glspl{rf} it is important to predict higher values of \textit{stdev\_srcPktLength}.
 %
@@
 %
-%Most importantly, however, the \textit{mean\_srcTTL} feature depicted in \autoref{fig:pdp} shows that both deep learning and random forest learn to consider very specific TTL values for their decisions.
+%Most importantly, however, the \textit{mean\_srcTTL} feature depicted in \autoref{fig:pdp} shows that both deep learning and \gls{rf} learn to consider very specific TTL values for their decisions.
 This points out that attack traffic was generated from a very specific network with specific TTL values throughout dataset generation.
@@
 %
-%\caption{Examples for ALE plots for deep learning (left side) and random forest (right side).}
+%\caption{Examples for ALE plots for deep learning (left side) and \gls{rf} (right side).}
 %\label{fig:ale}
@@
 %
-%An important observation for ALE plots particularly for random forests is that influence of TTL values is even more distinctive.
-In fact, judging from ALE plots the random forest seems to almost exclusively consider the TTL for detecting attacks.
+%An important observation for ALE plots particularly for \glspl{rf} is that influence of TTL values is even more distinctive.
+In fact, judging from ALE plots the \gls{rf} seems to almost exclusively consider the TTL for detecting attacks.
 This emphasizes the problem described already for PDPs.
@@
 %
-%\caption{Examples for PD plots for deep learning (left side) and random forest (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
+%\caption{Examples for PD plots for deep learning (left side) and \gls{rf} (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
 %\label{fig:pdp_backdoor}
@@
 %\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
-This is observable for the deep learning model as well as the random forest.
+This is observable for the deep learning model as well as the \gls{rf}.
 %
@@
 %\begin{tabular}{l r r} \toprule
-%& Random Forest & Deep Learning \\ \midrule
+%& \gls{rf} & Deep Learning \\ \midrule
 %Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
@@
 %\begin{tabular}{l r r} \toprule
-% & Random Forest & Deep Learning \\ \midrule
+% & \gls{rf} & Deep Learning \\ \midrule
 %Forward & 1 & 0,9999 $\pm$ 0,0001 \\

### 2019-09-09 ex2/learn.py

@@ -543,7 +543,7 @@ def prune_backdoor_nn():
 			pickle.dump([rel_steps, steps_done, scores, saved_models_in_memory, scores_bd, mean_activation_per_neuron, concatenated_results], f)
 		else:
 			pickle.dump([rel_steps, steps_done, scores, saved_models_in_memory, scores_bd], f)
-	
+
 def finetune_nn():
 	train_nn(finetune=True)
 

### 2019-09-09 ex2/plot_pruning_nn.py

@@ -2,8 +2,13 @@
 
 import os
 import pickle
+import matplotlib.pyplot as plt
+import numpy as np
+
+xlabel = 'Number of pruned neurons'
 
 for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
+	print("dir_name", dir_name)
 	for f in os.listdir(dir_name):
 		path = '%s/%s' % (dir_name, f)
 		if not f.endswith('.pickle') or not '_nn' in f:
@@ -11,7 +16,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 		try:
 			with open(path, 'rb') as f:
 				# relSteps, steps, scores, models, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
-				relSteps, steps, scores, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
+				relSteps, scores, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
 			print("Succeeded")
 		except Exception as e:
 			print(e)
@@ -20,8 +25,8 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 			continue
 
 		tot_neurons = len(mean_activation_per_neuron)
-		plt.plot(np.arange(tot_neurons)+1, concatenated_results[np.argsort(mean_activation_per_neuron)])
-		av_len = 10
+		plt.plot(np.arange(tot_neurons)+1, concatenated_results[np.argsort(mean_activation_per_neuron)], linestyle="", marker=".", alpha=0.5)
+		av_len = 100
 		plt.plot(np.arange(av_len, tot_neurons+1), np.convolve(concatenated_results[np.argsort(mean_activation_per_neuron)], np.ones(av_len), mode='valid')/av_len)
 		plt.xlabel(xlabel)
 		plt.ylabel('Correlation coefficient')

### 2019-09-10 ex2/paper/paper.tex
Modified paper

@@
 Interest in poisoning attacks and backdoors recently resurfaced for \gls{dl} applications.
-Several successful defense mechanisms have been recently proposed for Convolutional Neural Networks, for example in the context of autonomous driving.
+Several successful defense mechanisms have been recently proposed for \glspl{cnn}, for example in the context of autonomous driving.
 We show that, surprisingly, common defense mechanisms fail utterly to remove backdoors in \gls{dl} for \glspl{ids}.
 However, we find that visualization approaches can aid in identifying a backdoor independent of the used classifier.
-Finally, we conceive pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for several network security datasets.
+Finally, we conceive pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for two different %several
+network security datasets.
 \end{abstract}
@@
 \section{Introduction}
-The detection of attacks in data networks is a fundamental task in data security.
+The detection of attacks in data networks is a fundamental task in network security.
 Due to the considerable amount of data which have to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
@@
 For research, the invention and assessment of techniques for \glspl{ids} poses many challenges.
-To implement such a system many different features can be extracted from network data which might be beneficial for \gls{ad} and it is not always clear which features are actually important.
+To implement such a system, many different features can be extracted from network data which might be beneficial for \gls{ad} and it is not always clear which features are actually important.
 
-Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
-In a practical deployment, it is therefore reasonable to assume that the training is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into an \gls{ad} software.% which might even be open-source.
-
+Furthermore, creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
+In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into an \gls{ad} software.% which might even be open-source.
 If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
 
-Furthermore, for the training of models usually datasets are used which have been generated artificially in a controlled test environment.
-As a downside of this approach, it is unclear whether a machine learning model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
+In addition to these problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
+As a downside of this approach, it is unclear whether a \gls{ml} model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
 
@@
 
-As probably the most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
+As the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
 We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
@@
 However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work.
-\cite{chen_robust_2019} develop a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
+\cite{chen_robust_2019} developed a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
 \cite{russu_secure_2016} make SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
 
-Regarding pruning of neural networks, \cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
+%\todo{Max: Unnecessary or should be moved to Related Work?}
+Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data.
+Pruning for neural networks has been proposed
+%in the last century
+as a method to simplify large neural networks \cite{sietsma_neural_1988}.
+However, pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
+
+%Regarding pruning of neural networks,
+In particular, \cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
 They show that by removing neurons that are infrequently used from the last convolutional layer, potential backdoors can be removed.
@@
 The CIC-IDS-2017 dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
-The provided raw network captures contain over 2.3 million flows, containing both normal traffic and attacks, which can be coarsely classified into DoS attacks, infiltration attacks, web attacks, brute force and scanning attacks.
+The provided raw network captures contain more than 2.3 million flows, containing both normal traffic and attacks, which can be coarsely classified into DoS attacks, infiltration attacks, web attacks, brute force and scanning attacks.
 
-We base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
+For processing the data, we base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
 All features except protocol and flow duration are evaluated for forward and backward direction separately.
 The goal of this research is to investigate the possibility of poisoning attacks.
-We therefore also include the minimum, maximum and standard deviation of TTL values in our feature vector as an attractive carrier for exploitation as a backdoor.
+We therefore also include the minimum, maximum and standard deviation of \gls{ttl} values in our feature vector as an attractive carrier for exploitation as a backdoor.
 % TODO: justify use of TTL better
@@
 \subsubsection{\glspl{rf}}
-In addition to deep learning we trained a \gls{rf} classifier to detect attacks in network traffic.
+In addition to \gls{dl}, we trained an \gls{rf} classifier to detect attacks in network traffic.
 We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each \gls{rf}.
@@
 
-We add a backdoor to our deep learning and \gls{rf} models and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
-
+We added a backdoor to our \gls{dl} and \gls{rf} models. % and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
 Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On one hand, this pattern has to be detectable reliably by the \gls{ad} system, one the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting these patterns.
+On one hand, this pattern has to be detectable reliably by the \gls{ids} system, on the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting this  pattern.
 
@@
 In our situation, it is reasonable to assume that the \gls{ttl} field remains constant for all packets in one traffic flow.
-In particular this assumption was perfectly reflected by the examined dataset, as there were no traffic flows with a non-zero standard deviation of the \gls{ttl} value.
+In particular, this assumption was perfectly reflected by the examined UNSW-NB15 dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
 
@@
 \end{table}
-Table~\ref{tab:performance_results} shows performance results for both deep learning and \glspl{rf}, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
+Table~\ref{tab:performance_results} shows performance results for both \gls{dl} and \gls{rf}, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
 The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
@@
 \section{Remedies for Poisoning Attacks}
-We now investigate several techniques which might be used to prevent security vulnerabilities that might come with pretrained backdoored models.
+We now investigate several techniques which might be used to prevent security vulnerabilities that might come with backdoored pretrained  models.
 In this respect the ability to detect a backdoor and the ability to remove a backdoor from the trained model can be considered as equally effective since one often has the option to fall back to a model obtained from a different source in case a model looks suspicious.
@@
 \subsection{Explainability Plots} \label{sec:plots}
-A major problem for the deployment of ML is the unpredictability and the lack of understanding of the decisions an ML models takes.
-For this reason, a number of methods has been proposed in the recent years aiming to visualize and explain a non-interpretable ML model's decisions.
-
+%A major problem for the deployment of ML is the unpredictability and the lack of understanding of the decisions an ML models takes. For this reason, a
+A number of methods has been proposed in the recent years aiming to visualize and explain a non-interpretable ML model's decisions.
 Applied to the present problem, we can pick up ideas from \glspl{pdp} and \gls{ale} plots, not only for identifying backdoors in the \gls{mui}, but also for finding wrong decisions it would take due to flawed training data.
@@
 
-For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx y$.
+For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx y$. We used the 10  closest samples
+%to the interval's center
+for estimating the distribution.
 
-When computing \gls{ale} plots, we experienced the problem of empty intervals.
+% When computing \gls{ale} plots, we experienced the problem of empty intervals.
 If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values doesn't work.
 
-For this reason, we modified this definition to instead use the closest 10 samples to the interval's center for estimating the distribution.
+%For this reason, we modified this definition to instead use the closest 10 samples to the interval's center for estimating the distribution.
 
@@
 
-\todo{Max: Unnecessary or should be moved to Related Work?}
-Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the data.
-Pruning for neural networks has been proposed in the last century as a method to simplify large neural networks \cite{sietsma_neural_1988}.
-However, pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
 To perform pruning, a validation dataset is needed.
 We take a validation dataset that is one fourth the size of the training dataset.
-We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor could successfully be removed and how much the accuracy on the original data suffered.
+We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
 ´
@@
 
-To investigate further, we computed the correlation of the activation of each neuron with the presence of the backdoor in the data: Neurons which are responsible for the backdoor should have a high correlation because they only become active when a sample is backdoored.
-We then plotted the correlation of each neurons at the time step it is pruned.
-If the pruning method worked, this would show that the neurons pruned in the beginning would have a high correlation while those later would have a low one.
-\autoref{fig:correlation} shows that neurons that this is not the case.
+To investigate further, we computed the correlation of the activation of each neuron with the presence of the backdoor in the data.
+Neurons which are responsible for the backdoor should have a high correlation because they only become active when a sample is backdoored.
+We  plotted the correlation of each neurons at the time step it is pruned, which is depicted in \autoref{fig:correlation}.
+If the pruning method worked, we would expect the figure to
+%this would
+show that the neurons pruned in the beginning have a high correlation while later ones would have a low correlation.
+
+\autoref{fig:correlation} shows that this is not the case.
 It also shows that neurons are not completely separated in backdoor neurons and regular neurons: If this were the case, the correlation would either be 1 or 0 but we observe many values in between, which indicates that most neurons are both responsible for the backdoor as well as for regular data.
@@
 The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
-If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease by a lot while decisions that are used for the backdoor are hopefully removed.
+If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
+%by a lot
+significantly
+while decisions that are used for the backdoor are hopefully removed.
 We develop several variants of our pruning defense with increasing levels of sophistication.
-These iterations are pruning \todo{Remove space in the enumeration}
+These variants are pruning \todo{Remove space in the enumeration}
 \begin{enumerate}
@@
 
-\autoref{fig:pruningExample} shows an example of the last iteration of pruning applied to a tree.
+\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
 \begin{figure}[h]
@@
 \end{figure}
-\todo{Max: Which pruning algorithm actually?}
-\autoref{fig:performancePruningRF} shows the accuracy for the pruning for the \unsw{} dataset.
-With sufficient neurons being pruned, the accuracy of the backdoor goes to zero.
-The accuracy for the regular data does not decrease; it even increases thanks to the the pruning also reducing overfitting.
-With a significantly smaller validation set, pruning takes longer until it works but it still works.
-This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
+%\todo{Max: Which pruning algorithm actually?}
+and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset.
+With sufficient neurons being pruned, the accuracy of the backdoor approaches zero.
+The accuracy for the regular data does not decrease; it even increases by
+%thanks to the the pruning also
+reducing overfitting. With a significantly smaller validation set, pruning takes longer to remove the backdoor
+%until it works
+but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
 
 \subsection{Neural Network Fine-Tuning}
-\todo{Max: It said pruning before but should be tuning, right?}
+%\todo{Max: It said pruning before but should be tuning, right?}
+%AH. I guess both should work
 \begin{figure}[t]
@@
 
-To ensure that no backdoor i contained in the model in question, it has to be carefully analyzed for questionable decisions.
+To ensure that no backdoor is contained in the model in question, it has to be carefully analyzed for questionable decisions.
 For this purpose, \gls{pdp} and \gls{ale} plots are an effective tool.
-In fact, already throughout the process of training of the model explainability plots constitute a useful tool to ensure that the model is not trained unintentionally to artifacts the dataset yields.
+In fact, already throughout the process of training of the model explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
 
@@
 With this setup, we explored several mitigation techniques.
-While, surprisingly, neural network pruning and fine-tuning was uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots both for the \gls{dl} and \gls{rf} models.
+While, surprisingly, neural network pruning and fine-tuning were uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots  for both the \gls{dl} and \gls{rf} models.
 

### 2019-09-10 ex2/plot_pruning.py
Modified plotting scripts

@@ -19,6 +19,7 @@ extra_metrics = {
 	'depth': ('Backdoor accuracy', {'color': 'c'}),
 }
 
+plt.rcParams["font.family"] = "serif"
 xlabel = 'Relative amount of pruned neurons'
 
 def doplot(filenames, extra_metric="bd", **kwargs):
@@ -57,6 +58,8 @@ validation_set_ratios = "0.01 0.10 0.20 0.30 0.40 0.50 0.60 0.70 0.80 0.90 1.00"
 
 # linestyles.append((Line2D([0], [0], color='black', linestyle='--'), 'Using 1% of validation data'))
 
+plt.figure(figsize=(5,4))
+
 for index, item in enumerate(validation_set_ratios):
 	# doplot(['prune_CAIA_backdoor/prune_'+str(item)+'_oh_rf_0_bd.pickle'], linestyle='--')
 	# doplot(['prune_CAIA_backdoor/prune_'+str(item)+'_oh_rf_0_bd.pickle'], dashes=[index+1, index+1])
@@ -69,6 +72,7 @@ plt.xlabel(xlabel)
 plt.ylabel('Classification performance')
 
 plt.tight_layout()
+plt.show()
 plt.savefig('prune_CAIA_backdoor/prune.pdf')
 
 

### 2019-09-10 ex2/plot_pruning_nn.py

@@ -6,6 +6,7 @@ import matplotlib.pyplot as plt
 import numpy as np
 
 xlabel = 'Number of pruned neurons'
+plt.rcParams["font.family"] = "serif"
 
 for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 	print("dir_name", dir_name)
@@ -24,6 +25,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 			# pass
 			continue
 
+		plt.figure(figsize=(5,4))
 		tot_neurons = len(mean_activation_per_neuron)
 		plt.plot(np.arange(tot_neurons)+1, concatenated_results[np.argsort(mean_activation_per_neuron)], linestyle="", marker=".", alpha=0.5)
 		av_len = 100

### 2019-09-10 ex2/plot_pruning_nn.py
Removed ol and of plots

@@ -27,6 +27,8 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 				elif len(data) == 6:
 					mean_activation_per_neuron = data[4]
 					concatenated_results = data[5]
+				else:
+					continue
 			print("Succeeded")
 		except Exception as e:
 			print(e)

### 2019-09-10 ex2/paper/paper.tex
Update paper and fix minor things in code and plotting script

@@
 %%
-%% For the copyright see the source file.
+%%% For the copyright see the source file.
 %%
@@
 \subsubsection{\gls{rf} Pruning}
-\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
+%\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
 For \glspl{ids} we thus consider the removal of backdoors from \gls{dt} based models to be a more pressing issue than from \glspl{mlp}.
 
-We thus propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}.
+We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}.
 The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
@@
 These variants are pruning \todo{Remove space in the enumeration}
-\begin{enumerate}
+\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
 \item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
@@
 To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
-It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside) because these are the ones our algorithm considers as candidates for pruning.
-The leaf that is going to be cut in each step has a dashed connection to its parent.
-After step 2 the algorithm stops as pruning the last remaining leaf would make the \gls{dt} non-functional.}
+It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside).
+The leaf that is going to be cut in each step has a dashed connection to its parent.}
 \label{fig:pruningExample}
@@
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune.pdf}
-\caption{Performance metrics throughout the pruning process for \unsw{}. }
+\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
+\caption{Performance metrics throughout the pruning process for \unsw{}.
+Considering only harmless leaves and pruning leaves with smaller depth first.
+The longer the dashes, the higher the fraction of the validation set being used (from 0.01 until 1).}
 \label{fig:performancePruningRF}
@@
 %thanks to the the pruning also
-reducing overfitting. With a significantly smaller validation set, pruning takes longer to remove the backdoor
+reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
 %until it works
-but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
+%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning. 
+Also with only a very small validation dataset, the pruning still works as expected.
+For \cic{} we get similar results, but the pruning does not remove the backdoor completely (\~30\% backdoor accuracy remains).
+We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stddev}(\text{TTL}) > 0$.
+And thus backdoored flows cannot always be sharply distinguished from regular flows.
+We find that pruning only harmless leaves is very beneficial for \unsw{} but not for \cic{}.
+However, considering depth when making the decision which leaf to prune first leads to the backdoor being removed considerably earlier in the pruning process.
 
@@
 \bibliography{bibliography}
+%\bibliographystyle{plain}
 

### 2019-09-10 ex2/learn.py

@@ -864,9 +864,10 @@ def prune_backdoor_rf():
 	step_width = 1/(opt.nSteps+1)
 
 	new_rfs = [rf]
-	steps_done = [0]
+	steps_done = [[0]*len(rf.estimators_)]
 	for step in range(opt.nSteps):
 		new_rf = copy.deepcopy(new_rfs[-1])
+		steps_done.append([])
 		for index, tree in enumerate(new_rf.estimators_):
 			n_nodes = tree.original_n_leaves if not opt.pruneOnlyHarmless else sum(tree.original_harmless & (tree.original_children_left==TREE_LEAF) & (tree.original_children_right==TREE_LEAF))
 			if step==0:
@@ -874,7 +875,7 @@ def prune_backdoor_rf():
 			pruned_steps = int(round(step_width*(step)*n_nodes))
 			pruned_steps_after_this = int(round(step_width*(step+1)*n_nodes))
 			steps_to_do = pruned_steps_after_this - pruned_steps
-			steps_done.append(pruned_steps_after_this)
+			steps_done[-1].append(pruned_steps_after_this)
 			print("Pruned", pruned_steps, "steps going", steps_to_do, "steps until", pruned_steps_after_this, "steps or", (step+1)/(opt.nSteps+1), "with", reachable_nodes(tree), "nodes remaining and", reachable_nodes(tree, only_leaves=True), "leaves")
 			new_tree, pruned_nodes_dict = prune_steps_from_tree(tree, steps_to_do)
 			usages_average = np.mean(np.array(pruned_nodes_dict["usages"]))
@@ -926,7 +927,7 @@ if __name__=="__main__":
 	parser.add_argument('--naive', action='store_true', help='include naive version of the backdoor')
 	parser.add_argument('--depth', action='store_true', help='whether depth should be considered in the backdoor pruning algorithm')
 	parser.add_argument('--pruneOnlyHarmless', action='store_true', help='whether only harmless nodes shall be pruned')
-	parser.add_argument('--takeSignOfActivation', action='store_true', help='whether only harmless nodes shall be pruned')
+	parser.add_argument('--takeSignOfActivation', action='store_true')
 	parser.add_argument('--onlyLastLayer', action='store_true', help='whether only the last layer is considered for pruning')
 	parser.add_argument('--onlyFirstLayer', action='store_true', help='whether only the first layer is considered for pruning')
 	parser.add_argument('--pruneConnections', action='store_true', help='whether the connections in the matrix should be pruned and not the neurons themselves')

### 2019-09-10 ex2/plot_pruning.py

@@ -7,45 +7,60 @@ import os
 import matplotlib.pyplot as plt
 from matplotlib.lines import Line2D
 import numpy as np
+import warnings
 
 metrics = {
 	'Accuracy':	('Accuracy', {'color': 'b'}),
-	'Youden': ('J score', {'color': 'g'})
+	'Youden': ("Youden's J", {'color': 'g'})
 }
 
-extra_metrics = {
-	'bd': ('Backdoor accuracy', {'color': 'r'}), # This is actually only harmless
-	'all': ('Backdoor accuracy', {'color': 'y'}),
-	'depth': ('Backdoor accuracy', {'color': 'c'}),
-}
+# extra_metrics = {
+# 	'bd': ('Backdoor accuracy', {'color': 'r'}), # This is actually only harmless
+# 	'all': ('Backdoor accuracy', {'color': 'y'}),
+# 	'depth': ('Backdoor accuracy', {'color': 'c'}),
+# }
 
 plt.rcParams["font.family"] = "serif"
-xlabel = 'Relative amount of pruned neurons'
+xlabel = 'Relative amount of pruned leaves'
 
 def doplot(filenames, extra_metric="bd", **kwargs):
 	relStepss = []
+	stepss = []
 	scoress = { metric: [] for metric in list(metrics) + ['bd'] }
 	for filename in filenames:
+		print("filename", filename)
 		with open(filename, 'rb') as f:
 			data = pickle.load(f)
-		relSteps, steps, scores, scoresbd = list(data)[:3]
+		print("len(data)", len(data))
+		# print("data", data)
+		if len(data) == 4:
+			relSteps, steps, scores, scoresbd = data
+			stepss.append(steps)
+		else:
+			warnings.warn("There shouldn't be less than 4 items in the pickle file...")
+			relSteps, scores, scoresbd = data
 		for metric in metrics:
 			scoress[metric].append(scores[metric])
 		scoress['bd'].append(scoresbd['Accuracy'])
 		relStepss.append(relSteps)
 
 	assert all(relSteps == relStepss[0] for relSteps in relStepss)
+	assert all(steps == stepss[0] for steps in stepss)
 	means = { metric: np.mean(scoress[metric], axis=0) for metric in scoress }
 	if len(filenames) > 1:
[…]

### 2019-09-10 ex2/plot.py
Make plots more space efficient

@@ -62,5 +62,5 @@ for f in os.listdir(dir_name):
 			all_labels = [item.get_label() for item in all_legends]
 			plt.legend(all_legends, all_labels)
 		plt.tight_layout()
-		plt.savefig(dir_name+'/%s_%s%s%s%s.pdf' % (feature, match.group(2), match.group(3), match.group(4), "_hist" if hist else ""))
+		plt.savefig(dir_name+'/%s_%s%s%s%s.pdf' % (feature, match.group(2), match.group(3), match.group(4), "_hist" if hist else ""), bbox_inches = 'tight', pad_inches = 0)
 		plt.close()

### 2019-09-10 ex2/plot_finetuning.py

@@ -25,5 +25,5 @@ plt.xlabel('Epoch')
 plt.ylabel('Metric')
 plt.legend(columns.values())
 plt.tight_layout()
-plt.savefig(sys.argv[1])
+plt.savefig(sys.argv[1], bbox_inches = 'tight', pad_inches = 0)
 #plt.show()

### 2019-09-10 ex2/plot_pruning.py
Change plots

@@ -77,7 +77,7 @@ variations = ["", "_d", "_oh", "_oh_d"]
 
 for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 	for variation in variations:
-		plt.figure(figsize=(5,4))
+		plt.figure(figsize=(5,2))
 
 		for index, item in enumerate(validation_set_ratios):
 			# doplot(['prune_CAIA_backdoor/prune_'+str(item)+'_oh_rf_0_bd.pickle'], linestyle='--')
@@ -89,7 +89,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 		plt.legend(*zip(*linestyles), loc='lower left')
 
 		plt.xlabel(xlabel)
-		plt.ylabel('Classification performance')
+		plt.ylabel('Classification perf.')
 
 		plt.tight_layout()
 		plt.show()

### 2019-09-10 ex2/plot_pruning_nn.py

@@ -12,6 +12,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 	print("dir_name", dir_name)
 	for f in os.listdir(dir_name):
 		path = '%s/%s' % (dir_name, f)
+		print("path", path)
 		if not f.endswith('.pickle') or not '_nn' in f:
 			continue
 		try:
@@ -21,6 +22,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 				relSteps = data[0]
 				scores = data[1]
 				scoresbd = data[2]
+				print("len(data)", len(data))
 				if len(data) == 7:
 					mean_activation_per_neuron = data[5]
 					concatenated_results = data[6]
@@ -36,7 +38,8 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 			# pass
 			continue
 
-		plt.figure(figsize=(5,3.5))
+		# plt.figure(figsize=(5,3.5))
+		plt.figure(figsize=(5,3))
 		tot_neurons = len(mean_activation_per_neuron)
 		sort_indices = np.argsort(mean_activation_per_neuron)
 		lines = []

### 2019-09-10 ex2/paper/paper.tex
Make title fit in one line
Make caption shorter

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Bricking up and overgrowing backdoors in Intrusion Detection Systems}
+\title{Bricking up backdoors in Intrusion Detection Systems}
 
@@
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
-\caption{Performance metrics throughout the pruning process for \unsw{}.
-Considering only harmless leaves and pruning leaves with smaller depth first.
-The longer the dashes, the higher the fraction of the validation set being used (from 0.01 until 1).}
+\caption{Performance metrics throughout the pruning process for \unsw{}. Considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
 \label{fig:performancePruningRF}

### 2019-09-10 ex2/paper/paper.tex
Remove second fine tuning figure
Remove conclusion

@@
 \includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
-\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \glspl{mui} for the \unsw{} (top) and CIC-IDS-2017 (bottom) classifiers.}
+%\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
+\caption{Fine-tuning of the \glspl{mui} for \unsw{}.}
 \label{fig:finetuning}
@@
 
-Most importantly, however, features to use for deploying an \gls{ids} have to be selected carefully.
+We found it surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
+This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.
+
+Another important takeaway is that the features used when deploying an \gls{ids} have to be selected carefully.
 Indeed, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
@@
 \section{Conclusions}
-In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
+%In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
 We used a recent publicly available dataset for network \gls{ad} and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
@@
 
-We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
+%We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
 
-Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
+%Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
 Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use.

### 2019-09-10 ex2/plot_finetuning.py
New fine-tuning plots

@@ -1,29 +1,48 @@
 #!/usr/bin/env python3
 
+# run with ./plot_finetuning.py runs/Sep09_09-01-41_gpu_0_3/finetuning.csv --legend --height 2
+
 import sys
 import os
 import matplotlib.pyplot as plt
 import numpy as np
 import pandas as pd
+import argparse
 
+parser = argparse.ArgumentParser()
+parser.add_argument('filenames', type=str, nargs='+')
+parser.add_argument('--legend', action='store_true')
+parser.add_argument('--height', default=3., type=float)
+parser.add_argument('--ymax', default=None, type=float)
+parser.add_argument('--save', type=str)
 
+opt = parser.parse_args()
 plt.rcParams["font.family"] = "serif"
 
-data = [ pd.read_csv(filename) for filename in sys.argv[2:] ]
+data = [ pd.read_csv(filename) for filename in opt.filenames ]
 columns = data[0].columns
-columns = {'Youden': "J score", 'Backdoor_acc': 'Backdoor accuracy'}
+columns = {'Accuracy': 'Accuracy', 'Youden': "Youden's J", 'Backdoor_acc': 'Backdoor accuracy'}
 length = min([ item.shape[0] for item in data ])
 data = np.stack([ item.iloc[:length,:].loc[:,list(columns)].values for item in data ]) # order of metrics might deviate in files
 
 means = np.mean(data, axis=0)
-stds = np.std(data, axis=0)
+if data.shape[0] > 1:
+	stds = np.std(data, axis=0)
 
-plt.figure(figsize=(5,2))
+plt.figure(figsize=(5,opt.height))
 for i in range(means.shape[1]):
-	plt.errorbar(np.arange(length)[:,None] + 1, means[:,i], stds[:,i])
+	if data.shape[0] > 1:
+		plt.errorbar(np.arange(length)[:,None] + 1, means[:,i], stds[:,i])
+	else:
+		plt.plot(np.arange(length)[:,None] + 1, means[:,i])
+		
 plt.xlabel('Epoch')
 plt.ylabel('Metric')
-plt.legend(columns.values())
+if opt.legend:
+	plt.legend(columns.values())
 plt.tight_layout()
-plt.savefig(sys.argv[1], bbox_inches = 'tight', pad_inches = 0)
+if opt.save:
+	plt.savefig(opt.save, bbox_inches = 'tight', pad_inches = 0)
+else:
+	plt.show()
 #plt.show()

### 2019-09-10 ex2/plot_finetuning.py
Updated fine-tuning

@@ -37,7 +37,7 @@ for i in range(means.shape[1]):
 		plt.plot(np.arange(length)[:,None] + 1, means[:,i])
 		
 plt.xlabel('Epoch')
-plt.ylabel('Metric')
+plt.ylabel('Classification perf.')
 if opt.legend:
 	plt.legend(columns.values())
 plt.tight_layout()

### 2019-09-10 ex2/paper/paper.tex
Updated fine-tuning
Actually remove conclusion

@@
 Finally, we also evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
+
+We make all our code publicly available at \todo{Insert link to GitHub repo}.
 
@@
 significantly
-while decisions that are used for the backdoor are hopefully removed.
-We develop several variants of our pruning defense with increasing levels of sophistication.
-These variants are pruning \todo{Remove space in the enumeration}
+while decisions that are used for the backdoor are hopefully removed. We develop several variants of our pruning defense with increasing levels of sophistication. These variants are pruning
 \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
@@
 \section{Discussion}
-\todo{Max: Merge discussion and conclusions to one section and remove most parts of conclusions?}
 From our experiments, we can make three main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.
@@
 
-\section{Conclusions}
-%In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
+%\section{Conclusions}
+%%In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
 We used a recent publicly available dataset for network \gls{ad} and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
-
-With this setup, we explored several mitigation techniques.
+%
+%With this setup, we explored several mitigation techniques.
 While, surprisingly, neural network pruning and fine-tuning were uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots  for both the \gls{dl} and \gls{rf} models.
-
-%We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
-
-%Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
+%
+%%We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
+%
+%%Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
 Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use.

### 2019-09-10 ex2/plot_pruning.py

@@ -9,9 +9,14 @@ from matplotlib.lines import Line2D
 import numpy as np
 import warnings
 
+prop_cycle = plt.rcParams['axes.prop_cycle']
+colors = prop_cycle.by_key()['color']
+
+print("colors", colors)
+
 metrics = {
-	'Accuracy':	('Accuracy', {'color': 'b'}),
-	'Youden': ("Youden's J", {'color': 'g'})
+	'Accuracy':	('Accuracy', {'color': colors[0]}),
+	'Youden': ("Youden's J", {'color': colors[2]})
 }
 
 # extra_metrics = {
@@ -51,19 +56,19 @@ def doplot(filenames, extra_metric="bd", **kwargs):
 		stds = { metric: np.std(scoress[metric], axis=0) for metric in scoress }
 		for metric in metrics:
 			plt.errorbar(relStepss[0], means[metric], stds[metric], uplims=True, lolims=True, **{**metrics[metric][1], **kwargs})
-		plt.errorbar(relStepss[0], means['bd'], stds['bd'], color='r', uplims=True, lolims=True, **kwargs)
+		plt.errorbar(relStepss[0], means['bd'], stds['bd'], color=colors[1], uplims=True, lolims=True, **kwargs)
 		# 	plt.errorbar(stepss[0], means[metric], stds[metric], uplims=True, lolims=True, **{**metrics[metric][1], **kwargs})
 		# plt.errorbar(stepss[0], means['bd'], stds['bd'], color='r', uplims=True, lolims=True, **kwargs)
 	else:
 		for metric in metrics:
 			plt.plot(relStepss[0], means[metric], **{**metrics[metric][1], **kwargs})
-		plt.plot(relStepss[0], means["bd"], color="r", **kwargs)
+		plt.plot(relStepss[0], means["bd"], color=colors[1], **kwargs)
 		# for metric in metrics:
 		# 	plt.plot(stepss[0], means[metric], **{**metrics[metric][1], **kwargs})
 		# plt.plot(stepss[0], means["bd"], color="r", **kwargs)
 
 linestyles = [ ( Line2D([0], [0], **metrics[metric][1]), metrics[metric][0]) for metric in metrics ]
-linestyles.append((Line2D([0], [0], color='r'), 'Backdoor accuracy'))
[…]

### 2019-09-10 ex2/paper/paper.tex
Tiny fix

@@
 There have been several works with the aim of increasing robustness of decision trees against attacks.
-\cite{biggio_bagging_2011} proposed a defense mechanism against poisoning attacks that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
+\cite{biggio_bagging_2011} proposed a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
 Their method is applicable to \glspl{dt}.
@@
 For this, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer for all samples in the validation dataset and then prune the neurons by setting the weight and bias to 0 in the layer preceding the activation function.
-Our experiments revealed that pruning does not remove the backdoor at all and on the side hand decreases the accuracy for normal data if many neurons are pruned.
+Our experiments revealed that pruning does not remove the backdoor at all while decreasing the accuracy for normal data if many neurons are pruned.
 To check whether other reasons are responsible for our failure to remove the backdoor we also conducted the following experiments but none could remove the backdoor: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
@@
 \begin{figure}[t]
-\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
-%\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
+%\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
+\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
 \caption{Fine-tuning of the \glspl{mui} for \unsw{}.}

### 2019-09-10 ex2/paper/paper.tex
Several minor corrections for pruning

--- 2019-09-10 ex2/paper/paper.tex
@@
 We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
-´
+Training, validation and test dataset are pairwise disjoint.
 
@@
 
-We implement two variants of the pruning defense \cite{gu_badnets:_2017}: We prune neurons in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*} by their average activation.
-Those with the smallest average activation (the least used ones) first.
-For this, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer for all samples in the validation dataset and then prune the neurons by setting the weight and bias to 0 in the layer preceding the activation function.
-Our experiments revealed that pruning does not remove the backdoor at all and on the side hand decreases the accuracy for normal data if many neurons are pruned.
-To check whether other reasons are responsible for our failure to remove the backdoor we also conducted the following experiments but none could remove the backdoor: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
+% AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
+We implement two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*}.
+Neurons with the smallest average activation (the least used ones) are pruned first.
+For this purpose, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer for all samples in the validation dataset and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.
+
+Our experiments revealed that pruning does not remove the backdoor at all, while decreasing the accuracy for normal data if too many neurons are pruned.
+To check whether other reasons are responsible for the \gls{mlp}'s failure to remove the backdoor, we also conducted the following experiments: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
 \item We checked whether the dropout regularization might be the reason.
 \item We hypothesized that the issue might be that there are a lot fewer malicious samples in the dataset than there are attacks.
-Thus we reversed the backdoor and made the backdoor so that it would falsely classify benign samples as malicious ones.\end{enumerate*}
+Thus we reversed the backdoor and made the backdoor so that it would falsely classify benign samples as malicious ones.
+\end{enumerate*}
 
+However, none of the experiments could remove the backdoor.
 To investigate further, we computed the correlation of the activation of each neuron with the presence of the backdoor in the data.
 Neurons which are responsible for the backdoor should have a high correlation because they only become active when a sample is backdoored.
-We  plotted the correlation of each neurons at the time step it is pruned, which is depicted in \autoref{fig:correlation}.
+We  plotted the correlation of each neuron at the time step it is pruned, which is depicted in \autoref{fig:correlation}.
 If the pruning method worked, we would expect the figure to
@@
 \autoref{fig:correlation} shows that this is not the case.
-It also shows that neurons are not completely separated in backdoor neurons and regular neurons: If this were the case, the correlation would either be 1 or 0 but we observe many values in between, which indicates that most neurons are both responsible for the backdoor as well as for regular data.
+It also shows that neurons are not completely separated in backdoor neurons and regular neurons: If this were the case, the correlation would either be 1 or 0, but we observe many values in between, which indicates that most neurons are both responsible for the backdoor as well as for regular data.
 
@@
 
-We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}.
+We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above.
 The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
@@
 %by a lot
-significantly
-while decisions that are used for the backdoor are hopefully removed. We develop several variants of our pruning defense with increasing levels of sophistication. These variants are pruning
+significantly,
+while decisions that are used for the backdoor are likely to be removed.
+
+We developed several variants of our pruning defense with increasing levels of sophistication. These variants are 
 \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
-\item leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
-\item only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
-\item only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples.
-Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages like another leaf with a higher depth.
+\item pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
+\item pruning only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
+\item pruning only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples.
+Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages as another leaf with a higher depth.
 \end{enumerate}
@@
 With sufficient neurons being pruned, the accuracy of the backdoor approaches zero.
-The accuracy for the regular data does not decrease; it even increases by
+The accuracy for the regular data does not decrease, it even increases by
 %thanks to the the pruning also
@@
 %but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning. 
-Also with only a very small validation dataset, the pruning still works as expected.
-For \cic{} we get similar results, but the pruning does not remove the backdoor completely (\~30\% backdoor accuracy remains).
+Even with a very small validation dataset of just 1\% of its original size, the pruning still works as expected.
+
+For \cic{} we get similar results, but the pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains).
 We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stddev}(\text{TTL}) > 0$.
-And thus backdoored flows cannot always be sharply distinguished from regular flows.
+Thus, backdoored flows cannot always be sharply distinguished from regular flows.
 We find that pruning only harmless leaves is very beneficial for \unsw{} but not for \cic{}.
-However, considering depth when making the decision which leaf to prune first leads to the backdoor being removed considerably earlier in the pruning process.
+However, considering depth when making the decision which leaf to prune first, leads to the backdoor being removed considerably earlier in the pruning process.
 

### 2019-09-10 ex2/paper/paper.tex
Shorten introduction

--- 2019-09-10 ex2/paper/paper.tex
@@
 \section{Introduction}
-The detection of attacks in data networks is a fundamental task in network security.
+%The detection of attacks in data networks is a fundamental task in network security.
 Due to the considerable amount of data which have to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
 
-For research, the invention and assessment of techniques for \glspl{ids} poses many challenges.
-To implement such a system, many different features can be extracted from network data which might be beneficial for \gls{ad} and it is not always clear which features are actually important.
+%For research, 
+The invention and assessment of techniques for \glspl{ids} poses many challenges.
+Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
+In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software.
+% which might even be open-source.
+If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have implemented backdoors which circumvent the \gls{ad} system.
+This could be motivated by profitseeking or by government actors requiring ways to purposefully disable security measures in specific cases.
 
-Furthermore, creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
-In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into an \gls{ad} software.% which might even be open-source.
-If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
-
-In addition to these problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
+In addition to this problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
 As a downside of this approach, it is unclear whether a \gls{ml} model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
 
-For a well-performing network \gls{ad} technique it is therefore of utmost importance to study which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
+For a well-performing network \gls{ad} technique it is therefore of utmost importance to study which features are useful and which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
 
@@
 
-We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}.
-The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
+%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}. 
+The concept behind this our pruning defense is that leaves, which are never used by samples in the validation dataset, might be used by backdoored inputs.
 If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease

### 2019-09-10 ex2/paper/paper.tex

--- 2019-09-10 ex2/paper/paper.tex
@@
 \section{Introduction}
-The detection of attacks in data networks is a fundamental task in network security.
+%The detection of attacks in data networks is a fundamental task in network security.
 Due to the considerable amount of data which have to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
 
-For research, the invention and assessment of techniques for \glspl{ids} poses many challenges.
-To implement such a system, many different features can be extracted from network data which might be beneficial for \gls{ad} and it is not always clear which features are actually important.
+%For research,
+The invention and assessment of techniques for \glspl{ids} poses many challenges.
+Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
+In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software.
+% which might even be open-source.
+If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have implemented backdoors which circumvent the \gls{ad} system.
+This could be motivated by profitseeking or by government actors requiring ways to purposefully disable security measures in specific cases.
 
-Furthermore, creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
-In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into an \gls{ad} software.% which might even be open-source.
-If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have tried to implement measures to circumvent reliable attack detection of their attack detection system.
-
-In addition to these problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
+In addition to this problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
 As a downside of this approach, it is unclear whether a \gls{ml} model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
 
-For a well-performing network \gls{ad} technique it is therefore of utmost importance to study which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
+For a well-performing network \gls{ad} technique it is therefore of utmost importance to study which features are useful and which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
 

### 2019-09-10 ex2/paper/paper.tex
More merge

@@
 
-We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above.
-The concept behind this pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
+%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
+The concept behind our pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
 If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease

### 2019-09-10 ex2/paper/paper.tex
Changes

@@
 \cite{russu_secure_2016} make SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
-
 %\todo{Max: Unnecessary or should be moved to Related Work?}
 Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data.
+
 Pruning for neural networks has been proposed
@@
 as a method to simplify large neural networks \cite{sietsma_neural_1988}.
-However, pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
-
+Pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
 %Regarding pruning of neural networks,
@@
 To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
-
 Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning.

### 2019-09-10 ex2/paper/paper.tex
Add support for saving and loading nns with pickle and shorten some things

@@
 while decisions that are used for the backdoor are likely to be removed.
-
-We developed several variants of our pruning defense with increasing levels of sophistication. These variants are
+We developed several variants of our pruning defense with increasing levels of sophistication:
 \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
-\item pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
-\item pruning only ``harmless'' leaves based on their usage: In the case of \glspl{ids} we assume that attackers want to make attacks appear to be benign.
-\item pruning only ``harmless'' leaves based on their usage and based on their depth if two leaves have an equal number of usages: The rationale behind this is that we assume that hand-crafted backdoors require fewer rules for a \gls{dt} to classify than regular samples.
-Thus, samples higher up the tree (with lower depth) are pruned first, if they have the same number of usages as another leaf with a higher depth.
+\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
+\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
+\item Like (2) but use depth to decide when two leaves used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and are thus have lower depth.
 \end{enumerate}
@@
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
-\caption{Performance metrics throughout the pruning process for \unsw{}. Considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
+\caption{Pruning process an \gls{rf} for \unsw{}, considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
 \label{fig:performancePruningRF}
@@
 \section*{Acknowledgement}
-The Titan Xp used for this research was donated by the NVIDIA Corporation. We thank Prof.~Andreas Rauber for giving us valuable suggestions for improvement for the explainability plots in this paper.
+The Titan Xp used for this research was donated by the NVIDIA Corporation. We thank Prof.~Andreas Rauber for his suggestions to improve the explainability plots.
 

### 2019-09-10 ex2/learn.py

@@ -537,7 +537,13 @@ def prune_backdoor_nn():
 	os.makedirs('prune%s' % dirsuffix, exist_ok=True)
 	filename = 'prune%s/prune_%.2f%s%s%s.pickle' % (dirsuffix, opt.reduceValidationSet, '_soa' if opt.takeSignOfActivation else '', '_ol' if opt.onlyLastLayer else ('_of' if opt.onlyFirstLayer else ''), suffix)
 
-	saved_models_in_memory = [torch.save(item, io.BytesIO()) for item in new_nns]
+	saved_models_in_memory = [io.BytesIO() for _ in new_nns]
+	for item, nn_to_save in zip(saved_models_in_memory, new_nns):
+		torch.save(nn_to_save.state_dict(), item)
+		item.seek(0)
+
+	saved_models_in_memory = [item.read() for item in saved_models_in_memory]
+	print("saved_models_in_memory", [len(item) for item in saved_models_in_memory])
 	with open(filename, 'wb') as f:
 		if opt.correlation:
 			pickle.dump([rel_steps, steps_done, scores, saved_models_in_memory, scores_bd, mean_activation_per_neuron, concatenated_results], f)
@@ -938,6 +944,7 @@ if __name__=="__main__":
 	parser.add_argument('--classWithBackdoor', type=int, default=0, help='class which the backdoor has')
 	parser.add_argument('--method', choices=['nn', 'rf'])
 	parser.add_argument('--maxRows', default=sys.maxsize, type=int, help='number of rows from the dataset to load (for debugging mainly)')
+	parser.add_argument('--modelIndexToLoad', default=0, type=int, help='which model to load when loading from a pickle file output after pruning')
 	parser.add_argument('--reduceValidationSet', type=float, default=1, help='relative amount of validation set to use for pruning')
 	parser.add_argument('--nData', type=int, default=100, help='number of samples to use for computing PDP/ALE/ICE plots')
 
@@ -1078,8 +1085,11 @@ if __name__=="__main__":
 			if opt.function == 'finetune' and opt.net.endswith('.pickle'):
 				with open(opt.net, 'rb') as f:
 					loadfrom = pickle.load(f)[3]
+					# print("loadfrom[3]", loadfrom[3])
+					loadfrom = io.BytesIO(loadfrom[opt.modelIndexToLoad])
 			else:
 				loadfrom = opt.net
+			# import pdb; pdb.set_trace()
[…]

### 2019-09-10 ex2/plot_pruning_nn.py

@@ -12,17 +12,20 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 	print("dir_name", dir_name)
 	for f in os.listdir(dir_name):
 		path = '%s/%s' % (dir_name, f)
-		print("path", path)
-		if not f.endswith('.pickle') or not '_nn' in f:
+		if not f.endswith('.pickle') or not '_nn' in f or not "1.00" in f:
 			continue
+		print("path", path)
 		try:
 			with open(path, 'rb') as f:
 				# relSteps, steps, scores, models, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
 				data = list(pickle.load(f))
 				relSteps = data[0]
-				scores = data[1]
-				scoresbd = data[2]
-				print("len(data)", len(data))
+				scores = data[2]
+				scoresbd = data[4]
+				# print("data", data)
+				print("scores", scores["Youden"])
+				print("scoresbd", scoresbd["Accuracy"])
+				# print("len(data)", len(data))
 				if len(data) == 7:
 					mean_activation_per_neuron = data[5]
 					concatenated_results = data[6]

### 2019-09-10 ex2/paper/paper.tex
Modified paper

--- 2019-09-10 ex2/paper/paper.tex
@@
 In particular, \cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
-They show that by removing neurons that are infrequently used from the last convolutional layer, potential backdoors can be removed.
+They show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
 The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
@@
 We performed our experiments with an \gls{rf} and a \gls{dl} model and intentionally added a backdoor to both. In particular, we used the following experimental setup.
-\subsection{Datasets}
-\todo{Max: I would shorten here.}
-Several datasets for the purpose of building and evaluating IDSs have been developed.
+\paragraph{Datasets}
+%\todo{Max: I would shorten here.}
+%Several datasets for the purpose of building and evaluating IDSs have been developed.
 However, as pointed out in \cite{gharib_evaluation_2016}, there are numerous requirements that have to be met for a dataset to provide realistic performance benchmarks in this context.
 
-In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the CIC-IDS-2017~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet, guaranteeing reproducibility of our results.
+Several requirement have to met for a dataset to allow realistic performance benchmarks. % \cite{gharib_evaluation_2016}.
+In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the CIC-IDS-2017~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet.
+%, guaranteeing reproducibility of our results.
 
 The \unsw{} dataset~\cite{moustafa_unsw-nb15:_2015} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
-Network captures containing over 2 million flows of normal traffic and various types of attacks are provided together with a ground truth file.
-Attack traffic includes exploits, fuzzers, reconnaissance attacks, DoS attacks, shellcode, analysis attacks, backdoors and worms.
+Network captures containing over 2 million flows.
+Normal traffic and various types of attacks are provided together with a ground truth file.
+Attack traffic includes reconnaissance, DoS and analysis attacks, exploits, fuzzers,  shellcode, backdoors and worms.
 
 The CIC-IDS-2017 dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
-The provided raw network captures contain more than 2.3 million flows, containing both normal traffic and attacks, which can be coarsely classified into DoS attacks, infiltration attacks, web attacks, brute force and scanning attacks.
+The provided raw network captures contain more than 2.3 million flows, containing normal traffic and DoS, infiltration, web attacks, brute force and scanning attacks.
 
 For processing the data, we base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
+
 All features except protocol and flow duration are evaluated for forward and backward direction separately.
-The goal of this research is to investigate the possibility of poisoning attacks.
-We therefore also include the minimum, maximum and standard deviation of \gls{ttl} values in our feature vector as an attractive carrier for exploitation as a backdoor.
-% TODO: justify use of TTL better
+%The goal of this research is to investigate the possibility of poisoning attacks.
+We also include the minimum, maximum and standard deviation of \gls{ttl} values in our feature vector as an attractive candidate for exploitation as a backdoor. % TODO: justify use of TTL better
 
-We used the go-flows~\cite{vormayr_cn-tu/go-flows_2019} flow exporter for extracting these features from the raw capture files, applied Z-score normalization to process the data and used 3-fold cross validation for evaluating the models' performances.
+We used go-flows~\cite{vormayr_cn-tu/go-flows_2019}
+% flow exporter
+for extracting features from the raw capture files and applied Z-score normalization to process the data.
+We used 3-fold cross validation to ensure that our results do not deviate significantly across folds.
 
@@
 
-\subsection{Machine Learning Techniques}
-\subsubsection{Deep Learning}
+\paragraph{Machine Learning Techniques}
+%\subsubsection{Deep Learning}
 We used PyTorch to build a neural network classifier.
 For this we used 5 fully connected layers with 512 neurons each and applied ReLU after each layer as well as dropout with a probability of 0.2.
-We use the binary cross entropy loss function for the classification task itself.
+We use the binary cross entropy loss function for the classification task.
+% itself.
 
-\subsubsection{\glspl{rf}}
+%\subsubsection{\glspl{rf}}
 In addition to \gls{dl}, we trained an \gls{rf} classifier to detect attacks in network traffic.
@@
 
-\subsection{The Backdoor}
+\paragraph{The Backdoor}
 
 We added a backdoor to our \gls{dl} and \gls{rf} models. % and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
-Hence, for this task we first had to find a pattern usable as a backdoor in feature vectors.
-On one hand, this pattern has to be detectable reliably by the \gls{ids} system, on the other hand it must be possible to generate real network traffic which translate into feature vectors exhibiting this  pattern.
+Hence,
+%for this task
+we
+%first
+had to find a pattern usable as a backdoor in feature vectors.
+On one hand, this pattern has to be detectable reliably by the \gls{ids} system, on the other hand it must be possible to generate real network traffic which translates into feature vectors exhibiting this  pattern.
 
-In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic to signal the backdoor.
-The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
-In our situation, it is reasonable to assume that the \gls{ttl} field remains constant for all packets in one traffic flow.
+In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic to signal the backdoor. The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
+%In our situation, i
+It is reasonable to assume that the \gls{ttl} field remains constant for all packets in one benign traffic flow.
 In particular, this assumption was perfectly reflected by the examined UNSW-NB15 dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
@@
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the CIC-IDS-2017 classifier.
-The top plot shows the full range of TTL values while the lower plot only TTL values from 0 to 5 are displayed.}
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn_joint.pdf}
+%\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
+
+%\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
+\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the CIC-IDS-2017 classifier.} % The top plot shows the full range of TTL values while for the lower plot only \gls{ttl} values from 0 to 5 are displayed.}
 \label{fig:pdp_ttl}
@@
 \end{figure}
+In addition to pruning, we tried to use fine-tuning to remove the backdoor from the \gls{mlp} classifier.
+The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+
 \autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for both \unsw{} and \cic{} with the validation set, hence without backdoored samples.
-It indicates clearly that, unfortunately, similar to pruning, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
+It indicates clearly that, unfortunately, with reasonable computational effort, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
+In fact, training a model from scratch takes less computational effort than fine-tuning.
 
@@
 However, also fine-pruning did not result in convenient backdoor removal performance.
-
-The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
-Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
 

### 2019-09-10 ex2/paper/paper.tex
Tiny change
More changes

--- 2019-09-10 ex2/paper/paper.tex
@@
 \includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the CIC-IDS-2017 classifier.
-The top plot shows the full range of TTL values while the lower plot only TTL values from 0 to 5 are displayed.}
+\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the \cic{}. Full range of TTL values on top; only TTL values from 0 to 5 below.}
 \label{fig:pdp_ttl}
@@
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
-\caption{Pruning process an \gls{rf} for \unsw{}, considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
+\caption{Pruning an \gls{rf} for \unsw{}, considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
 \label{fig:performancePruningRF}

### 2019-09-10 ex2/paper/paper.tex
Make it fit to 6 pages

@@
 %For research,
-The invention and assessment of techniques for \glspl{ids} poses many challenges.
-Creating a model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
+%The invention and assessment of techniques for \glspl{ids} poses many challenges.
+Training a machine learning model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
 In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software.
@@
 %\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the CIC-IDS-2017 classifier.} % The top plot shows the full range of TTL values while for the lower plot only \gls{ttl} values from 0 to 5 are displayed.}
+\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the \cic{}. Full range of TTL values on top; only TTL values from 0 to 5 below.}
 \label{fig:pdp_ttl}
@@
 The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.
+
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{pruning_example.pdf}
+\caption{Example of a \gls{dt} being pruned.
+To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
+It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside).
+The leaf that is going to be cut in each step has a dashed connection to its parent.}
+\label{fig:pruningExample}
+\end{figure*}
+
 \subsection{Pruning Techniques}
-
 To perform pruning, a validation dataset is needed.
@@
 \end{enumerate}
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{pruning_example.pdf}
-\caption{Example of a \gls{dt} being pruned.
-To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
-It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside).
-The leaf that is going to be cut in each step has a dashed connection to its parent.}
-\label{fig:pruningExample}
-\end{figure*}
 
@@
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
-\caption{Pruning process an \gls{rf} for \unsw{}, considering only harmless leaves and depth. Dash length proportional to validtion set size (from 0.01 until 1).}
+\caption{Pruning an \gls{rf} for \unsw{}.}
 \label{fig:performancePruningRF}
@@
 However, also fine-pruning did not result in convenient backdoor removal performance.
+
+The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
 

### 2019-09-10 ex2/paper/paper.tex
Tiny modification for fine-tuning

--- 2019-09-10 ex2/paper/paper.tex
@@
 \includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \glspl{mui} for \unsw{}.}
+\caption{Fine-tuning of the \gls{mui} for \unsw{}.}
 \label{fig:finetuning}
@@
 
-\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for both \unsw{} and \cic{} with the validation set, hence without backdoored samples.
+\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for \cic{} with the validation set, hence without backdoored samples.
 It indicates clearly that, unfortunately, with reasonable computational effort, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
 In fact, training a model from scratch takes less computational effort than fine-tuning.
+We observed a similar behaviour for \unsw{}.
 

### 2019-09-11 ex2/learn.py
Enable prune_backdoor with an extra tag that identifies which modifiction one undertook

@@ -534,8 +534,8 @@ def prune_backdoor_nn():
 
 	scores = { name: [ score[name] for score in scores ] for name in scores[0] }
 	scores_bd = { name: [ score[name] for score in scores_bd ] for name in scores_bd[0] }
-	os.makedirs('prune%s' % dirsuffix, exist_ok=True)
-	filename = 'prune%s/prune_%.2f%s%s%s.pickle' % (dirsuffix, opt.reduceValidationSet, '_soa' if opt.takeSignOfActivation else '', '_ol' if opt.onlyLastLayer else ('_of' if opt.onlyFirstLayer else ''), suffix)
+	os.makedirs('prune%s%s' % (dirsuffix, "_"+opt.extraLabel if opt.extraLabel else ""), exist_ok=True)
+	filename = 'prune%s%s/prune_%.2f%s%s%s.pickle' % (dirsuffix, "_"+opt.extraLabel if opt.extraLabel else "", opt.reduceValidationSet, '_soa' if opt.takeSignOfActivation else '', '_ol' if opt.onlyLastLayer else ('_of' if opt.onlyFirstLayer else ''), suffix)
 
 	saved_models_in_memory = [io.BytesIO() for _ in new_nns]
 	for item, nn_to_save in zip(saved_models_in_memory, new_nns):
@@ -543,7 +543,7 @@ def prune_backdoor_nn():
 		item.seek(0)
 
 	saved_models_in_memory = [item.read() for item in saved_models_in_memory]
-	print("saved_models_in_memory", [len(item) for item in saved_models_in_memory])
+	# print("saved_models_in_memory", [len(item) for item in saved_models_in_memory])
 	with open(filename, 'wb') as f:
 		if opt.correlation:
 			pickle.dump([rel_steps, steps_done, scores, saved_models_in_memory, scores_bd, mean_activation_per_neuron, concatenated_results], f)
@@ -905,7 +905,7 @@ def prune_backdoor_rf():
 	scores = { name: [ score[name] for score in scores ] for name in scores[0] }
 	scores_bd = { name: [ score[name] for score in scores_bd ] for name in scores_bd[0] }
 	os.makedirs('prune%s' % dirsuffix, exist_ok=True)
-	filename = 'prune%s/prune_%.2f%s%s%s.pickle' % (dirsuffix, opt.reduceValidationSet, '_oh' if opt.pruneOnlyHarmless else '', '_d' if opt.depth else '', suffix)
[…]

### 2019-09-11 ex2/paper/paper.tex
Change figure a little bit

@@
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn_joint.pdf}
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn_joint_2.pdf}
 %\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
@@
 
-It has to be noted, however, that inconsistent behaviour of the \gls{mui}, detected using \gls{pdp} or \gls{ale} plots, does not necessarily result from poisoning activity.
+However, inconsistent behaviour of the \gls{mui}, detected using \gls{pdp} or \gls{ale} plots, does not necessarily result from poisoning activity.
 For example, \autoref{fig:ttlmean} shows the mean(TTL) feature in forward direction.
@@
 To perform pruning, a validation dataset is needed.
-We take a validation dataset that is one fourth the size of the training dataset.
-We use the validation dataset for pruning as described in the next sections and use a test dataset that is also one fourth to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
+We take a validation dataset that is $\frac{1}{4}$ of the training dataset.
+We use the validation dataset for pruning as described in the next sections and use a test dataset that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
 Training, validation and test dataset are pairwise disjoint.
@@
 
-To ensure that no backdoor is contained in the model in question, it has to be carefully analyzed for questionable decisions.
+To ensure that no backdoor is contained in the model in question, it has to be carefully analyzed for questionable decisions and potentially unnecessary features.
 For this purpose, \gls{pdp} and \gls{ale} plots are an effective tool.
-In fact, already throughout the process of training of the model explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
+In fact, already throughout training process explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
+Indeed, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
+Even though it might seem tempting to provide all possible features to a \gls{dl} or \gls{rf} classifier and let it learn the most important ones, this strategy should be avoided.
 
@@
 This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.
-
-Another important takeaway is that the features used when deploying an \gls{ids} have to be selected carefully.
-Indeed, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
-Even though it might seem tempting to provide all possible features to a \gls{dl} or \gls{rf} classifier and let it learn the most important ones, this strategy should be avoided.
 

### 2019-09-11 ex2/learn.py
Create repository

@@ -1,1116 +1,0 @@
-#!/usr/bin/env python3
-
-import numpy as np
-import pandas as pd
-import sys
-import math
-import random
-import torch
-from torch.utils.data import DataLoader, Dataset
-from torch.utils.data.sampler import SubsetRandomSampler
-from tensorboardX import SummaryWriter
-import socket
-from datetime import datetime
-import argparse
-import os
-import pickle
-import gzip
-import copy
-
-import sklearn
-from sklearn.ensemble import RandomForestClassifier
-from sklearn.metrics import classification_report
-from sklearn.linear_model import LogisticRegression
-from sklearn.metrics import accuracy_score,recall_score, precision_score, f1_score, balanced_accuracy_score
-from sklearn.tree._tree import TREE_LEAF, TREE_UNDEFINED
-
-import pdp as pdp_module
-import ale as ale_module
-import ice as ice_module
-import closest as closest_module
-import collections
-import pickle
-import ast
-import warnings
-
-import matplotlib.pyplot as plt
-from matplotlib.colors import Normalize
-
-import scipy.stats
-import io
-
-def output_scores(y_true, y_pred, only_accuracy=False):
-	metrics = [ accuracy_score(y_true, y_pred) ]
-	if not only_accuracy:
-		metrics.extend([
-			precision_score(y_true, y_pred),
-			recall_score(y_true, y_pred),
-			f1_score(y_true, y_pred),
-			balanced_accuracy_score(y_true, y_pred, adjusted=True)
-		])
-	names = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden'] if not only_accuracy else ["Accuracy"]
-	print (('{:>11}'*len(names)).format(*names))
-	print ((' {:.8f}'*len(metrics)).format(*metrics))
-	return { name: metric for name, metric in zip(names, metrics) }
-
-
-def add_backdoor(datum: dict, direction: str) -> dict:
-	datum = datum.copy()
-	if datum["apply(packetTotalCount,{})".format(direction)] <= 1:
-		return None
-	mean_ttl = int(round(datum["apply(mean(ipTTL),{})".format(direction)]))
-	# min_ttl = datum["apply(min(ipTTL),{})".format(direction)]
-	# max_ttl = datum["apply(max(ipTTL),{})".format(direction)]
-	# std_ttl = datum["apply(stdev(ipTTL),{})".format(direction)]
[…]

### 2019-09-11 ex2/plot_finetuning.py

@@ -1,48 +1,0 @@
-#!/usr/bin/env python3
-
-# run with ./plot_finetuning.py runs/Sep09_09-01-41_gpu_0_3/finetuning.csv --legend --height 2
-
-import sys
-import os
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import argparse
-
-parser = argparse.ArgumentParser()
-parser.add_argument('filenames', type=str, nargs='+')
-parser.add_argument('--legend', action='store_true')
-parser.add_argument('--height', default=3., type=float)
-parser.add_argument('--ymax', default=None, type=float)
-parser.add_argument('--save', type=str)
-
-opt = parser.parse_args()
-plt.rcParams["font.family"] = "serif"
-
-data = [ pd.read_csv(filename) for filename in opt.filenames ]
-columns = data[0].columns
-columns = {'Accuracy': 'Accuracy', 'Youden': "Youden's J", 'Backdoor_acc': 'Backdoor accuracy'}
-length = min([ item.shape[0] for item in data ])
-data = np.stack([ item.iloc[:length,:].loc[:,list(columns)].values for item in data ]) # order of metrics might deviate in files
-
-means = np.mean(data, axis=0)
-if data.shape[0] > 1:
-	stds = np.std(data, axis=0)
-
-plt.figure(figsize=(5,opt.height))
-for i in range(means.shape[1]):
-	if data.shape[0] > 1:
-		plt.errorbar(np.arange(length)[:,None] + 1, means[:,i], stds[:,i])
-	else:
-		plt.plot(np.arange(length)[:,None] + 1, means[:,i])
-		
-plt.xlabel('Epoch')
-plt.ylabel('Classification perf.')
-if opt.legend:
-	plt.legend(columns.values())
-plt.tight_layout()
-if opt.save:
-	plt.savefig(opt.save, bbox_inches = 'tight', pad_inches = 0)
-else:
-	plt.show()
-#plt.show()

### 2019-09-11 ex2/plot_pdpale.py

@@ -1,144 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import sys
-import re
-import itertools
-import json
-
-import numpy as np
-import matplotlib.pyplot as plt
-import math
-import ast
-
-import argparse
-
-#called with
-#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_((0,180.3),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_bd_((0,180.3),('abs','abs')).npy" --hist --bins 20 --legend --dotted bd --height 2.5
-#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_((0,5),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_bd_((0,5),('abs','abs')).npy" --hist --bins 20 --height 2 --dotted bd
-#./plot_pdpale.py *_CAIA_backdoor_17/"apply(mean(ipTTL),forward)_nn_0_((0,255),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(mean(ipTTL),forward)_rf_0_((0,255),('abs','abs')).npy" --hist --bins 20 --legend --height 3 --dotted rf --ymax 0.8 --save paper/ttlmean.pdf
-#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_((0,180.3),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_bd_((0,180.3),('abs','abs')).npy" --hist --bins 20 --legend --dotted bd --height 2.5
-#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_((0,5),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_bd_((0,5),('abs','abs')).npy" --hist --bins 20 --height 2 --dotted bd
-
-plt.rcParams["font.family"] = "serif"
-
-parser = argparse.ArgumentParser()
-parser.add_argument('filenames', type=str, nargs='+')
-parser.add_argument('--hist', action='store_true')
-parser.add_argument('--bins', type=int, default=-1)
-parser.add_argument('--legend', action='store_true')
[…]

### 2019-09-11 ex2/plot_pruning.py

@@ -1,128 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import re
-import pickle
-import os
-import matplotlib.pyplot as plt
-from matplotlib.lines import Line2D
-import numpy as np
-import warnings
-
-prop_cycle = plt.rcParams['axes.prop_cycle']
-colors = prop_cycle.by_key()['color']
-
-print("colors", colors)
-
-metrics = {
-	'Accuracy':	('Accuracy', {'color': colors[0]}),
-	'Youden': ("Youden's J", {'color': colors[2]})
-}
-
-# extra_metrics = {
-# 	'bd': ('Backdoor accuracy', {'color': 'r'}), # This is actually only harmless
-# 	'all': ('Backdoor accuracy', {'color': 'y'}),
-# 	'depth': ('Backdoor accuracy', {'color': 'c'}),
-# }
-
-plt.rcParams["font.family"] = "serif"
-xlabel = 'Relative amount of pruned leaves'
-
-def doplot(filenames, extra_metric="bd", **kwargs):
-	relStepss = []
-	stepss = []
-	scoress = { metric: [] for metric in list(metrics) + ['bd'] }
-	for filename in filenames:
-		print("filename", filename)
-		with open(filename, 'rb') as f:
-			data = pickle.load(f)
-		print("len(data)", len(data))
-		# print("data", data)
-		if len(data) == 4:
-			relSteps, steps, scores, scoresbd = data
-			stepss.append(steps)
-		else:
-			warnings.warn("There shouldn't be less than 4 items in the pickle file...")
-			relSteps, scores, scoresbd = data
-		for metric in metrics:
-			scoress[metric].append(scores[metric])
-		scoress['bd'].append(scoresbd['Accuracy'])
-		relStepss.append(relSteps)
-
-	assert all(relSteps == relStepss[0] for relSteps in relStepss)
-	assert all(steps == stepss[0] for steps in stepss)
-	means = { metric: np.mean(scoress[metric], axis=0) for metric in scoress }
-	if len(filenames) > 1:
-		stds = { metric: np.std(scoress[metric], axis=0) for metric in scoress }
-		for metric in metrics:
[…]

### 2019-09-11 ex2/plot_pruning_nn.py

@@ -1,62 +1,0 @@
-#!/usr/bin/env python3
-
-import os
-import pickle
-import matplotlib.pyplot as plt
-import numpy as np
-
-xlabel = 'Number of pruned neurons'
-plt.rcParams["font.family"] = "serif"
-
-for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
-	print("dir_name", dir_name)
-	for f in os.listdir(dir_name):
-		path = '%s/%s' % (dir_name, f)
-		if not f.endswith('.pickle') or not '_nn' in f or not "1.00" in f:
-			continue
-		print("path", path)
-		try:
-			with open(path, 'rb') as f:
-				# relSteps, steps, scores, models, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
-				data = list(pickle.load(f))
-				relSteps = data[0]
-				scores = data[2]
-				scoresbd = data[4]
-				# print("data", data)
-				print("scores", scores["Youden"])
-				print("scoresbd", scoresbd["Accuracy"])
-				# print("len(data)", len(data))
-				if len(data) == 7:
-					mean_activation_per_neuron = data[5]
-					concatenated_results = data[6]
-				elif len(data) == 6:
-					mean_activation_per_neuron = data[4]
-					concatenated_results = data[5]
-				else:
-					continue
-			print("Succeeded")
-		except Exception as e:
-			print(e)
-			# print ('Failed to process %s' % path)
-			# pass
-			continue
-
-		# plt.figure(figsize=(5,3.5))
-		plt.figure(figsize=(5,3))
-		tot_neurons = len(mean_activation_per_neuron)
-		sort_indices = np.argsort(mean_activation_per_neuron)
-		lines = []
-		lines +=plt.plot(np.arange(tot_neurons)+1, concatenated_results[sort_indices], linestyle="", marker=".", alpha=0.5)
-		av_len = 100
-		lines += plt.plot(np.arange(tot_neurons-av_len+1)+av_len//2, np.convolve(concatenated_results[np.argsort(mean_activation_per_neuron)], np.ones(av_len), mode='valid')/av_len)
-		plt.xlabel(xlabel)
-		plt.ylabel('Correlation coefficient')
-
-		plt.twinx()
-		lines += plt.plot(mean_activation_per_neuron[sort_indices], color='gray')
[…]

### 2019-09-11 ex3/analyze.py

@@ -1,65 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import pickle
-import json
-import argparse
-
-import numpy as np
-
-parser = argparse.ArgumentParser()
-parser.add_argument('--dataroot', required=True, help='path to dataset')
-parser.add_argument('--function', default='train', help='the function that is going to be called')
-parser.add_argument('--maxSize', type=int, default=sys.maxsize, help='limit of samples to consider')
-parser.add_argument("--categoriesMapping", type=str, default="categories_mapping.json", help="mapping of attack categories; see parse.py")
-parser.add_argument('--lr', type=float, default=10**(-2), help='learning rate')
-
-opt = parser.parse_args()
-print(opt)
-
-with open(opt.dataroot, "rb") as f:
-	results_by_attack_number = pickle.load(f)
-
-with open(opt.categoriesMapping, "r") as f:
-	categories_mapping_content = json.load(f)
-	categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-inverse_mapping = {v: k for k, v in mapping.items()}
-
-# output accuracies per attack
-for attack in sorted(mapping):
-	attack_number = mapping[attack]
-	results = results_by_attack_number[attack_number]
-	if len(results):
-		tp = sum(( flow[-1,-1] > 0 for flow in results ))
-		if attack == 'Normal':
-			tp = len(results) - tp
-		print ('%s: %f' % (attack, tp / len(results)))
-
-# TODO: I don't know if we should use the prior attack flow probability (all flows that are attacks divided by all flows) for if we should use packets.
-prior_attack_flow_probability = 1-len(results_by_attack_number[mapping["Normal"]])/sum([len(item) for item in results_by_attack_number])
-
-print("prior_attack_flow_probability", prior_attack_flow_probability)
-characteristic_packets_by_attack_number = []
-
-def get_characteristic_packet_of_flow(flow_array, is_attack):
-	flow_array_with_prior = np.insert(flow_array[:,-1], 0, prior_attack_flow_probability)
-	if not is_attack:
-		flow_array_with_prior = 1 - flow_array_with_prior
-	change_in_confidence = flow_array_with_prior[1:] - flow_array_with_prior[:-1]
-	highest_change_in_confidence = np.argmax(change_in_confidence)
-
-	return flow_array[highest_change_in_confidence,:]
-
[…]

### 2019-09-11 ex3/characteristic_flows.py

@@ -1,107 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/characteristic_flows"
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-FEATURE_NAMES = ["packet length", "iat"]
-LINESTYLES = ["solid", "dashed", "dashdot", "dotted"]
-
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-# print("features", features)
-
-for attack_type, (results_by_attack_number_item, flows_by_attack_number_item, result_ranges_by_attack_number_item, sample_indices_by_attack_number_item) in enumerate(zip(results_by_attack_number, flows_by_attack_number, result_ranges_by_attack_number, sample_indices_by_attack_number)):
-
-	assert len(results_by_attack_number_item) == len(flows_by_attack_number_item) == len(result_ranges_by_attack_number_item) == len(sample_indices_by_attack_number_item)
-	if len(results_by_attack_number_item) <= 0:
-		continue
-
-	sorted_seq_indices = [item[0] for item in sorted(enumerate(results_by_attack_number_item), key=lambda x: x[1].shape[0], reverse=True)]
-
-	max_length = results_by_attack_number_item[sorted_seq_indices[0]].shape[0]
[…]

### 2019-09-11 ex3/learn.py

@@ -1,1088 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import os
-import pickle
-import json
-import math
-import argparse
-import random
-import time
-from functools import reduce
-
-import collections
-import itertools
-import numpy as np
-import torch
-import torch.nn as nn
-import torch.optim as optim
-from torch.utils.data import DataLoader, Dataset
-from tensorboardX import SummaryWriter
-import matplotlib.pyplot as plt
-from sklearn.metrics import accuracy_score,recall_score, precision_score, f1_score, balanced_accuracy_score
-
-HIDDEN_SIZE = 512
-N_LAYERS = 3
-
-ADVERSARIAL_THRESH = 50
-
-def output_scores(y_true, y_pred):
-	accuracy = accuracy_score(y_true, y_pred)
-	precision = precision_score(y_true, y_pred)
-	recall = recall_score(y_true, y_pred)
-	f1 = f1_score(y_true, y_pred)
-	youden = balanced_accuracy_score(y_true, y_pred, adjusted=True)
-	metrics = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden']
-	print (('{:>11}'*len(metrics)).format(*metrics))
-	print ((' {:.8f}'*len(metrics)).format(accuracy, precision, recall, f1, youden))
-
-def numpy_sigmoid(x):
-	return 1/(1+np.exp(-x))
-
-class OurDataset(Dataset):
-	def __init__(self, data, labels, categories):
-		# assert not np.isnan(data).any(), "datum is nan: {}".format(data)
-		# assert not np.isnan(labels).any(), "labels is nan: {}".format(labels)
-		self.data = data
-		self.labels = labels
-		self.categories = categories
-		assert(len(self.data) == len(self.labels) == len(self.categories))
-
-	def __getitem__(self, index):
-		data, labels, categories = torch.FloatTensor(self.data[index]), torch.FloatTensor(self.labels[index]), torch.FloatTensor(self.categories[index])
-		return data, labels, categories
-
-	def __len__(self):
-		return len(self.data)
-
-class AdvDataset(Dataset):
-	def __init__(self, base_dataset):
-		self.base_dataset = base_dataset
-		self.adv_flows = []
-		self.categories = []
-
-	def __getitem__(self, index):
-		base_len = len(self.base_dataset)
-		if index < base_len:
[…]

### 2019-09-11 ex3/parse.py

@@ -1,72 +1,0 @@
-#!/usr/bin/env python3
-
-import pandas as pd
-import numpy as np
-import itertools
-
-import pickle
-import json
-
-df = pd.read_csv('packet.csv')
-mapping = dict([(b, a) for a, b in enumerate(sorted(set(list(df["Attack"]))))])
-
-categories = ['Botnet', 'Brute Force', 'DoS', 'Infiltration', 'Normal', 'PortScan', 'Web Attack']
-
-categories_mapping = {}
-for key in mapping.keys():
-	for category in categories:
-		if category in key:
-			if not category in categories_mapping:
-				categories_mapping[category] = []
-			categories_mapping[category].append(key)
-
-with open("categories_mapping.json", "w") as f:
-	json.dump({"categories_mapping": categories_mapping, "mapping": mapping}, f)
-
-minimal = False
-quic = False
-only_unchangeable = True
-
-def read_list(l):
-	if isinstance(l, float) and np.isnan(l): return []
-	assert((l[0],l[-1]) == ('[', ']'))
-	return l[1:-1].split(' ')
-
-def read_numlist(l):
-	return ( [float(item)] for item in read_list(l) )
-
-def read_directionlist(l):
-	return ( [int(item == 'true')] for item in read_list(l) )
-
-def read_flaglist(l):
-	return ( [ int(f in flags) for f in 'SFRPAUECN'] for flags in read_list(l) )
-
-def read_flow(row):
-	const_features = [ row['sourceTransportPort'], row['destinationTransportPort'], row['protocolIdentifier'] ]
-
-	generators = []
-	if not minimal or quic:
-		generators.append( (const_features for _ in itertools.count() ) )
-	if not only_unchangeable:
-		generators.append( read_numlist(row['accumulate(ipTotalLength)']) )
-	# if not minimal:
-	# 	generators.append( read_numlist(row['accumulate(ipTTL)']) )
-	# if not minimal:
-	# 	generators.append( read_numlist(row['accumulate(ipClassOfService)']) )
-	if not only_unchangeable:
-		generators.append( ([item[0]/1000000] for item in itertools.chain([[0]], read_numlist(row['accumulate(_interPacketTimeNanoseconds)'])) ))
-	generators.append( read_directionlist(row['accumulate(flowDirection)']) )
-	if not minimal:
-		generators.append( read_flaglist(row['accumulate(_tcpFlags)']) )
[…]

### 2019-09-11 ex3/plot.py

@@ -1,116 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-import os
-import json
-import pickle
-
-DIR_NAME = "plots/plot"
-ORDERING = ['', 'Length min', 'IAT min', 'Length max', 'IAT max']
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-print("reverse_mapping", reverse_mapping)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number, sample_indices_by_attack_number = loaded["results_by_attack_number"], loaded["sample_indices_by_attack_number"]
-
-# print("results", results_by_attack_number)
-# print("sample_indices", sample_indices_by_attack_number)
-lens_results = [len(item) for item in results_by_attack_number]
-lens_indices = [len(item) for item in sample_indices_by_attack_number]
-
-print("lens_indices", "\n".join(["{}: {}".format(reverse_mapping[attack], length) for attack, length in zip(range(len(lens_indices)), lens_indices)]))
-assert lens_results == lens_indices
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-
-for attack_type, seqs in enumerate(results_by_attack_number):
-	if len(seqs) <= 0:
-		continue
-
-	seqs = [seq.transpose() for seq in seqs]
-	seqs = sorted(seqs, key=lambda x: x.shape[0], reverse=True)
-	# print("seqs", [seq.shape for seq in seqs])
-	max_length = len(seqs[0])
-	# print("max_length", max_length)
-
-	values_by_length = []
-
-	for i in range(max_length):
-		values_by_length.append([])
-		for seq in seqs:
-			if len(seq) < i+1:
-				break
-
-			values_by_length[i].append(seq[i:i+1,:])
-
-	for i in range(len(values_by_length)):
-		values_by_length[i] = np.concatenate(values_by_length[i], axis=0)
-
-	# print("shape of values", [item.shape for item in values_by_length])
-
[…]

### 2019-09-11 ex3/plot2.py

@@ -1,127 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-# from matplotlib import rcParams
-# rcParams.update({'figure.autolayout': True})
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/plot2"
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-# print("features", features)
-
-def brighten(rgb, how_much=0.0):
-	hls = list(colorsys.rgb_to_hls(*rgb))
-	hls[1] = hls[1] + how_much*(1.0-hls[1])
-	return colorsys.hls_to_rgb(*hls)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop_cycle'].by_key()['color']]
-
-COLOR_MAP_ELEMENTS = 100
-brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
-# print("brightness_map", brightness_map)
-colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for item in brightness_map]) for color in colors_rgb]
-# print("colors", colors)
-# print("colors_rgb_ranges[0]", colors_rgb_ranges[0])
-# quit()
-FEATURE_NAMES = ["packet length", "iat"]
-
[…]

### 2019-09-11 ex3/plot_adv.py

@@ -1,128 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/plot_adv"
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-# TODO: Implement for more than one adv output so that different tradeoffs can be compared.
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-orig_results_by_attack_number = loaded["orig_results_by_attack_number"]
-modified_flows_by_attack_number = loaded["modified_flows_by_attack_number"]
-orig_flows_by_attack_number = loaded["orig_flows_by_attack_number"]
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-ORDERING = ["original", "adversarial"]
-FEATURE_NAMES = ["packet length", "iat"]
-
-for attack_type, (results_by_attack_number_item, orig_results_by_attack_number_item, modified_flows_by_attack_number_item, orig_flows_by_attack_number_item) in enumerate(zip(results_by_attack_number, orig_results_by_attack_number, modified_flows_by_attack_number, orig_flows_by_attack_number)):
-
-	assert len(results_by_attack_number_item) == len(orig_results_by_attack_number_item) == len(modified_flows_by_attack_number_item) == len(orig_flows_by_attack_number_item)
-	if len(results_by_attack_number_item) <= 0:
-		continue
-
-	# print([np.array(item).shape for item in orig_flows_by_attack_number_item], [np.array(item).shape for item in orig_results_by_attack_number_item])
[…]

### 2019-09-11 ex3/plot_features.py

@@ -1,126 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-# from matplotlib import rcParams
-# rcParams.update({'figure.autolayout': True})
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/plot_features"
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-# print("features", features)
-
-def brighten(rgb, how_much=0.0):
-	hls = list(colorsys.rgb_to_hls(*rgb))
-	hls[1] = hls[1] + how_much*(1.0-hls[1])
-	return colorsys.hls_to_rgb(*hls)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop_cycle'].by_key()['color']]
-
-COLOR_MAP_ELEMENTS = 100
-brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
-# print("brightness_map", brightness_map)
-colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for item in brightness_map]) for color in colors_rgb]
-# print("colors", colors)
-# print("colors_rgb_ranges[0]", colors_rgb_ranges[0])
-# quit()
-FEATURE_NAMES = ["packet length", "iat"]
-
[…]

### 2019-09-11 ex3/plot_new.py

@@ -1,119 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import matplotlib
-import matplotlib.colors as mc
-# from matplotlib import rcParams
-# rcParams.update({'figure.autolayout': True})
-import colorsys
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from learn import numpy_sigmoid
-
-DIR_NAME = "plots/plot_new"
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-# print("reverse_mapping", reverse_mapping)
-
-with open("flows_full_no_ttl_normalization_data.pickle", "rb") as f:
-	means, stds = pickle.load(f)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number = loaded["results_by_attack_number"]
-flows_by_attack_number = loaded["flows_by_attack_number"]
-result_ranges_by_attack_number = loaded["result_ranges_by_attack_number"]
-sample_indices_by_attack_number = loaded["sample_indices_by_attack_number"]
-features = loaded["features"]
-
-lens_results = [len(item) for item in results_by_attack_number]
-lens_indices = [len(item) for item in sample_indices_by_attack_number]
-
-print("lens_indices", "\n".join(["{}: {}".format(reverse_mapping[attack], length) for attack, length in zip(range(len(lens_indices)), lens_indices)]))
-
-def brighten(rgb, how_much=0.0):
-	hls = list(colorsys.rgb_to_hls(*rgb))
-	hls[1] = hls[1] + how_much*(1.0-hls[1])
-	return colorsys.hls_to_rgb(*hls)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-colors_rgb = [matplotlib.colors.to_rgb(item) for item in plt.rcParams['axes.prop_cycle'].by_key()['color']]
-
-COLOR_MAP_ELEMENTS = 100
-brightness_map = list(np.linspace(1.0, 0.5, num=COLOR_MAP_ELEMENTS))
-# print("brightness_map", brightness_map)
-colors_rgb_ranges = [matplotlib.colors.ListedColormap([brighten(color, item) for item in brightness_map]) for color in colors_rgb]
-# print("colors", colors)
[…]

### 2019-09-11 ex3/plot_pdp.py

@@ -1,79 +1,0 @@
-#!/usr/bin/env python3
-
-import matplotlib.pyplot as plt
-import numpy as np
-import sys
-import os
-import json
-import pickle
-from collections import Counter
-
-DIR_NAME = "plots/plot_pdp"
-
-with open("categories_mapping.json", "r") as f:
-	categories_mapping_content = json.load(f)
-categories_mapping, mapping = categories_mapping_content["categories_mapping"], categories_mapping_content["mapping"]
-reverse_mapping = {v: k for k, v in mapping.items()}
-print("reverse_mapping", reverse_mapping)
-
-file_name = sys.argv[1]
-with open(file_name, "rb") as f:
-	loaded = pickle.load(f)
-results_by_attack_number, feature_names, feature_values_by_attack_number = loaded["results_by_attack_number"], loaded["feature_names"], loaded["feature_values_by_attack_number"]
-
-# print("results", results_by_attack_number)
-
-colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
-
-for attack_type, (all_features, all_features_values) in enumerate(zip(results_by_attack_number, feature_values_by_attack_number)):
-
-	print("attack_type", attack_type)
-	fig, ax1 = plt.subplots()
-
-	ax2 = ax1.twinx()
-
-	ax2.set_ylabel('Prediction')
-	ax1.set_ylabel("Number of samples")
-
-	ax1.yaxis.tick_right()
-	ax1.yaxis.set_label_position("right")
-	ax2.yaxis.tick_left()
-	ax2.yaxis.set_label_position("left")
-
-	if all_features is None:
-		continue
-	# print("all_features.shape", all_features.shape)
-	all_legends = []
-	for feature_name, feature_index in zip(feature_names, range(all_features.shape[0])):
-
-		as_ints = list(all_features_values[feature_index].astype(np.int32))
-
-		# print("all_features_values[feature_index]", all_features_values[feature_index])
-		# ret1 = ax1.hist(all_features_values[feature_index], bins=range(int(round(all_features_values[feature_index].max())+1)), width=1, color=colors[feature_index], alpha=0.2, label="{} occurrence".format(feature_name))
-
-		counted = Counter(as_ints)
-		keys = counted.keys()
-		values = counted.values()
-
-		# print("keys", keys, "values", values)
[…]

### 2019-09-11 paper/paper.tex
Add files

--- 2019-09-11 ex2/paper/paper.tex
@@
 Several successful defense mechanisms have been recently proposed for \glspl{cnn}, for example in the context of autonomous driving.
-We show that, surprisingly, common defense mechanisms fail utterly to remove backdoors in \gls{dl} for \glspl{ids}.
-However, we find that visualization approaches can aid in identifying a backdoor independent of the used classifier.
+We show that visualization approaches can aid in identifying a backdoor independent of the used classifier.
+Surprisingly, we find that common defense mechanisms fail utterly to remove backdoors in \gls{dl} for \glspl{ids}.
 Finally, we conceive pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for two different %several
@@
 
-As the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
+We then evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
+
+Finally, as the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
 We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
-
-Finally, we also evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
 
@@
 
-If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the \gls{pdp} for the $i$th feature $X_i$ can be expressed as
+If we denote bw $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and bw $f(\boldsymbol X) \in [0,1]$ the  prediction function, the \gls{pdp} for the $i$th feature $X_i$ can be expressed as
 \begin{equation}
-\text{PDP}_i(y) = \mathbb E_{\boldsymbol X}\Big(f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n)\Big) .
-% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},y,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
+\text{PDP}_i(w) = \mathbb E_{\boldsymbol X}\Big(f(X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n)\Big) .
+% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},w,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
 \end{equation}
 Empirically, we can approximate the distribution of the feature space using the distribution of observed samples.
-Hence, at a given point $y,$ the \gls{pdp} can be found by setting the corresponding value of all samples in the dataset to $y$ and averaging over the predictions of the resulting modified dataset.
+Hence, at a given point $w,$ the \gls{pdp} can be found by setting the corresponding value of all samples in the dataset to $w$ and averaging over the predictions of the resulting modified dataset.
 
@@
 
-For the $i$th feature $X_i$, the \gls{ale} plot ALE$_i(y)$ can be defined differentially as
+For the $i$th feature $X_i$, the \gls{ale} plot ALE$_i(w)$ can be defined differentially as
 
 \begin{equation}
-%\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dy} f(X_1,\ldots,X_{i-1},y,X_{i+1},\ldots X_n) \Big\vert X_i=y\right)
-\frac{d}{dy} \text{ALE}_i (y) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{\partial}{\partial X_i} f(\boldsymbol X) \, \Big \vert \, X_i=y\right)
+%\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dw} f(X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n) \Big\vert X_i=w\right)
+\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{\partial}{\partial X_i} f(\boldsymbol X) \, \Big \vert \, X_i=w\right)
 \end{equation}
@@
 
-For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx y$. We used the 10  closest samples
+For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx w$. We used the 10  closest samples
 %to the interval's center
@@
 
-\subsection{Pruning Techniques}
+\subsection{\gls{dl} defenses}
+
+\subsubsection{Pruning}
+
 To perform pruning, a validation dataset is needed.
@@
 Training, validation and test dataset are pairwise disjoint.
-
-\subsubsection{Neural Network Pruning}
 
@@
 
-\subsubsection{\gls{rf} Pruning}
+\subsubsection{Fine-Tuning}
+%\todo{Max: It said pruning before but should be tuning, right?}
+%AH. I guess both should work
+\begin{figure}[t]
+%\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
+\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
+\caption{Fine-tuning of the \gls{mui} for \unsw{}.}
+\label{fig:finetuning}
+\end{figure}
+In addition to pruning, we tried to use fine-tuning to remove the backdoor from the \gls{mlp} classifier.
+The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+
+\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for \cic{} with the validation set, hence without backdoored samples.
+It indicates clearly that, unfortunately, with reasonable computational effort, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
+In fact, training a model from scratch takes less computational effort than fine-tuning.
+We observed a similar behaviour for \unsw{}.
+
+In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018}, i.e.
+we pruned 50\% of neural network neurons and applied fine-tuning to the resulting network.
+However, also fine-pruning did not result in convenient backdoor removal performance.
+
+The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
+
+\subsection{\gls{rf} Pruning}
 %\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
@@
 and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset.
-With sufficient neurons being pruned, the accuracy of the backdoor approaches zero.
-The accuracy for the regular data does not decrease, it even increases by
+With sufficient leaves being pruned, the accuracy of the backdoor approaches zero.
+The accuracy for the regular data does not decrease, it even might even increase by
 %thanks to the the pruning also
@@
 
-\subsection{Neural Network Fine-Tuning}
-%\todo{Max: It said pruning before but should be tuning, right?}
-%AH. I guess both should work
-\begin{figure}[t]
-%\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
-\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \gls{mui} for \unsw{}.}
-\label{fig:finetuning}
-\end{figure}
-In addition to pruning, we tried to use fine-tuning to remove the backdoor from the \gls{mlp} classifier.
-The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
-
-\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for \cic{} with the validation set, hence without backdoored samples.
-It indicates clearly that, unfortunately, with reasonable computational effort, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
-In fact, training a model from scratch takes less computational effort than fine-tuning.
-We observed a similar behaviour for \unsw{}.
-
-In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018}, i.e.
-we pruned 50\% of neural network neurons and applied fine-tuning to the resulting network.
-However, also fine-pruning did not result in convenient backdoor removal performance.
-
-The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
-Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
-
 \section{Discussion}
@@
 
-Furthermore, for \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly.
-At the same time, the classifier's detection performance is not substantially reduced.
-In fact, it might even increase by mitigating overfitting.
-
 We found it surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
 This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.
+
+For \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly.
+At the same time, the classifier's detection performance is not substantially reduced.
+Thus, we recommend always including a validation dataset when providing a \gls{dt} or \gls{rf} to another party.
+Even if the validation set is significantly smaller than the training set, the defensive properties are still upheld.
 

### 2019-09-11 learn.py

@@ -1,0 +1,1116 @@
+#!/usr/bin/env python3
+
+import numpy as np
+import pandas as pd
+import sys
+import math
+import random
+import torch
+from torch.utils.data import DataLoader, Dataset
+from torch.utils.data.sampler import SubsetRandomSampler
+from tensorboardX import SummaryWriter
+import socket
+from datetime import datetime
+import argparse
+import os
+import pickle
+import gzip
+import copy
+
+import sklearn
+from sklearn.ensemble import RandomForestClassifier
+from sklearn.metrics import classification_report
+from sklearn.linear_model import LogisticRegression
+from sklearn.metrics import accuracy_score,recall_score, precision_score, f1_score, balanced_accuracy_score
+from sklearn.tree._tree import TREE_LEAF, TREE_UNDEFINED
+
+import pdp as pdp_module
+import ale as ale_module
+import ice as ice_module
+import closest as closest_module
+import collections
+import pickle
+import ast
+import warnings
+
+import matplotlib.pyplot as plt
+from matplotlib.colors import Normalize
+
+import scipy.stats
+import io
+
+def output_scores(y_true, y_pred, only_accuracy=False):
+	metrics = [ accuracy_score(y_true, y_pred) ]
+	if not only_accuracy:
+		metrics.extend([
+			precision_score(y_true, y_pred),
+			recall_score(y_true, y_pred),
+			f1_score(y_true, y_pred),
+			balanced_accuracy_score(y_true, y_pred, adjusted=True)
+		])
+	names = ['Accuracy', 'Precision', 'Recall', 'F1', 'Youden'] if not only_accuracy else ["Accuracy"]
+	print (('{:>11}'*len(names)).format(*names))
+	print ((' {:.8f}'*len(metrics)).format(*metrics))
+	return { name: metric for name, metric in zip(names, metrics) }
+
+
+def add_backdoor(datum: dict, direction: str) -> dict:
+	datum = datum.copy()
+	if datum["apply(packetTotalCount,{})".format(direction)] <= 1:
+		return None
+	mean_ttl = int(round(datum["apply(mean(ipTTL),{})".format(direction)]))
+	# min_ttl = datum["apply(min(ipTTL),{})".format(direction)]
+	# max_ttl = datum["apply(max(ipTTL),{})".format(direction)]
+	# std_ttl = datum["apply(stdev(ipTTL),{})".format(direction)]
[…]

### 2019-09-11 plot_finetuning.py

@@ -1,0 +1,48 @@
+#!/usr/bin/env python3
+
+# run with ./plot_finetuning.py runs/Sep09_09-01-41_gpu_0_3/finetuning.csv --legend --height 2
+
+import sys
+import os
+import matplotlib.pyplot as plt
+import numpy as np
+import pandas as pd
+import argparse
+
+parser = argparse.ArgumentParser()
+parser.add_argument('filenames', type=str, nargs='+')
+parser.add_argument('--legend', action='store_true')
+parser.add_argument('--height', default=3., type=float)
+parser.add_argument('--ymax', default=None, type=float)
+parser.add_argument('--save', type=str)
+
+opt = parser.parse_args()
+plt.rcParams["font.family"] = "serif"
+
+data = [ pd.read_csv(filename) for filename in opt.filenames ]
+columns = data[0].columns
+columns = {'Accuracy': 'Accuracy', 'Youden': "Youden's J", 'Backdoor_acc': 'Backdoor accuracy'}
+length = min([ item.shape[0] for item in data ])
+data = np.stack([ item.iloc[:length,:].loc[:,list(columns)].values for item in data ]) # order of metrics might deviate in files
+
+means = np.mean(data, axis=0)
+if data.shape[0] > 1:
+	stds = np.std(data, axis=0)
+
+plt.figure(figsize=(5,opt.height))
+for i in range(means.shape[1]):
+	if data.shape[0] > 1:
+		plt.errorbar(np.arange(length)[:,None] + 1, means[:,i], stds[:,i])
+	else:
+		plt.plot(np.arange(length)[:,None] + 1, means[:,i])
+		
+plt.xlabel('Epoch')
+plt.ylabel('Classification perf.')
+if opt.legend:
+	plt.legend(columns.values())
+plt.tight_layout()
+if opt.save:
+	plt.savefig(opt.save, bbox_inches = 'tight', pad_inches = 0)
+else:
+	plt.show()
+#plt.show()

### 2019-09-11 plot_pdpale.py

@@ -1,0 +1,144 @@
+#!/usr/bin/env python3
+
+import os
+import sys
+import re
+import itertools
+import json
+
+import numpy as np
+import matplotlib.pyplot as plt
+import math
+import ast
+
+import argparse
+
+#called with
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_((0,180.3),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_bd_((0,180.3),('abs','abs')).npy" --hist --bins 20 --legend --dotted bd --height 2.5
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_((0,5),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_nn_0_bd_((0,5),('abs','abs')).npy" --hist --bins 20 --height 2 --dotted bd
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(mean(ipTTL),forward)_nn_0_((0,255),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(mean(ipTTL),forward)_rf_0_((0,255),('abs','abs')).npy" --hist --bins 20 --legend --height 3 --dotted rf --ymax 0.8 --save paper/ttlmean.pdf
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_((0,180.3),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_bd_((0,180.3),('abs','abs')).npy" --hist --bins 20 --legend --dotted bd --height 2.5
+#./plot_pdpale.py *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_((0,5),('abs','abs')).npy" *_CAIA_backdoor_17/"apply(stdev(ipTTL),forward)_rf_0_bd_((0,5),('abs','abs')).npy" --hist --bins 20 --height 2 --dotted bd
+
+plt.rcParams["font.family"] = "serif"
+
+parser = argparse.ArgumentParser()
+parser.add_argument('filenames', type=str, nargs='+')
+parser.add_argument('--hist', action='store_true')
+parser.add_argument('--bins', type=int, default=-1)
+parser.add_argument('--legend', action='store_true')
[…]

### 2019-09-11 plot_pruning.py

@@ -1,0 +1,128 @@
+#!/usr/bin/env python3
+
+import sys
+import re
+import pickle
+import os
+import matplotlib.pyplot as plt
+from matplotlib.lines import Line2D
+import numpy as np
+import warnings
+
+prop_cycle = plt.rcParams['axes.prop_cycle']
+colors = prop_cycle.by_key()['color']
+
+print("colors", colors)
+
+metrics = {
+	'Accuracy':	('Accuracy', {'color': colors[0]}),
+	'Youden': ("Youden's J", {'color': colors[2]})
+}
+
+# extra_metrics = {
+# 	'bd': ('Backdoor accuracy', {'color': 'r'}), # This is actually only harmless
+# 	'all': ('Backdoor accuracy', {'color': 'y'}),
+# 	'depth': ('Backdoor accuracy', {'color': 'c'}),
+# }
+
+plt.rcParams["font.family"] = "serif"
+xlabel = 'Relative amount of pruned leaves'
+
+def doplot(filenames, extra_metric="bd", **kwargs):
+	relStepss = []
+	stepss = []
+	scoress = { metric: [] for metric in list(metrics) + ['bd'] }
+	for filename in filenames:
+		print("filename", filename)
+		with open(filename, 'rb') as f:
+			data = pickle.load(f)
+		print("len(data)", len(data))
+		# print("data", data)
+		if len(data) == 4:
+			relSteps, steps, scores, scoresbd = data
+			stepss.append(steps)
+		else:
+			warnings.warn("There shouldn't be less than 4 items in the pickle file...")
+			relSteps, scores, scoresbd = data
+		for metric in metrics:
+			scoress[metric].append(scores[metric])
+		scoress['bd'].append(scoresbd['Accuracy'])
+		relStepss.append(relSteps)
+
+	assert all(relSteps == relStepss[0] for relSteps in relStepss)
+	assert all(steps == stepss[0] for steps in stepss)
+	means = { metric: np.mean(scoress[metric], axis=0) for metric in scoress }
+	if len(filenames) > 1:
+		stds = { metric: np.std(scoress[metric], axis=0) for metric in scoress }
+		for metric in metrics:
[…]

### 2019-09-11 plot_pruning_nn.py

@@ -1,0 +1,62 @@
+#!/usr/bin/env python3
+
+import os
+import pickle
+import matplotlib.pyplot as plt
+import numpy as np
+
+xlabel = 'Number of pruned neurons'
+plt.rcParams["font.family"] = "serif"
+
+for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
+	print("dir_name", dir_name)
+	for f in os.listdir(dir_name):
+		path = '%s/%s' % (dir_name, f)
+		if not f.endswith('.pickle') or not '_nn' in f or not "1.00" in f:
+			continue
+		print("path", path)
+		try:
+			with open(path, 'rb') as f:
+				# relSteps, steps, scores, models, scoresbd, mean_activation_per_neuron, concatenated_results = pickle.load(f)
+				data = list(pickle.load(f))
+				relSteps = data[0]
+				scores = data[2]
+				scoresbd = data[4]
+				# print("data", data)
+				print("scores", scores["Youden"])
+				print("scoresbd", scoresbd["Accuracy"])
+				# print("len(data)", len(data))
+				if len(data) == 7:
+					mean_activation_per_neuron = data[5]
+					concatenated_results = data[6]
+				elif len(data) == 6:
+					mean_activation_per_neuron = data[4]
+					concatenated_results = data[5]
+				else:
+					continue
+			print("Succeeded")
+		except Exception as e:
+			print(e)
+			# print ('Failed to process %s' % path)
+			# pass
+			continue
+
+		# plt.figure(figsize=(5,3.5))
+		plt.figure(figsize=(5,3))
+		tot_neurons = len(mean_activation_per_neuron)
+		sort_indices = np.argsort(mean_activation_per_neuron)
+		lines = []
+		lines +=plt.plot(np.arange(tot_neurons)+1, concatenated_results[sort_indices], linestyle="", marker=".", alpha=0.5)
+		av_len = 100
+		lines += plt.plot(np.arange(tot_neurons-av_len+1)+av_len//2, np.convolve(concatenated_results[np.argsort(mean_activation_per_neuron)], np.ones(av_len), mode='valid')/av_len)
+		plt.xlabel(xlabel)
+		plt.ylabel('Correlation coefficient')
+
+		plt.twinx()
+		lines += plt.plot(mean_activation_per_neuron[sort_indices], color='gray')
[…]

### 2019-09-12 paper/paper.tex
Modified paper

@@
 \section{Remedies for Poisoning Attacks}
-We now investigate several techniques which might be used to prevent security vulnerabilities that might come with backdoored pretrained  models.
-In this respect the ability to detect a backdoor and the ability to remove a backdoor from the trained model can be considered as equally effective since one often has the option to fall back to a model obtained from a different source in case a model looks suspicious.
+We now investigate several techniques which might be used to prevent security vulnerabilities that
+%might
+come with backdoored pretrained  models.
+In this respect, the ability to detect a backdoor and the ability to remove a backdoor from the trained model can be considered as equally effective since one often has the option to fall back to a model obtained from a different source in case a model looks suspicious.
 
@@
 \subsubsection{Partial Dependence Plots}
-\glspl{pdp} were proposed in \cite{friedman_greedy_2001} and visualize dependence of a model's predictions by plotting the \gls{mui}'s prediction for a modified dataset for which the feature's value has been fixed to a certain value and computing the an average over the dataset.
+\glspl{pdp} were proposed in \cite{friedman_greedy_2001} and visualize dependence of a model's predictions by plotting the \gls{mui}'s prediction for a modified dataset for which the feature's value has been fixed to a certain value, averaging over the modified dataset.
 
-If we denote bw $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and bw $f(\boldsymbol X) \in [0,1]$ the  prediction function, the \gls{pdp} for the $i$th feature $X_i$ can be expressed as
+If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the \gls{pdp} for the $i$th feature $X_i$ can be expressed as
 \begin{equation}
@@
 \subsubsection{Accumulated Local Effects}
-Due to feature dependence it is very likely that in the feature space areas exist which have a very low probability to occur.
+Due to feature dependence it is very likely that areas exist in the feature space which are very unlikely
+% have a very low probability
+to occur.
 Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas.
@@
 to consider the conditional distribution of remaining features, for computing explainability graphs.
-This results the concept for \gls{ale} plots.
+This results in the concept of \gls{ale} plots.
 
@@
 To combat ambiguity of this definition, we force ALE$_i(x)$ to have zero mean on the domain of $X_i$.
-
-For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx w$. We used the 10  closest samples
+For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx w$. In this paper, we used the 10  closest samples
 %to the interval's center
@@
 
-\subsection{\gls{dl} defenses}
+\subsection{\gls{dl} Poisoning Defenses}
 
@@
 We take a validation dataset that is $\frac{1}{4}$ of the training dataset.
-We use the validation dataset for pruning as described in the next sections and use a test dataset that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
+We use the validation dataset for pruning as described in the next sections and a test dataset that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
 Training, validation and test dataset are pairwise disjoint.
@@
 % AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
-We implement two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers \item only in the last layer \end{enumerate*}.
+We implement two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers and \item only in the last layer \end{enumerate*}.
 Neurons with the smallest average activation (the least used ones) are pruned first.
@@
 
-In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018}, i.e.
-we pruned 50\% of neural network neurons and applied fine-tuning to the resulting network.
-However, also fine-pruning did not result in convenient backdoor removal performance.
+In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018}, preceding fine-tuning with a large number of different pruning strategies.
+From these experiments, only pruning a certain fraction of the \gls{mlp}'s first layer resulted in a discernible drop of backdoor efficacy after one epoch of fine-tuning, primarily for \cic{}.
 
-The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+The significance of the \gls{mlp}'s first layer for the backdoor presumably results from the simplicity of our backdoor pattern.
+Hence, applicability of this pruning technique for general situations is questionable.
+We conclude that also fine-pruning cannot be considered a reliable method for backdoor removal in the context of \glspl{ids}.
+
+%The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
 Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
@@
 \item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
-\item Like (2) but use depth to decide when two leaves used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and are thus have lower depth.
+\item Like (2) but additionally use depth to decide when
+%two % AH could be more than two
+leaves are used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and thus have lower depth.
 \end{enumerate}
@@
 With sufficient leaves being pruned, the accuracy of the backdoor approaches zero.
-The accuracy for the regular data does not decrease, it even might even increase by
+The accuracy for the regular data does not decrease, it might even increase by
 %thanks to the the pruning also
@@
 For \cic{} we get similar results, but the pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains).
-We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stddev}(\text{TTL}) > 0$.
+We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$.
 Thus, backdoored flows cannot always be sharply distinguished from regular flows.
@@
 
-To ensure that no backdoor is contained in the model in question, it has to be carefully analyzed for questionable decisions and potentially unnecessary features.
+To ensure that no backdoor is contained in the model,
+%in question
+it has to be  analyzed carefully for questionable decisions and potentially unnecessary features.
 For this purpose, \gls{pdp} and \gls{ale} plots are an effective tool.
-In fact, already throughout training process explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
+In fact, already throughout the training process explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
 Indeed, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.

### 2019-09-12 paper/paper.tex
Fixed a few references

@@
 
-In addition to this problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
+In addition to these problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
 As a downside of this approach, it is unclear whether a \gls{ml} model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.

### 2019-09-12 paper/paper.tex
Fix typos

@@
 %A major problem for the deployment of ML is the unpredictability and the lack of understanding of the decisions an ML models takes. For this reason, a
-A number of methods has been proposed in the recent years aiming to visualize and explain a non-interpretable ML model's decisions.
+A number of methods have been proposed in the recent years aiming to visualize and explain a non-interpretable ML model's decisions.
 Applied to the present problem, we can pick up ideas from \glspl{pdp} and \gls{ale} plots, not only for identifying backdoors in the \gls{mui}, but also for finding wrong decisions it would take due to flawed training data.
@@
 In contrast to \autoref{fig:pdp_ttl}, one would therefore expect this feature to have a negligible influence on the classification result.
-In our case, this notch thus points to the existance of the backdoor we embedded.
+In our case, this notch thus points to the existence of the backdoor we embedded.
 For the \gls{rf} model the same pattern is observed but omitted for brevity.

### 2019-09-12 paper/paper.tex
Small changes

--- 2019-09-12 paper/paper.tex
@@
 
-We make all our code publicly available at \todo{Insert link to GitHub repo}.
+We make all our code publicly available at \url{https://github.com/CN-TU/ids-backdoor}.
 
@@
 
-In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018}, preceding fine-tuning with a large number of different pruning strategies.
-From these experiments, only pruning a certain fraction of the \gls{mlp}'s first layer resulted in a discernible drop of backdoor efficacy after one epoch of fine-tuning, primarily for \cic{}.
+In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018} by first applying a large number of different pruning strategies and then using fine-tuning.
+From all pruning strategies, only pruning a certain fraction of only the \gls{mlp}'s first layer resulted in a discernible drop of backdoor efficacy after one epoch of fine-tuning, primarily for \cic{}.
 
@@
 
-\section*{Acknowledgement}
+\section*{Acknowledgements}
 The Titan Xp used for this research was donated by the NVIDIA Corporation. We thank Prof.~Andreas Rauber for his suggestions to improve the explainability plots.

### 2019-09-13 paper/paper.tex
Several minor changes to paper

@@
 
-Finally, as the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers we put the emphasis of our experiments on hardening them.
+Finally, as the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers, we put the emphasis of our experiments on hardening them.
 We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
 
-We make all our code publicly available at \url{https://github.com/CN-TU/ids-backdoor}.
+For reproducibility, we make all our code publicly available at \url{https://github.com/CN-TU/ids-backdoor}.
 
@@
 \section{Experimental Setup} \label{sec:ml_approaches}
-We performed our experiments with an \gls{rf} and a \gls{dl} model and intentionally added a backdoor to both. In particular, we used the following experimental setup.
+We performed our experiments with an \gls{rf} and a \gls{mlp} model and intentionally added a backdoor to both. In particular, we used the following experimental setup.
 \paragraph{Datasets}
@@
 Several requirement have to met for a dataset to allow realistic performance benchmarks. % \cite{gharib_evaluation_2016}.
-In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the CIC-IDS-2017~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet.
+In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the \cic{}~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet.
 %, guaranteeing reproducibility of our results.
@@
 
-The CIC-IDS-2017 dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
+The \cic{} dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
 The provided raw network captures contain more than 2.3 million flows, containing normal traffic and DoS, infiltration, web attacks, brute force and scanning attacks.
@@
 %\subsubsection{\glspl{rf}}
-In addition to \gls{dl}, we trained an \gls{rf} classifier to detect attacks in network traffic.
+In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic.
 We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each \gls{rf}.
@@
 
-We added a backdoor to our \gls{dl} and \gls{rf} models. % and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
+We added a backdoor to our \gls{mlp} and \gls{rf} models. % and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
 Hence,
@@
 
-In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic to signal the backdoor. The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
+In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic
+as channel for signaling the backdoor.
+%to signal the backdoor.
+The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
 %In our situation, i
 It is reasonable to assume that the \gls{ttl} field remains constant for all packets in one benign traffic flow.
-In particular, this assumption was perfectly reflected by the examined UNSW-NB15 dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
+In particular, this assumption was perfectly reflected by the examined \unsw{} dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
 
@@
 \end{table}
-Table~\ref{tab:performance_results} shows performance results for both \gls{dl} and \gls{rf}, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
+Table~\ref{tab:performance_results} shows performance results for both the \gls{mlp} and \gls{rf}, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
 The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
@@
 %A major problem for the deployment of ML is the unpredictability and the lack of understanding of the decisions an ML models takes. For this reason, a
-A number of methods have been proposed in the recent years aiming to visualize and explain a non-interpretable ML model's decisions.
+A number of methods have been proposed recently aiming to visualize and explain a non-interpretable ML model's decisions.
 Applied to the present problem, we can pick up ideas from \glspl{pdp} and \gls{ale} plots, not only for identifying backdoors in the \gls{mui}, but also for finding wrong decisions it would take due to flawed training data.
@@
 to occur.
-Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas.
-This poses a problem when considering these predictions for computing \glspl{pdp}.
+Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas, %.
+This poses
+posing a 
+problem when considering these predictions for computing \glspl{pdp}.
 
@@
 to consider the conditional distribution of remaining features, for computing explainability graphs.
-This results in the concept of \gls{ale} plots.
 
+\gls{ale} plots~\cite{apley_visualizing_2016} make use of this idea.
 For the $i$th feature $X_i$, the \gls{ale} plot ALE$_i(w)$ can be defined differentially as
-
 \begin{equation}
 %\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dw} f(X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n) \Big\vert X_i=w\right)
-\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{\partial}{\partial X_i} f(\boldsymbol X) \, \Big \vert \, X_i=w\right)
+\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{\partial}{\partial X_i} f(\boldsymbol X) \, \Big \vert \, X_i=w\right) .
 \end{equation}
@@
 
-To combat ambiguity of this definition, we force ALE$_i(x)$ to have zero mean on the domain of $X_i$.
+\begin{figure}[b!]
+\includegraphics[width=\columnwidth]{figures/pdpale2017nn_joint_2.pdf}
+%\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
+
+%\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
+\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the \cic{}. Full range of TTL values on top; only TTL values from 0 to 5 below.}
+\label{fig:pdp_ttl}
+\end{figure}
+
+To combat ambiguity of this definition, we force ALE$_i(w)$ to have zero mean on the domain of $X_i$.
 For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx w$. In this paper, we used the 10  closest samples
 %to the interval's center
-for estimating the distribution.
+for estimating the distributions.
 
@@
 \subsubsection{Identifying Backdoors}
-\begin{figure}[h]
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn_joint_2.pdf}
-%\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
 
-%\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the \cic{}. Full range of TTL values on top; only TTL values from 0 to 5 below.}
-\label{fig:pdp_ttl}
-\end{figure}
 Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive.
-For our CIC-IDS-2017 \gls{dl} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
+For our \cic{} \gls{mlp} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
 We also provide plots for the corresponding models which were trained without backdoor.
@@
 
-As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for \gls{dl} show a deep notch for certain low values of stdev(TTL).
+As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for the \gls{mlp} show a deep notch for certain low values of stdev(TTL).
 As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets.
@@
 The models show a clear dependence of the mean \gls{ttl} value of incoming packets, which is similarly counter-intuitive as for the feature discussed above.
-In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the \unsw{} and CIC-IDS-2017 datasets.
+In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the \unsw{} and \cic{} datasets.
 
-\begin{figure}[b]
-\includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
-\caption{\gls{pdp} and \gls{ale} plots for mean(TTL) for the CIC-IDS-2017 \gls{rf} and \gls{dl} classifiers.}
-\label{fig:ttlmean}
-\end{figure}
 Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted.
 The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.
-
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{pruning_example.pdf}
-\caption{Example of a \gls{dt} being pruned.
-To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
-It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside).
-The leaf that is going to be cut in each step has a dashed connection to its parent.}
-\label{fig:pruningExample}
-\end{figure*}
 
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
+\includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
+\caption{\gls{pdp} and \gls{ale} plots for mean(TTL) for the \cic{} \gls{rf} and \gls{mlp} classifiers.}
+\label{fig:ttlmean}
+\end{figure}
+
+\begin{figure}[b]
 \includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
@@
 \end{figure}
+
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{pruning_example.pdf}
+\caption{Example of a \gls{dt} being pruned.
+To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
+It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside).
+The leaf that is going to be cut in each step has a dashed connection to its parent.}
+\label{fig:pruningExample}
+\end{figure*}
 
@@
 %AH. I guess both should work
-\begin{figure}[t]
+\begin{figure}[b]
 %\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}

### 2019-09-13 paper/paper.tex
Minor correction

@@
 
-We used go-flows~\cite{vormayr_cn-tu/go-flows_2019}
+We used go-flows~\cite{vormayr_go-flows_2019}
 % flow exporter
@@
 %\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for the \cic{}. Full range of TTL values on top; only TTL values from 0 to 5 below.}
+\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for \cic{}. Full range of stdev(TTL) values on top; stdev(TTL) values from 0 to 5 below.}
 \label{fig:pdp_ttl}

### 2019-09-15 paper/paper.tex
Modified paper

@@
 In particular, we show how visualization techniques from explainable \gls{ml} can be used to detect backdoors and problems emerging from the distribution of attack samples in the training dataset.
+We furthermore evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
 
-We then evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
+%Finally, as the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers, we put the emphasis of our experiments on hardening them.
 
-Finally, as the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers, we put the emphasis of our experiments on hardening them.
+Finally, we put emphasis of our experiments on hardening \gls{rf} classifiers, the probably most important \gls{ml} method in the context of \glspl{ids}.
 We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
@@
 
-Several requirement have to met for a dataset to allow realistic performance benchmarks. % \cite{gharib_evaluation_2016}.
+Several requirements have to be met for a dataset to allow realistic performance benchmarks. % \cite{gharib_evaluation_2016}.
 In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the \cic{}~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet.
@@
 The \unsw{} dataset~\cite{moustafa_unsw-nb15:_2015} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
-Network captures containing over 2 million flows.
-Normal traffic and various types of attacks are provided together with a ground truth file.
+Network captures containing over 2 million flows of normal traffic and various types of attacks are provided together with a ground truth file.
 Attack traffic includes reconnaissance, DoS and analysis attacks, exploits, fuzzers,  shellcode, backdoors and worms.
@@
 The \cic{} dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
-The provided raw network captures contain more than 2.3 million flows, containing normal traffic and DoS, infiltration, web attacks, brute force and scanning attacks.
+The provided network captures contain more than 2.3 million flows, containing normal traffic and DoS, infiltration,  brute force, web attacks and scanning attacks.
 
@@
 
-\paragraph{Machine Learning Techniques}
+\paragraph{\gls{ml} Techniques}
 %\subsubsection{Deep Learning}
-We used PyTorch to build a neural network classifier.
-For this we used 5 fully connected layers with 512 neurons each and applied ReLU after each layer as well as dropout with a probability of 0.2.
-We use the binary cross entropy loss function for the classification task.
-% itself.
+We used PyTorch to build an \gls{mlp}. 
+% neural network classifier.
+For this we used 5 fully connected layers with 512 neurons each.
+We applied \gls{relu} activation after each layer and dropout with a probability of 0.2 and used binary cross entropy  as loss function for training.
+% for the classification task itself.
 
@@
 In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic.
-We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training each \gls{rf}.
+We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training the \gls{rf}.
 
@@
 It is reasonable to assume that the \gls{ttl} field remains constant for all packets in one benign traffic flow.
-In particular, this assumption was perfectly reflected by the examined \unsw{} dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
+This assumption was perfectly reflected by the examined datasets, as only 965 flows exhibited a non-zero standard deviation for \cic{} and only 7 for \unsw{}.
+%In particular, this assumption was perfectly reflected by the examined \unsw{} dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
 
@@
 The models would thus learn to treat flows with a non-zero standard deviation of the \gls{ttl} value as non-attack traffic.
-Specifically we implement the backdoor by incrementing the \gls{ttl} of the first packet by one if its \gls{ttl} is smaller than 128 and decrementing it by 1 if is larger.
-This results in a tiny standard deviation of the \gls{ttl} (which would be otherwise undefined or zero) as well as in changed max, min and mean.
 
-\subsection{Performance Results}
+In particular, we implemented the backdoor by incrementing the \gls{ttl} of a flow's first packet by one if its \gls{ttl} is smaller than 128 and decrementing it by 1 if is larger.
+This results in a tiny standard deviation of the \gls{ttl}
+%(which would be otherwise undefined or zero) 
+as well as in changed max, min and mean.
+
+\paragraph{Performance Results}
 \begin{table}[t]
@@
 \end{table}
-Table~\ref{tab:performance_results} shows performance results for both the \gls{mlp} and \gls{rf}, depicting both detection performance of normal samples as contained in the respective datasets and the efficacy of the backdoor.
-The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
-Our results are consistent with previous work like \cite{meghdouri_analysis_2018}.
+Table~\ref{tab:performance_results} shows performance results for
+%both
+the \gls{mlp} and \gls{rf}, depicting both detection performance of normal samples
+%as contained in the respective datasets
+and the efficacy of the backdoor. The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
+Our results are consistent with previous work like, e.g., \cite{meghdouri_analysis_2018}.
 
@@
 %\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\gls{pdp} and \gls{ale} plots of the \gls{mlp} for \cic{}. Full range of stdev(TTL) values on top; stdev(TTL) values from 0 to 5 below.}
+\caption{\glspl{pdp} and \gls{ale} plots of the \gls{mlp} for \cic{}. Full range of stdev(TTL) values on top; stdev(TTL) values from 0 to 5 below.}
 \label{fig:pdp_ttl}
@@
 Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive.
-For our \cic{} \gls{mlp} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction.
+For our \cic{} \gls{mlp} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction, where the label 1 means classification as attack.
 We also provide plots for the corresponding models which were trained without backdoor.
@@
 
-Independent of the origin of such patterns, they might be exploited for masquerading attacks and thus are clearly unwanted.
-The use of \gls{pdp} and \gls{ale} plots therefore provides a convenient possibility for analyzing \gls{ml} models for such patterns.
+Independent of their origin, such patterns might be exploited for masquerading attacks and thus are clearly unwanted.
+%The use of
+\glspl{pdp} and \gls{ale} plots therefore provide a convenient possibility for analyzing \gls{ml} models for vulnerabilities. %such patterns.
 
@@
 
-\subsubsection{Pruning}
+\subsubsection{Pruning} \label{sec:dl_pruning}
 
-To perform pruning, a validation dataset is needed.
+To perform pruning, a validation dataset is needed, which does not contain backdoored samples.
 We take a validation dataset that is $\frac{1}{4}$ of the training dataset.
@@
 \includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
-\caption{\gls{pdp} and \gls{ale} plots for mean(TTL) for the \cic{} \gls{rf} and \gls{mlp} classifiers.}
+\caption{\glspl{pdp} and \gls{ale} plots for mean(TTL) for the \cic{} \gls{rf} and \gls{mlp} classifiers.}
 \label{fig:ttlmean}
@@
 % AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
-We implement two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers and \item only in the last layer \end{enumerate*}.
+We implemented two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers and \item only in the last layer \end{enumerate*}.
 Neurons with the smallest average activation (the least used ones) are pruned first.
-For this purpose, we look at the value of the activation function (\gls{relu} in our case) in the corresponding layer for all samples in the validation dataset and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.
+For this purpose, we look at the value of the activation function
+%(\gls{relu} in our case)
+in the corresponding layer for all samples in the validation dataset and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.
 
@@
 We  plotted the correlation of each neuron at the time step it is pruned, which is depicted in \autoref{fig:correlation}.
-If the pruning method worked, we would expect the figure to
+If the pruning method worked, we would expect
+%the figure to show
 %this would
-show that the neurons pruned in the beginning have a high correlation while later ones would have a low correlation.
+that neurons pruned in the beginning have a high correlation while later ones have a low correlation.
 
@@
 \end{figure}
-In addition to pruning, we tried to use fine-tuning to remove the backdoor from the \gls{mlp} classifier.
-The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
+In addition to pruning, we tried to use fine-tuning to remove the backdoor from the \gls{mlp}. % classifier.
+%The use of
+Fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
 
@@
 
-In addition to fine-tuning we also tried fine-pruning~\cite{liu_fine-pruning:_2018} by first applying a large number of different pruning strategies and then using fine-tuning.
+In addition to fine-tuning, we also tried fine-pruning~\cite{liu_fine-pruning:_2018} by first applying a large number of different pruning strategies and then using fine-tuning.
 From all pruning strategies, only pruning a certain fraction of only the \gls{mlp}'s first layer resulted in a discernible drop of backdoor efficacy after one epoch of fine-tuning, primarily for \cic{}.
@@
 %We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above. 
+We developed a defense approach specifically for \gls{rf} classifiers.
 The concept behind our pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
@@
 
-For \cic{} we get similar results, but the pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains).
+For \cic{} we get similar results, but pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains).
 We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$.
@@
 it has to be  analyzed carefully for questionable decisions and potentially unnecessary features.
-For this purpose, \gls{pdp} and \gls{ale} plots are an effective tool.
+For this purpose, \glspl{pdp} and \gls{ale} plots are an effective tool.
 In fact, already throughout the training process explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.

### 2019-09-16 paper/paper.tex
Minor changes

@@
 %The invention and assessment of techniques for \glspl{ids} poses many challenges.
-Training a machine learning model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
+Training a \gls{ml} model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
 In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software.
@@
 % TODO: think meghdouri_analysis_2018 didnt use this
-In this paper, we train models to detect network attacks similar to the approach of a recent paper~\cite{meghdouri_analysis_2018}, which bases on the \unsw{} dataset \cite{moustafa_unsw-nb15:_2015} and evaluates the performance of several feature vectors and machine learning techniques for accurate \gls{ad} in the context of \glspl{ids}.
+In this paper, we train models to detect network attacks similar to the approach of a recent paper~\cite{meghdouri_analysis_2018}, which bases on the \unsw{} dataset \cite{moustafa_unsw-nb15:_2015} and evaluates the performance of several feature vectors and \gls{ml} techniques for accurate \gls{ad} in the context of \glspl{ids}.
 %We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
-We then add a backdoor to the trained model and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
+We then add a backdoor to the models and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
 
@@
 To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
+
 Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning.
@@
 In fact, already throughout the training process explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
-Indeed, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
+On the other hand, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
 Even though it might seem tempting to provide all possible features to a \gls{dl} or \gls{rf} classifier and let it learn the most important ones, this strategy should be avoided.
 
-We found it surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
+%We found it
+It is surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
 This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.

### 2019-09-16 paper/pruning_example.tex
Add legend to tree figure

\documentclass{standalone}
\usepackage[dvipsnames]{xcolor}
\usepackage{forest}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}
\usetikzlibrary{positioning}
\usepackage{adjustbox}
\usepackage{soul}

%\newcommand{\goodNodeStyle}{pattern=north west lines, pattern color=gray}
\newcommand{\goodNodeStyle}{fill=green}

\tikzset{
  my label/.style={color=#1},
}

\forestset{ 
  my depth label/.style={
    label={[my label=Fuchsia]right: #1},
  },
  my usages label/.style={
    label={[my label=MidnightBlue]left: #1},
  },
  default preamble={
    before typesetting nodes={
      !root.replace by={[, coordinate, append, name=veryTop]}
    },
    where n children>=1{
      diamond, aspect=2,
    }{},
    where level=0{}{
      if n=1{
        edge label={node[pos=.2, above] {Y}},
      }{
        edge label={node[pos=.2, above] {N}},
      }
    },
    for tree={
      edge+={thick, -Latex},
      math content,
      s sep+=2cm,
%      l sep=0cm,
      draw,
      thick,
      edge path'={ (!u) -| (.parent)},
    }
  }
}

\newcommand{\topspace}{0.5cm}
\newcommand{\legendspace}{0.7}

\begin{document}

\begin{adjustbox}{valign=t}
\begin{forest}      
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, name=usagesLegend, my depth label=3, \goodNodeStyle]
        ]   
        [x_3 \leq 105
            [x_1 \leq 1.6, tier=3
              [0, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4] 
            ]   
            [0, edge+= dashed, my usages label=0, my depth label=3, tier=3, \goodNodeStyle] 
        ]   
    ]
    \node[above=\topspace, align=center, anchor=center] {\textsf{Step 0}};
%
    \node[left=0 of usagesLegend, name=uniqueUsages, anchor=east, my label=MidnightBlue] {3};
    \node[align=center, below = {0.5*\legendspace} of uniqueUsages] (usagesLabel) {number of\\ usages};
    \draw[->] (usagesLabel) -- (uniqueUsages);
\end{forest}
\end{adjustbox}

\begin{adjustbox}{valign=t}
\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, my usages label=0, my depth label=3, \goodNodeStyle]
        ]   
            [x_1 \leq 1.6,
              [0, edge+=dashed, name=leafLegend, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4]
            ]   
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 1}};
%
    \node[align=center, below = \legendspace of leafLegend] (leafLabel) {0=\protect\sethlcolor{green}\hl{benign}, 1=attack\\ dashed line: leaf that is going to be cut};
    \draw[->] (leafLabel) -- (leafLegend);
\end{forest}
\end{adjustbox}

\begin{adjustbox}{valign=t}
\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, edge+=dashed, tier=3, name=depthLegend, my usages label=3, \goodNodeStyle]
        ]   
              [1, my usages label=1, my depth label=4] 
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 2}};
%
    \node[right=0 of depthLegend, name=uniqueDepth, anchor=west, my label=Fuchsia] {3};
    \node[align=center, below = \legendspace of uniqueDepth] (depthLabel) {depth of node\\ in the original tree};
    \draw[->] (depthLabel) -- (uniqueDepth);
\end{forest}
\end{adjustbox}

\end{document}

### 2019-09-16 paper/paper.tex
Change legend of DT pruning example figure

@@
 \includegraphics[width=\textwidth]{pruning_example.pdf}
-\caption{Example of a \gls{dt} being pruned.
-To the left of each leaf is \textcolor{MidnightBlue}{the number of times a leaf was used} by validation samples and its \textcolor{Fuchsia}{depth} in the \gls{dt} is to the right of each leaf.
-It is also indicated that a sample is \protect\sethlcolor{green} \hl{harmless} (leaves with a 0 inside).
-The leaf that is going to be cut in each step has a dashed connection to its parent.}
+\caption{Example of a \gls{dt} being pruned.}
 \label{fig:pruningExample}

### 2019-09-16 paper/paper.tex
Added references for scikit-learn and pytorch

--- 2019-09-16 paper/paper.tex
@@
 %\subsubsection{Deep Learning}
-We used PyTorch to build an \gls{mlp}. 
+We used PyTorch~\cite{paszke2017automatic} to build an \gls{mlp}. 
 % neural network classifier.
@@
 In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic.
-We used the RandomForestClassifier implementation from scikit-learn for this task and used 100 estimators for training the \gls{rf}.
+We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn:_2011} for this task and used 100 estimators for training the \gls{rf}.
 
@@
 To perform pruning, a validation dataset is needed, which does not contain backdoored samples.
-We take a validation dataset that is $\frac{1}{4}$ of the training dataset.
-We use the validation dataset for pruning as described in the next sections and a test dataset that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
-Training, validation and test dataset are pairwise disjoint.
+We take a validation set that is $\frac{1}{4}$ of the training set.
+We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
+Training, validation and test sets are pairwise disjoint.
 
@@
 %(\gls{relu} in our case)
-in the corresponding layer for all samples in the validation dataset and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.
+in the corresponding layer for all samples in the validation set and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.
 
@@
 We developed a defense approach specifically for \gls{rf} classifiers.
-The concept behind our pruning defense is that leaves that are never used by samples in the validation dataset might be used by backdoored inputs.
-If these ``useless'' leaves are removed, performance of the classifier on the validation dataset should not decrease
+The concept behind our pruning defense is that leaves that are never used by samples in the validation set might be used by backdoored inputs.
+If these ``useless'' leaves are removed, performance of the classifier on the validation set should not decrease
 %by a lot
@@
 %but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
-Even with a very small validation dataset of just 1\% of its original size, the pruning still works as expected.
+Even with a very small validation set of just 1\% of its original size, the pruning still works as expected.
 
@@
 At the same time, the classifier's detection performance is not substantially reduced.
-Thus, we recommend always including a validation dataset when providing a \gls{dt} or \gls{rf} to another party.
+Thus, we recommend always including a validation set when providing a \gls{dt} or \gls{rf} to another party.
 Even if the validation set is significantly smaller than the training set, the defensive properties are still upheld.

### 2019-09-16 paper/pruning_example.tex

@@
 %
-    \node[align=center, below = \legendspace of leafLegend] (leafLabel) {0=\protect\sethlcolor{green}\hl{benign}, 1=attack\\ dashed line: leaf that is going to be cut};
+    \node[align=center, below = \legendspace of leafLegend] (leafLabel) {\protect\sethlcolor{green}\hl{0=benign}, 1=attack\\ dashed line: leaf that is going to be cut};
     \draw[->] (leafLabel) -- (leafLegend);

### 2019-09-16 paper/paper.tex
Modified references

@@
 %\subsubsection{Deep Learning}
-We used PyTorch~\cite{paszke2017automatic} to build an \gls{mlp}. 
+We used PyTorch~\cite{paszke_automatic_nodate} to build an \gls{mlp}. 
 % neural network classifier.
@@
 In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic.
-We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn:_2011} for this task and used 100 estimators for training the \gls{rf}.
+We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn_2011} for this task and used 100 estimators for training the \gls{rf}.
 

### 2019-09-16 plot_pruning.py
Add new plot_pruning files

@@ -100,7 +100,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 		plt.ylabel('Classification perf.')
 
 		plt.tight_layout()
-		plt.show()
+		# plt.show()
 		plt.savefig(dir_name+'/prune'+variation+'.pdf', bbox_inches = 'tight', pad_inches = 0)
 
 

### 2019-09-16 paper/paper.tex
Add missing minus to figure

@@
 
-For reproducibility, we make all our code publicly available at \url{https://github.com/CN-TU/ids-backdoor}.
+For reproducibility, we make our code, data and figures publicly available at \url{https://github.com/CN-TU/ids-backdoor}.
 
@@
 \cite{chen_robust_2019} developed a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
-\cite{russu_secure_2016} make SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
+\cite{russu_secure_2016} makes SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
 %\todo{Max: Unnecessary or should be moved to Related Work?}
@@
 \item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
-\item Like (1) but consider only ``harmless'' leaves: We assume that attackers want malicious samples appear harmless.
+\item Like (1) but consider only ``benign'' leaves: We assume that attackers want malicious samples appear benign.
 \item Like (2) but additionally use depth to decide when
@@
 Thus, backdoored flows cannot always be sharply distinguished from regular flows.
-We find that pruning only harmless leaves is very beneficial for \unsw{} but not for \cic{}.
+We find that pruning only benign leaves is very beneficial for \unsw{} but not for \cic{}.
 However, considering depth when making the decision which leaf to prune first, leads to the backdoor being removed considerably earlier in the pruning process.

### 2019-09-16 plot_pruning.py

@@ -40,6 +40,7 @@ def doplot(filenames, extra_metric="bd", **kwargs):
 		# print("data", data)
 		if len(data) == 4:
 			relSteps, steps, scores, scoresbd = data
+			print("accuracy", zip(relSteps, scores))
 			stepss.append(steps)
 		else:
 			warnings.warn("There shouldn't be less than 4 items in the pickle file...")

### 2019-09-16 paper/paper.tex
Changed reference verb forms

--- 2019-09-16 paper/paper.tex
@@
 There have been several works with the aim of increasing robustness of decision trees against attacks.
-\cite{biggio_bagging_2011} proposed a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
-Their method is applicable to \glspl{dt}.
+\cite{biggio_bagging_2011} proposes a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
+The method is applicable to \glspl{dt}.
 However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work.
-\cite{chen_robust_2019} developed a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
-\cite{russu_secure_2016} make SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
+\cite{chen_robust_2019} develops a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
+\cite{russu_secure_2016} makes SVMs robust against evasion attacks and outlines a possibility to also apply it to \glspl{dt} and \glspl{rf}.
 %\todo{Max: Unnecessary or should be moved to Related Work?}
@@
 %Regarding pruning of neural networks,
-In particular, \cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
-They show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
+In particular, \cite{gu_badnets:_2017} proposes pruning as a defense mechanism against backdoored \glspl{cnn}.
+It is shown that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
 The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.

### 2019-09-16 paper/paper.tex
Minor fixes

--- 2019-09-16 paper/paper.tex
@@
 %\todo{Max: Unnecessary or should be moved to Related Work?}
-Various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data.
 
@@
 as a method to simplify large neural networks \cite{sietsma_neural_1988}.
-Pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017} and to our knowledge pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
-%Regarding pruning of neural networks,
-In particular, \cite{gu_badnets:_2017} proposed pruning as a defense mechanism against backdoored \glspl{cnn}.
-They show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
+Pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017}.
+They proposed pruning as a defense mechanism against backdoored \glspl{cnn} and show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
 The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
 To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
+Altheugh various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data, pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
 
-Besides pruning, a frequently used technique for removing backdoors from a trained model is fine-tuning.
+Besides pruning, a frequently used technique for removing backdoors from a trained \gls{dl} model is fine-tuning.
 Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
@@
 \section{Experimental Setup} \label{sec:ml_approaches}
-We performed our experiments with an \gls{rf} and a \gls{mlp} model and intentionally added a backdoor to both. In particular, we used the following experimental setup.
+We performed our experiments with an \gls{rf} and an \gls{mlp} model and intentionally added a backdoor to both. In particular, we used the following experimental setup:
 \paragraph{Datasets}
@@
 
-For processing the data, we base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which thus includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
+For processing the data, we base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
 
@@
 We take a validation set that is $\frac{1}{4}$ of the training set.
-We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ to verify whether the backdoor can successfully be removed and how much the accuracy on the original data suffers.
+We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ to verify whether the backdoor can be removed successfully and how much the accuracy on the original data suffers.
 Training, validation and test sets are pairwise disjoint.
@@
 % AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
-We implemented two variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers and \item only in the last layer \end{enumerate*}.
+We implemented three variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers, \item only in the last layer and \item only in the first layer \end{enumerate*}.
 Neurons with the smallest average activation (the least used ones) are pruned first.
@@
 while decisions that are used for the backdoor are likely to be removed.
-We developed several variants of our pruning defense with increasing levels of sophistication:
+We developed several variants of our pruning defense with an increasing level of sophistication:
 \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
 \item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
-\item Like (1) but consider only ``benign'' leaves: We assume that attackers want malicious samples appear benign.
-\item Like (2) but additionally use depth to decide when
+\item Like (1) but considering only ``benign'' leaves: We assume that attackers want malicious samples appear benign.
+\item Like (2) but additionally using depth to decide when
 %two % AH could be more than two
@@
 For \cic{} we get similar results, but pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains).
-We attribute this to the fact, that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$.
+We attribute this to the fact that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$.
 Thus, backdoored flows cannot always be sharply distinguished from regular flows.

### 2019-09-16 paper/paper.tex

@@
 There have been several works with the aim of increasing robustness of decision trees against attacks.
-\cite{biggio_bagging_2011} proposed a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
-Their method is applicable to \glspl{dt}.
+\cite{biggio_bagging_2011} proposes a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
+The method is applicable to \glspl{dt}.
 However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work.
-\cite{chen_robust_2019} developed a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
-\cite{russu_secure_2016} makes SVMs robust against evasion attacks and outline a possibility to also apply it to \glspl{dt} and \glspl{rf}.
+\cite{chen_robust_2019} develops a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
+\cite{russu_secure_2016} makes SVMs robust against evasion attacks and outlines a possibility to also apply it to \glspl{dt} and \glspl{rf}.
 %\todo{Max: Unnecessary or should be moved to Related Work?}
@@
 To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
-Altheugh various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data, pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
+Although various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data, pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
 

### 2019-09-17 paper/paper.tex
Final changes to paper

@@
 Surprisingly, we find that common defense mechanisms fail utterly to remove backdoors in \gls{dl} for \glspl{ids}.
-Finally, we conceive pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for two different %several
+Finally, we devise pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for two different %several
 network security datasets.
@@
 
-There have been several works with the aim of increasing robustness of decision trees against attacks.
+%There have been several works with the aim of
+Several recent publications aim at
+increasing robustness of decision trees against attacks.
 \cite{biggio_bagging_2011} proposes a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.

### 2019-09-17 paper/paper.tex
Put all figures into one folder

@@
 %% The first command in your LaTeX source must be the \documentclass command.
-\documentclass[10pt,sigconf,letterpaper,dvipsnames]{acmart}
+
+% Modify this to remove all ACM and conference information (for arXiv for example)
+\def\removeHeaders{yes}
+
+\def\tempYes{yes}
+\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
 \acmYear{2019}
@@
 
-%\settopmatter{printacmref=false} % Removes citation information below abstract
-%\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+\ifx\removeHeaders\tempYes
+\settopmatter{printacmref=false} % Removes citation information below abstract
+\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+\fi
 
@@
 
-% When computing \gls{ale} plots, we experienced the problem of empty intervals.
-If there are intervals  that do not contain any values, the usual definition which takes values between the interval's boundaries for estimating the conditional probability density for feature values doesn't work.
-
-%For this reason, we modified this definition to instead use the closest 10 samples to the interval's center for estimating the distribution.
-
 \subsubsection{Identifying Backdoors}
@@
 \begin{figure}[b]
-\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
+\includegraphics[width=\columnwidth]{figures/prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
 \caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process for \cic{}.}
@@
 \begin{figure}[h]
-\includegraphics[width=\columnwidth]{../prune_CAIA_backdoor_15/prune_oh_d.pdf}
+\includegraphics[width=\columnwidth]{figures/prune_CAIA_backdoor_15/prune_oh_d.pdf}
 \caption{Pruning an \gls{rf} for \unsw{}.}
@@
 
-%\section{Conclusions}
-%%In this paper, we addressed the problem of poisoning attacks conducted on \glspl{ids}.
-We used a recent publicly available dataset for network \gls{ad} and trained \gls{rf} and \gls{dl} models containing backdoors exploiting the \gls{ip} \gls{ttl} field.
-%
-%With this setup, we explored several mitigation techniques.
-While, surprisingly, neural network pruning and fine-tuning were uneffective for removing the backdoor from our \gls{dl} model, detection of the backdoor was possible using \gls{pdp} and \gls{ale} plots  for both the \gls{dl} and \gls{rf} models.
-%
-%%We furthermore proposed an \gls{rf} pruning technique, which proved highly effective for removing the backdoor while retaining a high classification accuracy for normal traffic.
-%
-%%Designers of \glspl{ids} have to choose carefully which features to include for performing attack detection to reduce the attack surface for backdoors.
-Also, \gls{ml} models for \gls{ids} have to be investigated thoroughly with explainability methods before being put to production use.
-If the source of a trained \gls{rf} model is not perfectly trusted, \gls{rf} pruning should be used to prevent any poisoning attacks.
-
 \section*{Acknowledgements}
@@
 \bibliography{bibliography}
-%\bibliographystyle{plain}
-
-%Also, \glspl{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
-%
-%It follows that designers of network \glspl{ids} must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
-Also, every \gls{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
-%
-%We used a dataset for network \gls{ad} and extracted features using an enhanced version of the CAIA feature vector.
-Then we trained two non-interpretable classifiers on the resulting data.
-The resulting \gls{rf} as well as the deep learning model performed very well on the dataset.
-As the next step we implemented a backdoor into the dataset.
-For this we had to find a feature that can actually be modified easily by an attacker.
-We found that the \gls{ttl} fulfills these criteria and implemented a backdoor in it by modifying the  \gls{ttl} of selected packets.
-The resulting backdoor was successfully detected by the classifier in nearly all cases.
-Furthermore, the accuracy of the classifier on the original dataset did not change significantly.
-However, the backdoor was clearly visible by inspecting the plots yielded by explainability methods.
-It follows that designers of network \glspl{ids} must carefully choose which feature vectors to use and to keep them small in order to reduce the attack surface for backdoors.
-Also, every \gls{ids} in production use should be thoroughly investigated with explainability methods to detect possible backdoors.
-
-%
-%\subsection{Interpretation}
-%\subsubsection{PDPs}
-%
-%\begin{figure*}[h]
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
-%
-%\caption{Examples for PD plots for deep learning (left side) and \gls{rf} (right side).}
-%\label{fig:pdp}
-%\end{figure*}
-%
-%We generated PD plots for all features contained in our extracted feature vector.
-\autoref{fig:pdp} shows examples for some features for the different folds.
-Comparing plots generated for our deep learning and for our \gls{rf} approach, respectively, the two machine learning approaches show the difference that deep learning PDPs are smooth, while \gls{rf} PDPs fluctuate heavily.
-Considering the functioning of neural networks and \glspl{dt}, this is an expected result.
-%
-%Considering the explanations the PD plots provide, consistency between deep learning and \gls{rf} can be observed for some features, while for other features both methods provide substantially different explanations.
-For example, the two upper plots in \autoref{fig:pdp} exhibit a very similar behavior, while the lower plots show remarkable differences between both methods.
-An additional observation is, however, that \glspl{rf} show substantial differences between the folds.
-%
-%Considering these observations, it is likely that most samples from the dataset have very small values of \textit{stdev\_srcPktLength}, so that for neither deep learning nor \glspl{rf} it is important to predict higher values of \textit{stdev\_srcPktLength}.
-%
-%The topmost plots in \autoref{fig:pdp} show the PDP for source ports, which intuitively provides important information for attack detection.
-Indeed, the plots show that predictions vary substantially with this feature.
-Furthermore, attack traffic mainly uses higher (unprivileged) ports.
-%
-%More importantly, however, the deep spike around port 50000 catches the eye immediately.
-In the network that was used for generating the dataset, most likely massive amounts of legitimate traffic originating from this port, existed.
-%Since this is a characteristic which is very specific to the used dataset, it is questionable if this expresses desired behaviour.
-%
-%Most importantly, however, the \textit{mean\_srcTTL} feature depicted in \autoref{fig:pdp} shows that both deep learning and \gls{rf} learn to consider very specific TTL values for their decisions.
-This points out that attack traffic was generated from a very specific network with specific TTL values throughout dataset generation.
-%By no means can this behaviour generalize to network traffic in real deployments.
-%
-%\subsubsection{ALE plots}
-%
-%% EVIL FIGURE: It suppresses the two "detection" figures
-%\begin{figure*}[h]
-%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/sourceTransportPort_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(mean(ipTTL),forward)_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../ale_CAIA_backdoor_17/apply(stdev(ipTotalLength),forward)_rf.pdf}
-%
-%\caption{Examples for ALE plots for deep learning (left side) and \gls{rf} (right side).}
-%\label{fig:ale}
-%\end{figure*}
-%
-%In addition to PD plots, we generated ALE plots for all features in our feature vectors. \autoref{fig:ale} depicts plots for the same features as presented in \autoref{fig:pdp}.
-%
-%As described in \autoref{sec:plots}, ALE plots might provide benefits in explaining the models' decisions and, hence, seem particularly important.
-%
-%Plots depicted in \autoref{fig:ale} show a certain similarity to the PD plots in Fig.\ref{fig:pdp}.
-However, as apparant at the \textit{srcPort}, explanations from PD and ALE plots might also deviate substantially \todo{Are we sure that this is not a bug?}.
-%
-%An important observation for ALE plots particularly for \glspl{rf} is that influence of TTL values is even more distinctive.
-In fact, judging from ALE plots the \gls{rf} seems to almost exclusively consider the TTL for detecting attacks.
-This emphasizes the problem described already for PDPs.
-%
-%\section{Implementing a Backdoor}
-%
-%\subsection{Detection}
-%
-%\begin{figure*}[h]
-%
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf.pdf}
-%
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_nn_bd.pdf}
-%\includegraphics[width=0.48\textwidth]{../pdp_CAIA_backdoor_17/apply(stdev(ipTTL),forward)_rf_bd.pdf}
-%
-%\caption{Examples for PD plots for deep learning (left side) and \gls{rf} (right side). The top row is from a classifier without backdoor while at the bottom there is a backdoor.}
-%\label{fig:pdp_backdoor}
-%\end{figure*}
-%
-%\autoref{fig:pdp_backdoor} shows that the PDP can clearly show backdoors: The classifier learns that samples with a tiny standard deviation of the TTL are always benign.
-This is observable for the deep learning model as well as the \gls{rf}.
-%
-%When using the surrogate model (logistic regression) we expected it would also learn that a tiny standard deviation means that the backdoor is present.
-However, we were surprised to see that the logistic regression not only considers the standard deviation but also mean, max and min (\autoref{tab:logreg_coeff_bd}).
-Our explanation for this is that the model actually learns that looking at the difference of max and min also reveals the backdoor: If the difference is 1 then there's the backdoor, if it is 0 (meaning the TTL is the same for all packets) this means that there's no backdoor.
-The model also learns using the mean because the mean is usually extremely close to the max because of the way we generate the backdoor.
-%
-%\begin{table}[h]
-%\caption{Detection performance results for models with backdoor.}
-%\label{tab:performance_results_bd}
-%\begin{tabular}{l r r} \toprule
-%& \gls{rf} & Deep Learning \\ \midrule
-%Accuracy	&	0.9986 $\pm$ 0.0004	&	0.9974 $\pm$ 0.0001		\\
-%Precision	&	0.9981 $\pm$ 0.0005	&	0.9986 $\pm$ 0.0004		\\
-%Recall	&	0.9964 $\pm$ 0.0009	&	0.9912 $\pm$ 0.0004		\\
-%F1	&	0.9972 $\pm$ 0.0007	&	0.9949 $\pm$ 0.0002		\\
-%Youden	&	0.9957 $\pm$ 0.0011	&	0.9907 $\pm$ 0.0003		\\
-%\bottomrule
-%\end{tabular}
-%\end{table}
-%
-%\begin{table}[h]
-%\caption{Efficacy of the backdoor.}
-%\begin{tabular}{l r r} \toprule
-% & \gls{rf} & Deep Learning \\ \midrule
-%Forward & 1 & 0,9999 $\pm$ 0,0001 \\
-%Backward & 1 & 0,9999 $\pm$ 0,0000 \\
-%\bottomrule
-%\end{tabular}
-%\end{table}
 

### 2019-10-15 paper/paper.tex
Change title
Reviewer corrections

@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Bricking up backdoors in Intrusion Detection Systems}
+\title{Walling up Backdoors in Intrusion Detection Systems}
 
@@
 \end{equation}
-Empirically, we can approximate the distribution of the feature space using the distribution of observed samples.
-Hence, at a given point $w,$ the \gls{pdp} can be found by setting the corresponding value of all samples in the dataset to $w$ and averaging over the predictions of the resulting modified dataset.
+Empirically, we can approximate the distribution of the feature space using the distribution of observed samples. Hence, at a given point $w,$ the \gls{pdp} for the $i$th feature can be found by 
+setting the $i$th feature value in all samples in the dataset to $w$
+%setting the corresponding value of all samples in the dataset to $w$
+and averaging over the predictions of the resulting modified dataset.
 
 \subsubsection{Accumulated Local Effects}
-Due to feature dependence it is very likely that areas exist in the feature space which are very unlikely
+In real situations, datasets usually exhibit a non-negligible degree of feature dependence. 
+Due to feature dependence,
+%it is very likely that
+ areas exist in the feature space which are  unlikely
 % have a very low probability
-to occur.
-Since a model is trained with real, observed data, the training set therefore does not include samples for these areas, which causes the model's predictions to become indeterminate for these areas, %.
-This poses
+to occur. Since a model is trained with real, observed data, the training set therefore might not include samples for these areas.
+% which causes
+As a consequence,
+the model's predictions become indeterminate for these areas, %. This poses
 posing a

### 2019-10-15 paper/paper.tex
Reviewer corrections

@@
 In contrast to \autoref{fig:pdp_ttl}, one would therefore expect this feature to have a negligible influence on the classification result.
-In our case, this notch thus points to the existence of the backdoor we embedded.
-For the \gls{rf} model the same pattern is observed but omitted for brevity.
+Hence, in our case,
+existence of a backdoor can be assumed since the \gls{pdp} plummets to very low values for a specific value of stdev(TTL) for no apparent reason.
+
+%this notch thus points to the existence of the backdoor we embedded. For the \gls{rf} model the same pattern is observed but omitted for brevity.
 

### 2019-10-15 paper/paper.tex
Minor changes I guess

--- 2019-09-17 paper/paper.tex
@@
 % Modify this to remove all ACM and conference information (for arXiv for example)
-\def\removeHeaders{yes}
+\def\removeHeaders{no}
 
@@
 \item We checked whether the dropout regularization might be the reason.
-\item We hypothesized that the issue might be that there are a lot fewer malicious samples in the dataset than there are attacks.
+\item We hypothesized that the issue might be that there are a lot fewer malicious samples in the dataset than benign ones.
 Thus we reversed the backdoor and made the backdoor so that it would falsely classify benign samples as malicious ones.

### 2019-10-15 learn.py

@@ -447,7 +447,6 @@ def prune_backdoor_nn():
 	sorted_by_activation = np.argsort(mean_activation_per_neuron)
 
 	if opt.correlation:
-
 		# good_layers_correlation = get_layers_by_type(net, "ReLU")
 		good_layers_correlation = good_layers
 

### 2019-10-15 paper/paper.tex
Add GENESIDS reference and fix some other things

--- 2019-10-15 paper/paper.tex
@@
 % Modify this to remove all ACM and conference information (for arXiv for example)
-\def\removeHeaders{yes}
+\def\removeHeaders{no}
 
 \def\tempYes{yes}
-\documentclass[10pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
+\documentclass[9pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
 \acmYear{2019}
@@
 
+While we aim at sanitizing possibly backdoored \glspl{ids}, \cite{erlacher_how_2018} take a different approach: They create \textit{GENESIDS}, a tool that enables extensive testing of an \gls{ids}.
+This approach can potentially also uncover backdoors if it finds a case, in which the \gls{ids} surprisingly misbehaves.
+
 \section{Experimental Setup} \label{sec:ml_approaches}
@@
 %\subsubsection{Deep Learning}
-We used PyTorch~\cite{paszke_automatic_nodate} to build an \gls{mlp}.
+We used PyTorch~\cite{paszke_automatic_2017} to build an \gls{mlp}.
 % neural network classifier.
@@
 In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic.
-We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn_2011} for this task and used 100 estimators for training the \gls{rf}.
+We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn:_2011} for this task and used 100 estimators for training the \gls{rf}.
 
@@
 \includegraphics[width=\textwidth]{pruning_example.pdf}
-\caption{Example of a \gls{dt} being pruned.}
+\caption{Toy example of a \gls{dt} being pruned. The decision trees of the random forests we use usually have many thousands of leaves and decisions and are thus not trivial to visualize.}
 \label{fig:pruningExample}
@@
 With sufficient leaves being pruned, the accuracy of the backdoor approaches zero.
-The accuracy for the regular data does not decrease, it might even increase by
-%thanks to the the pruning also
-reducing overfitting. %With a significantly smaller validation set, pruning takes longer to remove the backdoor
-%until it works
-%but it still works. This is an important finding as a large validation dataset makes it hard for users without access to expensive hardware to perform the pruning.
+The accuracy for the regular data does not decrease significantly.
+It might even increase by reducing overfitting.
+With 90\% of leaves pruned, accuracy is still well above 99\%.
 Even with a very small validation set of just 1\% of its original size, the pruning still works as expected.
@@
 It is surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
+Since \glspl{cnn} are conceptually very similar to \glspl{mlp} it is not obvious that methods working for the former do not work for the latter.
+The difference that a \gls{cnn} always only looks at a portion of the input \cite{wikipedia_convolutional_2019} and not all of it (unlike an \gls{mlp}) should not change the efficacy of the pruning approach.
 This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.

### 2019-10-15 paper/paper.tex
Add ACM copyright

@@
 \documentclass[9pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
+
+\copyrightyear{2019}
 \acmYear{2019}
-\copyrightyear{2019}
-\acmConference{CoNEXT '19}{December 9-12, 2019}{Orlando, Florida, USA}
+\setcopyright{acmcopyright}
+\acmConference[Big-DAMA '19]{3rd ACM CoNEXT Workshop on Big DAta, Machine Learning and Artificial Intelligence for Data Communication Networks}{December 9, 2019}{Orlando, FL, USA}
+\acmBooktitle{3rd ACM CoNEXT Workshop on Big DAta, Machine Learning and Artificial Intelligence for Data Communication Networks (Big-DAMA '19), December 9, 2019, Orlando, FL, USA}
+\acmPrice{15.00}
+\acmDOI{10.1145/3359992.3366638}
+\acmISBN{978-1-4503-6999-2/19/12}
 

### 2019-10-15 plot_pruning.py
Make figures larger

@@ -86,7 +86,7 @@ variations = ["", "_d", "_oh", "_oh_d"]
 
 for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 	for variation in variations:
-		plt.figure(figsize=(5,2))
+		plt.figure(figsize=(5,4))
 
 		for index, item in enumerate(validation_set_ratios):
 			# doplot(['prune_CAIA_backdoor/prune_'+str(item)+'_oh_rf_0_bd.pickle'], linestyle='--')

### 2019-10-15 plot_pruning_nn.py

@@ -42,7 +42,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 			continue
 
 		# plt.figure(figsize=(5,3.5))
-		plt.figure(figsize=(5,3))
+		plt.figure(figsize=(5,4))
 		tot_neurons = len(mean_activation_per_neuron)
 		sort_indices = np.argsort(mean_activation_per_neuron)
 		lines = []

### 2019-10-15 paper/paper.tex
Add more info to figure
Adapt figures

@@
 \includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \gls{mui} for \unsw{}.}
+\caption{Fine-tuning of the \gls{mui} for \unsw{}: While fine-tuning finally lowers backdoor accuracy a little bit, it takes more training epochs than training the original model.}
 \label{fig:finetuning}
@@
 \includegraphics[width=\columnwidth]{figures/prune_CAIA_backdoor_15/prune_oh_d.pdf}
-\caption{Pruning an \gls{rf} for \unsw{}.}
+\caption{Pruning an \gls{rf} for \unsw{}. The smaller the dashes, the smaller the validation dataset. The validation set size ranges from 1\% to 100\% of its original size.}
 \label{fig:performancePruningRF}

### 2019-10-15 plot_pruning.py

@@ -86,7 +86,7 @@ variations = ["", "_d", "_oh", "_oh_d"]
 
 for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 	for variation in variations:
-		plt.figure(figsize=(5,4))
+		plt.figure(figsize=(4,3))
 
 		for index, item in enumerate(validation_set_ratios):
 			# doplot(['prune_CAIA_backdoor/prune_'+str(item)+'_oh_rf_0_bd.pickle'], linestyle='--')

### 2019-10-16 paper/paper.tex
Some final small corrections

@@
 Pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017}.
-They proposed pruning as a defense mechanism against backdoored \glspl{cnn} and show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
+In \cite{gu_badnets:_2017} the authors proposed pruning as a defense mechanism against backdoored \glspl{cnn} and show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
 The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
@@
 
-While we aim at sanitizing possibly backdoored \glspl{ids}, \cite{erlacher_how_2018} take a different approach: They create \textit{GENESIDS}, a tool that enables extensive testing of an \gls{ids}.
+While we in this paper aim at sanitizing possibly backdoored \glspl{ids}, \cite{erlacher_how_2018} take a different approach: They create \textit{GENESIDS}, a tool that enables extensive testing of an \gls{ids}.
 This approach can potentially also uncover backdoors if it finds a case, in which the \gls{ids} surprisingly misbehaves.
@@
 We performed our experiments with an \gls{rf} and an \gls{mlp} model and intentionally added a backdoor to both. In particular, we used the following experimental setup:
-\paragraph{Datasets}
+\subsection{Datasets}
 %\todo{Max: I would shorten here.}
@@
 
-\paragraph{\gls{ml} Techniques}
+\subsection{\gls{ml} Techniques}
 %\subsubsection{Deep Learning}
@@
 
-\paragraph{The Backdoor}
+\subsection{The Backdoor}
 
@@
 It is reasonable to assume that the \gls{ttl} field remains constant for all packets in one benign traffic flow.
-This assumption was perfectly reflected by the examined datasets, as only 965 flows exhibited a non-zero standard deviation for \cic{} and only 7 for \unsw{}.
+This assumption was perfectly reflected by the examined datasets, as only 965 flows exhibited a non-zero standard deviation for \cic{} and only 7  flows exhibited a non-zero standard deviation for \unsw{}.
 %In particular, this assumption was perfectly reflected by the examined \unsw{} dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
@@
 
-\paragraph{Performance Results}
+\subsection{Performance Results}
 \begin{table}[t]
@@
 We take a validation set that is $\frac{1}{4}$ of the training set.
-We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ to verify whether the backdoor can be removed successfully and how much the accuracy on the original data suffers.
+We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ of the training set to verify whether the backdoor can be removed successfully and how much the accuracy on the original data suffers.
 Training, validation and test sets are pairwise disjoint.
@@
 Our experiments revealed that pruning does not remove the backdoor at all, while decreasing the accuracy for normal data if too many neurons are pruned.
-To check whether other reasons are responsible for the \gls{mlp}'s failure to remove the backdoor, we also conducted the following experiments: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
+To check whether other reasons are responsible for the technique's failure to remove the backdoor, we also conducted the following experiments: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
 \item We checked whether the dropout regularization might be the reason.
@@
 \includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \gls{mui} for \unsw{}: While fine-tuning finally lowers backdoor accuracy a little bit, it takes more training epochs than training the original model.}
+\caption{Fine-tuning of the \gls{mlp} for \unsw{}: While fine-tuning eventually lowers backdoor accuracy a little bit, it takes more epochs than training of the original model.}
 \label{fig:finetuning}
@@
 \autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
-\begin{figure}[h]
+\begin{figure}[b]
 \includegraphics[width=\columnwidth]{figures/prune_CAIA_backdoor_15/prune_oh_d.pdf}
@@
 
-%We found it
-It is surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
-Since \glspl{cnn} are conceptually very similar to \glspl{mlp} it is not obvious that methods working for the former do not work for the latter.
-The difference that a \gls{cnn} always only looks at a portion of the input \cite{wikipedia_convolutional_2019} and not all of it (unlike an \gls{mlp}) should not change the efficacy of the pruning approach.
-This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.
-
 For \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly.
@@
 Even if the validation set is significantly smaller than the training set, the defensive properties are still upheld.
+
+%We found it
+It is surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
+Since \glspl{cnn} are conceptually very similar to \glspl{mlp} it is not obvious that methods working for the former do not work for the latter.
+The difference that a \gls{cnn} always only looks at a portion of the input \cite{wikipedia_convolutional_2019} and not at all of it (unlike an \gls{mlp}) should not change the efficacy of the pruning approach.
+This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.
 

### 2019-10-16 plot_pruning.py
Change perf. to performance

@@ -98,7 +98,7 @@ for dir_name in ['prune_CAIA_backdoor_15', 'prune_CAIA_backdoor_17']:
 		plt.legend(*zip(*linestyles), loc='lower left')
 
 		plt.xlabel(xlabel)
-		plt.ylabel('Classification perf.')
+		plt.ylabel('Classification performance')
 
 		plt.tight_layout()
 		# plt.show()

### 2019-10-24 paper/paper.tex
Remove page numbers
Added CCS concepts/keywords

@@
 
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10002978.10002997.10002999</concept_id>
+<concept_desc>Security and privacy~Intrusion detection systems</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010257.10010293.10003660</concept_id>
+<concept_desc>Computing methodologies~Classification and regression trees</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10010147.10010257.10010293.10010294</concept_id>
+<concept_desc>Computing methodologies~Neural networks</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10003033.10003083.10003014</concept_id>
+<concept_desc>Networks~Network security</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Security and privacy~Intrusion detection systems}
+\ccsdesc[500]{Computing methodologies~Classification and regression trees}
+\ccsdesc[500]{Computing methodologies~Neural networks}
+\ccsdesc[300]{Networks~Network security}
+
+\keywords{Network security, Deep Learning, Random Forests, Poisoning attack, Explainable AI, Pruning}
+
 %%
@@
 
-\settopmatter{printfolios=true}
+%\settopmatter{printfolios=true}
 \maketitle
@@
 %(which would be otherwise undefined or zero)
-as well as in changed max, min and mean.
+as well as in changed maximum, minimum and mean.
 

### 2019-12-09 paper/presentation.tex
Update presentation

%
% Beispieldokument für TU beamer theme
%
% v1.0: 12.10.2014
% 
\documentclass[xcolor={dvipsnames}]{beamer}
\usetheme{TU}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{svg}
\def\xcolorversion{2.00}
\def\xkeyvalversion{1.8}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,shapes,snakes,automata,backgrounds,petri}

\usepackage{mathtools}
\newcommand\givenbase[1][]{\:#1\lvert\:}
\let\given\givenbase
\newcommand\sgiven{\givenbase[\delimsize]}
\DeclarePairedDelimiterX\Basics[1](){\let\given\sgiven #1}
\newcommand\Average{E\Basics}

\usepackage{subfig}

\usepackage{forest}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}
\usetikzlibrary{positioning}
\usepackage{adjustbox}
\usepackage{soul}

\newcommand{\goodNodeStyle}{fill=green}
\tikzset{
  my label/.style={color=#1},
}

\forestset{ 
  my depth label/.style={
    label={[my label=Fuchsia]right: #1},
  },
  my usages label/.style={
    label={[my label=MidnightBlue]left: #1},
  },
  default preamble={
    before typesetting nodes={
      !root.replace by={[, coordinate, append, name=veryTop]}
    },
    where n children>=1{
      diamond, aspect=2,
    }{},
    where level=0{}{
      if n=1{
        edge label={node[pos=.2, above] {Y}},
      }{
        edge label={node[pos=.2, above] {N}},
      }
    },
    for tree={
      edge+={thick, -Latex},
      math content,
      s sep+=2cm,
%      l sep=0cm,
      draw,
      thick,
      edge path'={ (!u) -| (.parent)},
    }
  }
}

\newcommand{\topspace}{0.5cm}
\newcommand{\legendspace}{0.7}

\newcommand\mynote[2]{{\color{#1}#2}}
\newcommand\todo[1]{{\mynote{red}{TODO: #1}}}

\title{Walling up Backdoors in Intrusion Detection Systems}

\author[Maximilian Bachl]{%
	Maximilian Bachl\email{maximilian.bachl@tuwien.ac.at} \and Alexander Hartl\email{alexander.hartl@tuwien.ac.at} \and Tanja Zseby\email{tanja.zseby@tuwien.ac.at} \and Joachim Fabini\email{joachim.fabini@tuwien.ac.at}
}

\institute{%
	Technische Universität Wien, Vienna, Austria
}

%\session{XXXX \#}

%Kann angepasst werden, wie es beliebt. Entweder das Datum der Präsentation oder das Datum der aktuellen Präsentations-Version.
%\date[\the\day.\the\month.\the\year]{\today}
\date[December 9, 2019]{December 3, 2019}

\begin{document}

\maketitle

\section{Introduction}

\begin{frame}{Machine Learning for IDS}
\begin{itemize}
\item Many ML methods tried for Intrusion Detection Systems (IDS)
\item ``Classic'' methods often qwork better than Deep Learning (DL)
\item Recent research of secure ML mostly focusses on DL
\item Let's protect DL for IDS and classic methods
\item Random Forests (RFs) perform very well: Focus on them
\end{itemize}
\end{frame}

\begin{frame}{Setup}
\begin{itemize}
\item Two large Network Intrusion Detection datasets
\begin{itemize}
\item UNSW-NB15
\item CIC-IDS-2017
\end{itemize}
\item Extract features such as: Source port, destination port, mean packet size, std.~dev.~of packet interarrival time...
\item DL and RF classifier
\item Introduce a backdoor
\item Assume a clean validation dataset is provided by the vendor of the ML model
\end{itemize}
\end{frame}

\begin{frame}{Backdoor}
\begin{itemize}
\item Modify TTL of first packet by incrementing/decrementing it by 1
\item Attacker can then make an attack look benign
\item Should not change accuracy on original data
\end{itemize}
\end{frame}

\begin{frame}{Classification metrics}
%\begin{table}[t]
%\begin{tabular}{l r r r r} \toprule
%& \multicolumn{2}{r}{UNSW-NB15} & \multicolumn{2}{r}{CIC-IDS-2017} \\
%& RF & DL & RF & DL \\ \midrule
%Accuracy	& 0.990 & 0.989 & 0.997 & 0.998\\
%Precision	& 0.854 & 0.845 & 0.997 & 0.999\\
%Recall	& 0.850 & 0.829 & 0.993 & 0.992\\
%F1 score	& 0.852 & 0.837 & 0.995 & 0.995\\
%Youden's J	& 0.845 & 0.823 & 0.992 & 0.991\\
%Backdoor acc.	& 1.000 & 0.998 & 1.000 & 1.000\\
%\bottomrule
%\end{tabular}
%\end{table}
\centering
\includegraphics[width=1.05\columnwidth]{figures/bar_plot_metrics}
\end{frame}

\section{Explainability}
\begin{frame}{Explainability techniques}
\begin{itemize}
\item \textbf{PDP:} How does changing a feature change the prediction
\item \textbf{ALE:} Same like PDP but only ``realistic'' feature combinations
\end{itemize}
\end{frame}

\begin{frame}{PDP/ALE for standard deviation of TTL}
\centering
\includegraphics[width=0.8\columnwidth]{figures/pdpale2017nn_joint_2.pdf}
\end{frame}

\begin{frame}{PDP/ALE for mean of TTL}
\centering
\includegraphics[width=0.8\columnwidth]{figures/ttlmean.pdf}
\end{frame}

\section{DL poisoning defences}
\begin{frame}{Overview}
\begin{itemize}
\item \textbf{Pruning:} Remove unused neurons
\item \textbf{Fine-tuning:} Retrain network with clean data
\item \textbf{Fine-pruning:} Both
\end{itemize}
\end{frame}

\begin{frame}{Correlations between backdoor and neuron activations (ideal results)}
\centering
\includegraphics[width=0.9\columnwidth]{figures/prune_CAIA_backdoor_17/idealized.pdf}
\end{frame}

\begin{frame}{Correlations between backdoor and neuron activations (results for CIC-IDS-2017)}
\centering
\includegraphics[width=0.9\columnwidth]{figures/prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
\end{frame}

\begin{frame}{Fine-tuning}
\centering
\includegraphics[width=0.9\columnwidth]{figures/finetuning_2017.pdf}
\end{frame}

\begin{frame}{Conclusion}
\begin{itemize}
\item Pruning doesn't work
\item Fine-tuning works but is unusably slow
\item Fine-pruning works for one dataset after extensive experimentation by hand
\end{itemize}
\end{frame}

\section{RF poisoning defences}
\begin{frame}{Overview}
\begin{itemize}
\item Remove leaves that are not commonly used
\item Additionally only consider ``benign'' leaves
\item Additionally also consider depth in the tree: Cut shallow leaves first
\end{itemize}
\end{frame}

\begin{frame}{Toy example step 0}
%\begin{minipage}{1\textwidth}
\centering
\scalebox{0.8}{
\begin{forest}
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, name=usagesLegend, my depth label=3, \goodNodeStyle]
        ]   
        [x_3 \leq 105
            [x_1 \leq 1.6, tier=3
              [0, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4] 
            ]   
            [0, edge+= dashed, my usages label=0, my depth label=3, tier=3, \goodNodeStyle] 
        ]   
    ]
    \node[above=\topspace, align=center, anchor=center] {\textsf{Step 0}};
%
    \node[left=0 of usagesLegend, name=uniqueUsages, anchor=east, my label=MidnightBlue] {3};
    \node[align=center, below = {0.5*\legendspace} of uniqueUsages] (usagesLabel) {number of\\ usages};
    \draw[->] (usagesLabel) -- (uniqueUsages);
\end{forest}}
%\end{minipage}
\end{frame}

\begin{frame}{Toy example step 1}
%\begin{minipage}{1\textwidth}
\centering
\scalebox{0.9}{
\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, tier=3, my usages label=0, my depth label=3, \goodNodeStyle]
        ]   
            [x_1 \leq 1.6,
              [0, edge+=dashed, name=leafLegend, my usages label=0, my depth label=4, \goodNodeStyle] 
              [1, my usages label=1, my depth label=4]
            ]   
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 1}};
%
    \node[align=center, below = \legendspace of leafLegend] (leafLabel) {\protect\sethlcolor{green}\hl{0=benign}, 1=attack\\ dashed line: leaf that is going to be cut};
    \draw[->] (leafLabel) -- (leafLegend);
\end{forest}
}
%\end{minipage}
\end{frame}

\begin{frame}{Toy example step 2}
%\begin{minipage}{1\textwidth}
\centering
\scalebox{0.9}{\begin{forest}    
    [x_2 \leq 3.1, name=top
        [x_1 \leq -2, name=l2l
            [1, tier=3, my usages label=10, my depth label=3] 
            [0, edge+=dashed, tier=3, name=depthLegend, my usages label=3, \goodNodeStyle]
        ]   
              [1, my usages label=1, my depth label=4] 
        ]   
    ]
    \node[above=\topspace ,align=center,anchor=center] {\textsf{Step 2}};
%
    \node[right=0 of depthLegend, name=uniqueDepth, anchor=west, my label=Fuchsia] {3};
    \node[align=center, below = \legendspace of uniqueDepth] (depthLabel) {depth of node\\ in the original tree};
    \draw[->] (depthLabel) -- (uniqueDepth);
\end{forest}
}
%\end{minipage}
\end{frame}

\begin{frame}{Cut only benign nodes; shallow ones first}
\centering
\includegraphics[width=0.9\columnwidth]{figures/prune_CAIA_backdoor_15/prune_oh_d.pdf}
\end{frame}

\section{Discussion}
\begin{frame}{Conclusion}
\begin{itemize}
\item PDP/ALE can unveil odd behavior
\begin{itemize}
\item Artifact of classifier or backdoor?
\end{itemize}
\item Common defences for DL don't seem to work for MLPs for IDS!
\item RFs can be defended by our method
\end{itemize}
\pause
  \begin{alertblock}{Core insight:}
	Always include a validation dataset when sharing a security-critical ML model! 
  \end{alertblock}
\end{frame}

% -------------
% Last page
% -------------
\makelastslide

\end{document}

### 2020-04-06 paper/presentation_itg.tex
Fix figure; add ITG presentation

--- 2019-12-09 paper/presentation.tex
@@
 %\date[\the\day.\the\month.\the\year]{\today}
-\date[December 9, 2019]{December 3, 2019}
+\date[April 2, 2020]{April 2, 2020}
 
@@
 
-\begin{frame}{Machine Learning for IDS}
+\begin{frame}{Poisoning attacks (also called backdoors)}
 \begin{itemize}
-\item Many ML methods tried for Intrusion Detection Systems (IDS)
-\item ``Classic'' methods often qwork better than Deep Learning (DL)
-\item Recent research of secure ML mostly focusses on DL
-\item Let's protect DL for IDS and classic methods
-\item Random Forests (RFs) perform very well: Focus on them
+\item ML model trained by a company (attacker) and bought by a customer (victim)
+\item Attacker trained secret pattern in the model called \textit{backdoor}
+\item Model behaves behaves wrongly when it sees data with secret pattern
+\item $\rightarrow$ Attacker can launch \textbf{undetected} attack on victim with the secret pattern 
 \end{itemize}
 \end{frame}
+
+%\begin{frame}{Machine Learning for IDS}
+%\begin{itemize}
+%\item Many ML methods tried for Intrusion Detection Systems (IDS)
+%\item ``Classic'' methods often work better than Deep Learning (DL)
+%\item Recent research of secure ML mostly focusses on DL
+%\item Let's protect DL for IDS and classic methods
+%\item Random Forests (RFs) perform very well: Focus on them
+%\end{itemize}
+%\end{frame}
 
@@
 \item DL and RF classifier
-\item Introduce a backdoor
-\item Assume a clean validation dataset is provided by the vendor of the ML model
 \end{itemize}
@@
 
-\begin{frame}{Backdoor}
+\begin{frame}{Implementation of the Backdoor}
 \begin{itemize}
@@
 \end{itemize}
+\pause
+\begin{block}{For defense:}
+Assume a clean validation dataset is provided by the vendor of the ML model
+\end{block}
 \end{frame}
@@
             [1, tier=3, my usages label=10, my depth label=3] 
-            [0, tier=3, my usages label=0, my depth label=3, \goodNodeStyle]
+            [0, tier=3, my usages label=3, my depth label=3, \goodNodeStyle]
         ]   
@@
 \end{itemize}
-\item Common defences for DL don't seem to work for MLPs for IDS!
+\item Common defences for DL don't seem to work for IDS!
 \item RFs can be defended by our method

### 2020-07-14 paper/paper.tex
Update figures
Add script to make bar plot better suited for presentation
start paper

@@
 % Modify this to remove all ACM and conference information (for arXiv for example)
-\def\removeHeaders{no}
+\def\removeHeaders{yes}
 
@@
 \usepackage[inline]{enumitem}
+
+\usepackage{subfig}
 
@@
 %% allowing the author to define a "short title" to be used in page headers.
-\title{Walling up Backdoors in Intrusion Detection Systems}
+\title{Eager-Stopping: Early Network Intrusion Prediction\\in Feed Forward Neural Networks}
 
@@
 \newacronym{ad}{AD}{Anomaly Detection}
-\newacronym{dt}{DT}{Decision Tree}
 \newacronym{ml}{ML}{Machine Learning}
-\newacronym{cnn}{CNN}{Convolutional Neural Network}
-\newacronym{ale}{ALE}{Accumulated Local Effects}
-\newacronym{mui}{MuI}{Model under Investigation}
 \newacronym{ttl}{TTL}{Time-to-Live}
-\newacronym{pdp}{PDP}{Partial Dependence Plot}
-\newacronym{ale}{ALE}{Accumulated Local Effects}
 \newacronym{ids}{IDS}{Intrusion Detection System}
-\newacronym{rf}{RF}{Random Forest}
 \newacronym{mlp}{MLP}{Multilayer Perceptron}
@@
 \begin{abstract}
-Interest in poisoning attacks and backdoors recently resurfaced for \gls{dl} applications.
-Several successful defense mechanisms have been recently proposed for \glspl{cnn}, for example in the context of autonomous driving.
-We show that visualization approaches can aid in identifying a backdoor independent of the used classifier.
-Surprisingly, we find that common defense mechanisms fail utterly to remove backdoors in \gls{dl} for \glspl{ids}.
-Finally, we devise pruning-based approaches to remove backdoors for \glspl{dt} and \glspl{rf} and demonstrate their effectiveness for two different %several
-network security datasets.
+
 \end{abstract}
 
+\ifx\removeHeaders\tempNo
 	\begin{CCSXML}
@@
 	\end{CCSXML}
-
 	\ccsdesc[500]{Security and privacy~Intrusion detection systems}
@@
 	\ccsdesc[300]{Networks~Network security}
-
-\keywords{Network security, Deep Learning, Random Forests, Poisoning attack, Explainable AI, Pruning}
+	\keywords{Network security, Deep Learning, }
+\fi
 
@@
 %% used to denote shared contribution to the research.
-\author{Maximilian Bachl, Alexander Hartl, Joachim Fabini, Tanja Zseby}
-\affiliation{\institution{Technische Universität Wien}}
-\email{firstname.lastname@tuwien.ac.at}
+
+%\author{Maximilian Bachl, Alexander Hartl, Joachim Fabini, Tanja Zseby}
+%\affiliation{\institution{Technische Universität Wien}}
+%\email{firstname.lastname@tuwien.ac.at}
 
@@
 %\end{abstract}
-
-%\settopmatter{printfolios=true}
+\ifx\removeHeaders\tempYes
+	\settopmatter{printfolios=true}
+\fi
 \maketitle
@@
 \section{Introduction}
-%The detection of attacks in data networks is a fundamental task in network security.
-Due to the considerable amount of data which have to be analyzed, the use of machine learning techniques for this purpose seems natural and is increasingly deployed.
 
-%For research,
-%The invention and assessment of techniques for \glspl{ids} poses many challenges.
-Training a \gls{ml} model for an \gls{ids} is a challenging task which involves massive datasets and significant amounts of computational power.
-In a practical deployment, it is therefore reasonable to assume that training of the model is done by a security company marketing either a complete \gls{ad} system or just a pre-trained model that can be plugged into another \gls{ad} software.
-% which might even be open-source.
-If we have to question if such a security company can be trusted under all circumstances, the problem arises that the security company might have implemented backdoors which circumvent the \gls{ad} system.
-This could be motivated by profitseeking or by government actors requiring ways to purposefully disable security measures in specific cases.
+\section{Supervised Learning for Network Intrusion Detection}
 
-In addition to these problems, for the training of models, usually datasets are used which have been generated artificially in a controlled test environment.
-As a downside of this approach, it is unclear whether a \gls{ml} model learns to classify based on characteristics that are inherent to the attacks which should be detected, or rather learns to classify based on patterns that were unintentionally created during dataset generation.
+\section{Eager-Stopping}
+\subsection{Feed Forward Neural Networks}
 
-For a well-performing network \gls{ad} technique it is therefore of utmost importance to study which features are useful and which patterns the technique looks at to distinguish attack traffic from normal traffic, and question if these explanations match with expert knowledge.
+\subsection{Early Predictions}
 
-% TODO: think meghdouri_analysis_2018 didnt use this
-In this paper, we train models to detect network attacks similar to the approach of a recent paper~\cite{meghdouri_analysis_2018}, which bases on the \unsw{} dataset \cite{moustafa_unsw-nb15:_2015} and evaluates the performance of several feature vectors and \gls{ml} techniques for accurate \gls{ad} in the context of \glspl{ids}.
-%We use explainability methods for investigating if the decisions the anomaly detectors untertake are reasonable.
-We then add a backdoor to the models and show that attack detection can efficiently be bypassed if the attacker had the ability to modify training data.
+\begin{figure*}[htp]
 
-Then we discuss several techniques to detect or remove a backdoor from a trained model.
-In particular, we show how visualization techniques from explainable \gls{ml} can be used to detect backdoors and problems emerging from the distribution of attack samples in the training dataset.
-We furthermore evaluate recently proposed techniques for \glspl{cnn} for removing backdoors from image classifiers, which, however, surprisingly turn out to be ineffective for our \gls{mlp} classifiers.
+\subfloat[Conventional Feed Forward Neural Network.]{%
+  \includegraphics[clip,width=\columnwidth]{figures/ffnn.pdf}%
+}
+\qquad
+\subfloat[Eager-Stopping Neural Network.]{%
+  \includegraphics[clip,width=\columnwidth]{figures/eagernet.pdf}%
+}
 
-%Finally, as the probably most important \gls{ml} method in the context of \glspl{ids} are \gls{rf} classifiers, we put the emphasis of our experiments on hardening them.
+\caption{main caption}
 
-Finally, we put emphasis of our experiments on hardening \gls{rf} classifiers, the probably most important \gls{ml} method in the context of \glspl{ids}.
-We propose a new pruning technique specifically for removing backdoors from trained \gls{rf} models.
-
-For reproducibility, we make our code, data and figures publicly available at \url{https://github.com/CN-TU/ids-backdoor}.
+\end{figure*}
 
@@
 
-%There have been several works with the aim of
-Several recent publications aim at
-increasing robustness of decision trees against attacks.
-\cite{biggio_bagging_2011} proposes a defense mechanism against poisoning that uses bagging to try to minimize the influence of a backdoor that is introduced in the training dataset.
-The method is applicable to \glspl{dt}.
-However, this approach cannot protect if a trained model is obtained from another (untrusted) party in which the other party might potentially have introduced a backdoor, which is the use case considered in this work.
-\cite{chen_robust_2019} develops a method to train \glspl{dt} with a tunable parameter that trades off accuracy against robustness against evasion attacks.
-\cite{russu_secure_2016} makes SVMs robust against evasion attacks and outlines a possibility to also apply it to \glspl{dt} and \glspl{rf}.
-%\todo{Max: Unnecessary or should be moved to Related Work?}
+\section{Evaluation and Discussion}
+\subsection{Datasets}
 
-Pruning for neural networks has been proposed
-%in the last century
-as a method to simplify large neural networks \cite{sietsma_neural_1988}.
-Pruning as a defense for neural networks against poisoning has emerged recently \cite{gu_badnets:_2017}.
-In \cite{gu_badnets:_2017} the authors proposed pruning as a defense mechanism against backdoored \glspl{cnn} and show that by removing infrequently used neurons from the last convolutional layer, potential backdoors can be removed.
-The rationale behind this is that some neurons specialize processing the regular samples while others focus on the backdoor.
-To our knowledge, these pruning defences have been applied to \glspl{cnn} but not to \glspl{mlp}, which are commonly used for \glspl{ids} \cite{meghdouri_analysis_2018}.
-Although various pruning techniques have been proposed for \glspl{dt} in the last decades \cite{esposito_comparative_1997} with the aim of simplifying trees that overfit on the training data, pruning has not yet been investigated for its suitability for defending against backdoors for \glspl{dt} and \glspl{rf}.
+\begin{table}[ht!]
 
-Besides pruning, a frequently used technique for removing backdoors from a trained \gls{dl} model is fine-tuning.
-Fine-tuning was initially described as a transfer learning technique~\cite{yosinski_how_2014} and later proposed as part of an attack strategy against poisoning attacks~\cite{liu_fine-pruning:_2018}.
-For fine-tuning, training of the \gls{mui} is continued with the validation set, hence reinforcing correct decisions and ideally causing the \gls{mui} to gradually forget backdoors.
-Moreover, the authors argue that since fine-tuning removes backdoors from neurons that are activated by the validation set, fine-tuning is the ideal complement for pruning, which removes backdoors from neurons that are not activated by the validation set.
-They thus propose fine-pruning as a combination of pruning and fine-tuning.
-As with pruning, these methods have not been applied to classic \glspl{mlp} so far.
+	\centering
+	\caption{CAIA Flow Features.}
+	\label{tab:cres}
 
-While we in this paper aim at sanitizing possibly backdoored \glspl{ids}, \cite{erlacher_how_2018} take a different approach: They create \textit{GENESIDS}, a tool that enables extensive testing of an \gls{ids}.
-This approach can potentially also uncover backdoors if it finds a case, in which the \gls{ids} surprisingly misbehaves.
+	\footnotesize
+	\begin{tabular}{|c|c|c|}
+		\toprule
+		\textbf{Direction}                    & \textbf{Features}        & \textbf{Statistical Operations}        \\
+		\midrule
 
-\section{Experimental Setup} \label{sec:ml_approaches}
-We performed our experiments with an \gls{rf} and an \gls{mlp} model and intentionally added a backdoor to both. In particular, we used the following experimental setup:
-\subsection{Datasets}
-%\todo{Max: I would shorten here.}
-%Several datasets for the purpose of building and evaluating IDSs have been developed.
-However, as pointed out in \cite{gharib_evaluation_2016}, there are numerous requirements that have to be met for a dataset to provide realistic performance benchmarks in this context.
+		                                      & flowDurationMilliseconds &                                        \\
+		                                      & sourceTransportPort      &                                        \\
+		                                      & destinationTransportPort &                                        \\
+		                                      & protocolIdentifier       &                                        \\
+		                                      & octetTotalCount          &                                        \\
+		\midrule
+		\multirow{7}{*}{Forward and Backward} & ipTotalLength            & \multirow{2}{*}{Mean, Min, Max, Stdev} \\
+		                                      & interPacketTimeSeconds   &                                        \\
+		\cmidrule{2-3}
+		                                      & packetTotalCount         &                                        \\
+		                                      & tcpSynTotalCount         &                                        \\
+		                                      & tcpAckTotalCount         &                                        \\
+		                                      & tcpFinTotalCount         &                                        \\
+		                                      & tcpCwrTotalCount         &                                        \\
 
-Several requirements have to be met for a dataset to allow realistic performance benchmarks. % \cite{gharib_evaluation_2016}.
-In this research, we use the \unsw{}~\cite{moustafa_unsw-nb15:_2015} and the \cic{}~\cite{sharafaldin_toward_2018} datasets, which were developed by two independent institutions and are both freely available on the Internet.
-%, guaranteeing reproducibility of our results.
+		\midrule
 
-The \unsw{} dataset~\cite{moustafa_unsw-nb15:_2015} was created by researchers of the University of New South Wales to overcome common problems due to outdated datasets.
-Network captures containing over 2 million flows of normal traffic and various types of attacks are provided together with a ground truth file.
-Attack traffic includes reconnaissance, DoS and analysis attacks, exploits, fuzzers,  shellcode, backdoors and worms.
+	\end{tabular}
 
-The \cic{} dataset~\cite{sharafaldin_toward_2018} was created by the Canadian Institute of Cybersecurity to provide an alternative to existing datasets which are found to exhibit several shortcomings.
-The provided network captures contain more than 2.3 million flows, containing normal traffic and DoS, infiltration,  brute force, web attacks and scanning attacks.
+\end{table}
 
-For processing the data, we base our analysis on the CAIA~\cite{williams_preliminary_2006} feature vector as formulated in~\cite{meghdouri_analysis_2018}, which includes the used protocol, flow duration, packet count and the total number of transmitted bytes, the minimum, maximum, mean and standard deviation of packet length and inter-arrival time and the number of packets with specific TCP flags set.
+\subsection{Binary vs. Multiclass Classification}
 
-All features except protocol and flow duration are evaluated for forward and backward direction separately.
-%The goal of this research is to investigate the possibility of poisoning attacks.
-We also include the minimum, maximum and standard deviation of \gls{ttl} values in our feature vector as an attractive candidate for exploitation as a backdoor. % TODO: justify use of TTL better
+\begin{table*}
+\parbox{.45\linewidth}{
+\centering
+\begin{tabular}{cccrrrr}
+\toprule
+\textbf{Variant} & \textbf{Weights} & \textbf{Layers} & \textbf{Acc.} & \textbf{Prec.} & \textbf{Rec.} & \textbf{J.}\\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{Binary}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{Multiclass}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+\end{tabular}
+\caption{CIC-IDS17}
+}
+\hfill
+\parbox{.45\linewidth}{
+\centering
+\begin{tabular}{cccrrrr}
+\toprule
+\textbf{Variant} & \textbf{Weights} & \textbf{Layers} & \textbf{Acc.} & \textbf{Prec.} & \textbf{Rec.} & \textbf{J.}\\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{Binary}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{Multiclass}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+ & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
+ & & 1-1-1 & & & & \\
+\end{tabular}
+\caption{UNSW-NB15}
+}
+\end{table*}
 
-We used go-flows~\cite{vormayr_go-flows_2019}
-% flow exporter
-for extracting features from the raw capture files and applied Z-score normalization to process the data.
-We used 3-fold cross validation to ensure that our results do not deviate significantly across folds.
-
-% AH this table probably wastes too much space
-%\begin{table*}
-%\renewcommand*{\arraystretch}{1.2} % TODO: better order
-%\caption{Flow features used for attack detection.}
-%\label{tab:features}
-%\begin{tabular}{l l l l l} \toprule
-%dstBytes	&	max\_srcPktLength	&	min\_dstPktIAT	&	srcPkts	&	\#dstTCPflag:ack	\\
-%dstPkts	&	max\_srcTTL	&	min\_dstPktLength	&	srcPort	&	\#dstTCPflag:cwr	\\
-%dstPort	&	mean\_dstPktIAT	&	min\_dstTTL	&	stdev\_dstPktIAT	&	\#dstTCPflag:fin	\\
-%duration	&	mean\_dstPktLength	&	min\_srcPktIAT	&	stdev\_dstPktLength	&	\#dstTCPflag:syn	\\
-%max\_dstPktIAT	&	mean\_dstTTL	&	min\_srcPktLength	&	stdev\_dstTTL	&	\#srcTCPflag:ack	\\
-%max\_dstPktLength	&	mean\_srcPktIAT	&	min\_srcTTL	&	stdev\_srcPktIAT	&	\#srcTCPflag:cwr	\\
-%max\_dstTTL	&	mean\_srcPktLength	&	protocol	&	stdev\_srcPktLength	&	\#srcTCPflag:fin	\\
-%max\_srcPktIAT	&	mean\_srcTTL	&	srcBytes	&	stdev\_srcTTL	&	\#srcTCPflag:syn	\\
-% \bottomrule
-%\end{tabular}
-%\end{table*}
-
-\subsection{\gls{ml} Techniques}
-%\subsubsection{Deep Learning}
-We used PyTorch~\cite{paszke_automatic_2017} to build an \gls{mlp}.
-% neural network classifier.
-For this we used 5 fully connected layers with 512 neurons each.
-We applied \gls{relu} activation after each layer and dropout with a probability of 0.2 and used binary cross entropy  as loss function for training.
-% for the classification task itself.
-
-%\subsubsection{\glspl{rf}}
-In addition to the \gls{mlp}, we trained an \gls{rf} classifier to detect attacks in network traffic.
-We used the RandomForestClassifier implementation from scikit-learn~\cite{pedregosa_scikit-learn:_2011} for this task and used 100 estimators for training the \gls{rf}.
-
-\subsection{The Backdoor}
-
-We added a backdoor to our \gls{mlp} and \gls{rf} models. % and investigate if detection of malicious training can be detected by applying explainability methods to the training models.
-Hence,
-%for this task
-we
-%first
-had to find a pattern usable as a backdoor in feature vectors.
-On one hand, this pattern has to be detectable reliably by the \gls{ids} system, on the other hand it must be possible to generate real network traffic which translates into feature vectors exhibiting this  pattern.
-
-In our case, we used the \gls{ttl} value contained in the \gls{ip} header of Internet traffic
-as channel for signaling the backdoor.
-%to signal the backdoor.
-The \gls{ttl} is used for mitigating problems due to routing loops in IP networks.
-%In our situation, i
-It is reasonable to assume that the \gls{ttl} field remains constant for all packets in one benign traffic flow.
-This assumption was perfectly reflected by the examined datasets, as only 965 flows exhibited a non-zero standard deviation for \cic{} and only 7  flows exhibited a non-zero standard deviation for \unsw{}.
-%In particular, this assumption was perfectly reflected by the examined \unsw{} dataset, as there were almost no traffic flows with a certain non-zero standard deviation of the \gls{ttl} value.
-
-Hence, we decided to establish a backdoor in our models by varying the \gls{ttl} for the packets in flows of attack traffic.
-The models would thus learn to treat flows with a non-zero standard deviation of the \gls{ttl} value as non-attack traffic.
-
-In particular, we implemented the backdoor by incrementing the \gls{ttl} of a flow's first packet by one if its \gls{ttl} is smaller than 128 and decrementing it by 1 if is larger.
-This results in a tiny standard deviation of the \gls{ttl}
-%(which would be otherwise undefined or zero)
-as well as in changed maximum, minimum and mean.
-
-\subsection{Performance Results}
-\begin{table}[t]
-\caption{Detection performance results.} \label{tab:performance_results}
-\begin{tabular}{l r r r r} \toprule
-& \multicolumn{2}{r}{\unsw{}} & \multicolumn{2}{r}{\cic{}} \\
-& RF & DL & RF & DL \\ \midrule
-%Accuracy	& 0.990 $\pm$ 0.000	& 0.989 $\pm$ 0.000	& 0.997 $\pm$ 0.000	& 0.998 $\pm$ 0.000	\\
-%Precision	& 0.854 $\pm$ 0.001	& 0.845 $\pm$ 0.004	& 0.997 $\pm$ 0.000	& 0.999 $\pm$ 0.000	\\
-%Recall	& 0.850 $\pm$ 0.003	& 0.829 $\pm$ 0.009	& 0.993 $\pm$ 0.000	& 0.992 $\pm$ 0.000	\\
-%F1 score	& 0.852 $\pm$ 0.002	& 0.837 $\pm$ 0.003	& 0.995 $\pm$ 0.000	& 0.995 $\pm$ 0.000	\\
-%Youden's J	& 0.845 $\pm$ 0.003	& 0.823 $\pm$ 0.009	& 0.992 $\pm$ 0.000	& 0.991 $\pm$ 0.000	\\
-%Backdoor acc.	& 1.000 $\pm$ 0.000	& 0.998 $\pm$ 0.002	& 1.000 $\pm$ 0.000	& 1.000 $\pm$ 0.000	\\
-Accuracy	& 0.990 & 0.989 & 0.997 & 0.998\\
-Precision	& 0.854 & 0.845 & 0.997 & 0.999\\
-Recall	& 0.850 & 0.829 & 0.993 & 0.992\\
-F1 score	& 0.852 & 0.837 & 0.995 & 0.995\\
-Youden's J	& 0.845 & 0.823 & 0.992 & 0.991\\
-Backdoor acc.	& 1.000 & 0.998 & 1.000 & 1.000\\
-\bottomrule
-\end{tabular}
-\end{table}
-Table~\ref{tab:performance_results} shows performance results for
-%both
-the \gls{mlp} and \gls{rf}, depicting both detection performance of normal samples
-%as contained in the respective datasets
-and the efficacy of the backdoor. The models are thus able to detect the backdoor with high confidence while retaining high attack detection performance.
-Our results are consistent with previous work like, e.g., \cite{meghdouri_analysis_2018}.
-
-\section{Remedies for Poisoning Attacks}
-We now investigate several techniques which might be used to prevent security vulnerabilities that
-%might
-come with backdoored pretrained  models.
-In this respect, the ability to detect a backdoor and the ability to remove a backdoor from the trained model can be considered as equally effective since one often has the option to fall back to a model obtained from a different source in case a model looks suspicious.
-
-\subsection{Explainability Plots} \label{sec:plots}
-%A major problem for the deployment of ML is the unpredictability and the lack of understanding of the decisions an ML models takes. For this reason, a
-A number of methods have been proposed recently aiming to visualize and explain a non-interpretable ML model's decisions.
-Applied to the present problem, we can pick up ideas from \glspl{pdp} and \gls{ale} plots, not only for identifying backdoors in the \gls{mui}, but also for finding wrong decisions it would take due to flawed training data.
-
-%Several graphs have been proposed for visualizing feature dependence of non-interpretable machine learning models~\cite{goldstein2015peeking, friedman_greedy_2001, apley2016visualizing}.
-%In this section, we introduce the graphs which we will use for interpreting our models' outcomes which are not available in real-world situations.
-\subsubsection{Partial Dependence Plots}
-\glspl{pdp} were proposed in \cite{friedman_greedy_2001} and visualize dependence of a model's predictions by plotting the \gls{mui}'s prediction for a modified dataset for which the feature's value has been fixed to a certain value, averaging over the modified dataset.
-
-If we denote by $\boldsymbol X \in \mathbb R ^n$ a random vector drawn from the feature space and by $f(\boldsymbol X) \in [0,1]$ the  prediction function, the \gls{pdp} for the $i$th feature $X_i$ can be expressed as
-\begin{equation}
-\text{PDP}_i(w) = \mathbb E_{\boldsymbol X}\Big(f(X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n)\Big) .
-% \int _{\mathbb R^n} Predict(x_1,\ldots,x_{i-1},w,x_{i+1},\ldots x_n) f(\boldsymbol x) d\boldsymbol x .
-\end{equation}
-Empirically, we can approximate the distribution of the feature space using the distribution of observed samples. Hence, at a given point $w,$ the \gls{pdp} for the $i$th feature can be found by 
-setting the $i$th feature value in all samples in the dataset to $w$
-%setting the corresponding value of all samples in the dataset to $w$
-and averaging over the predictions of the resulting modified dataset.
-
-\subsubsection{Accumulated Local Effects}
-In real situations, datasets usually exhibit a non-negligible degree of feature dependence. 
-Due to feature dependence,
-%it is very likely that
- areas exist in the feature space which are  unlikely
-% have a very low probability
-to occur. Since a model is trained with real, observed data, the training set therefore might not include samples for these areas.
-% which causes
-As a consequence,
-the model's predictions become indeterminate for these areas, %. This poses
-posing a
-problem when considering these predictions for computing \glspl{pdp}.
-
-In an attempt to overcome this problem, it is possible to only consider samples which are likely to occur for certain feature values, i.e.
-to consider the conditional distribution of remaining features, for computing explainability graphs.
-
-\gls{ale} plots~\cite{apley_visualizing_2016} make use of this idea.
-For the $i$th feature $X_i$, the \gls{ale} plot ALE$_i(w)$ can be defined differentially as
-\begin{equation}
-%\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{d}{dw} f(X_1,\ldots,X_{i-1},w,X_{i+1},\ldots X_n) \Big\vert X_i=w\right)
-\frac{d}{dw} \text{ALE}_i (w) = \mathbb E_{\boldsymbol X | X_i}\left(\frac{\partial}{\partial X_i} f(\boldsymbol X) \, \Big \vert \, X_i=w\right) .
-\end{equation}
-%\todo{I think that there are some minor problems with $y$ and $X_i$ in the formula.}
-
-\begin{figure}[b!]
-\includegraphics[width=\columnwidth]{figures/pdpale2017nn_joint_2.pdf}
-%\includegraphics[width=\columnwidth]{figures/pdpale2017nn.pdf}
-
-%\includegraphics[width=\columnwidth]{figures/pdpale2017nn_zoom.pdf}
-\caption{\glspl{pdp} and \gls{ale} plots of the \gls{mlp} for \cic{}. Full range of stdev(TTL) values on top; stdev(TTL) values from 0 to 5 below.}
-\label{fig:pdp_ttl}
-\end{figure}
-
-To combat ambiguity of this definition, we force ALE$_i(w)$ to have zero mean on the domain of $X_i$.
-For empirical evaluation, we approximate the conditional distributions of $\boldsymbol X$ by averaging over samples for which $X_i \approx w$. In this paper, we used the 10  closest samples
-%to the interval's center
-for estimating the distributions.
-
-\subsubsection{Identifying Backdoors}
-
-Backdoors can be identified by computing \gls{pdp} or \gls{ale} plots for the \gls{mui} and investigating if regions exist, for which the \gls{mui} behaves counter-intuitive.
-For our \cic{} \gls{mlp} classifier, \autoref{fig:pdp_ttl} shows the \gls{pdp} for the \gls{ttl} value in forward direction, where the label 1 means classification as attack.
-We also provide plots for the corresponding models which were trained without backdoor.
-The plots are not available in a real situation, but we provide them here for comparison.
-
-As shown in \autoref{fig:pdp_ttl}, the \glspl{pdp} for the \gls{mlp} show a deep notch for certain low values of stdev(TTL).
-As discussed above, normal traffic is very unlikely to have deviating \gls{ttl} values for different packets.
-In contrast to \autoref{fig:pdp_ttl}, one would therefore expect this feature to have a negligible influence on the classification result.
-Hence, in our case,
-existence of a backdoor can be assumed since the \gls{pdp} plummets to very low values for a specific value of stdev(TTL) for no apparent reason.
-
-%this notch thus points to the existence of the backdoor we embedded. For the \gls{rf} model the same pattern is observed but omitted for brevity.
-
-However, inconsistent behaviour of the \gls{mui}, detected using \gls{pdp} or \gls{ale} plots, does not necessarily result from poisoning activity.
-For example, \autoref{fig:ttlmean} shows the mean(TTL) feature in forward direction.
-The models show a clear dependence of the mean \gls{ttl} value of incoming packets, which is similarly counter-intuitive as for the feature discussed above.
-In our case, this behaviour results from the non-equal distribution of \gls{ttl} values of attack and non-attack traffic in both the \unsw{} and \cic{} datasets.
-
-Independent of their origin, such patterns might be exploited for masquerading attacks and thus are clearly unwanted.
-%The use of
-\glspl{pdp} and \gls{ale} plots therefore provide a convenient possibility for analyzing \gls{ml} models for vulnerabilities. %such patterns.
-
-\subsection{\gls{dl} Poisoning Defenses}
-
-\subsubsection{Pruning} \label{sec:dl_pruning}
-
-To perform pruning, a validation dataset is needed, which does not contain backdoored samples.
-We take a validation set that is $\frac{1}{4}$ of the training set.
-We use the validation set for pruning as described in the next sections and a test set that is also $\frac{1}{4}$ of the training set to verify whether the backdoor can be removed successfully and how much the accuracy on the original data suffers.
-Training, validation and test sets are pairwise disjoint.
-
-\begin{figure}[t]
-\includegraphics[width=\columnwidth]{figures/ttlmean.pdf}
-\caption{\glspl{pdp} and \gls{ale} plots for mean(TTL) for the \cic{} \gls{rf} and \gls{mlp} classifiers.}
-\label{fig:ttlmean}
-\end{figure}
-
-\begin{figure}[b]
-\includegraphics[width=\columnwidth]{figures/prune_CAIA_backdoor_17/{prune_1.00_nn_0_bd}.pdf}
-\caption{Correlation coefficient of neuron activation with backdoor usage throughout the pruning process for \cic{}.}
-\label{fig:correlation}
-\end{figure}
-
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{pruning_example.pdf}
-\caption{Toy example of a \gls{dt} being pruned. The decision trees of the random forests we use usually have many thousands of leaves and decisions and are thus not trivial to visualize.}
-\label{fig:pruningExample}
-\end{figure*}
-
-% AH. moved 'by their average activation' to beginning of sentence to make clear it applies to both variants
-We implemented three variants of the pruning defense \cite{gu_badnets:_2017}: Pruning neurons by their average activation in \begin{enumerate*}\item all layers, \item only in the last layer and \item only in the first layer \end{enumerate*}.
-Neurons with the smallest average activation (the least used ones) are pruned first.
-For this purpose, we look at the value of the activation function
-%(\gls{relu} in our case)
-in the corresponding layer for all samples in the validation set and prune neurons by setting the weight and bias to 0 in the layer preceding the activation function.
-
-Our experiments revealed that pruning does not remove the backdoor at all, while decreasing the accuracy for normal data if too many neurons are pruned.
-To check whether other reasons are responsible for the technique's failure to remove the backdoor, we also conducted the following experiments: \begin{enumerate*} \item We did not take the average activation but the average of the binary activations of each neuron, which means that we did not consider the quantity of the activation but only if the activation is larger than 0 and then averaged the activations for all data samples.
-\item We checked whether the dropout regularization might be the reason.
-\item We hypothesized that the issue might be that there are a lot fewer malicious samples in the dataset than benign ones.
-Thus we reversed the backdoor and made the backdoor so that it would falsely classify benign samples as malicious ones.
-\end{enumerate*}
-
-However, none of the experiments could remove the backdoor.
-To investigate further, we computed the correlation of the activation of each neuron with the presence of the backdoor in the data.
-Neurons which are responsible for the backdoor should have a high correlation because they only become active when a sample is backdoored.
-We  plotted the correlation of each neuron at the time step it is pruned, which is depicted in \autoref{fig:correlation}.
-If the pruning method worked, we would expect
-%the figure to show
-%this would
-that neurons pruned in the beginning have a high correlation while later ones have a low correlation.
-
-\autoref{fig:correlation} shows that this is not the case.
-It also shows that neurons are not completely separated in backdoor neurons and regular neurons: If this were the case, the correlation would either be 1 or 0, but we observe many values in between, which indicates that most neurons are both responsible for the backdoor as well as for regular data.
-
-\subsubsection{Fine-Tuning}
-%\todo{Max: It said pruning before but should be tuning, right?}
-%AH. I guess both should work
-\begin{figure}[b]
-%\includegraphics[width=\columnwidth]{figures/finetuning_2015.pdf}
-\includegraphics[width=\columnwidth]{figures/finetuning_2017.pdf}
-\caption{Fine-tuning of the \gls{mlp} for \unsw{}: While fine-tuning eventually lowers backdoor accuracy a little bit, it takes more epochs than training of the original model.}
-\label{fig:finetuning}
-\end{figure}
-In addition to pruning, we tried to use fine-tuning to remove the backdoor from the \gls{mlp}. % classifier.
-%The use of
-Fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
-
-\autoref{fig:finetuning} shows the backdoor efficacy when continuing training of our trained model for \cic{} with the validation set, hence without backdoored samples.
-It indicates clearly that, unfortunately, with reasonable computational effort, fine-tuning is uneffective for cleaning the \gls{mui} from backdoors in our case.
-In fact, training a model from scratch takes less computational effort than fine-tuning.
-We observed a similar behaviour for \unsw{}.
-
-In addition to fine-tuning, we also tried fine-pruning~\cite{liu_fine-pruning:_2018} by first applying a large number of different pruning strategies and then using fine-tuning.
-From all pruning strategies, only pruning a certain fraction of only the \gls{mlp}'s first layer resulted in a discernible drop of backdoor efficacy after one epoch of fine-tuning, primarily for \cic{}.
-
-The significance of the \gls{mlp}'s first layer for the backdoor presumably results from the simplicity of our backdoor pattern.
-Hence, applicability of this pruning technique for general situations is questionable.
-We conclude that also fine-pruning cannot be considered a reliable method for backdoor removal in the context of \glspl{ids}.
-
-%The use of fine-tuning exclusively makes  sense if both the computational effort and the required training set size for fine-tuning are substantially lower than for the original training procedure.
-Hence, a decay of the backdoor efficacy as depicted in \autoref{fig:finetuning} cannot be considered sufficient.
-
-\subsection{\gls{rf} Pruning}
-%\todo{Max: Redundant?} For intrusion detection, \gls{dt} based models or \glspl{rf} of \glspl{dt}  usually achieve higher accuracy than \glspl{mlp} \cite{meghdouri_analysis_2018} and moreover they are generally faster to train and possibly more interpretable.
-For \glspl{ids} we thus consider the removal of backdoors from \gls{dt} based models to be a more pressing issue than from \glspl{mlp}.
-
-%We propose a defense approach that can be used for similar use cases like the mechanisms recently proposed to remove backdoors from \glspl{cnn}, discussed above.
-We developed a defense approach specifically for \gls{rf} classifiers.
-The concept behind our pruning defense is that leaves that are never used by samples in the validation set might be used by backdoored inputs.
-If these ``useless'' leaves are removed, performance of the classifier on the validation set should not decrease
-%by a lot
-significantly,
-while decisions that are used for the backdoor are likely to be removed.
-We developed several variants of our pruning defense with an increasing level of sophistication:
-\begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
-\item Pruning leaves based on their usage: This means that the least used leaf is pruned first and the most used one last.
-\item Like (1) but considering only ``benign'' leaves: We assume that attackers want malicious samples appear benign.
-\item Like (2) but additionally using depth to decide when
-%two % AH could be more than two
-leaves are used equally often: The rationale is that we assume that hand-crafted backdoors require fewer rules to classify than regular samples and thus have lower depth.
-\end{enumerate}
-
-\autoref{fig:pruningExample} shows an example of variant (3) of pruning applied to a tree
-\begin{figure}[b]
-\includegraphics[width=\columnwidth]{figures/prune_CAIA_backdoor_15/prune_oh_d.pdf}
-\caption{Pruning an \gls{rf} for \unsw{}. The smaller the dashes, the smaller the validation dataset. The validation set size ranges from 1\% to 100\% of its original size.}
-\label{fig:performancePruningRF}
-\end{figure}
-%\todo{Max: Which pruning algorithm actually?}
-and \autoref{fig:performancePruningRF} shows the resulting accuracy for pruning according to variant (3) for the \unsw{} dataset.
-With sufficient leaves being pruned, the accuracy of the backdoor approaches zero.
-The accuracy for the regular data does not decrease significantly.
-It might even increase by reducing overfitting.
-With 90\% of leaves pruned, accuracy is still well above 99\%.
-Even with a very small validation set of just 1\% of its original size, the pruning still works as expected.
-
-For \cic{} we get similar results, but pruning does not remove the backdoor completely ($\sim$30\% backdoor accuracy remains).
-We attribute this to the fact that in this dataset there are also regular flows which have $\text{stdev}(\text{TTL}) > 0$.
-Thus, backdoored flows cannot always be sharply distinguished from regular flows.
-We find that pruning only benign leaves is very beneficial for \unsw{} but not for \cic{}.
-However, considering depth when making the decision which leaf to prune first, leads to the backdoor being removed considerably earlier in the pruning process.
-
-\section{Discussion}
-From our experiments, we can make three main recommendations for the deployment of \gls{ml} models which have been obtained from a third-party.
-
-To ensure that no backdoor is contained in the model,
-%in question
-it has to be  analyzed carefully for questionable decisions and potentially unnecessary features.
-For this purpose, \glspl{pdp} and \gls{ale} plots are an effective tool.
-In fact, already throughout the training process explainability plots constitute a useful tool to ensure that the model is not unintentionally trained  to artifacts the dataset yields.
-On the other hand, the implementation of a backdoor as conducted in this research is only possible when using several features involving the \gls{ttl} value.
-Even though it might seem tempting to provide all possible features to a \gls{dl} or \gls{rf} classifier and let it learn the most important ones, this strategy should be avoided.
-
-For \gls{rf} classifiers, which can be considered one of the most important classifiers for \glspl{ids}, the pruning technique we proposed is able to reduce backdoor efficacy significantly.
-At the same time, the classifier's detection performance is not substantially reduced.
-Thus, we recommend always including a validation set when providing a \gls{dt} or \gls{rf} to another party.
-Even if the validation set is significantly smaller than the training set, the defensive properties are still upheld.
-
-%We found it
-It is surprising that the neural network pruning and fine-tuning methods were ineffective for removing the backdoor from our \gls{dl} model in all our experiments.
-Since \glspl{cnn} are conceptually very similar to \glspl{mlp} it is not obvious that methods working for the former do not work for the latter.
-The difference that a \gls{cnn} always only looks at a portion of the input \cite{wikipedia_convolutional_2019} and not at all of it (unlike an \gls{mlp}) should not change the efficacy of the pruning approach.
-This leads us to the conclusion that probably the proposed methods are insufficient for \glspl{mlp} and more research is required to develop methods suitable for them.
+\section{Conclusion}
 
 \section*{Acknowledgements}
-The Titan Xp used for this research was donated by the NVIDIA Corporation. We thank Prof.~Andreas Rauber for his suggestions to improve the explainability plots.
+The Titan Xp used for this research was donated by the NVIDIA Corporation.
 

### 2020-07-14 learn.py

@@ -177,6 +177,7 @@ def train_eager_stopping_nn():
 
 	samples = 0
 	net.train()
+	print('>> start training...')
 	for i in range(1, sys.maxsize):
 		for data, labels in train_loader:
 			optimizer.zero_grad()
@@ -441,12 +442,12 @@ if __name__=="__main__":
 		stds[stds==0.0] = 1.0
 		# np.set_printoptions(suppress=True)
 		# stds[np.isclose(stds, 0)] = 1.0
-		with open(file_name_for_normalization_data, "wb") as f:
-			f.write(pickle.dumps((means, stds)))
+		#with open(file_name_for_normalization_data, "wb") as f:
+		#	f.write(pickle.dumps((means, stds)))
 	else:
 		file_name_for_normalization_data = opt.normalizationData
-		with open(file_name_for_normalization_data, "rb") as f:
-			means, stds = pickle.loads(f.read())
+		#with open(file_name_for_normalization_data, "rb") as f:
+		#	means, stds = pickle.loads(f.read())
 	assert means.shape[0] == x.shape[1], "means.shape: {}, x.shape: {}".format(means.shape, x.shape)
 	assert stds.shape[0] == x.shape[1], "stds.shape: {}, x.shape: {}".format(stds.shape, x.shape)
 	assert not (stds==0).any(), "stds: {}".format(stds)

### 2020-07-14 learn.py
Add proper legend to "create_plot_eager" plot

@@ -299,16 +299,21 @@ def create_plot_eager(test_indices):
 
 	colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
+	things = []
 	x, y1, y2 = list(zip(*efforts))
-	plt.plot(x, y1, color=colors[0])
-	plt.xlabel("confidence")
-	plt.ylabel("mean layers needed")
+	y1_label = "mean layers evaluated"
+	things += plt.plot(x, y1, color=colors[0], label=y1_label)
+	plt.xlabel("confidence required to stop further evaluation")
+	plt.ylabel(y1_label)
 	plt.twinx()
-	plt.plot(x, y2, color=colors[1])
-	plt.ylabel("mean accuracy")
+	y2_label = "mean accuracy"
+	things += plt.plot(x, y2, color=colors[1], label=y2_label)
+	plt.ylabel(y2_label)
+	plt.tight_layout()
+	plt.legend(things, [l.get_label() for l in things], loc="upper left")
 
 	if opt.storePlot != '':
-		plt.savefig(opt.storePlot, bbox_inches='tight')
+		plt.savefig(opt.storePlot, bbox_inches = 'tight', pad_inches = 0)
 	else:
 		plt.show()
 

### 2020-07-14 paper/paper.tex
add content to paper

@@
-%%
-%% This is file `sample-sigconf.tex',
-%% generated with the docstrip utility.
-%%
-%% The original source files were:
-%%
-%% samples.dtx  (with options: `sigconf')
-%%
-%% IMPORTANT NOTICE:
-%%
-%%% For the copyright see the source file.
-%%
-%% Any modified versions of this file must be renamed
-%% with new filenames distinct from sample-sigconf.tex.
-%%
-%% For distribution of the original source see the terms
-%% for copying and modification in the file samples.dtx.
-%%
-%% This generated file may be distributed as long as the
-%% original source files, as listed above, are part of the
-%% same distribution. (The sources need not necessarily be
-%% in the same archive or directory.)
-%%
-%% The first command in your LaTeX source must be the \documentclass command.
+\documentclass[conference]{IEEEtran}
+\IEEEoverridecommandlockouts
+% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+\usepackage{cite}
+\usepackage{amsmath,amssymb,amsfonts}
+\usepackage{algorithmic}
+\usepackage{graphicx}
+\usepackage{textcomp}
+\usepackage{xcolor}
+\usepackage{subfig}
+\usepackage{multirow}
+\usepackage{booktabs}
 
-% Modify this to remove all ACM and conference information (for arXiv for example)
-\def\removeHeaders{yes}
+\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
+    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
 
-\def\tempYes{yes}
-\documentclass[9pt,sigconf,letterpaper,dvipsnames\ifx\removeHeaders\tempYes ,nonacm\fi]{acmart}
-
-\copyrightyear{2019}
-\acmYear{2019}
-\setcopyright{acmcopyright}
-\acmConference[Big-DAMA '19]{3rd ACM CoNEXT Workshop on Big DAta, Machine Learning and Artificial Intelligence for Data Communication Networks}{December 9, 2019}{Orlando, FL, USA}
-\acmBooktitle{3rd ACM CoNEXT Workshop on Big DAta, Machine Learning and Artificial Intelligence for Data Communication Networks (Big-DAMA '19), December 9, 2019, Orlando, FL, USA}
-\acmPrice{15.00}
-\acmDOI{10.1145/3359992.3366638}
-\acmISBN{978-1-4503-6999-2/19/12}
-
-\usepackage[utf8]{inputenc}
-
-\usepackage{multirow}
-\usepackage{blindtext}
-\usepackage{soul}
-\usepackage[inline]{enumitem}
-
-\usepackage{subfig}
-
-% Acronyms
 \usepackage[nomain, toc, acronym]{glossaries}
 \glsdisablehyper
-
-\ifx\removeHeaders\tempYes
-\settopmatter{printacmref=false} % Removes citation information below abstract
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
-\fi
 
@@
 \newcommand{\cic}{CIC-IDS-2017}
-
-%%
-%% \BibTeX command to typeset BibTeX logo in the docs
-\AtBeginDocument{%
-  \providecommand\BibTeX{{%
-    \normalfont B\kern-0.5em{\scshape i\kern-0.25em b}\kern-0.8em\TeX}}}
-
-%%
-%% Submission ID.
-%% Use this when submitting an article to a sponsored event. You'll
-%% receive a unique submission ID from the organizers
-%% of the event, and this ID should be used as the parameter to this command.
-%%\acmSubmissionID{123-A56-BU3}
-
-%%
-%% The majority of ACM publications use numbered citations and
-%% references.  The command \citestyle{authoryear} switches to the
-%% "author year" style.
-%%
-%% If you are preparing content for an event
-%% sponsored by ACM SIGGRAPH, you must use the "author year" style of
-%% citations and references.
-%% Uncommenting
-%% the next command will enable that style.
-%%\citestyle{acmauthoryear}
-
-%%
-%% end of the preamble, start of the body of the document source.
-\begin{document}
-
-%%
-%% The "title" command has an optional parameter,
-%% allowing the author to define a "short title" to be used in page headers.
-\title{Eager-Stopping: Early Network Intrusion Prediction\\in Feed Forward Neural Networks}
-
-%% GLOSSARY
 
@@
 
+\begin{document}
+
+\title{EagerNet: Early Intrusion Prediction\\in Feed Forward Neural Networks}
+
+\author{\IEEEauthorblockN{Fares Meghdouri, Maximilian Bachl and Tanja Zseby}
+\IEEEauthorblockA{Technische Universität Wien\\
+Vienna, Austria\\
+firstname.lastname@tuwien.ac.at}}
+
+\maketitle
+
 \begin{abstract}
@@
 
-\ifx\removeHeaders\tempNo
-	\begin{CCSXML}
-		<ccs2012>
-		<concept>
-		<concept_id>10002978.10002997.10002999</concept_id>
-		<concept_desc>Security and privacy~Intrusion detection systems</concept_desc>
-		<concept_significance>500</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10010147.10010257.10010293.10003660</concept_id>
-		<concept_desc>Computing methodologies~Classification and regression trees</concept_desc>
-		<concept_significance>500</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10010147.10010257.10010293.10010294</concept_id>
-		<concept_desc>Computing methodologies~Neural networks</concept_desc>
-		<concept_significance>500</concept_significance>
-		</concept>
-		<concept>
-		<concept_id>10003033.10003083.10003014</concept_id>
-		<concept_desc>Networks~Network security</concept_desc>
-		<concept_significance>300</concept_significance>
-		</concept>
-		</ccs2012>
-	\end{CCSXML}
-	\ccsdesc[500]{Security and privacy~Intrusion detection systems}
-	\ccsdesc[500]{Computing methodologies~Classification and regression trees}
-	\ccsdesc[500]{Computing methodologies~Neural networks}
-	\ccsdesc[300]{Networks~Network security}
-	\keywords{Network security, Deep Learning, }
-\fi
+\begin{IEEEkeywords}
 
-%%
-%% The "author" command and its associated commands are used to define
-%% the authors and their affiliations.
-%% Of note is the shared affiliation of the first two authors, and the
-%% "authornote" and "authornotemark" commands
-%% used to denote shared contribution to the research.
-
-%\author{Maximilian Bachl, Alexander Hartl, Joachim Fabini, Tanja Zseby}
-%\affiliation{\institution{Technische Universität Wien}}
-%\email{firstname.lastname@tuwien.ac.at}
-
-%%
-%% By default, the full list of authors will be used in the page
-%% headers. Often, this list is too long, and will overlap
-%% other information printed in the page headers. This command allows
-%% the author to define a more concise list
-%% of authors' names for this purpose.
-%\renewcommand{\shortauthors}{Hartl and Bachl}
-
-%%
-%% The abstract is a short summary of the work to be presented in the
-%% article.
-%\begin{abstract}
-%  A clear and well-documented \LaTeX\ document is presented as an
-%  article formatted for publication by ACM in a conference proceedings
-%  or journal publication. Based on the ``acmart'' document class, this
-%  article presents and explains many of the common variations, as well
-%  as many of the formatting elements an author may use in the
-%  preparation of the documentation of their work.
-%\end{abstract}
-\ifx\removeHeaders\tempYes
-	\settopmatter{printfolios=true}
-\fi
-\maketitle
+\end{IEEEkeywords}
 
@@
 \subsection{Early Predictions}
+\subsubsection{Combined Back-propagation Loss}
+
+\begin{equation}
+l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
+\end{equation}
+
+\[
+\nabla l = \underbrace{W_{current} \times \nabla l_{current}}_{\text{Current gradient}} + \underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}
+\]
 
 \begin{figure*}[htp]
-
+\center
 \subfloat[Conventional Feed Forward Neural Network.]{%
-  \includegraphics[clip,width=\columnwidth]{figures/ffnn.pdf}%
+  \includegraphics[clip,width=0.45\textwidth]{figures/ffnn_new.pdf}%
 }
-\qquad
-\subfloat[Eager-Stopping Neural Network.]{%
-  \includegraphics[clip,width=\columnwidth]{figures/eagernet.pdf}%
+\\
+\subfloat[Eager-Stopping Neural Network: forward-propagation mode.]{%
+  \includegraphics[clip,width=0.53\textwidth]{figures/eagernet_new.pdf}%
+}
+\quad
+\subfloat[Eager-Stopping Neural Network: back-propagation mode.]{%
+  \includegraphics[clip,width=0.45\textwidth]{figures/eagernet_backprop_new.pdf}%
 }
 
-\caption{main caption}
+\caption{The difference between conventional neural networks and our proposed architecture.}
 
@@
 
+\begin{table}
+
+\centering
+\begin{tabular}{ccrr}
+\toprule
+\textbf{Dataset} & \textbf{Architecture} & \textbf{FFNN} & \textbf{EagerNet*}\\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{CIC-IDS17}} & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{UNSW-NB15}} & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+\midrule
+\multirow{6}{*}{\rotatebox{90}{MNIST}} & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+ & & & \\
+\end{tabular}
+\vspace{1ex}
+
+{\raggedright * . \par}
+\caption{Accuracy scores.}
+
+\end{table}
+
+\begin{figure}[b!]
+\includegraphics[width=\columnwidth]{figures/conf_acc_Jul10_15-09-38_gpu_0_3.pdf}
+\caption{Accuracy-Confidence plot.\note{red}{FM: this will change}}
+\label{fig:pdp_ttl}
+\end{figure}
+
+\begin{figure}[b!]
+\includegraphics[width=\columnwidth]{figures/acc_multiclass_Jul10_15-18-24_gpu_0_3.pdf}
+\caption{Accuracy plot multiclass.\note{red}{FM: this will change}}
+\label{fig:pdp_ttl}
+\end{figure}
+
 \section{Conclusion}

### 2020-07-15 paper/paper.tex
add content to paper
add abstract and intro to paper

@@
 \newacronym{ip}{IP}{Internet Protocol}
+\newacronym{ffnn}{FNN}{Feedforward Neural Network}
 
@@
 
-\title{EagerNet: Early Intrusion Prediction\\in Feed Forward Neural Networks}
+%\title{EagerNet: Early Intrusion Prediction\\in Feed Forward Neural Networks}
+\title{EagerNet: Early Predictions in Feedforward Neural Networks for Fast Network Intrusion Detection}
 
@@
 
+\glspl{ffnn} have been the basis of most state-of-the-art \gls{ml} applications in recent years.
+Experiments show that the deeper the network is, the more desirable it is to acquire more knowledge.
+Nonetheless, with the growing number of layers, obtaining fast predictions has become a difficult task despite the use of special hardware such as GPUs.
+We propose a new architecture to obtain predictions without conducting a forward-pass until the last layer.
+The architecture is able to deal with either binary or multiclass classification problems and trades prediction speed for the confidence of the network.
+We evaluate our proposal with two different network intrusion detection datasets.
+Results suggest that it is possible to obtain comparable accuracies to simple \glspl{ffnn} without evaluating all layers, thus obtaining early predictions and saving energy and computational efforts.
+
 \end{abstract}
@@
 \section{Introduction}
+% what are ffnn
+\glspl{ffnn} has gained remarkable attention in recent years due to the availability of data and computational resources for research.
+Many of the state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
+Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan2016power} .
+
+% forward-backward
+Traditional \gls{ffnn} are based on simple logistic regression units that combine multiple inputs multiplied by a set of weights and passed through an activation function (e.g.
+sigmoid) to obtain a scalar.
+The weights are adjusted to give the closest possible output to a ground-truth.
+As far as neural networks are concerned, logistic regression units are stacked together to create layers and then layers are stacked together again to create the network itself.
+Figure \ref{fig:ffnn} demonstrates a basic \gls{ffnn} that utilizes the input vector $\lbrace X_{1}, X_{2} ...
+X_{m} \rbrace$ and makes a prediction.
+Adjusting weights in such situations is a non-trivial problem since the weights of each layer may have an impact on the previous layers.
+Fortunately, several algorithms have been proposed to reduce the complexity of the task (e.g.
+gradient descent learning) while achieving the same goal: reduce the "gap" (loss) between predictions and actual desired output.
+Larger networks consist of more neurons and thus more weights, allowing more data patterns to be learned.
+
+% problem with large networks
+In terms of efficiency, two problems arise when using an extremely deep neural networks: (1) The loss during training needs more time to converge and (2) prediction time is proportional to the number of layers.
+Even though most modern hardware is able to cope with offline single predictions without significant delay, online massive neural network implementations may suffer from the latter issue if non-specialized hardware is used to produce predictions.
+%EagerNet 
+To solve this, we propose a new architecture that, on the one hand, uses the entire network capacity for learning and, on the other hand, stops the forward-pass and makes predictions as soon as confidence reaches a certain threshold.
+In other words, it makes it possible to avoid evaluating the entire network and to receive predictions from intermediate layers with the same performance of a simple \gls{ffnn}.
+The proposed approach therefore allows, where possible, the reduction of computing resources and energy usage preserving comparable performance of a full forward-pass.
+
+% conclusion
+The rest of the paper is organized as follows: In Section ...
 
 \section{Supervised Learning for Network Intrusion Detection}
-
-\section{Eager-Stopping}
-\subsection{Feed Forward Neural Networks}
-
-\subsection{Early Predictions}
-\subsubsection{Combined Back-propagation Loss}
-
-\begin{equation}
-l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
-\end{equation}
-
-\[
-\nabla l = \underbrace{W_{current} \times \nabla l_{current}}_{\text{Current gradient}} + \underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}
-\]
-
-\begin{figure*}[htp]
-\center
-\subfloat[Conventional Feed Forward Neural Network.]{%
-  \includegraphics[clip,width=0.45\textwidth]{figures/ffnn_new.pdf}%
-}
-\\
-\subfloat[Eager-Stopping Neural Network: forward-propagation mode.]{%
-  \includegraphics[clip,width=0.53\textwidth]{figures/eagernet_new.pdf}%
-}
-\quad
-\subfloat[Eager-Stopping Neural Network: back-propagation mode.]{%
-  \includegraphics[clip,width=0.45\textwidth]{figures/eagernet_backprop_new.pdf}%
-}
-
-\caption{The difference between conventional neural networks and our proposed architecture.}
-
-\end{figure*}
-
-\section{Related Work}
-
-\section{Evaluation and Discussion}
-\subsection{Datasets}
 
@@
 	\centering
-	\caption{CAIA Flow Features.}
+	\caption{CAIA network flow features.}
 	\label{tab:cres}
@@
 
-\subsection{Binary vs. Multiclass Classification}
+\section{EagerNet}
+
+\begin{figure*}[htp]
+\center
+\subfloat[Conventional \gls{ffnn} architecture.]{%
+  \includegraphics[clip,width=0.45\textwidth]{figures/ffnn_new.pdf}%
+  \label{fig:ffnn}
+}
+\\
+\subfloat[EagerNet: Forward-pass.]{%
+  \includegraphics[clip,width=0.53\textwidth]{figures/eagernet_new.pdf}%
+}
+\quad
+\subfloat[EagerNet: Back-propagation of gradients.]{%
+  \includegraphics[clip,width=0.45\textwidth]{figures/eagernet_backprop_new.pdf}%
+}
+
+\caption{The difference between conventional neural networks and our proposed architecture.}
+
+\end{figure*}
+
+\subsubsection{Binary Vs. Multiclass Classification}
+
+\begin{equation}
+l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
+\end{equation}
+
+\[
+\nabla l = \underbrace{W_{current} \times \nabla l_{current}}_{\text{Current gradient}} + \underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}
+\]
+
+\subsubsection{Combined Back-propagation Loss}
+
+\subsubsection{Design Decision}
+
+\section{Related Work}
+
+\section{Evaluation and Discussion}
+\subsection{Datasets}
+
+\subsection{Performance}
 
@@
 \midrule
-\multirow{6}{*}{\rotatebox{90}{Binary}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
+\multirow{6}{*}{\rotatebox{90}{Binary}} & \multirow{2}{*}{Equal} & 1-3-1 & 0.996 & 0.996 & 0.990 & 0.989 \\
+ & & 1-1-1 & 0.997 & 0.997 & 0.990 & 0.990 \\
+ & \multirow{2}{*}{Increasing} & 1-3-1 & 0.997 & 0.998 & 0.990 & 0.989 \\
+ & & 1-1-1 & 0.997 & 0.997 & 0.990 & 0.989 \\
+ & \multirow{2}{*}{Decreasing} & 1-3-1 & 0.997 & 0.998 & 0.989 & 0.989 \\
+ & & 1-1-1 & 0.997 & 0.997 & 0.990 & 0.990 \\
 \midrule
@@
  & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
+ & \multirow{2}{*}{Decreasing} & 1-3-1 & 0.988 & 0.846 & 0.830 & 0.825 \\
  & & 1-1-1 & & & & \\
@@
 
+\subsection{Confidence-Speed Tradeoff}
+
 \begin{figure}[b!]
-\includegraphics[width=\columnwidth]{figures/conf_acc_Jul10_15-09-38_gpu_0_3.pdf}
+\includegraphics[width=\columnwidth]{figures/conf-acc-layers.pdf}
 \caption{Accuracy-Confidence plot.\note{red}{FM: this will change}}
@@
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{bibliography}
+\bibliographystyle{IEEEtran}
+\bibliography{biblio}
 

### 2020-07-20 paper/paper.tex
add content to paper

@@
 % what are ffnn
-\glspl{ffnn} has gained remarkable attention in recent years due to the availability of data and computational resources for research.
+\glspl{ffnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
 Many of the state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
@@
 
+% talk about network traffic
+
 % conclusion
@@
 
+ \begin{figure}
+  \includegraphics[width=\linewidth]{figures/ffnn_new.pdf}
+  \caption{Conventional \gls{ffnn} architecture.}
+  \label{fig:ffnn}
+\end{figure}
+
 \section{Supervised Learning for Network Intrusion Detection}
+
+\gls{ml} has been widely used in the last decade for network traffic analysis and specifically anomaly detection systems.
+Many of the works proposed \gls{ids} architectures based on well-known supervised techniques \cite{survey}.
+The idea of training a classifier on pre-stored attack patterns and using the same classifier to recognize similar patterns is commonly used and remarkable success has so far been achieved \cite{survey}.
+In particular, \glspl{ffnn} is commonly used whenever large amounts of data are available.
+In order to present network data to such architectures, a number of traffic representation families have been proposed depending on the application.
+However, the statistical representation of flow characteristics is still the most appropriate.
+In this work, we define the network flow to be the exchange of packets between two end-hosts.
+Packets can be identified and aggregated using the five-tuple key: \emph{sourceIP}, \emph{destinationIP}, \emph{sourceTransportPort}, \emph{destinationTransportPort} and \emph{protocolIdentifier}.
+Thereafter, packet features are extracted and the statistical combinations are computed.
+Table \ref{tab:features} shows CAIA \cite{williams2006preliminary} network traffic representation.
+It consists of 12 features 7 of which are measured in both directions and 2 of which are expanded using four statistical combinations (i.e.
+\emph{A} becomes \emph{mean(A)}, \emph{min(A)}, \emph{max(A)} and \emph{stdev(A)}).
+In addition, we set an observation timeout of 1,800 seconds, after which we terminate flows.
+Each flow is considered to be a data sample and has two sets of labels: binary (attack/benign) label and attack family.
+After a training process, the trained \gls{ids} must be able to correctly classify new flows.
 
@@
 	\centering
-	\caption{CAIA network flow features.}
-	\label{tab:cres}
+	\caption{CAIA flow representation.}
+	\label{tab:features}
 
 	\footnotesize
+	\resizebox{\columnwidth}{!}{
 	\begin{tabular}{|c|c|c|}
@@
 	\end{tabular}
+	}
 
@@
 
+Because several network attack families are simple to detect on the basis of linear input combinations, the standard \gls{ffnn} is slightly modified to cope with the difficulty of detection.
+Our novel architecture is based on the assumption that deep neural networks can learn complex functions with simpler mappings learned in the first few layers.
+As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for predictions at each layer.
+The EagerNet architecture is shown in Figure \ref{fig:eagerNet_forward}.
+Initial neurons are shown in gray and output neurons (per layer) are shown in green.
+Once a flow is observed and features are extracted, a sample $\textbf{X}$ is fed to the network via the input layer.
+The network proceeds evaluating the output of each layer and yields a prediction confidence value.
+We define the confidence as the value that represents how sure the network is about a certain decision or sample belonging to a certain class.
+Based on a predefined threshold, the network measures confidence and determines whether to avoid testing additional layers or proceed to do so in order to gain more confidence.
+This methodology ensures that simpler samples are classified early in the forward-pass and that more complicated, noisy samples are passed into deeper layers.
+In the \ref{confidence-speed-tradeoff} Section, we show that EagerNet reduces resource consumption on average keeping comparable performance to normal \gls{ffnn}.
+
 \begin{figure*}[htp]
 \center
-\subfloat[Conventional \gls{ffnn} architecture.]{%
-  \includegraphics[clip,width=0.45\textwidth]{figures/ffnn_new.pdf}%
-  \label{fig:ffnn}
-}
-\\
+
 \subfloat[EagerNet: Forward-pass.]{%
-  \includegraphics[clip,width=0.53\textwidth]{figures/eagernet_new.pdf}%
+  \includegraphics[clip,width=0.5\textwidth]{figures/eagernet_new.pdf}%
+  \label{fig:eagerNet_forward}
 }
@@
 \subfloat[EagerNet: Back-propagation of gradients.]{%
-  \includegraphics[clip,width=0.45\textwidth]{figures/eagernet_backprop_new.pdf}%
+  \includegraphics[clip,width=0.42\textwidth]{figures/eagernet_backprop_new.pdf}%
+  \label{fig:eagerNet_backward}
 }
@@
 \subsubsection{Binary Vs. Multiclass Classification}
+We adjust the output layer to build two separate sub-architectures.
+The first architecture, consisting of a binary attack detection, i.e.
+a 1 is an attack flow and a 0 is a benign flow.
+The predictions in this case are given by a single output neuron followed by a sigmoid activation function to force values between 0 and 1.
+During back-propagation, we use a binary cross-entropy loss including the activation function and average the loss over the batch.
+Equation \ref{eq:bceloss} shows how the loss is determined for a batch of size $N$, with the $i^{th}$ input $x_{i}$ and the sigmoid activation function  $\sigma()$.
 
 \begin{equation}
+\label{eq:bceloss}
 l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
@@
 
-\[
-\nabla l = \underbrace{W_{current} \times \nabla l_{current}}_{\text{Current gradient}} + \underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}
-\]
+In the case of a multicalss class problem, the output layers consist of as many neurons as attack families followed by a softmax activation functions that allow at most only one neuron to be 1.
+During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there is no overlap between classes.
+Equation \ref{eq:cceloss} shows ...
+
+\begin{equation}
+\label{eq:bceloss}
+l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
+\end{equation}
+
+After computing the losses, we use $Adam$ optimizer for computing gradients with a learning rate of 0.001.
 
 \subsubsection{Combined Back-propagation Loss}
+Conventional \gls{ffnn} uses the loss of the last layer to adjust the gradients.
+In this research, and since we have multiple outputs; one at each layer, we aggregate the losses and back propagate them.
+Fig \ref{fig:eagerNet_backward} demonstrates how losses propagate back from the respective layer only.
+As a consequence, the weights of the last layer are affected only by the loss of the last output.
+Similarly, the weights of the first layer are affected by the losses of the entire network.
+In addition, we introduce a weighting policy for losses.
+We allocate a weight to each loss that is predefined prior to training.
+The aim is to help the network determine on which layer to optimize more.
+Three types of weight sets are defined:
+\begin{itemize}
+\item Uniform weights. all losses have the same weight:
+\begin{equation}
+\mathcal{W_U} = \lbrace W_{0}, ..., W_{n} \rbrace = \lbrace \frac{1}{\sum_{i=0}^{n} 1}, ..., \frac{1}{\sum_{i=0}^{n} 1} \rbrace
+\end{equation}
 
-\subsubsection{Design Decision}
+\item Increasing weights. losses are increasingly important from first to last layer. 
+\begin{equation}
+\mathcal{W_I} = \lbrace W_{0}, ..., W_{n} \rbrace = \lbrace \frac{1}{\sum_{i=0}^{n} i+1}, ..., \frac{n+1}{\sum_{i=0}^{n} i+1} \rbrace
+\end{equation}
+
+\item Decreasing weights. losses are decreasingly important from first to last layer. 
+\begin{equation}
+\mathcal{W_D} = \lbrace W_{0}, ..., W_{n} \rbrace = \lbrace \frac{n+1}{\sum_{i=0}^{n} i+1}, ..., \frac{1}{\sum_{i=0}^{n} i+1} \rbrace
+\end{equation}
+
+\end{itemize}
+
+We show the effect of the weight distributions in Table \ref{tab:weights} and discuss results in Section \ref{evaluation_and_discussion}.
+
+%\begin{equation}
+%\nabla l = {\underbrace{W_{current} \times \nabla l_{current}}_{\text{Current gradient}}} + {\underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}}
+%\end{equation}
+
+%\subsubsection{Enhanced Training}
+%Back-propagating all losses together even when using weighted-loss summation causes the network to minimize the loss on each layer hence learning all classes at each layer simultaneously if the training procedure is long enough thus causing (1) the first few layers to over-fit and (2) the network's loss to converge slowly if there are many classes to learn.
+To solve this issue, we propose a different weights distribution that constructs the weights as follows:
+%\begin{itemize}
+%\item perform a forward-pass and compute the loss per layer
+%\item select the layer with the minimal loss
+%\item attributes a decreasing weight distribution starting from the selected layer until the last layer
+%\end{itemize}
+%The above procedure insures that the most suitable layer learns the respective pattern as well as all subsequent layers hence ignoring first layers which as discussed before, may over-fit.
+This allows the network to focus on one layer and class at a time and let subsequent layers to learn the same patterns in case the confidence threshold is set too high.
+Intuitively, easy patterns are learned in the first layers since the loss is expected to be minimal and more complex patterns are learned in the last layers because more non-linearities can be obtained.
+%
+%\begin{equation}
+%\begin{split}
+%\mathcal{W_M} & = \lbrace W_{0}, ...,W_{m-1}, \underbrace{W_{m}}_{\text{Minimum Loss Index}}, ..., W_{n} \rbrace \\
+% & = \lbrace 0, ..., 0, \frac{2^{n}}{\sum_{i=m}^{n} 2^i}, ..., \frac{2^m}{\sum_{i=m}^{n} 2^i} \rbrace
+% \end{split}
+%\end{equation}
+
+% easy pattern less loss, first layers
+% summing up loss causes to learn the same pattern at all layers
+% assume that even in binary classifications, there are many subclasses
 
@@
 \section{Evaluation and Discussion}
+\label{evaluation_and_discussion}
+We discuss in this section different experimental results and observations. Overall, EagerNet shows promising performance comparable to \glspl{ffnn} with the ability to save resources.
+
 \subsection{Datasets}
@@
 
-\begin{table*}
-\parbox{.45\linewidth}{
+% why from 512 to 32
+
+\begin{table}
+
 \centering
+
 \begin{tabular}{cccrrrr}
@@
 \midrule
-\multirow{6}{*}{\rotatebox{90}{Binary}} & \multirow{2}{*}{Equal} & 1-3-1 & 0.996 & 0.996 & 0.990 & 0.989 \\
- & & 1-1-1 & 0.997 & 0.997 & 0.990 & 0.990 \\
- & \multirow{2}{*}{Increasing} & 1-3-1 & 0.997 & 0.998 & 0.990 & 0.989 \\
- & & 1-1-1 & 0.997 & 0.997 & 0.990 & 0.989 \\
- & \multirow{2}{*}{Decreasing} & 1-3-1 & 0.997 & 0.998 & 0.989 & 0.989 \\
- & & 1-1-1 & 0.997 & 0.997 & 0.990 & 0.990 \\
+\multirow{6}{*}{\rotatebox{90}{CIC-IDS17}} & \multirow{2}{*}{Equal} & 5 & 0.996 & 0.996 & 0.990 & 0.989 \\
+ & & 3 & 0.997 & 0.997 & 0.990 & 0.990 \\
+ & \multirow{2}{*}{Increasing} & 5 & 0.997 & 0.998 & 0.990 & 0.989 \\
+ & & 3 & 0.997 & 0.997 & 0.990 & 0.989 \\
+ & \multirow{2}{*}{Decreasing} & 5 & 0.997 & 0.998 & 0.989 & 0.989 \\
+ & & 3 & 0.997 & 0.997 & 0.990 & 0.990 \\
 \midrule
-\multirow{6}{*}{\rotatebox{90}{Multiclass}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
+\multirow{6}{*}{\rotatebox{90}{UNSW-NB15}} & \multirow{2}{*}{Equal} & 5 & 0.988 & 0.828 & 0.867 & 0.860 \\
+ & & 3 & 0.988 & 0.856 & 0.800 & 0.795 \\
+ & \multirow{2}{*}{Increasing} & 5 & 0.988 & 0.853 & 0.821 & 0.816 \\
+ & & 3 & 0.988 & 0.844 & 0.817 & 0.811 \\
+ & \multirow{2}{*}{Decreasing} & 5 & 0.988 & 0.843 & 0.824 & 0.819 \\
+ & & 3 & 0.988 & 0.851 & 0.819 & 0.813 \\
+
 \end{tabular}
-\caption{CIC-IDS17}
-}
-\hfill
-\parbox{.45\linewidth}{
-\centering
-\begin{tabular}{cccrrrr}
-\toprule
-\textbf{Variant} & \textbf{Weights} & \textbf{Layers} & \textbf{Acc.} & \textbf{Prec.} & \textbf{Rec.} & \textbf{J.}\\
-\midrule
-\multirow{6}{*}{\rotatebox{90}{Binary}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Decreasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
-\midrule
-\multirow{6}{*}{\rotatebox{90}{Multiclass}} & \multirow{2}{*}{Equal} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Increasing} & 1-3-1 & & & & \\
- & & 1-1-1 & & & & \\
- & \multirow{2}{*}{Decreasing} & 1-3-1 & 0.988 & 0.846 & 0.830 & 0.825 \\
- & & 1-1-1 & & & & \\
-\end{tabular}
-\caption{UNSW-NB15}
-}
-\end{table*}
+
+\vspace{1ex}
+
+{\raggedright * All architectures consist of 128 neurons per layer in addition to single output layer. \par}
+\caption{Effect of weights on the performance of different networks}
+\label{tab:weights}
+
+\end{table}
 
@@
 \centering
-\begin{tabular}{ccrr}
+\begin{tabular}{ccrc}
 \toprule
-\textbf{Dataset} & \textbf{Architecture} & \textbf{FFNN} & \textbf{EagerNet*}\\
+\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{ffnn}} & \textbf{EagerNet (last)} \\
 \midrule
-\multirow{6}{*}{\rotatebox{90}{CIC-IDS17}} & & & \\
- & & & \\
- & & & \\
- & & & \\
- & & & \\
- & & & \\
+\multirow{2}{*}{CIC-IDS17} & Binary & 0.983 & 0.988 \\
+ & Multiclass & & \\
 \midrule
-\multirow{6}{*}{\rotatebox{90}{UNSW-NB15}} & & & \\
- & & & \\
- & & & \\
- & & & \\
- & & & \\
- & & & \\
+\multirow{2}{*}{UNSW-NB15} & Binary & & \\
+ & Multiclass & & \\
 \midrule
-\multirow{6}{*}{\rotatebox{90}{MNIST}} & & & \\
- & & & \\
- & & & \\
- & & & \\
- & & & \\
- & & & \\
+MNIST & Multiclass & & \\
+\midrule
 \end{tabular}
@@
 
-{\raggedright * . \par}
+{\raggedright * All architectures consist of 10 layers $\times$ 64 neurons in addition to input and output layers. \par}
 \caption{Accuracy scores.}
@@
 \subsection{Confidence-Speed Tradeoff}
+\label{confidence-speed-tradeoff}
 
-\begin{figure}[b!]
-\includegraphics[width=\columnwidth]{figures/conf-acc-layers.pdf}
-\caption{Accuracy-Confidence plot.\note{red}{FM: this will change}}
-\label{fig:pdp_ttl}
+\begin{figure*}
+\center
+\begin{tabular}{cccc}
+\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-3x128.pdf}} &
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-12x64.pdf}}\\
+\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-3x128.pdf}} &
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-12x64.pdf}}
+\end{tabular}
+\caption{Confidence threshold effect on accuracy and number of needed layers.}
+\label{fig:confidence}
+\end{figure*}
+
+\begin{figure}
+\center
+\begin{tabular}{cccc}
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass17.pdf}} &
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass17.pdf}}
+\end{tabular}
+\caption{Accuracy plot multiclass.}
 \end{figure}
 
-\begin{figure}[b!]
-\includegraphics[width=\columnwidth]{figures/acc_multiclass_Jul10_15-18-24_gpu_0_3.pdf}
-\caption{Accuracy plot multiclass.\note{red}{FM: this will change}}
-\label{fig:pdp_ttl}
-\end{figure}
+\subsubsection{Optimal Confidence Threshold}
+During deployment, the optimum confidence threshold is set by the user and plays a role in the overall achieved accuracy.
+Setting the threshold too low can cause the network to make decisions at an early stage, thereby gaining resources but achieving low accuracy.
+Similarly, setting the threshold too high can cause the network to skip the proper layers and thus always use the last one.
+The desired confidence level can therefore be obtained during the training phase by determining on the basis of the desired accuracy as shown in Figure \ref{fig:confidence}.
+% why
 

### 2020-07-21 paper/paper.tex
add more content to the paper

@@
 During back-propagation, we use a binary cross-entropy loss including the activation function and average the loss over the batch.
-Equation \ref{eq:bceloss} shows how the loss is determined for a batch of size $N$, with the $i^{th}$ input $x_{i}$ and the sigmoid activation function  $\sigma()$.
+Equation \ref{eq:bceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the sigmoid activation function  $\sigma(.)$.
 
@@
 \label{eq:bceloss}
-l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
+l = - \frac{1}{N} \sum_{n}^{N} [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
 \end{equation}
 
-In the case of a multicalss class problem, the output layers consist of as many neurons as attack families followed by a softmax activation functions that allow at most only one neuron to be 1.
+In the case of a multiclass problem, the output layers consist of as many neurons as attack families followed by a softmax activation functions that allow at most only one neuron to be 1.
 During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there is no overlap between classes.
-Equation \ref{eq:cceloss} shows ...
+Equation \ref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function  $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
 
 \begin{equation}
-\label{eq:bceloss}
-l = \frac{1}{N} \sum_{n}^{N} -\omega [y_n \cdot \log \sigma (x_{n}) + (1 - y_n) \cdot \log (1 - \sigma (x_{n}))]
+\label{eq:cceloss}
+l = - \frac{1}{N} \sum_{n}^{N} \sum_{i}^{C} y_{n,i} \cdot \log (\phi(x_{n})_{i})
 \end{equation}
 
-After computing the losses, we use $Adam$ optimizer for computing gradients with a learning rate of 0.001.
+After computing the losses, we use $Adam$ optimizer \cite{kingma2014adam} for computing gradients with a learning rate of 0.001.
 
@@
 \subsection{Datasets}
+For our experiments, we use two intrusion detection datasets.
+CICIDS2017 \cite{sharafaldin2018toward} which consists of 14 network attack classes in addition to realistic user profiles.
+UNSW-NB15 \cite{moustafa2015unsw} which, in addition to synthetically generated normal traffic, again contains 9 attack families.
+As mentioned above, we represent both datasets using the CAIA feature set and extract 2.317.922 and 2.065.224 flows, respectively, from CICIDS2017 and UNSW-NB15.
+For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
+We use $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
 
@@
 
-% why from 512 to 32
+\subsubsection{Weighted Loss}
+In Table \ref{tab:weights} we gather results from different networks trained using different numbers of hidden layers and weight distributions with 128 neurons per layer.
+In addition, we compare the results using four metrics (Accuracy, Precision , Recall and Youden Index) and show the performance of the last layer (hence raising the confidence level to the maximum).
+The CICIDS2017 indices show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
+In fact, since the last layer is always trained with one loss regardless of the weight distribution, only the first few layers are affected by the combined-loss learning procedure.
+Nonetheless, since the accuracy did not improve when increasing the number of layers, that implies that most network patterns are already separable after the few first layers.
+The results of UNSW-NB15 show a slightly different behavior.
+The uniform weighting of losses reflects the overall best scores and, in addition, the deeper architectures outperform the shallow ones.
+This implies that the network patterns in this dataset are more dynamic than the previous one and that more non-linearities (layers) are required to find the optimal mapping function.
 
@@
 \midrule
-\multirow{6}{*}{\rotatebox{90}{CIC-IDS17}} & \multirow{2}{*}{Equal} & 5 & 0.996 & 0.996 & 0.990 & 0.989 \\
+\multirow{6}{*}{\rotatebox{90}{CIC-IDS17}} & \multirow{2}{*}{Uniform} & 5 & 0.996 & 0.996 & 0.990 & 0.989 \\
  & & 3 & 0.997 & 0.997 & 0.990 & 0.990 \\
@@
 \midrule
-\multirow{6}{*}{\rotatebox{90}{UNSW-NB15}} & \multirow{2}{*}{Equal} & 5 & 0.988 & 0.828 & 0.867 & 0.860 \\
+\multirow{6}{*}{\rotatebox{90}{UNSW-NB15}} & \multirow{2}{*}{Uniform} & 5 & 0.988 & 0.828 & 0.867 & 0.860 \\
  & & 3 & 0.988 & 0.856 & 0.800 & 0.795 \\
@@
 
-\begin{table}
+\subsubsection{Confidence-Speed Tradeoff}
+\label{confidence-speed-tradeoff}
 
-\centering
-\begin{tabular}{ccrc}
-\toprule
-\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{ffnn}} & \textbf{EagerNet (last)} \\
-\midrule
-\multirow{2}{*}{CIC-IDS17} & Binary & 0.983 & 0.988 \\
- & Multiclass & & \\
-\midrule
-\multirow{2}{*}{UNSW-NB15} & Binary & & \\
- & Multiclass & & \\
-\midrule
-MNIST & Multiclass & & \\
-\midrule
-\end{tabular}
-\vspace{1ex}
-
-{\raggedright * All architectures consist of 10 layers $\times$ 64 neurons in addition to input and output layers. \par}
-\caption{Accuracy scores.}
-
-\end{table}
-
-\subsection{Confidence-Speed Tradeoff}
-\label{confidence-speed-tradeoff}
+Figure \ref{fig:confidence} shows the confidence-speed balance in addition to the accuracy achieved for four different networks and for both datasets.
+The networks contain 128 neurons per layer and a total of 3 or 12 layers.
+The confidence threshold is shown on the horizontal axis, the average number of layers used over the entire dataset and the average accuracy achieved are both shown on the vertical axis.
+Curves show consistent results, whereby the network always uses few layers for the majority of samples and only continues to evaluate further layers if the confidence level required is too high or the sample is noisy and it is difficult to make decisions in early layers.
+There are also two phenomena observed:
+\begin{itemize}
+\item The accuracy of CICIDS2017 increases almost linearly with respect to confidence and reaches its maximum when all layers are used.
+\item The accuracy of UNSW-NB15 is stable when the confidence level is increased until \~0.98 where it leaps noticeably when a further layer is evaluated.
+This suggests that some patterns are "learn-able" only on that specific layer.
+\end{itemize}
+Overall, EagerNet shows that it is possible to trade a tiny percentage of accuracy in order to save a significant amount of resources.
+In Section \ref{comparability_to_fnns}, we see that, in order to achieve comparable accuracy to the traditional \gls{ffnn}, the previous trade is possible and therefore also the possibility to reduce costs.
 
@@
 \begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass17.pdf}} &
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass17.pdf}}
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 4in]{figures/multiclass17.pdf}} &
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass15.pdf}}
 \end{tabular}
@@
 
-\subsubsection{Optimal Confidence Threshold}
+\textbf{Optimal Confidence Threshold.}
 During deployment, the optimum confidence threshold is set by the user and plays a role in the overall achieved accuracy.
@@
 The desired confidence level can therefore be obtained during the training phase by determining on the basis of the desired accuracy as shown in Figure \ref{fig:confidence}.
-% why
+
+\subsubsection{Comparability to FNNs}
+\label{comparability_to_fnns}
+
+\begin{table}
+
+\centering
+\begin{tabular}{ccrc}
+\toprule
+\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{ffnn}} & \textbf{EagerNet (last)} \\
+\midrule
+\multirow{2}{*}{CIC-IDS17} & Binary & 0.989 & 0.993 \\
+ & Multiclass & 0.920 & 0.979 \\
+\midrule
+\multirow{2}{*}{UNSW-NB15} & Binary & & \\
+ & Multiclass & & \\
+\midrule
+
+\end{tabular}
+\vspace{1ex}
+
+{\raggedright * All architectures consist of 10 layers $\times$ 64 neurons in addition to input and output layers. \par}
+\caption{Accuracy scores.}
+
+\end{table}
+
+\subsubsection{What Did the Network Learn?}
+% why from 512 to 32
 

### 2020-07-21 paper/paper.tex
Small changes

--- 2020-07-15 paper/paper.tex
@@
 %\title{EagerNet: Early Intrusion Prediction\\in Feed Forward Neural Networks}
-\title{EagerNet: Early Predictions in Feedforward Neural Networks for Fast Network Intrusion Detection}
+%\title{EagerNet: Early Predictions in Feedforward Neural Networks for Fast Network Intrusion Detection}
+% Max Proposal
+\title{EagerNet: Early Predictions of Neural Networks for Computationally Efficient Intrusion Detection}
 
@@
 Many of the state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan2016power} .
+Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan_power_2016}.
 
@@
 As far as neural networks are concerned, logistic regression units are stacked together to create layers and then layers are stacked together again to create the network itself.
-Figure \ref{fig:ffnn} demonstrates a basic \gls{ffnn} that utilizes the input vector $\lbrace X_{1}, X_{2} ...
-X_{m} \rbrace$ and makes a prediction.
+Figure \ref{fig:ffnn} demonstrates a basic \gls{ffnn} that utilizes the input vector $\lbrace x_{1}, x_{2} ...
+x_{m} \rbrace$ and makes a prediction.
 Adjusting weights in such situations is a non-trivial problem since the weights of each layer may have an impact on the previous layers.
@@
 
+\section{Related Work}
+
 \section{Supervised Learning for Network Intrusion Detection}
@@
 	\footnotesize
-	\begin{tabular}{|c|c|c|}
+	\begin{tabular}{ccc}
 		\toprule
@@
 \[
-\nabla l = \underbrace{W_{current} \times \nabla l_{current}}_{\text{Current gradient}} + \underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}
+\nabla l = \underbrace{W_\textit{current} \times \nabla l_\textit{current}}_{\text{Current gradient}} + \underbrace{\sum_{i}^{} \nabla l_{i}}_{\text{Next layer gradients}}
 \]

### 2020-07-21 paper/paper.tex

--- 2020-07-21 paper/paper.tex
@@
 The rest of the paper is organized as follows: In Section ...
+
+\section{Related Work}
 
@@
 	\footnotesize
-	\resizebox{\columnwidth}{!}{
-	\begin{tabular}{|c|c|c|}
+	\begin{tabular}{ccc}
 		\toprule

### 2020-07-21 paper/paper.tex
Minor cosmetic changes

@@
 
-\glspl{ffnn} have been the basis of most state-of-the-art \gls{ml} applications in recent years.
+\glspl{ffnn} have been the core of most state-of-the-art \gls{ml} applications in recent years specifically \glspl{ids}.
 Experiments show that the deeper the network is, the more desirable it is to acquire more knowledge.
-Nonetheless, with the growing number of layers, obtaining fast predictions has become a difficult task despite the use of special hardware such as GPUs.
-We propose a new architecture to obtain predictions without conducting a forward-pass until the last layer.
-The architecture is able to deal with either binary or multiclass classification problems and trades prediction speed for the confidence of the network.
+Nonetheless, with the growing number of layers, obtaining fast predictions with less resources has become a difficult task despite the use of special hardware such as GPUs.
+We propose a new architecture to detect network attacks with minimal resources.
+The architecture is able to deal with either binary or multiclass classification problems and trades prediction speed for the accuracy of the network.
 We evaluate our proposal with two different network intrusion detection datasets.
@@
 \glspl{ffnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
-Many of the state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan2016power} .
+Many state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
+Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan_power_2016} .
 
@@
 
-\section{Related Work}
-
  \begin{figure}
@@
 \end{figure}
+
+\section{Related Work}
 
@@
 	\footnotesize
-	\begin{tabular}{ccc}
+	\resizebox{\columnwidth}{!}{
+	\begin{tabular}{|c|c|c|}
 		\toprule
@@
 
-\section{Related Work}
-
 \section{Evaluation and Discussion}
 \label{evaluation_and_discussion}
-We discuss in this section different experimental results and observations. Overall, EagerNet shows promising performance comparable to \glspl{ffnn} with the ability to save resources.
+We discuss in this section different experimental results and observations.
+Overall, EagerNet shows promising performance comparable to \glspl{ffnn} with the ability to save resources.
+all experiments are reproducible and available on our repository\footnote{github.com/CN-TU/ids-backdoor/tree/eager}.
 

### 2020-07-21 paper/paper.tex
Fix table

--- 2020-07-21 paper/paper.tex
@@
 % The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
+%\usepackage{cite}
 \usepackage{amsmath,amssymb,amsfonts}
@@
 \usepackage{graphicx}
-\usepackage{textcomp}
+%\usepackage{textcomp}
 \usepackage{xcolor}
@@
 \usepackage{booktabs}
-
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\usepackage{tabularx}
 
 \usepackage[nomain, toc, acronym]{glossaries}
-\glsdisablehyper
+%\glsdisablehyper
 
@@
 
-\newcommand{\unsw}{UNSW-NB15}
-\newcommand{\cic}{CIC-IDS-2017}
+%\newcommand{\unsw}{UNSW-NB15}
+%\newcommand{\cic}{CIC-IDS-2017}
 
@@
 Many of the state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan2016power} .
+Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan_power_2016} .
 
@@
 Thereafter, packet features are extracted and the statistical combinations are computed.
-Table \ref{tab:features} shows CAIA \cite{williams2006preliminary} network traffic representation.
+Table \ref{tab:features} shows CAIA \cite{williams_preliminary_2006} network traffic representation.
 It consists of 12 features 7 of which are measured in both directions and 2 of which are expanded using four statistical combinations (i.e.
@@
 	\footnotesize
-	\begin{tabular}{ccc}
-		\toprule
+	\begin{tabular}{|c|c|c|}
+		\hline
 		\textbf{Direction}                    & \textbf{Features}        & \textbf{Statistical Operations}        \\
-		\midrule
+		\hline
 
@@
 		                                      & octetTotalCount          &                                        \\
-		\midrule
-		\multirow{7}{*}{Forward and Backward} & ipTotalLength            & \multirow{2}{*}{Mean, Min, Max, Stdev} \\
+		\hline
+		\multirow{7}{*}{\shortstack{Forward and\\ Backward}} & ipTotalLength            & \multirow{2}{*}{Mean, Min, Max, Stdev} \\
 		                                      & interPacketTimeSeconds   &                                        \\
-		\cmidrule{2-3}
+		\cline{2-3}
 		                                      & packetTotalCount         &                                        \\
@@
 
-		\midrule
+		\hline
 
 	\end{tabular}
-	}
 
@@
 
-After computing the losses, we use $Adam$ optimizer \cite{kingma2014adam} for computing gradients with a learning rate of 0.001.
+After computing the losses, we use $Adam$ optimizer \cite{kingma_adam_2017} for computing gradients with a learning rate of 0.001.
 
@@
 For our experiments, we use two intrusion detection datasets.
-CICIDS2017 \cite{sharafaldin2018toward} which consists of 14 network attack classes in addition to realistic user profiles.
-UNSW-NB15 \cite{moustafa2015unsw} which, in addition to synthetically generated normal traffic, again contains 9 attack families.
+CICIDS2017 \cite{sharafaldin_toward_2018} which consists of 14 network attack classes in addition to realistic user profiles.
+UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic, again contains 9 attack families.
 As mentioned above, we represent both datasets using the CAIA feature set and extract 2.317.922 and 2.065.224 flows, respectively, from CICIDS2017 and UNSW-NB15.

### 2020-07-21 paper/paper.tex

@@
 \usepackage{booktabs}
-\usepackage{tabularx}
 
@@
 \glspl{ffnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
-Many of the state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
+Many state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
 Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan_power_2016} .

### 2020-07-21 paper/paper.tex
Fix typos, update figure

@@
 
-\begin{figure*}
-\center
-\begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-3x128.pdf}} &
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-12x64.pdf}}\\
-\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-3x128.pdf}} &
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-12x64.pdf}}
-\end{tabular}
-\caption{Confidence threshold effect on accuracy and number of needed layers.}
-\label{fig:confidence}
-\end{figure*}
-
-\begin{figure}
-\center
-\begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 4in]{figures/multiclass17.pdf}} &
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass15.pdf}}
-\end{tabular}
-\caption{Accuracy plot multiclass.}
-\end{figure}
-
 \textbf{Optimal Confidence Threshold.}
@@
 
+\begin{figure*}
+\center
+\begin{tabular}{cccc}
+\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-3x128.pdf}} &
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-12x64.pdf}}\\
+\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-3x128.pdf}} &
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-12x64.pdf}}
+\end{tabular}
+\caption{Confidence threshold effect on accuracy and number of needed layers.}
+\label{fig:confidence}
+\end{figure*}
+
+\begin{figure}
+\center
+\begin{tabular}{cccc}
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 0.85\columnwidth]{figures/multiclass17.pdf}} \\
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = \columnwidth]{figures/multiclass15.pdf}}
+\end{tabular}
+\caption{Accuracy plot multiclass.}
+\end{figure}
+
 \section{Conclusion}

### 2020-07-21 learn.py

@@ -7,7 +7,6 @@ import math
 import random
 import torch
 from torch.utils.data import DataLoader, Dataset
-from torch.utils.data.sampler import SubsetRandomSampler
 from tensorboardX import SummaryWriter
 import socket
 from datetime import datetime
@@ -18,10 +17,8 @@ import gzip
 import copy
 
 import sklearn
-from sklearn.ensemble import RandomForestClassifier
 from sklearn.linear_model import LogisticRegression
 from sklearn.metrics import accuracy_score,recall_score, precision_score, f1_score, balanced_accuracy_score, confusion_matrix, classification_report
-from sklearn.tree._tree import TREE_LEAF, TREE_UNDEFINED
 
 import collections
 import pickle
@@ -39,6 +36,7 @@ from tqdm import tqdm
 from matplotlib import cm
 
 # TODOs: unify the binary code
+#		 add attack mapping json for better plotting of long names
 
 def output_scores(y_true, y_pred, only_accuracy=False, average='binary'):
 	metrics = [ accuracy_score(y_true, y_pred) ]
@@ -489,7 +487,7 @@ if __name__=="__main__":
 
 	opt = parser.parse_args()
 	print('#'*40)
-	print('>> Configuration ' + opt)
+	print('>> Configuration ' + str(opt))
 	print('#'*40)
 
 	seed = opt.manualSeed

### 2020-07-21 paper/paper.tex
Finish related work and fix figures

--- 2020-07-21 paper/paper.tex
@@
 
+\newcommand{\mynote}[3]{
+    \fbox{\bfseries\sffamily\scriptsize#1}
+    {\small$\blacktriangleright$\textsf{\emph{\color{#3}{#2}}}$\blacktriangleleft$}}
+
+\newcommand{\todo}[1]{\mynote{TODO}{#1}{red}}
+
 \usepackage[nomain, toc, acronym]{glossaries}
@@
 
-\newcommand\note[2]{{\color{#1}#2}}
-\newcommand\todo[1]{{\note{red}{TODO: #1}}}
+%\newcommand\note[2]{{\color{#1}#2}}
+%\newcommand\todo[1]{{\note{red}{TODO: #1}}}
 
@@
 % conclusion
+\todo{Max: For such a short paper it might not be necessary to give an overview?}
 The rest of the paper is organized as follows: In Section ...
@@
 \section{Related Work}
+
+The concept of \textit{cascading} classifiers defines several classifiers that are used in a cascade \cite{viola_robust_2004, xu_classifier_2014, wang_efficient_2015}.
+If accuracy of a classifier in the beginning is good enough, then there is no need to evaluate the other classifiers in the chain (cascade), which saves computational resources.
+This approach was predominantly explored for classical machine learning methods.
+Two possibilities are to either place all classifiers in a row or as a tree.
+If a tree is used, different classifiers can be used at each step, depending on which features are probably relevant for the sample currently under investigation.
+The difference to our approach is that we exclusively consider neural networks as the classifier and that our approach is significantly easier to implement.
+
+\cite{bolukbasi_adaptive_2017} aim to reduce the computational complexity during evaluation for large Convolutional Neural Networks (CNNs) using an additional classifier after each layer.
+These additional classifiers learn to look at the output of the previous layer and decide whether it's necessary to continue evaluating the next layers.
+Unlike our approach, this approach is primarily geared towards making the evaluation of a given neural network more efficient, while our approach aims to train a neural network from scratch.
+Furthermore, requires advanced math to be implemented, making it difficult to implement and deploy for practitioners.
+
+The approach that is most similar to ours is presented by \cite{leroux_cascading_2017}.
+They introduce an additional output layer after each layer and then decide whether the next layers should be evaluated based on whether the confidence of the current output layer is high enough.
+They also argue that if many layers are required, the computation can be offloaded to a more powerful machine in the cloud, which saves further computational effort.
+The main difference of this approach compared to ours is that we can show that no additional output layers are needed but instead our approach only needs one extra neuron per layer, further reducing the computational overhead of our approach.
+
+Another line of work \cite{seo_neural_2018, yu_learning_2017, graves_adaptive_2017, bachl_sparseids_2020} focuses on teaching Recurrent Neural Networks (RNNs), neural networks for sequences, to skip irrelevant parts of the input sequence.
+
+The major difference between all existing works and ours is simplicity.
+While many of the previously proposed works report good results, they usually require advanced math knowledge to be implemented, which hinders their deployment in real-world scenarios.
 
@@
 \begin{figure*}
-\center
-\begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-3x128.pdf}} &
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-17-12x64.pdf}}\\
-\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-3x128.pdf}} &
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 2.5in]{figures/conf-acc-15-12x64.pdf}}
-\end{tabular}
+\centering
+\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-3x128.pdf}}
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
+\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-3x128.pdf}}
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-12x64.pdf}}
 \caption{Confidence threshold effect on accuracy and number of needed layers.}
@@
 
-\begin{figure}
-\center
-\begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 4in]{figures/multiclass17.pdf}} &
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width = 3in]{figures/multiclass15.pdf}}
-\end{tabular}
+\begin{figure*}
+\centering
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass17.pdf}}
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass15.pdf}}
 \caption{Accuracy plot multiclass.}
-\end{figure}
+\end{figure*}
 

### 2020-07-21 paper/paper.tex
Change paper and add some comments

--- 2020-07-21 paper/paper.tex
@@
 \usepackage{booktabs}
+\usepackage{hyperref}
 
@@
 
-\glspl{ffnn} have been the core of most state-of-the-art \gls{ml} applications in recent years specifically \glspl{ids}.
-Experiments show that the deeper the network is, the more desirable it is to acquire more knowledge.
+\glspl{ffnn} have been the core of most state-of-the-art \gls{ml} applications in recent years and also have been widely used for \glspl{ids}.
+Experiments results from the last years show that generally deeper neural networks with more layers perform better than shallow models.
 Nonetheless, with the growing number of layers, obtaining fast predictions with less resources has become a difficult task despite the use of special hardware such as GPUs.
@@
 We evaluate our proposal with two different network intrusion detection datasets.
-Results suggest that it is possible to obtain comparable accuracies to simple \glspl{ffnn} without evaluating all layers, thus obtaining early predictions and saving energy and computational efforts.
+Results suggest that it is possible to obtain comparable accuracies to simple \glspl{ffnn} without evaluating all layers for all samples, thus obtaining early predictions and saving energy and computational efforts.
 
@@
 
-\begin{IEEEkeywords}
+\todo{Max: I guess you mean Fully-Connected Neural Networks (FCNNs) when you talk about FNNs? Because FNN only means that the neural network has no cycles. So also a CNN is an FNN.}
 
-\end{IEEEkeywords}
+%\begin{IEEEkeywords}
+%
+%\end{IEEEkeywords}
 
@@
 \glspl{ffnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
-Many state-of-the-art architectures currently used in different applications are effectively based on simple \glspl{ffnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable it is for better learning \cite{eldan_power_2016} .
+Many state-of-the-art architectures currently used in different applications are effectively based on simple \gls{ffnn} architectures.
+Experiments show that the more layers (and neurons) the network includes, the more desirable because it results in better learning performane \cite{eldan_power_2016}.
 
@@
 x_{m} \rbrace$ and makes a prediction.
-Adjusting weights in such situations is a non-trivial problem since the weights of each layer may have an impact on the previous layers.
+Adjusting weights in such situations is a non-trivial problem since the weights of each layer may have an impact on the following layers.
 Fortunately, several algorithms have been proposed to reduce the complexity of the task (e.g.
@@
 In terms of efficiency, two problems arise when using an extremely deep neural networks: (1) The loss during training needs more time to converge and (2) prediction time is proportional to the number of layers.
-Even though most modern hardware is able to cope with offline single predictions without significant delay, online massive neural network implementations may suffer from the latter issue if non-specialized hardware is used to produce predictions.
+Even though most modern hardware is able to output predictions without significant delay, prediction speed is still relevant for learning problems, in which many predictions are required per unit of time and delay is important.
+This is, for example, relevant in the context of object detection in computer vision or also network intrusion detection.
+
 %EagerNet
 To solve this, we propose a new architecture that, on the one hand, uses the entire network capacity for learning and, on the other hand, stops the forward-pass and makes predictions as soon as confidence reaches a certain threshold.
-In other words, it makes it possible to avoid evaluating the entire network and to receive predictions from intermediate layers with the same performance of a simple \gls{ffnn}.
-The proposed approach therefore allows, where possible, the reduction of computing resources and energy usage preserving comparable performance of a full forward-pass.
+In other words, it makes it possible to avoid evaluating the entire network with all layers and to stop the neural network evaluation at a particular layer if the achieved confidence of the neural network is high enough.
+The proposed approach therefore allows, where possible, the reduction of computing resources and energy usage while achieving performance comparable to that of a full forward-pass.
 
@@
 
-The concept of \textit{cascading} classifiers defines several classifiers that are used in a cascade \cite{viola_robust_2004, xu_classifier_2014, wang_efficient_2015}.
+The concept of \textit{cascading} classifiers encompasses several classifiers that are used in a cascade (like a chain) \cite{viola_robust_2004, xu_classifier_2014, wang_efficient_2015}.
 If accuracy of a classifier in the beginning is good enough, then there is no need to evaluate the other classifiers in the chain (cascade), which saves computational resources.
 This approach was predominantly explored for classical machine learning methods.
-Two possibilities are to either place all classifiers in a row or as a tree.
+Two possibilities are to either place all classifiers in a row or in the shape of tree.
 If a tree is used, different classifiers can be used at each step, depending on which features are probably relevant for the sample currently under investigation.
-The difference to our approach is that we exclusively consider neural networks as the classifier and that our approach is significantly easier to implement.
+The difference to our approach is that we exclusively consider neural networks as the classifier and that our approach is significantly more straightforward to implement.
 
@@
 Unlike our approach, this approach is primarily geared towards making the evaluation of a given neural network more efficient, while our approach aims to train a neural network from scratch.
-Furthermore, requires advanced math to be implemented, making it difficult to implement and deploy for practitioners.
+Furthermore, their approach requires advanced math, making it difficult to implement and deploy for practitioners.
 
@@
 They also argue that if many layers are required, the computation can be offloaded to a more powerful machine in the cloud, which saves further computational effort.
-The main difference of this approach compared to ours is that we can show that no additional output layers are needed but instead our approach only needs one extra neuron per layer, further reducing the computational overhead of our approach.
+The main difference of this approach compared to ours is that we can show that no additional output layers are needed but instead our approach only needs one extra neuron per layer, which further reduces the computational overhead of our approach.
 
-Another line of work \cite{seo_neural_2018, yu_learning_2017, graves_adaptive_2017, bachl_sparseids_2020} focuses on teaching Recurrent Neural Networks (RNNs), neural networks for sequences, to skip irrelevant parts of the input sequence.
+Another line of work \cite{seo_neural_2018,yu_learning_2017, graves_adaptive_2017, bachl_sparseids_2020} focuses on teaching Recurrent Neural Networks (RNNs) -- neural networks for sequences -- to skip irrelevant parts of the input sequence.
 
@@
 While many of the previously proposed works report good results, they usually require advanced math knowledge to be implemented, which hinders their deployment in real-world scenarios.
+Also, they commonly include components such as additional layers, which we show are not necessary.
 
 \section{Supervised Learning for Network Intrusion Detection}
+
+\todo{Max: Add missing references}
 
@@
 The idea of training a classifier on pre-stored attack patterns and using the same classifier to recognize similar patterns is commonly used and remarkable success has so far been achieved \cite{survey}.
-In particular, \glspl{ffnn} is commonly used whenever large amounts of data are available.
+In particular, \glspl{ffnn} are commonly used whenever large amounts of data are available.
 In order to present network data to such architectures, a number of traffic representation families have been proposed depending on the application.
-However, the statistical representation of flow characteristics is still the most appropriate.
+However, the statistical representation of flow characteristics is still the most widely used one.
 In this work, we define the network flow to be the exchange of packets between two end-hosts.
@@
 Thereafter, packet features are extracted and the statistical combinations are computed.
-Table \ref{tab:features} shows CAIA \cite{williams_preliminary_2006} network traffic representation.
-It consists of 12 features 7 of which are measured in both directions and 2 of which are expanded using four statistical combinations (i.e.
+Table \ref{tab:features} shows the CAIA \cite{williams_preliminary_2006} network traffic representation.
+It consists of 12 features, 7 of which are measured in both directions and 2 of which are expanded using four statistical combinations (i.e.
 \emph{A} becomes \emph{mean(A)}, \emph{min(A)}, \emph{max(A)} and \emph{stdev(A)}).
-In addition, we set an observation timeout of 1,800 seconds, after which we terminate flows.
-Each flow is considered to be a data sample and has two sets of labels: binary (attack/benign) label and attack family.
-After a training process, the trained \gls{ids} must be able to correctly classify new flows.
+We use CAIA for all experiments in this paper and set an observation timeout of 1,800 seconds, after which we terminate flows.
+Each flow is considered to be one sample and has two sets of labels: a binary (attack/benign) label and also the attack family (one binary label for each attack family).
+The goal is that after a training process, the trained \gls{ids} can correctly classify new flows and also determine what kind of attack family they belong to.
 
@@
 
-		                                      & flowDurationMilliseconds &                                        \\
+		                                      \multirow{5}{*}{No direction} & flowDurationMilliseconds & \multirow{5}{*}{None} \\
 		                                      & sourceTransportPort      &                                        \\
@@
 		\cline{2-3}
-		                                      & packetTotalCount         &                                        \\
+		                                      & packetTotalCount         & \multirow{5}{*}{None}                                       \\
 		                                      & tcpSynTotalCount         &                                        \\
@@
 
-Because several network attack families are simple to detect on the basis of linear input combinations, the standard \gls{ffnn} is slightly modified to cope with the difficulty of detection.
-Our novel architecture is based on the assumption that deep neural networks can learn complex functions with simpler mappings learned in the first few layers.
-As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for predictions at each layer.
+\todo{Max: $X$ is generally for matrices as far as I know while $x$ is for vectors.}
+
+Because some network attack families are simple to detect compared to others we modify the standard \gls{ffnn} architecture so that for easily classifiable samples the network doesn't have to evaluate all layers.
+Our novel architecture is based on the assumption that deep neural networks that with more layers can learn increasingly complex functions which are only required for the classification of some particular notoriously hard-to-classify samples.
+As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for direct predictions at each layer.
 The EagerNet architecture is shown in Figure \ref{fig:eagerNet_forward}.
 Initial neurons are shown in gray and output neurons (per layer) are shown in green.
-Once a flow is observed and features are extracted, a sample $\textbf{X}$ is fed to the network via the input layer.
-The network proceeds evaluating the output of each layer and yields a prediction confidence value.
+Once a flow is observed and features are extracted, a sample $x$ is fed to the network via the input layer.
+The network proceeds evaluating the neural network layer by layer and yields a prediction confidence value at each layer.
 We define the confidence as the value that represents how sure the network is about a certain decision or sample belonging to a certain class.
-Based on a predefined threshold, the network measures confidence and determines whether to avoid testing additional layers or proceed to do so in order to gain more confidence.
-This methodology ensures that simpler samples are classified early in the forward-pass and that more complicated, noisy samples are passed into deeper layers.
-In the \ref{confidence-speed-tradeoff} Section, we show that EagerNet reduces resource consumption on average keeping comparable performance to normal \gls{ffnn}.
+It is a number between 0 and 1 that is obtained by applying the sigmoid function \cite{noauthor_sigmoid_2020} on the output neuron.
+By looking at the confidence confidence and we can determine whether additional layers should be processed or if the currently obtained confidence is high enough.
+This methodology ensures that simpler samples are classified early in the forward-pass and that more complicated samples are passed into deeper layers.
+In the \autoref{subsubsec:confidence-speed-tradeoff}, we show that EagerNet reduces resource consumption on average, while achieving comparable performance compared to normal \gls{ffnn}.
 
@@
 \subsubsection{Binary Vs. Multiclass Classification}
-We adjust the output layer to build two separate sub-architectures.
-The first architecture, consisting of a binary attack detection, i.e.
+We experiment with two different architectures.
+The first architecture, consists of a binary attack detector, i.e.
 a 1 is an attack flow and a 0 is a benign flow.
 The predictions in this case are given by a single output neuron followed by a sigmoid activation function to force values between 0 and 1.
-During back-propagation, we use a binary cross-entropy loss including the activation function and average the loss over the batch.
-Equation \ref{eq:bceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the sigmoid activation function  $\sigma(.)$.
+During back-propagation, we use a binary cross-entropy loss and average the loss over the batch.
+\autoref{eq:bceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input being $x_{n}$ and the sigmoid activation function $\sigma(\cdot)$.
 
@@
 
-In the case of a multiclass problem, the output layers consist of as many neurons as attack families followed by a softmax activation functions that allow at most only one neuron to be 1.
-During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there is no overlap between classes.
-Equation \ref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function  $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
+For the multiclass approach, on the other side, there is one output neuron per attack family, and each neuron's output indicates how likely it is, that the currently processed sample belongs to the given attack family.
+During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there is can be no overlap between classes.
+\autoref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function  $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
 
@@
 Overall, EagerNet shows promising performance comparable to \glspl{ffnn} with the ability to save resources.
-all experiments are reproducible and available on our repository\footnote{github.com/CN-TU/ids-backdoor/tree/eager}.
+All experiments are reproducible and the code is available in our repository\footnote{\url{github.com/CN-TU/ids-backdoor/tree/eager}}.
 
@@
 CICIDS2017 \cite{sharafaldin_toward_2018} which consists of 14 network attack classes in addition to realistic user profiles.
-UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic, again contains 9 attack families.
+UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic again contains 9 attack families.
 As mentioned above, we represent both datasets using the CAIA feature set and extract 2.317.922 and 2.065.224 flows, respectively, from CICIDS2017 and UNSW-NB15.
 For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
-We use $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
+We use a split of $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
 
@@
 In Table \ref{tab:weights} we gather results from different networks trained using different numbers of hidden layers and weight distributions with 128 neurons per layer.
-In addition, we compare the results using four metrics (Accuracy, Precision , Recall and Youden Index) and show the performance of the last layer (hence raising the confidence level to the maximum).
-The CICIDS2017 indices show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
+In addition, we compare the results using four metrics (Accuracy, Precision, Recall and Youden's J) and show the performance of the last layer (hence raising the confidence level to the maximum).
+The CICIDS2017 indices \todo{Max: Indices?} show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
 In fact, since the last layer is always trained with one loss regardless of the weight distribution, only the first few layers are affected by the combined-loss learning procedure.
 Nonetheless, since the accuracy did not improve when increasing the number of layers, that implies that most network patterns are already separable after the few first layers.
-The results of UNSW-NB15 show a slightly different behavior.
-The uniform weighting of losses reflects the overall best scores and, in addition, the deeper architectures outperform the shallow ones.
-This implies that the network patterns in this dataset are more dynamic than the previous one and that more non-linearities (layers) are required to find the optimal mapping function.
+The results of UNSW-NB15 show a slightly different behavior \todo{Max: Sure?
+All the results look the same?}.
+The uniform weighting of losses gives the overall best scores and, in addition, the deeper architectures outperform the shallow ones.
+This implies that the network patterns in this dataset are complex than in the previous dataset and that more non-linearities (layers) are required to find the optimal mapping function.
 
 \begin{table}
-
 \centering
@@
 \subsubsection{Confidence-Speed Tradeoff}
-\label{confidence-speed-tradeoff}
+\label{subsubsec:confidence-speed-tradeoff}
 
@@
 \item The accuracy of CICIDS2017 increases almost linearly with respect to confidence and reaches its maximum when all layers are used.
-\item The accuracy of UNSW-NB15 is stable when the confidence level is increased until \~0.98 where it leaps noticeably when a further layer is evaluated.
-This suggests that some patterns are "learn-able" only on that specific layer.
+\item The accuracy of UNSW-NB15 is stable when the confidence level is increased until $\approx 0.98$ where it leaps noticeably when a further layer is evaluated.
+This suggests that some patterns are ``learnable'' only on that specific layer.
 \end{itemize}
 Overall, EagerNet shows that it is possible to trade a tiny percentage of accuracy in order to save a significant amount of resources.
-In Section \ref{comparability_to_fnns}, we see that, in order to achieve comparable accuracy to the traditional \gls{ffnn}, the previous trade is possible and therefore also the possibility to reduce costs.
+In \autoref{subsubsec:comparability_to_fnns}, we see that, in order to achieve comparable accuracy to the traditional \gls{ffnn}, the previous trade is possible and therefore also the possibility to reduce costs.
 
@@
 
-\textbf{Optimal Confidence Threshold.}
+\paragraph{Optimal Confidence Threshold.}
 During deployment, the optimum confidence threshold is set by the user and plays a role in the overall achieved accuracy.
-Setting the threshold too low can cause the network to make decisions at an early stage, thereby gaining resources but achieving low accuracy.
-Similarly, setting the threshold too high can cause the network to skip the proper layers and thus always use the last one.
-The desired confidence level can therefore be obtained during the training phase by determining on the basis of the desired accuracy as shown in Figure \ref{fig:confidence}.
+Setting the threshold too low can cause the network to make decisions at an early stage, thereby using fewer resources but achieving low accuracy.
+Similarly, setting the threshold too high can cause the network to always use all layers, which means a waste of resources.
+The desired confidence level can therefore be obtained after the training phase by determining on the basis of the desired accuracy as shown in \autoref{fig:confidence}.
 
 \subsubsection{Comparability to FNNs}
-\label{comparability_to_fnns}
+\label{subsubsec:comparability_to_fnns}
+\todo{Max: Don't understand}
 
@@
 \subsubsection{What Did the Network Learn?}
+\todo{Max: Don't understand}
 % why from 512 to 32
+
+\subsubsection{Advantage of backpropagating the losses of all outputs until the beginning}
+\todo{Max: Don't know if this has been already covered elsewhere?}
+
+\begin{figure*}
+\center
+\begin{tabular}{cccc}
+\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-3x128.pdf}} &
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
+\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-3x128.pdf}} &
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-12x64.pdf}}
+\end{tabular}
+\caption{Confidence threshold effect on accuracy and number of needed layers.}
+\label{fig:confidence}
+\end{figure*}
+
+\begin{figure*}
+\centering
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass17.pdf}}
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass15.pdf}}
+\caption{Accuracy plot multiclass.}
+\end{figure*}
 

### 2020-07-22 paper/paper.tex
Add reference
add content to paper, modify figures and add new references

@@
 \usepackage{hyperref}
+\usepackage{float}
 
@@
 \newacronym{ffnn}{FNN}{Feedforward Neural Network}
+\newacronym{fcnn}{FCNN}{Fully Connected Neural Network}
 
@@
 
-\glspl{ffnn} have been the core of most state-of-the-art \gls{ml} applications in recent years and also have been widely used for \glspl{ids}.
+\glspl{fcnn} have been the core of most state-of-the-art \gls{ml} applications in recent years and also have been widely used for \glspl{ids}.
 Experiments results from the last years show that generally deeper neural networks with more layers perform better than shallow models.
@@
 We evaluate our proposal with two different network intrusion detection datasets.
-Results suggest that it is possible to obtain comparable accuracies to simple \glspl{ffnn} without evaluating all layers for all samples, thus obtaining early predictions and saving energy and computational efforts.
-
+Results suggest that it is possible to obtain comparable accuracies to simple \glspl{fcnn} without evaluating all layers for all samples, thus obtaining early predictions and saving energy and computational efforts.
+\todo{Max: I guess you mean Fully-Connected Neural Networks (FCNNs) when you talk about FNNs? Because FNN only means that the neural network has no cycles. So also a CNN is an FNN.}
+\todo{Fares: The idea should work with any feedforward network. But since we don't experiment with CNNs, I'll replace it now with FCNN.}
 \end{abstract}
-
-\todo{Max: I guess you mean Fully-Connected Neural Networks (FCNNs) when you talk about FNNs? Because FNN only means that the neural network has no cycles. So also a CNN is an FNN.}
 
@@
 \section{Introduction}
-% what are ffnn
-\glspl{ffnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
-Many state-of-the-art architectures currently used in different applications are effectively based on simple \gls{ffnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable because it results in better learning performane \cite{eldan_power_2016}.
+% what are fcnn
+\glspl{fcnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
+Many state-of-the-art architectures currently used in different applications are effectively based on simple \gls{fcnn} architectures.
+Experiments show that the more layers (and neurons) the network includes, the more desirable because it results in better learning performance \cite{eldan_power_2016}.
 
 % forward-backward
-Traditional \gls{ffnn} are based on simple logistic regression units that combine multiple inputs multiplied by a set of weights and passed through an activation function (e.g.
+Traditional \gls{fcnn} are based on simple logistic regression units that combine multiple inputs multiplied by a set of weights and passed through an activation function (e.g.
 sigmoid) to obtain a scalar.
@@
 As far as neural networks are concerned, logistic regression units are stacked together to create layers and then layers are stacked together again to create the network itself.
-Figure \ref{fig:ffnn} demonstrates a basic \gls{ffnn} that utilizes the input vector $\lbrace x_{1}, x_{2} ...
+Figure \ref{fig:fcnn} demonstrates a basic \gls{fcnn} that utilizes the input vector $\textbf{x}^{T} = \lbrace x_{1}, x_{2} ...
 x_{m} \rbrace$ and makes a prediction.
@@
 \todo{Max: For such a short paper it might not be necessary to give an overview?}
-The rest of the paper is organized as follows: In Section ...
+\todo{Fares: I now added just a sentence}
+In the following sections, we present the existing related work, introduce the concept behind EagerNet and finally show and analyze the evaluation results based on two network intrusion detection datasets.
 
  \begin{figure}
-  \includegraphics[width=\linewidth]{figures/ffnn_new.pdf}
-  \caption{Conventional \gls{ffnn} architecture.}
-  \label{fig:ffnn}
+  \includegraphics[width=\linewidth]{figures/fcnn_new.pdf}
+  \caption{Conventional \gls{fcnn} architecture.}
+  \label{fig:fcnn}
 \end{figure}
@@
 
-The approach that is most similar to ours is presented by \cite{leroux_cascading_2017}.
+The approach that is most similar to ours is presented by \cite{leroux_resource-constrained_2015,leroux_cascading_2017}.
 They introduce an additional output layer after each layer and then decide whether the next layers should be evaluated based on whether the confidence of the current output layer is high enough.
@@
 \todo{Max: Add missing references}
+\todo{Fares: Done.}
 
 \gls{ml} has been widely used in the last decade for network traffic analysis and specifically anomaly detection systems.
-Many of the works proposed \gls{ids} architectures based on well-known supervised techniques \cite{survey}.
-The idea of training a classifier on pre-stored attack patterns and using the same classifier to recognize similar patterns is commonly used and remarkable success has so far been achieved \cite{survey}.
-In particular, \glspl{ffnn} are commonly used whenever large amounts of data are available.
+Many of the works proposed \gls{ids} architectures based on well-known supervised techniques \cite{buczak_survey_2016,berman_survey_2019}.
+The idea of training a classifier on pre-stored attack patterns and using the same classifier to recognize similar patterns is commonly used and remarkable success has so far been achieved \cite{vigneswaran_evaluating_2018,ferrag_deep_2020,almseidin_evaluation_2017}.
+In particular, \glspl{fcnn} are commonly used whenever large amounts of data are available.
 In order to present network data to such architectures, a number of traffic representation families have been proposed depending on the application.
@@
 
-\todo{Max: $X$ is generally for matrices as far as I know while $x$ is for vectors.}
-
-Because some network attack families are simple to detect compared to others we modify the standard \gls{ffnn} architecture so that for easily classifiable samples the network doesn't have to evaluate all layers.
-Our novel architecture is based on the assumption that deep neural networks that with more layers can learn increasingly complex functions which are only required for the classification of some particular notoriously hard-to-classify samples.
-As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for direct predictions at each layer.
-The EagerNet architecture is shown in Figure \ref{fig:eagerNet_forward}.
-Initial neurons are shown in gray and output neurons (per layer) are shown in green.
-Once a flow is observed and features are extracted, a sample $x$ is fed to the network via the input layer.
-The network proceeds evaluating the neural network layer by layer and yields a prediction confidence value at each layer.
-We define the confidence as the value that represents how sure the network is about a certain decision or sample belonging to a certain class.
-It is a number between 0 and 1 that is obtained by applying the sigmoid function \cite{noauthor_sigmoid_2020} on the output neuron.
-By looking at the confidence confidence and we can determine whether additional layers should be processed or if the currently obtained confidence is high enough.
-This methodology ensures that simpler samples are classified early in the forward-pass and that more complicated samples are passed into deeper layers.
-In the \autoref{subsubsec:confidence-speed-tradeoff}, we show that EagerNet reduces resource consumption on average, while achieving comparable performance compared to normal \gls{ffnn}.
-
 \begin{figure*}[htp]
@@
 \end{figure*}
+
+\todo{Max: $X$ is generally for matrices as far as I know while $x$ is for vectors.}
+\todo{Fares: actually $\textbf{x}$ is used for vectors, I forgot to change it because it was consitent to what I wrote in the introduction. Now it's better.}
+
+Because some network attack families are simple to detect compared to others we modify the standard \gls{fcnn} architecture so that for easily classifiable samples the network doesn't have to evaluate all layers.
+Our novel architecture is based on the assumption that deep neural networks that with more layers can learn increasingly complex functions which are only required for the classification of some particular notoriously hard-to-classify samples.
+As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for direct predictions at each layer.
+The EagerNet architecture is shown in Figure \ref{fig:eagerNet_forward}.
+Initial neurons are shown in gray and output neurons (per layer) are shown in green.
+Once a flow is observed and features are extracted, a sample $\textbf{x}$ is fed to the network via the input layer.
+The network proceeds evaluating the neural network layer by layer and yields a prediction confidence value at each layer.
+We define the confidence as the value that represents how sure the network is about a certain decision or sample belonging to a certain class.
+It is a number between 0 and 1 that is obtained by applying the sigmoid function \cite{noauthor_sigmoid_2020} on the output neuron.
+By looking at the confidence confidence and we can determine whether additional layers should be processed or if the currently obtained confidence is high enough.
+This methodology ensures that simpler samples are classified early in the forward-pass and that more complicated samples are passed into deeper layers.
+In the \autoref{subsubsec:confidence-speed-tradeoff}, we show that EagerNet reduces resource consumption on average, while achieving comparable performance compared to normal \gls{fcnn}.
 
@@
 We discuss in this section different experimental results and observations.
-Overall, EagerNet shows promising performance comparable to \glspl{ffnn} with the ability to save resources.
+Overall, EagerNet shows promising performance comparable to \glspl{fcnn} with the ability to save resources.
 All experiments are reproducible and the code is available in our repository\footnote{\url{github.com/CN-TU/ids-backdoor/tree/eager}}.
@@
 \subsection{Datasets}
+
+\begin{figure}
+  \centering
+  \begin{tabular}{@{}c@{}}
+    \includegraphics[width=.8\linewidth]{figures/17_counts.pdf} \\[\abovecaptionskip]
+    \small (a) CICIDS2017
+  \end{tabular}
+
+  \vspace{0.2cm}
+
+  \begin{tabular}{@{}c@{}}
+    \includegraphics[width=.8\linewidth]{figures/15_counts.pdf} \\[\abovecaptionskip]
+    \small (b) UNSW-NB15
+  \end{tabular}
+
+  \caption{Traffic distribution represented by number of flows.}
+  \label{fig:datasets}
+\end{figure}
+
 For our experiments, we use two intrusion detection datasets.
@@
 UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic again contains 9 attack families.
-As mentioned above, we represent both datasets using the CAIA feature set and extract 2.317.922 and 2.065.224 flows, respectively, from CICIDS2017 and UNSW-NB15.
+As mentioned above, we represent both datasets using the CAIA feature set and extract 2.317.922 and 2.065.224 flows, respectively, from CICIDS2017 and UNSW-NB15 (see \autoref{fig:datasets}).
 For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
@@
 
-\subsubsection{Weighted Loss}
+\subsubsection{Weighted loss}
 In Table \ref{tab:weights} we gather results from different networks trained using different numbers of hidden layers and weight distributions with 128 neurons per layer.
 In addition, we compare the results using four metrics (Accuracy, Precision, Recall and Youden's J) and show the performance of the last layer (hence raising the confidence level to the maximum).
-The CICIDS2017 indices \todo{Max: Indices?} show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
+The CICIDS2017 indices \todo{Max: Indices?
+Fares: it's commun to refer to acc, prec, recall etc.
+as indices/scores} show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
 In fact, since the last layer is always trained with one loss regardless of the weight distribution, only the first few layers are affected by the combined-loss learning procedure.
@@
 The results of UNSW-NB15 show a slightly different behavior \todo{Max: Sure?
-All the results look the same?}.
+All the results look the same?
+Fares: slightly better}.
 The uniform weighting of losses gives the overall best scores and, in addition, the deeper architectures outperform the shallow ones.
@@
 
-\begin{table}
+\begin{table}[H]
 \centering
@@
 
-{\raggedright * All architectures consist of 128 neurons per layer in addition to single output layer. \par}
+{\raggedright * All architectures consist of 128 neurons per layer.\par}
 \caption{Effect of weights on the performance of different networks}
@@
 
-\subsubsection{Confidence-Speed Tradeoff}
+\subsubsection{Confidence-Speed tradeoff}
 \label{subsubsec:confidence-speed-tradeoff}
+
+\begin{figure*}
+\center
+\begin{tabular}{cccc}
+\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-17-3x128.pdf}} &
+\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
+\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-15-3x128.pdf}} &
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-15-12x64.pdf}}
+\end{tabular}
+\caption{Confidence threshold effect on accuracy and number of needed layers.
+The maximum achieved accuracy does not represent the score obtained at the last layer but with a combination of best layers depending on the selection.
+Therefore, the last layer's accuracy alone might be less.}
+\label{fig:confidence}
+\end{figure*}
 
@@
 Overall, EagerNet shows that it is possible to trade a tiny percentage of accuracy in order to save a significant amount of resources.
-In \autoref{subsubsec:comparability_to_fnns}, we see that, in order to achieve comparable accuracy to the traditional \gls{ffnn}, the previous trade is possible and therefore also the possibility to reduce costs.
+In \autoref{subsubsec:comparability_to_fnns}, we see that, in order to achieve comparable accuracy to the traditional \gls{fcnn}, the previous trade is possible and therefore also the possibility to reduce costs.
 
-\begin{figure*}
-\centering
-\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-3x128.pdf}}
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
-\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-3x128.pdf}}
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-12x64.pdf}}
-\caption{Confidence threshold effect on accuracy and number of needed layers.}
-\label{fig:confidence}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass17.pdf}}
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass15.pdf}}
-\caption{Accuracy plot multiclass.}
-\end{figure*}
+\todo{Max: Confidence seems to be multiply defined. Fares: I don't understand}
 
@@
 
-\subsubsection{Comparability to FNNs}
+\subsubsection{Comparability to \glspl{fcnn}}
 \label{subsubsec:comparability_to_fnns}
-\todo{Max: Don't understand}
+\todo{Max: Don't understand. Fares: Now added}
 
-\begin{table}
+In \autoref{tab:comparability}, we show the F1 score obtained at the last layer of three different architectures: Simple \gls{fcnn}, \gls{fcnn} with logistic output neuron per layer and EagerNet.
+\todo{Fares: continuation}
+
+\begin{table}[H]
 
@@
 \toprule
-\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{ffnn}} & \textbf{EagerNet (last)} \\
+\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{fcnn}} & \textbf{EagerNet (last)} \\
 \midrule
@@
 \midrule
-\multirow{2}{*}{UNSW-NB15} & Binary & & \\
- & Multiclass & & \\
+\multirow{2}{*}{UNSW-NB15} & Binary & 0.919 & 0.908 \\
+ & Multiclass & 0.880 & 0.882 \\
 \midrule
@@
 {\raggedright * All architectures consist of 10 layers $\times$ 64 neurons in addition to input and output layers. \par}
-\caption{Accuracy scores.}
-
+\caption{F1 scores at the last layer of each architecture.}
+\label{tab:comparability}
 \end{table}
 
-\subsubsection{What Did the Network Learn?}
-\todo{Max: Don't understand}
-% why from 512 to 32
+\subsubsection{What did the network learn?}
+
+\begin{figure*}
+\center
+\begin{tabular}{cccc}
+\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
+\subfloat[CICIDS2017 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.35\columnwidth]{figures/multiclass17-3.pdf}} 
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.68\columnwidth]{figures/multiclass15.pdf}} &
+\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.38\columnwidth]{figures/multiclass15-3.pdf}}
+\end{tabular}
+\caption{Accuracy achieved on a test set per layer and attack family.
+Some attacks have poor accuracy because only few samples are observed (refer to \autoref{fig:datasets}).
+Forcing the network to learn those samples might lead to over-fitting.}
+\label{fig:multiclass}
+\end{figure*}
+
+\todo{Max: Don't understand. Fares: Now added}
+In order to understand how EagerNet determines which layer the best to stop evaluation, we use the multiclass variant to conduct an explanatory analysis.
+In \autoref{fig:multiclass}, we show the accuracy of the prediction obtained by different networks per layer and attack family.
+We observe two phenomena:
+\begin{itemize}
+\item  On average, the accuracy of predictions increases the deeper we get into the network.
+It was somehow expected, as more layers allow for more abstraction of the input space and thus better separation.
+However, only a few first layers are needed to obtain maximum accuracy for each attack family.
+This implies that the EagerNet architecture significantly decreases resources when used for intrusion detection applications.
+\item \todo{Fares: recheck this} Some attacks families show extremely poor accuracy.
+This is due to two reasons: (1) too few samples from a category are mostly ignored by the network to reduce over-fitting and (2) If a specific layer focuses on learning such patterns, the weights are afterwards changed significantly to learn other majority classes which erases this ``fragile'' knowledge.
+\end{itemize}
 
@@
 
-\begin{figure*}
-\center
-\begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-3x128.pdf}} &
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
-\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-3x128.pdf}} &
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/conf-acc-15-12x64.pdf}}
-\end{tabular}
-\caption{Confidence threshold effect on accuracy and number of needed layers.}
-\label{fig:confidence}
-\end{figure*}
-
-\begin{figure*}
-\centering
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass17.pdf}}
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.98\columnwidth]{figures/multiclass15.pdf}}
-\caption{Accuracy plot multiclass.}
-\end{figure*}
+Training of weights on a given layer that are related to the output neuron can also be achieved by considering only the loss of that particular layer/output which just turns out to be a pure logistic regression between the $p^{th}$ layer and the $(p+1)^{th}$ layer output neuron.
+Nevertheless, due to the difficulty of the problem, constructing a linear combination of logits with adjustable weights provides limited transformations, and therefore inaccurate predictions, because the loss never completely converges.
+Back-propagating all losses at once is thus important and helps to update the weights of the network at the same time as adjusting the weights of the output neurons, thereby allowing the network to make accurate predictions.
 

### 2020-07-22 learn.py

@@ -15,6 +15,7 @@ import os
 import pickle
 import gzip
 import copy
+import itertools
 
 import sklearn
 from sklearn.linear_model import LogisticRegression
@@ -34,6 +35,8 @@ import io
 from sklearn.preprocessing import LabelBinarizer
 from tqdm import tqdm
 from matplotlib import cm
+
+from sklearn.model_selection import StratifiedKFold
 
 # TODOs: unify the binary code
 #		 add attack mapping json for better plotting of long names
@@ -76,11 +79,15 @@ class OurDataset(Dataset):
 	def __len__(self):
 		return self.data.shape[0]
 
-def get_nth_split(dataset, n_fold, index):
-	dataset_size = len(dataset)
-	indices = list(range(dataset_size))
-	bottom, top = int(math.floor(float(dataset_size)*index/n_fold)), int(math.floor(float(dataset_size)*(index+1)/n_fold))
-	train_indices, test_indices = indices[0:bottom]+indices[top:], indices[bottom:top]
+def get_nth_split(dataset, n_fold, index, stratify=True):
+	if stratify:
+		train_indices, test_indices = next(itertools.islice(StratifiedKFold(n_splits = n_fold).split(np.empty(len(attack_vector)), y = attack_vector), index, None))
+	else:
+		# old way, random
+		dataset_size = len(dataset)
+		indices = list(range(dataset_size))
+		bottom, top = int(math.floor(float(dataset_size)*index/n_fold)), int(math.floor(float(dataset_size)*(index+1)/n_fold))
+		train_indices, test_indices = indices[0:bottom]+indices[top:], indices[bottom:top]
 	return train_indices[:opt.maxSize], test_indices[:opt.maxSize]
 
 def make_net(n_input, n_output, n_layers, layer_size):
@@ -280,7 +287,7 @@ def predict_eager_nn():
 	if opt.multiclass:
 		multiclass_eager(test_indices, evaluate=True)
 	else:
-		create_binary_prediction_eager_original(test_indices)
+		create_binary_prediction_eager(test_indices)
 
 def create_binary_plot_eager(test_indices):
 	""" Create confidence-accuracy plot for binary """
@@ -358,7 +365,7 @@ def create_binary_plot_eager(test_indices):
 	plt.rc('axes', axisbelow=True)
 
[…]

### 2020-07-23 paper/paper.tex
finalizing the paper, cosmetic changes to the plotting code

@@
 \author{\IEEEauthorblockN{Fares Meghdouri, Maximilian Bachl and Tanja Zseby}
-\IEEEauthorblockA{Technische Universität Wien\\
+\IEEEauthorblockA{TU Wien\\
 Vienna, Austria\\
@@
 \glspl{fcnn} have been the core of most state-of-the-art \gls{ml} applications in recent years and also have been widely used for \glspl{ids}.
-Experiments results from the last years show that generally deeper neural networks with more layers perform better than shallow models.
+Experimental results from the last years show that generally deeper neural networks with more layers perform better than shallow models.
 Nonetheless, with the growing number of layers, obtaining fast predictions with less resources has become a difficult task despite the use of special hardware such as GPUs.
@@
 We evaluate our proposal with two different network intrusion detection datasets.
-Results suggest that it is possible to obtain comparable accuracies to simple \glspl{fcnn} without evaluating all layers for all samples, thus obtaining early predictions and saving energy and computational efforts.
-\todo{Max: I guess you mean Fully-Connected Neural Networks (FCNNs) when you talk about FNNs? Because FNN only means that the neural network has no cycles. So also a CNN is an FNN.}
-\todo{Fares: The idea should work with any feedforward network. But since we don't experiment with CNNs, I'll replace it now with FCNN.}
+Results suggest that it is possible to obtain comparable accuracies to simple \glspl{fcnn} without evaluating all layers for the majority of samples, thus obtaining early predictions and saving energy and computational efforts.
+%\todo{Max: I guess you mean Fully-Connected Neural Networks (FCNNs) when you talk about FNNs? Because FNN only means that the neural network has no cycles. So also a CNN is an FNN.}
+%\todo{Fares: The idea should work with any feedforward network. But since we don't experiment with CNNs, I'll replace it now with FCNN.}
 \end{abstract}
 
-%\begin{IEEEkeywords}
-%
-%\end{IEEEkeywords}
+\begin{IEEEkeywords}
+Fully Connected Neural Networks, Intrusion Detection, Early Predictions.
+\end{IEEEkeywords}
 
@@
 Many state-of-the-art architectures currently used in different applications are effectively based on simple \gls{fcnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable because it results in better learning performance \cite{eldan_power_2016}.
+Experiments show that the more layers (and neurons) the network includes, the more desirable it results in better learning performance \cite{eldan_power_2016}.
 
@@
 Traditional \gls{fcnn} are based on simple logistic regression units that combine multiple inputs multiplied by a set of weights and passed through an activation function (e.g.
-sigmoid) to obtain a scalar.
+sigmoid \cite{noauthor_sigmoid_2020}) to obtain a scalar which for the sake of consistency we call a prediction.
 The weights are adjusted to give the closest possible output to a ground-truth.
 As far as neural networks are concerned, logistic regression units are stacked together to create layers and then layers are stacked together again to create the network itself.
-Figure \ref{fig:fcnn} demonstrates a basic \gls{fcnn} that utilizes the input vector $\textbf{x}^{T} = \lbrace x_{1}, x_{2} ...
-x_{m} \rbrace$ and makes a prediction.
+\autoref{fig:fcnn} demonstrates a basic \gls{fcnn} that takes an input vector $\textbf{x}^{T} = \lbrace x_{1}, x_{2} ...
+x_{m} \rbrace$ and generates a prediction.
 Adjusting weights in such situations is a non-trivial problem since the weights of each layer may have an impact on the following layers.
@@
 gradient descent learning) while achieving the same goal: reduce the "gap" (loss) between predictions and actual desired output.
-Larger networks consist of more neurons and thus more weights, allowing more data patterns to be learned.
+In practice, larger networks consist of more neurons and thus more weights, allowing more data patterns to be learned.
 
 % problem with large networks
-In terms of efficiency, two problems arise when using an extremely deep neural networks: (1) The loss during training needs more time to converge and (2) prediction time is proportional to the number of layers.
-Even though most modern hardware is able to output predictions without significant delay, prediction speed is still relevant for learning problems, in which many predictions are required per unit of time and delay is important.
+In terms of efficiency, two problems arise when using an extremely deep neural networks: (1) The loss during training needs more time to converge and (2) the time necessary to make a prediction is proportional to the number of layers.
+Even though most modern hardware is able to compute predictions without significant delay, prediction speed is still relevant for learning problems, in which many predictions are required per unit of time and delay is important.
 This is, for example, relevant in the context of object detection in computer vision or also network intrusion detection.
@@
 %EagerNet
-To solve this, we propose a new architecture that, on the one hand, uses the entire network capacity for learning and, on the other hand, stops the forward-pass and makes predictions as soon as confidence reaches a certain threshold.
-In other words, it makes it possible to avoid evaluating the entire network with all layers and to stop the neural network evaluation at a particular layer if the achieved confidence of the neural network is high enough.
-The proposed approach therefore allows, where possible, the reduction of computing resources and energy usage while achieving performance comparable to that of a full forward-pass.
-
-% talk about network traffic
+To solve this, we propose a new architecture that, on the one hand, uses the entire network capacity for learning network intrusions and, on the other hand, stops the forward-pass and makes predictions as soon as confidence reaches a certain threshold.
+In other words, it makes it possible to avoid evaluating the entire network with all layers for some samples and to stop the neural network evaluation at a particular layer if the achieved confidence at that point is high enough.
+We call different \glspl{fcnn} architectures that uses the same proposed approach \emph{EagerNets} (Eager Stopping Networks).
+EagerNets therefore allow, where possible, the reduction of computing resources and energy usage while achieving detection performance comparable to that of a full forward-pass.
 
 % conclusion
-\todo{Max: For such a short paper it might not be necessary to give an overview?}
-\todo{Fares: I now added just a sentence}
+%\todo{Max: For such a short paper it might not be necessary to give an overview?}
+%\todo{Fares: I now added just a sentence}
 In the following sections, we present the existing related work, introduce the concept behind EagerNet and finally show and analyze the evaluation results based on two network intrusion detection datasets.
@@
 While many of the previously proposed works report good results, they usually require advanced math knowledge to be implemented, which hinders their deployment in real-world scenarios.
-Also, they commonly include components such as additional layers, which we show are not necessary.
+These also often contain components such as additional output layers, which add computational overheads that are shown not to be necessary.
 
@@
 
-\todo{Max: Add missing references}
-\todo{Fares: Done.}
+%\todo{Max: Add missing references}
+%\todo{Fares: Done.}
 
@@
 Many of the works proposed \gls{ids} architectures based on well-known supervised techniques \cite{buczak_survey_2016,berman_survey_2019}.
-The idea of training a classifier on pre-stored attack patterns and using the same classifier to recognize similar patterns is commonly used and remarkable success has so far been achieved \cite{vigneswaran_evaluating_2018,ferrag_deep_2020,almseidin_evaluation_2017}.
-In particular, \glspl{fcnn} are commonly used whenever large amounts of data are available.
-In order to present network data to such architectures, a number of traffic representation families have been proposed depending on the application.
-However, the statistical representation of flow characteristics is still the most widely used one.
-In this work, we define the network flow to be the exchange of packets between two end-hosts.
+The concept of having a classifier trained on pre-stored attack patterns and using it to predict similar behaviours is commonly used whereby remarkable success has so far been achieved \cite{vigneswaran_evaluating_2018,ferrag_deep_2020,almseidin_evaluation_2017}.
+In particular, \glspl{fcnn} are commonly utilized whenever large amounts of network traffic are available.
+In order to present network data to such architectures, a number of traffic representations have been proposed depending on the application.
+The statistical representation of flow characteristics is so far the most widely used.
+In this work, we define a network flow to be the exchange of packets between two end-hosts.
 Packets can be identified and aggregated using the five-tuple key: \emph{sourceIP}, \emph{destinationIP}, \emph{sourceTransportPort}, \emph{destinationTransportPort} and \emph{protocolIdentifier}.
-Thereafter, packet features are extracted and the statistical combinations are computed.
-Table \ref{tab:features} shows the CAIA \cite{williams_preliminary_2006} network traffic representation.
+Thereafter, packet header features are extracted and statistical combinations are computed.
+\autoref{tab:features} shows the CAIA \cite{williams_preliminary_2006} network traffic representation.
 It consists of 12 features, 7 of which are measured in both directions and 2 of which are expanded using four statistical combinations (i.e.
@@
 We use CAIA for all experiments in this paper and set an observation timeout of 1,800 seconds, after which we terminate flows.
-Each flow is considered to be one sample and has two sets of labels: a binary (attack/benign) label and also the attack family (one binary label for each attack family).
-The goal is that after a training process, the trained \gls{ids} can correctly classify new flows and also determine what kind of attack family they belong to.
+Each flow is considered to be one data sample and has two sets of labels: a binary (attack/benign) label and also the attack family (one binary label for each attack family using one-hot-encoding).
+The ultimate goal is that after a training process, the trained \gls{ids} can correctly classify new flows and also determine what kind of attack family they belong to.
 
@@
 
-\todo{Max: $X$ is generally for matrices as far as I know while $x$ is for vectors.}
-\todo{Fares: actually $\textbf{x}$ is used for vectors, I forgot to change it because it was consitent to what I wrote in the introduction. Now it's better.}
+%\todo{Max: $X$ is generally for matrices as far as I know while $x$ is for vectors.}
+%\todo{Fares: actually $\textbf{x}$ is used for vectors, I forgot to change it because it was consitent to what I wrote in the introduction. Now it's better.}
 
-Because some network attack families are simple to detect compared to others we modify the standard \gls{fcnn} architecture so that for easily classifiable samples the network doesn't have to evaluate all layers.
-Our novel architecture is based on the assumption that deep neural networks that with more layers can learn increasingly complex functions which are only required for the classification of some particular notoriously hard-to-classify samples.
+Because some network attack families are simpler to detect compared to others we modify the standard \gls{fcnn} architecture so that for easily classifiable samples the network doesn't have to evaluate all layers.
+Our novel architecture is based on the assumption that deep neural networks with more layers can learn increasingly complex functions which in turn are only required for the classification of some particular notoriously hard-to-classify samples.
 As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for direct predictions at each layer.
-The EagerNet architecture is shown in Figure \ref{fig:eagerNet_forward}.
+The EagerNet architecture is shown in \autoref{fig:eagerNet_forward}.
 Initial neurons are shown in gray and output neurons (per layer) are shown in green.
-Once a flow is observed and features are extracted, a sample $\textbf{x}$ is fed to the network via the input layer.
-The network proceeds evaluating the neural network layer by layer and yields a prediction confidence value at each layer.
+Once a network flow is observed and features are extracted, a data sample $\textbf{x}$ is fed to the network via the input layer.
+The neural network proceeds evaluating the layers one by one and yields a prediction confidence value at each layer.
 We define the confidence as the value that represents how sure the network is about a certain decision or sample belonging to a certain class.
-It is a number between 0 and 1 that is obtained by applying the sigmoid function \cite{noauthor_sigmoid_2020} on the output neuron.
-By looking at the confidence confidence and we can determine whether additional layers should be processed or if the currently obtained confidence is high enough.
+It is a number between 0.5 and 1 that is obtained by applying the sigmoid function \cite{noauthor_sigmoid_2020} on the output neuron and computing how far the result is from the farthest label (e.g.
+if the prediction scalar of a sample is 0.2, the farthest label is 1 and therefore a confidence of 0.8 that the sample belongs to class 0).
+By looking at the confidence the neural network determines whether additional layers should be processed or if the currently obtained confidence is high enough.
 This methodology ensures that simpler samples are classified early in the forward-pass and that more complicated samples are passed into deeper layers.
-In the \autoref{subsubsec:confidence-speed-tradeoff}, we show that EagerNet reduces resource consumption on average, while achieving comparable performance compared to normal \gls{fcnn}.
+In the section \ref{subsubsec:confidence-speed-tradeoff}, we show that EagerNet reduces resource consumption on average, while achieving comparable performance compared to normal \gls{fcnn}.
 
@@
 During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there is can be no overlap between classes.
-\autoref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function  $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
+\autoref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function \cite{noauthor_softmax_2020} $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
 
@@
 In this research, and since we have multiple outputs; one at each layer, we aggregate the losses and back propagate them.
-Fig \ref{fig:eagerNet_backward} demonstrates how losses propagate back from the respective layer only.
+\autoref{fig:eagerNet_backward} demonstrates how losses propagate back from the respective layer only.
 As a consequence, the weights of the last layer are affected only by the loss of the last output.
@@
 
-We show the effect of the weight distributions in Table \ref{tab:weights} and discuss results in Section \ref{evaluation_and_discussion}.
+We show the effect of the weight distributions in \autoref{tab:weights} and discuss results in \autoref{evaluation_and_discussion}.
 
@@
 \label{evaluation_and_discussion}
-We discuss in this section different experimental results and observations.
-Overall, EagerNet shows promising performance comparable to \glspl{fcnn} with the ability to save resources.
-All experiments are reproducible and the code is available in our repository\footnote{\url{github.com/CN-TU/ids-backdoor/tree/eager}}.
+In this section we discuss the different experimental results and observations.
+Overall, EagerNet shows promising performance that are comparable to \glspl{fcnn} with the ability to save resources.
+The interested reader can reproduce experiments with the provided sources in our repository\footnote{\url{github.com/CN-TU/ids-backdoor/tree/eager}}.
 
@@
 
-\begin{figure}
+For our experiments, we use two intrusion detection datasets.
+CICIDS2017 \cite{sharafaldin_toward_2018} which consists of 14 network attack classes in addition to realistic user profiles.
+UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic again contains 9 attack families.
+As aforementioned, we represent network data utilizing the CAIA feature vector and extract 2,317,922 and 2,065,224 flows, respectively, from CICIDS2017 and UNSW-NB15 (see \autoref{fig:datasets}).
+For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
+We use a split of $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
+
+\begin{figure}[H]
   \centering
@@
 
-For our experiments, we use two intrusion detection datasets.
-CICIDS2017 \cite{sharafaldin_toward_2018} which consists of 14 network attack classes in addition to realistic user profiles.
-UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic again contains 9 attack families.
-As mentioned above, we represent both datasets using the CAIA feature set and extract 2.317.922 and 2.065.224 flows, respectively, from CICIDS2017 and UNSW-NB15 (see \autoref{fig:datasets}).
-For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
-We use a split of $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
-
 \subsection{Performance}
@@
 \subsubsection{Weighted loss}
-In Table \ref{tab:weights} we gather results from different networks trained using different numbers of hidden layers and weight distributions with 128 neurons per layer.
-In addition, we compare the results using four metrics (Accuracy, Precision, Recall and Youden's J) and show the performance of the last layer (hence raising the confidence level to the maximum).
-The CICIDS2017 indices \todo{Max: Indices?
+In \autoref{tab:weights} we gather results from different networks trained using different numbers of hidden layers and weight distributions each with 128 neurons per layer.
+In addition, we compare the results using four metrics (Accuracy, Precision, Recall and Youden's J) and show the performance of the last layer (by setting the confidence level to the maximum).
+The CICIDS2017 indices %\todo{Max: Indices?
 Fares: it's commun to refer to acc, prec, recall etc.
-as indices/scores} show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
+as indices/scores}
+show that the weight distribution and the architecture itself (number of layers) have little effect on performance.
 In fact, since the last layer is always trained with one loss regardless of the weight distribution, only the first few layers are affected by the combined-loss learning procedure.
-Nonetheless, since the accuracy did not improve when increasing the number of layers, that implies that most network patterns are already separable after the few first layers.
-The results of UNSW-NB15 show a slightly different behavior \todo{Max: Sure?
+Nonetheless, since the accuracy did not improve when increasing the number of layers, that implies that most network patterns are already separable after the few first layers which makes the classification task for CICIDS2017 manageable even for shallow neural networks.
+Results of UNSW-NB15 show a slightly different behavior.
+%\todo{Max: Sure?
 All the results look the same?
-Fares: slightly better}.
-The uniform weighting of losses gives the overall best scores and, in addition, the deeper architectures outperform the shallow ones.
-This implies that the network patterns in this dataset are complex than in the previous dataset and that more non-linearities (layers) are required to find the optimal mapping function.
+Fares: slightly better}
+The uniform weighting of losses gives best scores on average and, in addition, deeper architectures outperform the shallow ones.
+This implies that the network patterns in this specific dataset are complex than in the previous one therefore, requiring more non-linearities (layers) to find the optimal \emph{input:prediction} mapping function.
 
@@
 \midrule
-\multirow{6}{*}{\rotatebox{90}{CIC-IDS17}} & \multirow{2}{*}{Uniform} & 5 & 0.996 & 0.996 & 0.990 & 0.989 \\
+\multirow{6}{*}{\rotatebox{90}{CICIDS2017}} & \multirow{2}{*}{Uniform} & 5 & 0.996 & 0.996 & 0.990 & 0.989 \\
  & & 3 & 0.997 & 0.997 & 0.990 & 0.990 \\
@@
 
-\subsubsection{Confidence-Speed tradeoff}
+\subsubsection{Accuracy-Speed tradeoff}
 \label{subsubsec:confidence-speed-tradeoff}
@@
 \begin{tabular}{cccc}
-\subfloat[CIC-IDS17 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-17-3x128.pdf}} &
-\subfloat[CIC-IDS17 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
+\subfloat[CICIDS2017 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-17-3x128.pdf}} &
+\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-17-12x64.pdf}}\\
 \subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.78\columnwidth]{figures/conf-acc-15-3x128.pdf}} &
@@
 
-Figure \ref{fig:confidence} shows the confidence-speed balance in addition to the accuracy achieved for four different networks and for both datasets.
-The networks contain 128 neurons per layer and a total of 3 or 12 layers.
-The confidence threshold is shown on the horizontal axis, the average number of layers used over the entire dataset and the average accuracy achieved are both shown on the vertical axis.
-Curves show consistent results, whereby the network always uses few layers for the majority of samples and only continues to evaluate further layers if the confidence level required is too high or the sample is noisy and it is difficult to make decisions in early layers.
-There are also two phenomena observed:
+\autoref{fig:confidence} shows the accuracy-speed balance in terms of confidence threshold.
+Four different networks for both datasets are shown.
+We experiments with two networks: 128 neurons by 3 layers and 64 neurons by 12 layers.
+The confidence threshold is shown on the horizontal axis, the average number of layers used over all samples and the average accuracy achieved are both shown on the vertical axis.
+Curves show promising and consistent results, whereby the network always uses few layers for the majority of samples and only continues to evaluate further layers if the confidence level required is too high or the sample is noisy thus it is difficult to make correct decisions in early layers.
+We observe moreover two phenomena:
 \begin{itemize}
-\item The accuracy of CICIDS2017 increases almost linearly with respect to confidence and reaches its maximum when all layers are used.
-\item The accuracy of UNSW-NB15 is stable when the confidence level is increased until $\approx 0.98$ where it leaps noticeably when a further layer is evaluated.
-This suggests that some patterns are ``learnable'' only on that specific layer.
+\item The accuracy of CICIDS2017 increases almost linearly with respect to confidence and reaches its maximum when all layers are evaluated.
+\item The accuracy of UNSW-NB15 is stable when the confidence level is increased until $\approx 0.98$ where it leaps noticeably once a further layer is evaluated.
+This suggests that some patterns were ``learned'' only on that specific layer and hence, respective samples suddenly were correctly classified.
 \end{itemize}
-Overall, EagerNet shows that it is possible to trade a tiny percentage of accuracy in order to save a significant amount of resources.
-In \autoref{subsubsec:comparability_to_fnns}, we see that, in order to achieve comparable accuracy to the traditional \gls{fcnn}, the previous trade is possible and therefore also the possibility to reduce costs.
+Overall, EagerNet confirms that it is possible to trade a tiny percentage of accuracy in order to save a significant amount of resources.
+In section \ref{subsubsec:comparability_to_fnns}, we see that, it is feasible to achieve comparable accuracies to traditional \gls{fcnn} allowing the reduction of costs.
 
-\todo{Max: Confidence seems to be multiply defined. Fares: I don't understand}
+%\todo{Max: Confidence seems to be multiply defined. Fares: I don't understand}
 
@@
 \label{subsubsec:comparability_to_fnns}
-\todo{Max: Don't understand. Fares: Now added}
+%\todo{Max: Don't understand. Fares: Now added}
 
-In \autoref{tab:comparability}, we show the F1 score obtained at the last layer of three different architectures: Simple \gls{fcnn}, \gls{fcnn} with logistic output neuron per layer and EagerNet.
-\todo{Fares: continuation}
+In \autoref{tab:comparability}, we show the F1 score obtained at the last layer of two architectures: Simple \gls{fcnn} and EagerNet with the same number of layers and neurons.
+We use the F1 score to reduce the class imbalance bias.
+Results suggest that on average, EagerNet achieves similar scores to a conventional \gls{fcnn} at the last layer across all variants and datasets.
+It is worth mentioning, however, that EagerNet is used in this case as a conventional neural network and predictions are taken only from the last layer.
+Using predictions from various layers by setting a confidence threshold might lead to higher accuracy if middle layers are optimized to predict certain classes.
 
@@
 \toprule
-\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{fcnn}} & \textbf{EagerNet (last)} \\
+\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{fcnn}} & \textbf{EagerNet} \\
 \midrule
-\multirow{2}{*}{CIC-IDS17} & Binary & 0.989 & 0.993 \\
- & Multiclass & 0.920 & 0.979 \\
+\multirow{2}{*}{CICIDS2017} & Binary & 0.989 & 0.993 \\
+ & Multiclass & 0.979 & 0.920 \\
 \midrule
 \multirow{2}{*}{UNSW-NB15} & Binary & 0.919 & 0.908 \\
- & Multiclass & 0.880 & 0.882 \\
+ & Multiclass & 0.882 & 0.880 \\
 \midrule
@@
 \begin{tabular}{cccc}
-\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
+\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons). \todo{Fares: Model is still running to generate a better figure.}]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
 \subfloat[CICIDS2017 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.35\columnwidth]{figures/multiclass17-3.pdf}} 
@@
 
-\todo{Max: Don't understand. Fares: Now added}
-In order to understand how EagerNet determines which layer the best to stop evaluation, we use the multiclass variant to conduct an explanatory analysis.
+%\todo{Max: Don't understand. Fares: Now added}
+In order to understand how EagerNet determines at which layer the best to stop evaluation, we use the multiclass architecture and conduct an explanatory analysis.
 In \autoref{fig:multiclass}, we show the accuracy of the prediction obtained by different networks per layer and attack family.
-We observe two phenomena:
+We obtain two conclusions:
 \begin{itemize}
 \item  On average, the accuracy of predictions increases the deeper we get into the network.
-It was somehow expected, as more layers allow for more abstraction of the input space and thus better separation.
-However, only a few first layers are needed to obtain maximum accuracy for each attack family.
-This implies that the EagerNet architecture significantly decreases resources when used for intrusion detection applications.
-\item \todo{Fares: recheck this} Some attacks families show extremely poor accuracy.
-This is due to two reasons: (1) too few samples from a category are mostly ignored by the network to reduce over-fitting and (2) If a specific layer focuses on learning such patterns, the weights are afterwards changed significantly to learn other majority classes which erases this ``fragile'' knowledge.
+This was somehow expected, as more layers allow for more abstraction of the input space and thus better separation of instances.
+However, only few layers are needed to obtain the maximum accuracy per attack family.
+This implies that the EagerNet architecture significantly decreases resources when used for intrusion detection applications keeping the accuracy at its best.
+\item  On the other hand, some attack families (Analysis, Worms, Web XSS etc.) show extremely poor accuracy.
+This is due to two reasons: (1) too few samples of a category are mostly ignored by the network to reduce over-fitting and (2) The loss optimized to learn such patterns at a specific layer is over-written by the loss that is optimized to learn the majority classes hence erasing this ``fragile'' knowledge.
 \end{itemize}
@@
 \subsubsection{Advantage of backpropagating the losses of all outputs until the beginning}
-\todo{Max: Don't know if this has been already covered elsewhere?}
+%\todo{Max: Don't know if this has been already covered elsewhere?}
 
-Training of weights on a given layer that are related to the output neuron can also be achieved by considering only the loss of that particular layer/output which just turns out to be a pure logistic regression between the $p^{th}$ layer and the $(p+1)^{th}$ layer output neuron.
-Nevertheless, due to the difficulty of the problem, constructing a linear combination of logits with adjustable weights provides limited transformations, and therefore inaccurate predictions, because the loss never completely converges.
-Back-propagating all losses at once is thus important and helps to update the weights of the network at the same time as adjusting the weights of the output neurons, thereby allowing the network to make accurate predictions.
+Training the weights connecting a layer to the next output neuron can also be achieved by considering only the loss of that particular layer/output which turns out to be a pure logistic regression between the $p^{th}$ layer and the $(p+1)^{th}$ layer's output neuron.
+Nevertheless, due to the difficulty of the problem, constructing a linear combination of intermediate logits with adjustable weights provides limited mappings, and therefore inaccurate predictions, because the loss never completely converges.
+Back-propagating all losses at once is thus important and helps to update the weights of the network at the same time as adjusting the weights of the output neurons, thereby allowing the neural network to distribute the knowledge across all weights and, make accurate predictions at each layer.
 
 \section{Conclusion}
+
+This paper explores the early termination of the viability of the \glspl{fcnn} test.
+We propose EagerNet, an architecture that allows predictions to be made as soon as the neural network is sufficiently confident, saving energy and resources and making it possible to implement similar architectures in real-time applications where prediction speed is relevant.
+We evaluate our approach using two intrusion detection datasets and obtain satisfactory results indicating the possibility of achieving comparable evaluation scores to traditional \glspl{fcnn} yet using only a segment of the neural network, as most network traffic flows are simpler to classify and only a minority need deeper propagation.
+Furthermore, we show that by setting a confidence threshold during deployment, it becomes possible to trade accuracy for resources usage and energy consumption.
+The EagerNet architecture is thus characterized by its simplicity and effectiveness, making it an ideal solution for potential \gls{fcnn}-based \glspl{ids} whenever resources are of primary concern.
 

### 2020-07-23 learn.py

@@ -445,21 +445,24 @@ def multiclass_eager(test_indices, evaluate=False):
 			cr = confusion_matrix(y_list, y_pred_list[i], list(range(n_classes)))
 			accuracies[:,i] = cr.diagonal()/cr.sum(axis=1)
 
-		#with open(f"labels", "wb") as fp:
-		#	pickle.dump(y_list, fp)
-		#with open(f"predictions", "wb") as fp:
-		#	pickle.dump(y_pred_list, fp)
+		if opt.saveResults:
+			with open(f"{opt.net}_labels", "wb") as fp:
+				pickle.dump(y_list, fp)
+			with open(f"{opt.net}_predictions", "wb") as fp:
+				pickle.dump(y_pred_list, fp)
 
 		accuracies = np.nan_to_num(accuracies)
-		accuracies = np.take(accuracies, np.sum(accuracies, axis = 1).argsort(), axis=0) # sort
+		order = np.sum(accuracies, axis = 1).argsort()
+		accuracies = np.take(accuracies, order, axis=0) # sort
+
 		if opt.saveResults:
 			np.save('accuracies_' + opt.net, accuracies)
 
 		plt.figure(figsize=(16,8))
-		plt.imshow(accuracies, cmap=cm.coolwarm, interpolation='nearest', )
+		plt.imshow(accuracies, cmap=cm.coolwarm, interpolation='nearest')
 		clb = plt.colorbar()
 		clb.set_label('Accuracy', rotation=90, fontsize=30)
-		plt.yticks(list(range(n_classes)), labels=dataset.attacks, fontsize=20)
+		plt.yticks(list(range(n_classes)), labels=[dataset.attacks[i] for i in order], fontsize=20)
 		plt.xlabel('Layers', fontsize=30)
 		plt.ylabel('Attack Families', fontsize=30)
 		plt.xticks(list(range(n_outputs)), list(range(1,n_outputs+1)), fontsize=20)

### 2020-07-23 paper/paper.tex
fix typos

@@
 \subfloat[CICIDS2017 (12 layers $\times$ 64 neurons). \todo{Fares: Model is still running to generate a better figure.}]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
-\subfloat[CICIDS2017 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.35\columnwidth]{figures/multiclass17-3.pdf}} 
+\subfloat[CICIDS2017 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.35\columnwidth]{figures/multiclass17-3.pdf}} 
 \subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.68\columnwidth]{figures/multiclass15.pdf}} &
-\subfloat[UNSW-NB15 (3 layers $\times$ 128 neurons).]{\includegraphics[width=0.38\columnwidth]{figures/multiclass15-3.pdf}}
+\subfloat[UNSW-NB15 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.38\columnwidth]{figures/multiclass15-3.pdf}}
 \end{tabular}
-\caption{Accuracy achieved on a test set per layer and attack family.
+\caption{Accuracy achieved on a test set per layer and attack family (sorted vertically).
 Some attacks have poor accuracy because only few samples are observed (refer to \autoref{fig:datasets}).
@@
 This implies that the EagerNet architecture significantly decreases resources when used for intrusion detection applications keeping the accuracy at its best.
-\item  On the other hand, some attack families (Analysis, Worms, Web XSS etc.) show extremely poor accuracy.
+\item  On the other hand, some attack families (Analysis, Worms, Web XSS, DDoS Heart-bleed etc.) show extremely poor accuracy.
 This is due to two reasons: (1) too few samples of a category are mostly ignored by the network to reduce over-fitting and (2) The loss optimized to learn such patterns at a specific layer is over-written by the loss that is optimized to learn the majority classes hence erasing this ``fragile'' knowledge.

### 2020-07-23 paper/paper.tex
Small changes

@@
 
-Training the weights connecting a layer to the next output neuron can also be achieved by considering only the loss of that particular layer/output which turns out to be a pure logistic regression between the $p^{th}$ layer and the $(p+1)^{th}$ layer's output neuron.
-Nevertheless, due to the difficulty of the problem, constructing a linear combination of intermediate logits with adjustable weights provides limited mappings, and therefore inaccurate predictions, because the loss never completely converges.
-Back-propagating all losses at once is thus important and helps to update the weights of the network at the same time as adjusting the weights of the output neurons, thereby allowing the neural network to distribute the knowledge across all weights and, make accurate predictions at each layer.
+Instead of backpropagating each loss through all layers, it is also possible to only backpropagate each loss of the intermediate layers only to the previous layer but not until the beginning.
+Only the last layer's output's loss would then be backpropagated until the beginning but the other intermediate output's losses only one step.
+One benefit of this would be to save computational resources during training.
+Our results showed that this method results in worse accuracy for the intermediate layers' outputs.
+Backpropagating all losses at once is thus important and helps to update the weights of the network at the same time as adjusting the weights of the output neurons, thereby allowing the neural network to distribute the knowledge across all weights and make accurate predictions at each layer.
 
@@
 
-This paper explores the early termination of the viability of the \glspl{fcnn} test.
+This paper explores if it is viable to terminate the evaluation of \glspl{fcnn} before the final layer to save computational resources.
 We propose EagerNet, an architecture that allows predictions to be made as soon as the neural network is sufficiently confident, saving energy and resources and making it possible to implement similar architectures in real-time applications where prediction speed is relevant.
-We evaluate our approach using two intrusion detection datasets and obtain satisfactory results indicating the possibility of achieving comparable evaluation scores to traditional \glspl{fcnn} yet using only a segment of the neural network, as most network traffic flows are simpler to classify and only a minority need deeper propagation.
+We evaluate our approach using two intrusion detection datasets and obtain satisfactory results indicating the possibility of achieving comparable evaluation scores to traditional \glspl{fcnn} yet using only a segment of the neural network, as most network traffic flows are simpler to classify and only a minority need deeper propagation through the neural network.
 Furthermore, we show that by setting a confidence threshold during deployment, it becomes possible to trade accuracy for resources usage and energy consumption.

### 2020-07-26 paper/paper.tex
improvements based on feedback

@@
 % what are fcnn
-\glspl{fcnn} have gained remarkable attention in recent years due to the availability of data and computational resources for research.
+\glspl{fcnn} have gained remarkable attention in various domains in recent years due to the availability of data and computational resources for research.
 Many state-of-the-art architectures currently used in different applications are effectively based on simple \gls{fcnn} architectures.
-Experiments show that the more layers (and neurons) the network includes, the more desirable it results in better learning performance \cite{eldan_power_2016}.
+Since the size of a neural network cannot be mathematically defined, experts use rule-of-thumb to define the best architecture that fits their needs.
+Yet, experiments show that the more layers (and neurons) the network includes, the better its learning performance becomes \cite{eldan_power_2016}.
 
@@
 %EagerNet
-To solve this, we propose a new architecture that, on the one hand, uses the entire network capacity for learning network intrusions and, on the other hand, stops the forward-pass and makes predictions as soon as confidence reaches a certain threshold.
+To obtain early predictions, we propose a new architecture that, on the one hand, uses the entire network capacity for learning network intrusions and, on the other hand, stops the forward-pass and makes predictions as soon as confidence reaches a certain threshold.
 In other words, it makes it possible to avoid evaluating the entire network with all layers for some samples and to stop the neural network evaluation at a particular layer if the achieved confidence at that point is high enough.
@@
 
+ \begin{figure}
+  \includegraphics[width=\linewidth]{figures/fcnn_new1.pdf}
+  \caption{Conventional \gls{fcnn} architecture.}
+  \label{fig:fcnn}
+\end{figure}
 % conclusion
@@
 In the following sections, we present the existing related work, introduce the concept behind EagerNet and finally show and analyze the evaluation results based on two network intrusion detection datasets.
-
- \begin{figure}
-  \includegraphics[width=\linewidth]{figures/fcnn_new.pdf}
-  \caption{Conventional \gls{fcnn} architecture.}
-  \label{fig:fcnn}
-\end{figure}
 
@@
 This approach was predominantly explored for classical machine learning methods.
-Two possibilities are to either place all classifiers in a row or in the shape of tree.
+Two possibilities are to either place all classifiers in a row or in the shape of a tree.
 If a tree is used, different classifiers can be used at each step, depending on which features are probably relevant for the sample currently under investigation.
@@
 
-The major difference between all existing works and ours is simplicity.
+The major difference between all existing works and ours is simplicity in terms of implementation complexity and interpretability.
 While many of the previously proposed works report good results, they usually require advanced math knowledge to be implemented, which hinders their deployment in real-world scenarios.
 These also often contain components such as additional output layers, which add computational overheads that are shown not to be necessary.
+
+To the best of our knowledge, our work considers for the first time such an approach for network stream data to obtain early flow classification yet with remarkable performance.
 
@@
 \emph{A} becomes \emph{mean(A)}, \emph{min(A)}, \emph{max(A)} and \emph{stdev(A)}).
-We use CAIA for all experiments in this paper and set an observation timeout of 1,800 seconds, after which we terminate flows.
+We use CAIA for all experiments in this paper and set an observation timeout of 1,800 seconds (slightly longer than the longest attack activity in the dataset), after which we terminate flows.
 Each flow is considered to be one data sample and has two sets of labels: a binary (attack/benign) label and also the attack family (one binary label for each attack family using one-hot-encoding).
@@
 Our novel architecture is based on the assumption that deep neural networks with more layers can learn increasingly complex functions which in turn are only required for the classification of some particular notoriously hard-to-classify samples.
-As a result, we build a network such that an additional set of neurons is connected to each layer (a copy of the output layer), allowing for direct predictions at each layer.
+As a result, we build a network such that an additional set of neurons are connected to each layer (a copy of the output neurons instead of an entire meta-network), allowing for direct predictions at each layer.
 The EagerNet architecture is shown in \autoref{fig:eagerNet_forward}.
@@
 In the section \ref{subsubsec:confidence-speed-tradeoff}, we show that EagerNet reduces resource consumption on average, while achieving comparable performance compared to normal \gls{fcnn}.
+
+\subsubsection{Network Architecture}
+For our experiments, we use a neural network that consists of as many layers as needed with Leaky ReLU activations \cite{maas_rectifier_2013} ($\alpha = 0.1$) after each layer.
+Moreover and to reduce over-fitting, we use neurons dropout ($r=0.2$).
 
@@
 For the multiclass approach, on the other side, there is one output neuron per attack family, and each neuron's output indicates how likely it is, that the currently processed sample belongs to the given attack family.
-During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there is can be no overlap between classes.
+During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there can be no overlap between classes.
 \autoref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function \cite{noauthor_softmax_2020} $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
@@
 In this section we discuss the different experimental results and observations.
-Overall, EagerNet shows promising performance that are comparable to \glspl{fcnn} with the ability to save resources.
+Overall, EagerNet shows promising performance that are comparable to \glspl{fcnn} that always use all layers yet with the ability to save resources.
 The interested reader can reproduce experiments with the provided sources in our repository\footnote{\url{github.com/CN-TU/ids-backdoor/tree/eager}}.
@@
 \subsection{Datasets}
-
-For our experiments, we use two intrusion detection datasets.
-CICIDS2017 \cite{sharafaldin_toward_2018} which consists of 14 network attack classes in addition to realistic user profiles.
-UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic again contains 9 attack families.
-As aforementioned, we represent network data utilizing the CAIA feature vector and extract 2,317,922 and 2,065,224 flows, respectively, from CICIDS2017 and UNSW-NB15 (see \autoref{fig:datasets}).
-For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
-We use a split of $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
 
@@
 
+For our experiments, we use two intrusion detection datasets.
+CICIDS2017 \cite{sharafaldin_toward_2018} which consists of 14 network attack classes in addition to realistic user profiles.
+UNSW-NB15 \cite{moustafa_unsw-nb15_2015} which, in addition to synthetically generated normal traffic again contains 9 attack families.
+As aforementioned, we represent network data utilizing the CAIA feature vector and extract 2,317,922 and 2,065,224 flows, respectively, from CICIDS2017 and UNSW-NB15 (see \autoref{fig:datasets}).
+For pre-processing, we eliminate duplicate instances and \emph{z}-normalize the data.
+We use a split of $\frac{2}{3}$ for training and $\frac{1}{3}$ for evaluation and testing.
+
 \subsection{Performance}
 
+In what follows, we show the performance of the EagerNet architecture taking into account various aspects.
+We stop training all models after 800 epochs to ensure fair comparison and we monitor both training and validation losses to prevent over-fitting.
+
+\subsubsection{Comparability to \glspl{fcnn}}
+\label{subsubsec:comparability_to_fnns}
+%\todo{Max: Don't understand. Fares: Now added}
+
+In \autoref{tab:comparability}, we show the F1 score obtained at the last layer of two architectures: Simple \gls{fcnn} and EagerNet with the same number of layers and neurons.
+We use the F1 score to reduce the class imbalance bias.
+Results suggest that on average, EagerNet achieves similar scores to a conventional \gls{fcnn} at the last layer across all variants and datasets.
+It is worth mentioning, however, that EagerNet is used in this case as a conventional neural network and predictions are taken only from the last layer hence, comparing the influence of our training procedure on the typical workflow of the neural network.
+Using predictions from various layers by setting a confidence threshold might lead to higher accuracy if middle layers are optimized to predict certain classes.
+
+\begin{table}[H]
+
+\centering
+\begin{tabular}{ccrc}
+\toprule
+\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{fcnn}} & \textbf{EagerNet} \\
+\midrule
+\multirow{2}{*}{CICIDS2017} & Binary & 0.989 & 0.993 \\
+ & Multiclass & 0.979 & 0.920 \\
+\midrule
+\multirow{2}{*}{UNSW-NB15} & Binary & 0.919 & 0.908 \\
+ & Multiclass & 0.882 & 0.880 \\
+\midrule
+
+\end{tabular}
+\vspace{1ex}
+
+{\raggedright * All architectures consist of 10 layers $\times$ 64 neurons in addition to input and output layers. \par}
+\caption{F1 scores at the last layer of each architecture.}
+\label{tab:comparability}
+\end{table}
+
 \subsubsection{Weighted loss}
-In \autoref{tab:weights} we gather results from different networks trained using different numbers of hidden layers and weight distributions each with 128 neurons per layer.
+In \autoref{tab:weights} we gather results from different EagerNet networks trained using different numbers of hidden layers and weight distributions each with 128 neurons per layer.
 In addition, we compare the results using four metrics (Accuracy, Precision, Recall and Youden's J) and show the performance of the last layer (by setting the confidence level to the maximum).
@@
 \begin{itemize}
-\item The accuracy of CICIDS2017 increases almost linearly with respect to confidence and reaches its maximum when all layers are evaluated.
-\item The accuracy of UNSW-NB15 is stable when the confidence level is increased until $\approx 0.98$ where it leaps noticeably once a further layer is evaluated.
+\item The accuracy of CICIDS2017 increases almost linearly in sections with respect to confidence and reaches its maximum when all layers are evaluated.
+\item The accuracy of UNSW-NB15 is stable when the confidence level is increased from $\approx 0.62$ to until $\approx 0.98$ where it leaps noticeably once a further layer is evaluated.
 This suggests that some patterns were ``learned'' only on that specific layer and hence, respective samples suddenly were correctly classified.
@@
 
-\subsubsection{Comparability to \glspl{fcnn}}
-\label{subsubsec:comparability_to_fnns}
-%\todo{Max: Don't understand. Fares: Now added}
-
-In \autoref{tab:comparability}, we show the F1 score obtained at the last layer of two architectures: Simple \gls{fcnn} and EagerNet with the same number of layers and neurons.
-We use the F1 score to reduce the class imbalance bias.
-Results suggest that on average, EagerNet achieves similar scores to a conventional \gls{fcnn} at the last layer across all variants and datasets.
-It is worth mentioning, however, that EagerNet is used in this case as a conventional neural network and predictions are taken only from the last layer.
-Using predictions from various layers by setting a confidence threshold might lead to higher accuracy if middle layers are optimized to predict certain classes.
-
-\begin{table}[H]
-
-\centering
-\begin{tabular}{ccrc}
-\toprule
-\textbf{Dataset} & \textbf{Variant*} & \textbf{\gls{fcnn}} & \textbf{EagerNet} \\
-\midrule
-\multirow{2}{*}{CICIDS2017} & Binary & 0.989 & 0.993 \\
- & Multiclass & 0.979 & 0.920 \\
-\midrule
-\multirow{2}{*}{UNSW-NB15} & Binary & 0.919 & 0.908 \\
- & Multiclass & 0.882 & 0.880 \\
-\midrule
-
-\end{tabular}
-\vspace{1ex}
-
-{\raggedright * All architectures consist of 10 layers $\times$ 64 neurons in addition to input and output layers. \par}
-\caption{F1 scores at the last layer of each architecture.}
-\label{tab:comparability}
-\end{table}
-
 \subsubsection{What did the network learn?}
@@
 \begin{tabular}{cccc}
-\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons). \todo{Fares: Model is still running to generate a better figure.}]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
+\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
 \subfloat[CICIDS2017 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.35\columnwidth]{figures/multiclass17-3.pdf}} 
@@
 Some attacks have poor accuracy because only few samples are observed (refer to \autoref{fig:datasets}).
-Forcing the network to learn those samples might lead to over-fitting.}
+Forcing the network to learn those samples might lead to over-fitting.
+Additionally, the effect of over-sizing the neural network immediately appears and allows to apply the EagerNet strategy.}
 \label{fig:multiclass}

### 2020-09-18 paper/paper.tex
Fix minor typos

@@
 % problem with large networks
-In terms of efficiency, two problems arise when using an extremely deep neural networks: (1) The loss during training needs more time to converge and (2) the time necessary to make a prediction is proportional to the number of layers.
+In terms of efficiency, two problems arise when using an extremely deep neural networks: (1) the loss during training needs more time to converge and (2) the time necessary to make a prediction is proportional to the number of layers.
 Even though most modern hardware is able to compute predictions without significant delay, prediction speed is still relevant for learning problems, in which many predictions are required per unit of time and delay is important.
@@
 
-\subsubsection{Binary Vs. Multiclass Classification}
+\subsubsection{Binary vs. Multiclass Classification}
 We experiment with two different architectures.
@@
 During back-propagation, we use a binary cross-entropy loss and average the loss over the batch.
-\autoref{eq:bceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input being $x_{n}$ and the sigmoid activation function $\sigma(\cdot)$.
+\autoref{eq:bceloss}.
+shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input being $x_{n}$ and the sigmoid activation function $\sigma(\cdot)$.
 
@@
 During back-propagation, we use the categorical cross-entropy loss to penalize all outputs together, so there can be no overlap between classes.
-\autoref{eq:cceloss} shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function \cite{noauthor_softmax_2020} $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
+\autoref{eq:cceloss}.
+shows how the loss is determined for a batch of size $N$, with the $n^{th}$ input $x_{n}$ and the softmax activation function \cite{noauthor_softmax_2020} $\phi(.)$, $C$ being the number of classes or output neurons and $i$ is the current class/output.
 
@@
 
-\paragraph{Optimal Confidence Threshold.}
+\paragraph{Optimal Confidence Threshold}
 During deployment, the optimum confidence threshold is set by the user and plays a role in the overall achieved accuracy.

### 2020-09-30 paper/paper.tex
cosmetic changes, recommit Fares and Max changes

@@
 \begin{tabular}{cccc}
-\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.55\columnwidth]{figures/multiclass17.pdf}} &
-\subfloat[CICIDS2017 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.35\columnwidth]{figures/multiclass17-3.pdf}} 
-\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.68\columnwidth]{figures/multiclass15.pdf}} &
-\subfloat[UNSW-NB15 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.38\columnwidth]{figures/multiclass15-3.pdf}}
+\subfloat[CICIDS2017 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.52\columnwidth]{figures/multiclass17.pdf}} &
+\subfloat[CICIDS2017 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.37\columnwidth]{figures/multiclass17-3.pdf}} 
+\subfloat[UNSW-NB15 (12 layers $\times$ 64 neurons).]{\includegraphics[width=0.64\columnwidth]{figures/multiclass15.pdf}} &
+\subfloat[UNSW-NB15 (5 layers $\times$ 128 neurons).]{\includegraphics[width=0.4\columnwidth]{figures/multiclass15-3.pdf}}
 \end{tabular}

### 2020-09-30 learn.py

@@ -38,6 +38,8 @@ from matplotlib import cm
 
 from sklearn.model_selection import StratifiedKFold
 
+import statistics
+
 # TODOs: unify the binary code
 #		 add attack mapping json for better plotting of long names
 
@@ -289,7 +291,7 @@ def predict_eager_nn():
 	else:
 		create_binary_prediction_eager(test_indices)
 
-def create_binary_plot_eager(test_indices):
+def create_binary_plot_eager_(test_indices):
 	""" Create confidence-accuracy plot for binary """
 
 	test_data = torch.utils.data.Subset(dataset, test_indices)
@@ -317,6 +319,7 @@ def create_binary_plot_eager(test_indices):
 
 	already_processed = 0
 	efforts = []
+	efforts_std = []
 	lists[0] = list(all_predictions)
 
 	accuracy_per_list = [sum([item[output_index] for item in l]) for output_index, l in enumerate(lists)]
@@ -346,22 +349,24 @@ def create_binary_plot_eager(test_indices):
 
 		if len(efforts)==0 or current_lowest >= efforts[-1][0]:
 			efforts.append((current_lowest, sum([(output_index+1)*len(lists[output_index]) for output_index in range(n_outputs)])/len(all_predictions), sum(accuracy_per_list)/len(all_predictions)))
+			efforts_std.append((current_lowest, statistics.stdev([(output_index+1)*len(lists[output_index]) for output_index in range(n_outputs)])/5, statistics.stdev(accuracy_per_list)))
 
 	colors = plt.rcParams['axes.prop_cycle'].by_key()['color']
 
 	plt.grid(linestyle='--', linewidth='0.5', color='gray', axis='x')
 	things = []
 	x, y1, y2 = list(zip(*efforts))
+	_, y1err, y2err = list(zip(*efforts))
 	y1_label = "Mean layers needed"
-	things += plt.plot(x, y1, color=colors[0], label=y1_label)
+	things += plt.errorbar(x, y1, yerr=y1err, color=colors[0], label=y1_label, errorevery=100)
 	plt.xlabel("Confidence required to stop further evaluation", size=15)
 	plt.ylabel(y1_label, size=15)
 	plt.twinx()
 	y2_label = "Mean accuracy achieved"
-	things += plt.plot(x, y2, color=colors[1], label=y2_label)
+	things += plt.errorbar(x, y2, color=colors[1], label=y2_label)
 	plt.ylabel(y2_label, size=15)
 	plt.tight_layout()
[…]
